### 2023-02-22 docs/htw_exam.qmd
commit primary files

---
title: HTW EXAM Simulations
code-fold: true
execute: 
  warning: false
  eval: false
---

```{r}
pacman::p_load(tidyverse,data.table,igraph,ggraph,kableExtra)

d <- readRDS('dPrune-01-19-23.rds')

dtest <- d %>% filter(expMode %in% c("test-Nf","test-train-nf")) %>% group_by(id,lowBound) %>% 
  mutate(nBand=n(),band=bandInt,id=factor(id)) %>% group_by(id) %>% mutate(nd=n_distinct(lowBound))
# unique(dtest[dtest$nd==4,]$sbjCode) # 7 in wrong condition
dtest <- dtest %>% group_by(id,lowBound) %>% filter(nBand>=5 & nd==6)
# for any id that has at least 1 nBand >=5, remove all rows with that id. 
dtest <- dtest %>% group_by(id) %>% filter(!id %in% unique(dtest$id[dtest$nBand<5]))

dtestAgg <- dtest %>% group_by(id,condit,catOrder,feedbackType,vb,band,lowBound,highBound,input) %>% mutate(vxCapped=ifelse(vx>1600,1600,vx)) %>%
  summarise(vxMean=mean(vx),devMean=mean(dist),vxMed=median(vx),devMed=median(dist),
            vxMeanCap=mean(vxCapped),.groups = "keep")

# select first row for each id in d, then create histogram for nTrain
#  d  %>% group_by(id) %>% slice(1) %>% ggplot(aes(nTrain)) + geom_histogram() + facet_wrap(~condit)

ds <- d %>% filter(expMode %in% c("train","train-Nf","test-Nf","test-train-nf")) %>% 
filter(!id %in% unique(dtest$id[dtest$nBand<5])) %>% select(id,condit,catOrder,feedbackType,expMode,trial,vb,band,lowBound,highBound,input,vx,dist,vxb)

dst <- ds %>% filter(expMode=="train",catOrder=="orig")

```

```{r}

alm.response <- function(input=1,c) {
input.activation <- exp(-c*(input.layer - input)^2)
input.activation <<- input.activation/sum(input.activation)
#print(length(input.activation)); print(dim(weight.mat))
output.activation <<- weight.mat %*% input.activation
output.probability <<- output.activation/sum(output.activation)
mean.response <<- sum(output.layer * output.probability)
mean.response
}

alm.update <- function(corResp,c,lr){
  fz <- exp(-c*(output.layer - corResp)^2)
  teacherSignal <- (fz - output.activation)*lr
  #print(length(teacherSignal)); print(length(fz))
  wChange <- teacherSignal %*% t(input.activation)
  weight.mat <<- weight.mat + (wChange)
  weight.mat[weight.mat<0]=0
 # weight.mat[weight.mat>1]=1
  weight.mat <<- weight.mat
}

alm.trial <- function(input, corResp,c,lr){
  alm.response(input,c)
  alm.update(corResp,c,lr)
  print(paste0("input=",input,"; corResp=",corResp,"; mean.response=",mean.response))
  mean.response
}

htw_exam_response <- function(input,c){
  # Find the index of the input node with the highest activation
  trainVec = c(0,sort(unique(xt))) # add 0 to the training vector
  nearestTrain <- trainVec[which.min(abs(input - trainVec))]
  aresp <- alm.response(nearestTrain,c)
  #max.index <- which.max(input.activation)
  xUnder = ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver = ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])

  mUnder <- alm.response(xUnder,c)
  mOver <- alm.response(xOver,c)

   exam.output = round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (input - nearestTrain), 3)
  # Determine the input nodes and associated weights for computing the slope
  exam.output
}

# simulation function
htw_alm_sim <- function(dat, c, lr,testRange=seq(0,100,.5)){
input.layer <<- matrix(seq(1,7,1) ) # half step units for inputs, from 0 to 100
output.layer <<- matrix(seq(50,1600,50)) # single step units for outputs, from 0 to 250
weight.mat <<- matrix(0.00001,nrow=length(output.layer),ncol=length(input.layer )) # weights initialized to 0 (as in Delosh 1997)
xt<<-dat$x
# run training
st <- map2_dbl(dat$x, dat$y, ~alm.trial(.x,.y,c,lr))
# append training data to the data frame
dat <- dat %>% mutate(almResp = st)
# test final weights on the full range of inputs
almResp <- map_dbl(testRange, ~alm.response(.x,c)) 
examResp <- map_dbl(testRange,~exam.response(.x,c))

return(list(d=dat, alm=almResp,exam=examResp,wm=weight.mat,c=c,lr=lr)) # final weightmat is probs incorrect for all but last
}

simOrganize <- function(simOut){

  dat <- simOut$d
  weight.mat <<- simOut$wm
  c <- simOut$c
  lr <- simOut$lr

  almResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
 mutate(model="ALM",resp = alm.response(x,c))

examResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
 mutate(model="EXAM",resp = exam.response(x,c))

 bind_rows(almResp,examResp) %>% 
 mutate(type=first(dat$type),c=c,lr=lr,
 type=factor(type,levels=c("linear","exponential","quadratic")))

}

```

We can model the relation between reaction time (in seconds) and the number of practice trials as a power law function. Let $f: \mathbb{N} \rightarrow \mathbb{R}^+$ be a function that maps the number of trials to reaction times. We write

$$
f_p(N) = \alpha + \beta N^{-r} \enspace ,
$$

where $\alpha$ is a lower bound (one cannot respond faster than that due to processing and motor control limits); $\beta$ is the learning gain from practice with respect to the first trial ($N = 1$); $N$ indexes the particular trial; and $r$ is the learning rate. Similarly, we can write

$$
f_e(N) = \alpha + \beta e^{-rN} \enspace ,
$$

where the parameters have the same interpretation, except that $\beta$ is the learning gain from practice compared to no practice ($N = 0$).[^2]

What is the main difference between those two functions? *The exponential model assumes a constant learning rate, while the power model assumes diminishing returns*. To see this, let $\alpha = 0$, and ignore for the moment that $N$ is discrete. Taking the derivative for the power law model results in

$$
\frac{\partial f_p(N)}{\partial N} = -r\beta N^{-r - 1} = (-r/N) \, \beta N^{-r} = (-r/N) \, f_p(N) \enspace ,
$$

which shows that the *local learning rate* --- the change in reaction time as a function of $N$ --- is $-r/N$; it depends on how many trials have been completed previously. The more one has practiced, the smaller the local learning rate $-r / N$. The exponential function, in contrast, shows no such dependency on practice:

$$
\frac{\partial f_e(N)}{\partial N} = -r\beta e^{-rN} = -r \, f_e(N) \enspace .
$$
# simulate data
sim_power <- function(N, alpha, beta, r, t0 = 0, sdlog = 1) {
  t0 + (alpha + beta * N^(-r)) * rlnorm(length(N), 0, sdlog)
}
sim_exp <- function(N, alpha, beta, r, t0 = 0, sdlog = 1) {
  t0 + (alpha + beta * exp(-r*N)) * rlnorm(length(N), 0, sdlog)
}

negative_llh_exp <- function(y, N, alpha, beta, r, sigma_e, tau = 0) {
  -sum(dnorm(y, alpha + beta * (tau + 1) / (tau + exp(r*N)), sigma_e, log = TRUE))
}

```{r}

# create learning models for condit and varied groups. Aggregatign over ids in dst. The models predict dist as an exponential decay function of trial number. Band is an additional predictor.

# fit exponential decay model as a function of trial number

fit_exp <- function(trial,dist,input){
    # fit exponential decay model as a function of trial number, band is an additional predictor
    fit <- nls(dist ~ yf + (y0-yf) * exp(-r*trial) + beta2*input, start = list(yf = 300, y0 = 364, beta2=0, r = .1), data = data.frame(trial=trial,dist=dist,input=input))

    # extract parameters
    alpha <- coef(fit)[1]
    beta <- coef(fit)[2]
    beta2 <- coef(fit)[3]
    r <- coef(fit)[4]
    sigma_e <- summary(fit)$sigma

    # compute negative log likelihood
    nllh <- negative_llh_exp(dist, trial, alpha, beta, r, sigma_e)

    # return parameters and negative log likelihood
    return(list(alpha=alpha,beta=beta,beta2=beta2,r=r,sigma_e=sigma_e,nllh=nllh))
}

# Compute group averages for dist over trial and band. dst

avgTrain <- dst %>% group_by(id,condit,trial,band,input) %>% summarise(dist=mean(dist)) %>% ungroup() %>% group_by(condit,trial,band) %>% summarise(dist=mean(dist)) %>% ungroup()

# plot group averages
ggplot(avgTrain,aes(x=trial,y=dist)) + geom_line(aes(group=band,color=band)) +facet_grid(~condit)

avgTrain %>% filter(condit=="Constant") %>% nls(dist ~ yf + (y0-yf) * exp(-r*trial), start = list(yf = 120, y0 = 364, r = .1), data = .) %>% summary()

avgTrain %>% filter(condit=="Constant") %>% nls(dist ~ SSasymp(trial, yf, y0, log_alpha),data=.)

# fit exponential decay model for each condit
fit_condit <- avgTrain %>% group_by(condit) %>% do(fit_exp(trial=.$trial,dist=.$dist,input=.$input))

# fit exponential model

```