### 2023-04-06 DeLosh97_Sim.qmd

---
title: Simulating DeLosh 1997
code-fold: true
code-tools: true
execute: 
  warning: false
---

```{r}
#lapply(c('tidyverse','data.table','igraph','ggraph','kableExtra'),library,character.only=TRUE))
pacman::p_load(tidyverse,data.table,igraph,ggraph,kableExtra)

```

```{r}
#https://nrennie.rbind.io/blog/2022-06-06-creating-flowcharts-with-ggplot2/
inNodes <- seq(1,6,1) %>% as.integer()
outNodes <- seq(300,1000,50)%>% as.integer()
stim <- "Stim"
resp <- "Response"
inFlow <- tibble(expand.grid(from=stim,to=inNodes)) %>% mutate_all(as.character)
outFlow <- tibble(expand.grid(from=outNodes,to=resp)) %>% mutate_all(as.character)

gd <- tibble(expand.grid(from=inNodes,to=outNodes)) %>% mutate_all(as.character) %>%
  rbind(inFlow,.) %>% rbind(.,outFlow)

g = graph_from_data_frame(gd,directed=TRUE)
coords2=layout_as_tree(g)
colnames(coords2)=c("y","x")
odf <- as_tibble(coords2) %>% 
  mutate(label=vertex_attr(g,"name"),
         type=c("stim",rep("Input",length(inNodes)),rep("Output",length(outNodes)),"Resp"),
         x=x*-1) %>%
  mutate(y=ifelse(type=="Resp",0,y),xmin=x-.05,xmax=x+.05,ymin=y-.35,ymax=y+.35)

plot_edges = gd %>% mutate(id=row_number()) %>%
  pivot_longer(cols=c("from","to"),names_to="s_e",values_to=("label")) %>%
                 mutate(label=as.character(label)) %>% 
  group_by(id) %>%
  mutate(weight=sqrt(rnorm(1,mean=0,sd=10)^2)/10) %>%
  left_join(odf,by="label") %>%
  mutate(xmin=xmin+.02,xmax=xmax-.02)

ggplot() + geom_rect(data = odf,
            mapping = aes(xmin = xmin, ymin = ymin, 
                          xmax = xmax, ymax = ymax, 
                          fill = type, colour = type),alpha = 0.5) +
  geom_text(data=odf,aes(x=x,y=y,label=label,size=3)) +
  geom_path(data=plot_edges,mapping=aes(x=x,y=y,group=id,alpha=weight)) +
  # geom_rect(aes(xmin=-1.05,xmax=-.95,ymin=-10,ymax=5),color="red",alpha=.1)+
  # geom_rect(aes(xmin=-0.05,xmax=.05,ymin=-10,ymax=5),color="blue",alpha=.1) +
  theme_void()

```

## ALM Definition

::: {.column-margin}
###### Input Activation
$$
a_i(X)=\exp \left|-\gamma \cdot\left[X-X_i\right]^2\right|
$$

###### Output activation
$$
o_j(X)=\Sigma_{i=1, M} w_{j i} \cdot a_i(X) 
$$

###### Output Probability
$$
P\left[Y_j \mid X\right]=o_j(X) / \Sigma_{k=1, L} o_k(X) 
$$

###### Mean Response
$$
m(X)=\Sigma_{j=1, L} Y_j \cdot P\left[Y_j \mid X\right] 
$$

:::

### Generate Response
```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
alm.response <- function(input=1,c) {
input.activation <- exp(-c*(input.layer - input)^2)
input.activation <<- input.activation/sum(input.activation)
#print(length(input.activation)); print(dim(weight.mat))
output.activation <<- weight.mat %*% input.activation
output.probability <<- output.activation/sum(output.activation)
mean.response <<- sum(output.layer * output.probability)
mean.response
}
```

\ \ \ \ \

### Update Weights Based on Feedback
```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
#| code-overflow: wrap
alm.update <- function(corResp,c,lr){
  fz <- exp(-c*(output.layer - corResp)^2)
  teacherSignal <- (fz - output.activation)*lr
  #print(length(teacherSignal)); print(length(fz))
  wChange <- teacherSignal %*% t(input.activation)
  weight.mat <<- weight.mat + (wChange)
  weight.mat[weight.mat<0]=0 # prevent negative values
 # weight.mat[weight.mat>1]=1
  weight.mat <<- weight.mat
}

alm.trial <- function(input, corResp,c,lr){
  alm.response(input,c)
  alm.update(corResp,c,lr)
 # print(paste0("input=",input,"; corResp=",corResp,"; mean.response=",mean.response))
  mean.response
}
```

::: {.column-margin}
###### Feedback Signal

$$
f_j(Z)=e^{-c\cdot(Z-Y_j)^2}
$$

###### Weight Updates
$$
w_{ji}(t+1)=w_{ji}(t)+\alpha \cdot {f_i(Z(t))-O_j(X(t))} \cdot a_i(X(t))
$$
:::

### Exam Generalization
```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
exam.response <- function(input,c){
  # Find the index of the input node with the highest activation
  trainVec = sort(unique(xt))
  nearestTrain <- trainVec[which.min(abs(input - trainVec))]
  aresp <- alm.response(nearestTrain,c)
  #max.index <- which.max(input.activation)
  xUnder = ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver = ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])

  mUnder <- alm.response(xUnder,c)
  mOver <- alm.response(xOver,c)

   exam.output = round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (input - nearestTrain), 3)
  # Determine the input nodes and associated weights for computing the slope
  exam.output
}
```

::: {.column-margin}
###### Input node actvation
$$
P[X_i|X] = \frac{a_i(X)}{\\sum_{k=1}^Ma_k(X)}
$$

###### Slope Computation
$$
E[Y|X_i]=m(X_i) + \bigg[\frac{m(X_{i+1})-m(X_{i-1})}{X_{i+1} - X_{i-1}} \bigg]\cdot[X-X_i]
$$
:::

### Prepare Simulation Data

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
# function to generate data, from either linear, exponential, quadratic or sinusoidal functions

generate.data <- function(x, type = "linear", noise = NA) {
  if (type == "linear") {
    y <- round(2.2*x + 30,0)
  }
  else if (type == "exponential") {
    y <- round(200*(1-exp(-x/25)),0)
  }
  else if (type == "sinusoidal") {
    y <- sin(2 * pi * x) 
  }
  else if (type == "quadratic") {
    y <- round(210 - ((x-50)^2)/12,0)
  }
  else {
    stop("type must be linear, exponential, quadratic, or sinusoidal")
  }
  # if noise is specified, add noise to the y values
  if(!is.na(noise)) {
    y <- y + round(rnorm(length(y), 0, noise),2)
  }
  data.frame(x, y,type)
}

envTypes <- c("linear", "exponential", "quadratic")

lowDensityTrainBlock <- c(30.5, 36.0, 41.0, 46.5, 53.5, 59.0, 64.0, 69.5)
medDensityTrainBlock <- c(
    30.0, 31.5, 33.0, 34.5, 36.5, 38.5, 41.0, 43.5, 46.0,
    48.5, 51.5, 54.0, 56.5, 59.0, 61.5, 63.5, 65.5, 67.0, 68.5, 70.0
)
highDensityTrainBlock <- c(
    30.0, 30.5, 31.0, 32.0, 33.0, 33.5, 34.5, 35.5,
    36.5, 37.0, 38.0, 38.5, 39.5, 40.5, 41.5, 42.0, 43.0,
    43.5, 44.5, 45.5, 46.5, 47.0, 48.0, 48.5, 49.0, 51.0, 51.5, 52.0,
    53.0, 53.5, 54.5, 55.5, 56.5, 57.0, 58.0, 58.5, 59.5, 60.5, 61.5, 
    62.0, 63.0,63.5, 64.5, 65.5, 66.5, 67.0, 68.0, 69.0, 69.5, 70.0
)
# low density has 25 training blocks, medium has 10 blocks, high has 4 blocks.

# generate training data, for each combination of environment type and density. Use purrr map functions. Rep each dataset by its number of blocks.
lowTrain <- map_dfr(envTypes, ~ generate.data(rep(lowDensityTrainBlock,25), type = .x)) %>% group_by(type) %>% mutate(block = rep(1:25, each = 8),trial=seq(1,200))

medTrain <- map_dfr(envTypes, ~ generate.data(rep(medDensityTrainBlock,10), type = .x)) %>% group_by(type) %>% mutate(block = rep(1:10, each = 20),trial=seq(1,200))

highTrain <- map_dfr(envTypes, ~ generate.data(rep(highDensityTrainBlock,4), type = .x)) %>% group_by(type) %>% mutate(block = rep(1:4, each = 50),trial=seq(1,200))

# nTrainExamples <- 10
# trainLowBound <-20; trainHIghBound <- 80
# nBlock <- 5 # number of times each training example is presented
# sample training data from function, nTrainExamples., between trainLowBound and trainHIghBound
# xt <- runif(nTrainExamples, trainLowBound, trainHIghBound)
# trainVec <- rep(xt, nBlock)

```

### Simulation Functions

```{r}
#| code-fold: show
#| 
# simulation function
alm.sim <- function(dat, c, lr,testRange=seq(0,100,.5)){

input.layer <<- matrix(seq(0,100,.5) ) # half step units for inputs, from 0 to 100
output.layer <<- matrix(seq(0,250,1)) # single step units for outputs, from 0 to 250
weight.mat <<- matrix(0.0000000,nrow=length(output.layer),ncol=length(input.layer )) # weights initialized to 0 (as in Delosh 1997)

xt<<-dat$x
# run training
st <- map2_dbl(dat$x, dat$y, ~alm.trial(.x,.y,c,lr))
# append training data to the data frame
dat <- dat %>% mutate(almResp = st)

return(list(d=dat,wm=weight.mat,c=c,lr=lr)) # final weightmat is probs incorrect for all but last
}

simOrganize <- function(simOut){
  dat <- simOut$d
  weight.mat <<- simOut$wm
  c <- simOut$c
  lr <- simOut$lr

  trainX <- unique(dat$x)
  xt <<- trainX

 almResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
 mutate(model="ALM",resp = alm.response(x,c))

 examResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
 mutate(model="EXAM",resp = exam.response(x,c))

 bind_rows(almResp,examResp) %>% 
 mutate(type=first(dat$type),
        c=c,lr=lr,
 type=factor(type,levels=c("linear","exponential","quadratic"))) %>%
 # compute test_region, equal to "train" if x is within trainX, interpolate if within trainX range, else "extrapolate"
  mutate(test_region = ifelse(x %in% trainX, "train", ifelse(x > min(trainX) & x < max(trainX), "interpolate", "extrapolate")))

}

```

### Simulate Training:
```{r fig.width=10, fig.height=8}
#| column: page-inset-right
# split by type, then send each training dataset to simulation function
lowSim <- map(envTypes, ~ alm.sim(lowTrain %>% filter(type == .x), c = 1.4, lr = .4))

medSim <- map(envTypes, ~ alm.sim(medTrain %>% filter(type == .x), c = 1.4, lr = .4))

highSim <- map(envTypes, ~ alm.sim(highTrain %>% filter(type == .x), c = 1.4, lr = .4))

simAll <- rbind(bind_rows(lowSim %>% map("d")) %>% mutate(density = "low"), 
                bind_rows(medSim %>% map("d")) %>% mutate(density = "med"), 
                bind_rows(highSim %>% map("d")) %>% mutate(density = "high"))

simAll <- simAll %>% mutate(stage=as.numeric(cut(trial,breaks=20,labels=seq(1,20))),
                                      dev=sqrt((y-almResp)^2),
                                  #reorder density factor levels
                            density=factor(density,levels=c("low","med","high")),
                            type=factor(type,levels=c("linear","exponential","quadratic"))) %>%
                            dplyr::relocate(density,type,stage)

simAll %>% ggplot(aes(x=block,y=dev,color=type)) + stat_summary(geom="line",fun=mean,alpha=.4)+
  stat_summary(geom="point",fun=mean,alpha=.4)+
  stat_summary(geom="errorbar",fun.data=mean_cl_normal,alpha=.4)+facet_wrap(~density, scales="free_x")

```

Predictions for Generalization
```{r fig.width=11,fig.height=10}
#| column: page-inset-right
#| 
lowSimTest <- map_dfr(lowSim,simOrganize) %>% mutate(density = "low")
medSimTest <- map_dfr(medSim,simOrganize) %>% mutate(density = "med")
highSimTest <- map_dfr(highSim,simOrganize) %>% mutate(density = "high")

simTestAll <- rbind(lowSimTest,medSimTest,highSimTest) %>% group_by(type,density,model) %>%
  mutate(type=factor(type,levels=c("linear","exponential","quadratic")),
         density=factor(density,levels=c("low","med","high"))) %>%
  dplyr::relocate(density,type,test_region)

simTestAll %>% ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  #geom_point(aes(x=x,y=y,color=test_region),alpha=.2)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
 # geom_point(data=simTestAll %>% filter(test_region %in% c("interpolate","extrapolate")),aes(x=x,y=y,color=test_region),alpha=.6) +
  # geom_point(data=simTestAll %>% filter(test_region=="extrapolate"),aes(x=x,y=y),color="purple",alpha=.3) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

#lowSimTest %>% filter(model=="EXAM" & type=="linear")
#rm(list= ls()[sapply(ls(), function(x) class(get(x))) != 'function'])
#rm(weight.mat,input.activation,output.probability,output.activation,mean.response,xt)
# 
# simAll %>% ggplot(aes(x=block,y=dev)) + stat_summary(geom="line",fun=mean,alpha=.3)+stat_summary(geom="point",fun=mean)+
#   stat_summary(geom="errorbar",fun.data=mean_cl_normal)+facet_wrap(density~type, scales="free_x")

```

Collpasing Across Density Levels gives us: 
```{r}
#| eval: false
#| column: page-inset-right
simTestAll %>% group_by(type,model,x,y) %>% summarise(resp=mean(resp))  %>% ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  facet_grid(~type) + 
  theme_bw() + theme(legend.position="bottom")
```

### Master Function for full simulation

```{r}
#| eval: false
# Function that goes through every step of generating data, simulating training, and simulating generalization
full.sim <- function(c,lr,noise)
{

envTypes <- c("linear", "exponential", "quadratic")
lowDensityTrainBlock <- c(30.5, 36.0, 41.0, 46.5, 53.5, 59.0, 64.0, 69.5)
medDensityTrainBlock <- c(
    30.0, 31.5, 33.0, 34.5, 36.5, 38.5, 41.0, 43.5, 46.0,
    48.5, 51.5, 54.0, 56.5, 59.0, 61.5, 63.5, 65.5, 67.0, 68.5, 70.0
)
highDensityTrainBlock <- c(
    30.0, 30.5, 31.0, 32.0, 33.0, 33.5, 34.5, 35.5,
    36.5, 37.0, 38.0, 38.5, 39.5, 40.5, 41.5, 42.0, 43.0,
    43.5, 44.5, 45.5, 46.5, 47.0, 48.0, 48.5, 49.0, 51.0, 51.5, 52.0,
    53.0, 53.5, 54.5, 55.5, 56.5, 57.0, 58.0, 58.5, 59.5, 60.5, 61.5, 
    62.0, 63.0,63.5, 64.5, 65.5, 66.5, 67.0, 68.0, 69.0, 69.5, 70.0
)
# low density has 25 training blocks, medium has 10 blocks, high has 4 blocks. 
# generate training data, for each combination of environment type and density. Use purrr map functions. Rep each dataset by its number of blocks.
lowTrain <- map_dfr(envTypes, ~ generate.data(rep(lowDensityTrainBlock,25), type = .x, noise)) %>% group_by(type) %>% mutate(block = rep(1:25, each = 8),trial=seq(1,200))
medTrain <- map_dfr(envTypes, ~ generate.data(rep(medDensityTrainBlock,10), type = .x, noise)) %>% group_by(type) %>% mutate(block = rep(1:10, each = 20),trial=seq(1,200))
highTrain <- map_dfr(envTypes, ~ generate.data(rep(highDensityTrainBlock,4), type = .x, noise)) %>% group_by(type) %>% mutate(block = rep(1:4, each = 50),trial=seq(1,200))

lowSim <- map(envTypes, ~ alm.sim(lowTrain %>% filter(type == .x), c = 1.4, lr = .4))
medSim <- map(envTypes, ~ alm.sim(medTrain %>% filter(type == .x), c = 1.4, lr = .4))
highSim <- map(envTypes, ~ alm.sim(highTrain %>% filter(type == .x), c = 1.4, lr = .4))

simAll <- rbind(bind_rows(lowSim %>% map("d")) %>% mutate(density = "low"), 
                bind_rows(medSim %>% map("d")) %>% mutate(density = "med"), 
                bind_rows(highSim %>% map("d")) %>% mutate(density = "high"))

simAll <- simAll %>% mutate(stage=as.numeric(cut(trial,breaks=20,labels=seq(1,20))),
                                      dev=sqrt((y-almResp)^2),
                                  #reorder density factor levels
                            density=factor(density,levels=c("low","med","high")),
                            type=factor(type,levels=c("linear","exponential","quadratic"))) %>%
                            dplyr::relocate(density,type,stage)

lowSimTest <- map_dfr(lowSim,simOrganize) %>% mutate(density = "low")
medSimTest <- map_dfr(medSim,simOrganize) %>% mutate(density = "med")
highSimTest <- map_dfr(highSim,simOrganize) %>% mutate(density = "high")

simTestAll <- rbind(lowSimTest,medSimTest,highSimTest) %>% group_by(type,density,model) %>%
  mutate(type=factor(type,levels=c("linear","exponential","quadratic")),
         density=factor(density,levels=c("low","med","high"))) %>%
  dplyr::relocate(density,type,test_region)

return(list(simAll=list(simAll),simTestAll=list(simTestAll)))

}
```

### Simulations with noise
```{r fig.height=10,fig.width=11}
#| eval: false
k = full.sim(c=1.4,lr=.4,noise=2.0)
k4 = full.sim(c=1.4,lr=.4,noise=4.0)

# run simulation with noise=10, 3 times, average results together. 
k10 = map_dfr(1:3, ~ full.sim(c=1.4,lr=.4,noise=10.0)) %>% group_by(type,density,model) %>%
  mutate(type=factor(type,levels=c("linear","exponential","quadratic")),
         density=factor(density,levels=c("low","med","high"))) %>%
  dplyr::relocate(density,type,test_region)

k %>% pluck("simAll") %>% ggplot(aes(x=block,y=dev,color=type)) + stat_summary(geom="line",fun=mean,alpha=.4)+
  stat_summary(geom="point",fun=mean,alpha=.4)+
  stat_summary(geom="errorbar",fun.data=mean_cl_normal,alpha=.4)+facet_wrap(~density, scales="free_x")

k4 %>% pluck("simAll") %>% ggplot(aes(x=block,y=dev,color=type)) + stat_summary(geom="line",fun=mean,alpha=.4)+
  stat_summary(geom="point",fun=mean,alpha=.4)+
  stat_summary(geom="errorbar",fun.data=mean_cl_normal,alpha=.4)+facet_wrap(~density, scales="free_x")

k %>% pluck("simTestAll") %>%ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

k4 %>% pluck("simTestAll") %>%ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

k10 %>% pluck("simTestAll") %>%ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

```

```{r}
#| eval: false

# label each each simulation with its density level (low, med, high), then combine all 3
lowSimTest <- lowSimTest %>% mutate(density = "low")
medSimTest <- medSimTest %>% mutate(density = "med")
highSimTest <- highSimTest %>% mutate(density = "high")

simTest <- bind_rows(lowSimTest,medSimTest,highSimTest)

# extract element d from each list, and bind rows, remove Nan's, group by type, mutate new variable "Stage", which is set to first for block 1, last for final block, and middle for all other glocks, pipe to ggplot, plotting y and almResp in different colors, facet by type and stage (only first and last block)
ls2=bind_rows(lowSim %>% map("d")) %>% filter(!is.na(almResp)) %>% 
  group_by(type) %>% 
  mutate(stage = ifelse(block == 1, "first", ifelse(block == 25, "last", "middle")),
         stage2=cut(trial,breaks=20,labels=seq(1,20)),
         dev=sqrt((y-almResp)^2))

ls2 %>% ggplot() + geom_point(aes(x=stage2,y=dev))+facet_grid(~type)

ls2 %>% filter(stage %in% c("first","last")) %>% ggplot() + geom_point(aes(x=x,y=y),color="red",alpha=.3) + 
geom_point(aes(x=x,y=almResp),color="blue",alpha=.4)+ 
facet_grid(type~stage) + theme_bw() + theme(legend.position="bottom")

lowSimTest %>% ggplot() + geom_point(aes(x=x,y=resp,color=model)) + geom_line(aes(x=x,y=y),alpha=.3)+ facet_grid(~type) + theme_bw() + theme(legend.position="bottom")

# make grid of each combination of train dataset and envType
simGrid <- expand.grid(envTypes=envTypes,trainData=c("lowTrain","medTrain","highTrain"))

# map over the grid, for each row of simGrid, filter the training dataset by the envType, and then run the simulation function, and then bind rows of the output data frame.
simOut <- map2(c("lowTrain","medTrain","highTrain"),envTypes, ~ alm.sim(get(.x) %>% filter(type == .y), c = 1.4, lr = .4))

# simOut <- map(c("lowTrain","medTrain","highTrain"), ~ map_dfr(envTypes, ~ alm.sim(get(.) %>% filter(type == .x), c = 1.4, lr = .4) %>% simOrganize))
# 
# simOut <- map_dfr(simGrid, ~ alm.sim(get(.x$trainData) %>% filter(type == .x$envTypes), c = 1.4, lr = .4) %>% simOrganize)
# 
# for each training dataset in c(lowTrain,medTrain,highTrain), run simulation function, separately for each type, and then bind rows of the output data frame.

```

### Primary Functions

```{r}
#| code-fold: show
#| include: false
#| eval: false
alm.response <- function(input=1,c) {
input.activation <- exp(-c*(input.layer - input)^2)
input.activation <<- input.activation/sum(input.activation)
#print(length(input.activation)); print(dim(weight.mat))
output.activation <<- weight.mat %*% input.activation
output.probability <<- output.activation/sum(output.activation)
mean.response <<- sum(output.layer * output.probability)
mean.response
}

alm.update <- function(corResp,c,lr){
  fz <- exp(-c*(output.layer - corResp)^2)
  teacherSignal <- (fz - output.activation)*lr
  #print(length(teacherSignal)); print(length(fz))
  wChange <- teacherSignal %*% t(input.activation)
  weight.mat <<- weight.mat + (wChange)
  weight.mat[weight.mat<0]=0
 # weight.mat[weight.mat>1]=1
  weight.mat <<- weight.mat
}

alm.trial <- function(input, corResp,c,lr){
  alm.response(input,c)
  alm.update(corResp,c,lr)
  print(paste0("input=",input,"; corResp=",corResp,"; mean.response=",mean.response))
  mean.response
}

exam.response <- function(input,c){
  # Find the index of the input node with the highest activation
  trainVec = sort(unique(xt))
  nearestTrain <- trainVec[which.min(abs(input - trainVec))]
  aresp <- alm.response(nearestTrain,c)
  #max.index <- which.max(input.activation)
  xUnder = ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver = ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])

  mUnder <- alm.response(xUnder,c)
  mOver <- alm.response(xOver,c)

   exam.output = round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (input - nearestTrain), 3)
  # Determine the input nodes and associated weights for computing the slope
  exam.output
}

```