### 2023-04-06 benchmarks.qmd

---
title: Benchmarking 
code-fold: show
code-tools: true
execute: 
  warning: false
  eval: false
---

Furrr and standard version both work equally fast, tested with up to 120 simulation repetitions
```{r}
library(furrr)
furrr::furrr_options(seed = TRUE)
plan(multisession, workers = parallel::detectCores())

parmVec <- tibble(crossing(c = c(0.1,.5), lr = c(0.4), noise = c(500), trainRep = c(20), lossFun = list("RMSE", "RMSE.blocked"), simNum = 1:30))
tf <- system.time(
sdpf <- parmVec %>% 
  mutate(d = future_pmap(list(c, lr, noise, trainRep), ~sim_data(c = ..1, lr = ..2, noise = ..3, trainRep = ..4),
                         .options = furrr_options(seed = T)),
                           almTrainDat = map(d, "almTrain"),
                          almTestDat = map(d, "almPred"),
                          examTestDat = map(d, "examPred"),
                          td = map(trainRep, ~gen_train(trainRep = .)),
                          fitO = map2(td, lossFun, ~wrap_optim(.x, .y)),
                          fitG = map2(td, lossFun, ~wrap_grid(.x, .y)),
                          cFitO = map_dbl(fitO, "c"),
                          lrFitO = map_dbl(fitO, "lr"),
                          optimValO = map_dbl(fitO, "Value"),
                          cFitG = map_dbl(fitG, "c"),
                          lrFitG = map_dbl(fitG, "lr"),
                          optimValG = map_dbl(fitG, "Value"))
)
tf

ts <- system.time(
sdp <- parmVec %>% mutate(d = pmap(list(c, lr, noise, trainRep), ~sim_data(c = ..1, lr = ..2, noise = ..3, trainRep = ..4)),
                          almTrainDat = map(d, "almTrain"),
                          almTestDat = map(d, "almPred"),
                          examTestDat = map(d, "examPred"),
                          td = map(trainRep, ~gen_train(trainRep = .)),
                          fitO = map2(td, lossFun, ~wrap_optim(.x, .y)),
                          fitG = map2(td, lossFun, ~wrap_grid(.x, .y)),
                          cFitO = map_dbl(fitO, "c"),
                          lrFitO = map_dbl(fitO, "lr"),
                          optimValO = map_dbl(fitO, "Value"),
                          cFitG = map_dbl(fitG, "c"),
                          lrFitG = map_dbl(fitG, "lr"),
                          optimValG = map_dbl(fitG, "Value"))
)
ts

```

M1 Max times: 
tf
   user  system elapsed 
221.326   4.517 226.233

ts
   user  system elapsed 
221.140   4.288 225.330

Mac Pro times: 
 tf
    user   system  elapsed 
1125.102   76.859 1474.612

ts
    user   system  elapsed 
1132.716   76.260 1493.154

# Benchmarking ALM Model Fit Functions

```{r}
#| code-fold: true
pacman::p_load(tidyverse,data.table,microbenchmark())

d <- readRDS('dPrune-01-19-23.rds')

dtest <- d %>% filter(expMode %in% c("test-Nf","test-train-nf")) %>% group_by(id,lowBound) %>% 
  mutate(nBand=n(),band=bandInt,id=factor(id)) %>% group_by(id) %>% mutate(nd=n_distinct(lowBound))
# unique(dtest[dtest$nd==4,]$sbjCode) # 7 in wrong condition
dtest <- dtest %>% group_by(id,lowBound) %>% filter(nBand>=5 & nd==6)
# for any id that has at least 1 nBand >=5, remove all rows with that id. 
dtest <- dtest %>% group_by(id) %>% filter(!id %in% unique(dtest$id[dtest$nBand<5]))

dtestAgg <- dtest %>% group_by(id,condit,catOrder,feedbackType,vb,band,lowBound,highBound,input) %>% mutate(vxCapped=ifelse(vx>1600,1600,vx)) %>%
  summarise(vxMean=mean(vx),devMean=mean(dist),vxMed=median(vx),devMed=median(dist),
            vxMeanCap=mean(vxCapped),.groups = "keep")

# select first row for each id in d, then create histogram for nTrain
#  d  %>% group_by(id) %>% slice(1) %>% ggplot(aes(nTrain)) + geom_histogram() + facet_wrap(~condit)

ds <- d %>% filter(expMode %in% c("train","train-Nf","test-Nf","test-train-nf")) %>% 
filter(!id %in% unique(dtest$id[dtest$nBand<5])) %>% 
select(id,condit,catOrder,feedbackType,expMode,trial,gt.train,vb,band,bandInt,lowBound,highBound,input,vx,dist,vxb)

dst <- ds %>% filter(expMode=="train",catOrder=="orig")

vTrainTrial <- dst %>% filter(condit=="Varied",gt.train<=84) %>% group_by(gt.train,vb) %>% summarise(sdVx=sd(vx),vx=mean(vx),sdDist=sd(dist),dist=mean(dist)) %>% 
  group_by(vb) %>% mutate(gt.trainBin=cut(gt.train,breaks=5,labels=c(1:5)))

binTrainTrial <- dst %>% filter(gt.train<=83) %>% group_by(gt.train,vb,condit) %>% summarise(sdVx=sd(vx),vx=mean(vx),sdDist=sd(dist),dist=mean(dist)) %>% 
  group_by(vb) %>% mutate(gt.trainBin=cut(gt.train,breaks=6,labels=c(1:6)))

tMax=84
bandVec <- rep(c(800,1000,1200),each=tMax/3)
bandVec <- bandVec[sample(1:length(bandVec),tMax,replace=FALSE)]

trainTrials <- dst %>% filter(gt.train<=tMax) %>% group_by(condit,gt.train,vb,bandInt,input) %>% summarise(vx=mean(vx))

input.activation<-function(x.target, c){
  return(exp(-1*c*(x.target-inputNodes)^2))
}

output.activation<-function(x.target, weights, c){
  return(weights%*%input.activation(x.target, c))
}

mean.prediction<-function(x.target, weights, association.parameter){
  probability<-output.activation(x.target, weights, c)/sum(output.activation(x.target, weights, c))
  return(outputNodes%*%probability) # integer prediction
}
# function to generate exam predictions
exam.prediction<-function(x.target, weights, c,trainVec){
  #trainVec = sort(unique(x.learning))
  nearestTrain = trainVec[which.min(abs(trainVec-x.target))]
  aresp = mean.prediction(nearestTrain, weights, c)
  xUnder = ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver = ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])
  mUnder = mean.prediction(xUnder, weights, c)
  mOver = mean.prediction(xOver, weights, c)
  exam.output = round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (x.target - nearestTrain), 3)
  exam.output
}

update.weights<-function(x.new, y.new, weights, c, lr){
  y.feedback.activation<-exp(-1*c*(y.new-outputNodes)^2)
  x.feedback.activation<-output.activation(x.new, weights, c)
  return(weights+lr*(y.feedback.activation-x.feedback.activation)%*%t(input.activation(x.new, c)))
}

train.alm<-function(dat, c=0.05, lr=0.5, weights){
   alm.train<-rep(NA,nrow(dat))  
  for (i in 1:nrow(dat)){
    weights <- update.weights(dat$input[i], dat$vx[i], weights, c, lr)
    resp = mean.prediction(dat$input[i], weights, c)
    alm.train[i]=resp
    weights[weights<0]=0
  }
  alm.train
}

wrap_alm <- function(parms,dat, weights,lossFun){
    c=parms[1]; lr=parms[2]
   pred=train.alm(dat, c=c, lr=lr, weights=weights)
   #sqrt(mean((dat$vx -pred)^2))
   lossFun(dat$vx,pred)
}

wrap_optim <- function(dat,wm,lossFun){
  bounds_lower <- c(.0000001, .00001)
  bounds_upper <- c(5, 5)

 optim(c(.1, .2),
   fn = wrap_alm,
   dat = dat, weights = wm,lossFun=lossFun,
   method = "L-BFGS-B",
   lower = bounds_lower,
   upper = bounds_upper,
   control = list(maxit = 1e4, pgtol = 0, factr = 0)
 )
}

RMSE <- function(x,y){
  sqrt(mean((x-y)^2))
}

## First average observed and predicted data into blocks, then compute RMSE
RMSE.tb <- function(x,y,blocks=6){
  data.frame(x,y) %>% mutate(t=row_number(),fitBins=cut(t,breaks=blocks,labels=c(1:blocks))) %>%
    group_by(fitBins) %>% 
    summarise(predMean=mean(x),obsMean=mean(y)) %>% 
    summarise(RMSE(predMean,obsMean)) %>% as.numeric()
}

## Recode RMSE.tb using data.table functions rather than dplyr
RMSE.tb2 <- function(x,y,blocks=6){
  data.table(x=x,y=y,t=seq(1,length(x))) %>% 
    .[, `:=`(fitBins = cut(t, breaks = ..blocks, labels = c(1:..blocks)))] %>%
    .[, .(predMean = mean(x), obsMean = mean(y)), keyby = .(fitBins)] %>%
    .[, RMSE(predMean,obsMean)] %>% as.numeric()
}

```

### dplyr RMSE vs data.table RMSE
```{r}
dpVsDt=microbenchmark(
  dplyrMethod={
  fitVaried <- tv %>% filter(condit=="Varied") %>% wrap_optim(.,wm,lossFun=RMSE.tb);
  fitConstant <- tv %>% filter(condit=="Constant") %>% wrap_optim(.,wm,lossFun=RMSE.tb)},
  dtMethod={
  fitVaried2 <- tv %>% filter(condit=="Varied") %>% wrap_optim(.,wm,lossFun=RMSE.tb2);
  fitConstant2 <- tv %>% filter(condit=="Constant") %>% wrap_optim(.,wm,lossFun=RMSE.tb2)},
  times=5
)
knitr::kable(summary(dpVsDt),format="markdown")
```
|expr        |       min|        lq|      mean|    median|        uq|       max| neval|cld |
|:-----------|---------:|---------:|---------:|---------:|---------:|---------:|-----:|:---|
|dplyrMethod | 11.282291| 12.996857| 13.558623| 13.970367| 14.658515| 14.885086|     5|a   |
|dtMethod    |  4.933999|  5.306232|  5.681235|  5.515397|  6.131193|  6.519355|     5|b   |

The data.table version seems to be consistently more than 2x faster

### Separate fits, vs. nesting method vs. split method
```{r}
nestSplit<-microbenchmark(
separate={fitVaried <- tv %>% filter(condit=="Varied") %>% wrap_optim(.,wm,lossFun=RMSE.tb2);
fitConstant <- tv %>% filter(condit=="Constant") %>% wrap_optim(.,wm,lossFun=RMSE.tb2) },
nestGroups = tv %>% group_by(condit) %>% nest() %>% mutate(fit=map(data,~wrap_optim(.,wm,RMSE.tb2))),
splitGroups = tv %>% split(.$condit) %>% map(~wrap_optim(.,wm,RMSE.tb2)),
times=5
)
knitr::kable(summary(nestSplit),format="markdown") # 03/02/23 - Mac Pro
```
|expr        |      min|       lq|     mean|   median|       uq|      max| neval|cld |
|:-----------|--------:|--------:|--------:|--------:|--------:|--------:|-----:|:---|
|separate    | 6.394459| 6.483235| 6.709882| 6.508353| 7.059420| 7.103941|     5|a   |
|nestGroups  | 6.253850| 6.457956| 6.789962| 6.756942| 7.044965| 7.436094|     5|a   |
|splitGroups | 6.117184| 6.455866| 6.605814| 6.461572| 6.640798| 7.353648|     5|a   |

Not much of an effect for nesting method.

### Computing RMSE over Raw trials vs. RMSE of blocked training performance
```{r}
trialBlock<-microbenchmark(
  trialFit = tv %>% split(.$condit) %>% map(~wrap_optim(.,wm,RMSE)),
  blockFit = tv %>% split(.$condit) %>% map(~wrap_optim(.,wm,RMSE.tb2)),
  times=5
)
knitr::kable(summary(trialBlock),format="markdown")
```

|expr     |      min|       lq|     mean|   median|       uq|      max| neval|cld |
|:--------|--------:|--------:|--------:|--------:|--------:|--------:|-----:|:---|
|trialFit | 2.665184| 2.674186| 2.806476| 2.803491| 2.843806| 3.045712|     5|a   |
|blockFit | 5.169167| 5.264218| 5.350769| 5.285151| 5.491828| 5.543483|     5|b   |

The models are fit about 2x faster when computing RMSE over trials. This shouldn't be surprising, since
fitting to blocked data requires several additional computations (e.g. grouping, computing means)