### 2023-05-23 Model/abc_htw.qmd
using here() to reference images

---
title: Approximate Bayesian Fitting
date: last-modified
categories: [Modeling, ALM, R, Bayesian]
image: images/abc_varied.png
page-layout: full
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

```{r}
# load and view data
pacman::p_load(tidyverse,data.table,lme4,future,furrr,abc,mvtnorm,patchwork,here)
purrr::walk(here(c("Functions/alm_functions.R","Functions/Display_Functions.R","Functions/Noisy_Functions.R")),source)

select <- dplyr::select; mutate <- dplyr::mutate 
d <- readRDS(here("data/dPrune-01-19-23.rds"))
levels(d$condit)
# Prepare the data for analysis
dtest <- d %>% filter(expMode %in% c("test-Nf","test-train-nf")) %>% group_by(id,lowBound) %>% 
  mutate(nBand=n(),band=bandInt,id=factor(id)) %>% group_by(id) %>% mutate(nd=n_distinct(lowBound))
dtest <- dtest %>% group_by(id,lowBound) %>% filter(nBand>=5 & nd==6)
dtest <- dtest %>% group_by(id) %>% filter(!id %in% unique(dtest$id[dtest$nBand<5]))
ds <- d %>% filter(expMode %in% c("train","train-Nf","test-Nf","test-train-nf")) %>% 
filter(!id %in% unique(dtest$id[dtest$nBand<5])) %>% 
select(id,condit,catOrder,feedbackType,expMode,trial,gt.train,vb,band,bandInt,lowBound,highBound,input,vx,dist,vxb) 
dst <- ds %>% filter(expMode=="train")
dst <- dst %>%
  group_by(id, vb) %>%
  mutate(trial_band = row_number())
```

```{r}
dst <- dst %>% filter(expMode=="train",catOrder=="orig")
vst <- dst %>% filter(condit=="Varied",gt.train<=84) %>% group_by(gt.train,vb,input) %>% summarise(sdVx=sd(vx),vx=mean(vx),sdDist=sd(dist),dist=mean(dist)) %>% 
  group_by(vb) %>% mutate(gt.trainBin=cut(gt.train,breaks=5,labels=c(1:5)),
                          trial=gt.train) 
cst <- dst %>% filter(condit=="Constant",gt.train<=84) %>% group_by(gt.train,vb,input) %>% summarise(sdVx=sd(vx),vx=mean(vx),sdDist=sd(dist),dist=mean(dist)) %>% 
  group_by(vb) %>% mutate(gt.trainBin=cut(gt.train,breaks=5,labels=c(1:5)),
                          trial=gt.train)

ggplot(dst %>% filter(gt.train<=84), aes(x = gt.train, y = vx,color=vb)) +
  geom_point(aes(color = vb), stat = "summary", fun = mean) + 
  stat_summary(aes(color = vb), geom = "line", fun = mean) +
  stat_summary(geom="errorbar",fun.data=mean_se,width=.4,alpha=.7)+facet_wrap(~condit)

# ggplot(vst, aes(x = gt.trainBin, y = vx,color=vb)) +
  # geom_point(aes(color = vb), stat = "summary", fun = mean) + 
  # stat_summary(aes(color = vb), geom = "line", fun = mean) +
  # stat_summary(geom="errorbar",fun.data=mean_se,width=.4,alpha=.7)+facet_wrap(~condit)

```

## ABC Functions
```{r}
# c=1; lr=.5; noise_sd=.001; inNodes=7; outNodes=32
fit_alm_sim <- function(data, c, lr, noise_sd, inNodes, outNodes) {
  train_data <- data[, c("gt.train", "vb", "vx","input")] %>% rename("trial" = gt.train)

  sim_result <- sim_train(
    dat = train_data,
    c = c,
    lr = lr,
    inNodes = inNodes,
    outNodes = outNodes,
    noise_sd = noise_sd
  )

  train_data$almTrain <- sim_result$almTrain
  return(train_data$almTrain)
}

generate_prior <- function(n) {
  prior_samples <- tibble(
    c = runif(n, 0.001, 5),
    lr = runif(n, 0.001, 3),
    noise_sd = runif(n, 0, 0.0001),
    inNodes = sample(c( 3,7, 14,21,28,35), size = n, replace = TRUE),
    outNodes = sample(c(8,16, 32,48,64,80,96), size = n, replace = TRUE)
  )
  return(prior_samples)
}

```

### Varied ABC
```{r fig.width=12, fig.height=10}
#| eval: false

n_prior_samples <- 1000
prior_samples <- generate_prior(n_prior_samples)

# Replace gt with dst in the following line:
simulated_data <- future_map_dfc(seq_len(nrow(prior_samples)), function(idx) {
  params <- prior_samples[idx, ]
  fit_alm_sim(vst, params$c, params$lr, params$noise_sd, params$inNodes, params$outNodes)
},.options = furrr_options(seed = T))

# Replace gt_obs with dst_obs and gt with dst in the following lines:
dst_obs <- vst$vx
tolerance <- 0.1 * sd(dst_obs)*500
abc_result <- abc(
  target = dst_obs,
  param = prior_samples,
  sumstat = do.call(rbind, simulated_data),
  tol = .1,
  method = "rejection",
  names=colnames(dst_obs)
)

posterior_samples <- abc_result$unadj.values
colnames(posterior_samples) <- c("c", "lr", "noise_sd", "inNodes", "outNodes")
posterior_samples_long <- tidyr::pivot_longer(as.data.frame(posterior_samples), everything())

postV <- ggplot(posterior_samples_long, aes(x=value)) +
  geom_density() +
  facet_wrap(~name, scales="free") +
  theme_minimal() +
  labs(x="Value", y="Density", title="Posterior Density Plots")

summary_statistics <- data.frame(mean = apply(posterior_samples, 2, mean),
  median = apply(posterior_samples, 2, median))
( summary_statistics=rownames_to_column(summary_statistics,var="parameter") )

s = sim_train(dat=mutate(vst,trial=gt.train),c=summary_statistics$mean[1],lr=summary_statistics$mean[2],inNodes=summary_statistics$mean[4],outNodes=summary_statistics$mean[5],noise_sd=summary_statistics$mean[3])

ggp <- vst %>% ungroup() %>% mutate(pred=s$almTrain,input=as.factor(input))
gvp <- ggp %>% ggplot(aes(x = trial, y = pred, color = input)) +
  geom_line() + ylim(c(0,1600))+ggtitle("ALM Predictions (ABC estimation)")
gvo <- ggp  %>% ggplot(aes(x = trial, y = vx, color = input)) +
  geom_line() + ylim(c(0,1600))+ggtitle("HTW Observed Training data")
ggpv <- postV/(gvp + gvo)

#ggsave("images/abc_varied.png",ggpv,width=12,height=10)

```

```{r}
knitr::include_graphics(here("images/abc_varied.png"))
#![](images/abc_varied.png)
```

### Constant ABC
```{r fig.width=12, fig.height=10}
#| eval: false
# c=1; lr=.5; noise_sd=.001; inNodes=7; outNodes=32

n_prior_samples <- 1000
prior_samples <- generate_prior(n_prior_samples)

# Replace gt with dst in the following line:
simulated_dataC <- future_map_dfc(seq_len(nrow(prior_samples)), function(idx) {
  params <- prior_samples[idx, ]
  fit_alm_sim(cst, params$c, params$lr, params$noise_sd, params$inNodes, params$outNodes)
},.options = furrr_options(seed = T))

# Replace gt_obs with dst_obs and gt with dst in the following lines:
cst_obs <- cst$vx
tolerance <- 0.1 * sd(dst_obs)*500
abc_result <- abc(
  target = cst_obs,
  param = prior_samples,
  sumstat = do.call(rbind, simulated_dataC),
  tol = .1,
  method = "rejection",
  names=colnames(cst_obs)
)

posterior_samples <- abc_result$unadj.values
colnames(posterior_samples) <- c("c", "lr", "noise_sd", "inNodes", "outNodes")
posterior_samples_long <- tidyr::pivot_longer(as.data.frame(posterior_samples), everything())

postC <- ggplot(posterior_samples_long, aes(x=value)) +
  geom_density() +
  facet_wrap(~name, scales="free") +
  theme_minimal() +
  labs(x="Value", y="Density", title="Posterior Density Plots")

summary_statisticsC <- data.frame(mean = apply(posterior_samples, 2, mean),
  median = apply(posterior_samples, 2, median))
 summary_statisticsC=rownames_to_column(summary_statisticsC,var="parameter") 
summary_statisticsC

sC = sim_train(dat=mutate(cst,trial=gt.train),c=summary_statisticsC$mean[1],lr=summary_statisticsC$mean[2],inNodes=summary_statisticsC$mean[4],outNodes=summary_statisticsC$mean[5],noise_sd=summary_statisticsC$mean[3])

ggpC <- cst %>% ungroup() %>% mutate(pred=sC$almTrain,input=as.factor(input))
gcp <- ggpC %>% ggplot(aes(x = trial, y = pred, color = input)) +
  geom_line() + ylim(c(0,1600))+ggtitle("ALM Predictions (ABC estimation)")
gco <- ggpC  %>% ggplot(aes(x = trial, y = vx, color = input)) +
  geom_line() + ylim(c(0,1600))+ggtitle("HTW Observed Constant Training data")
postC/(gcp + gco)

ggsave("images/abc_constant.png",postC/(gcp + gco),width=12,height=10)

```

`r knitr::include_graphics(here("images/abc_constant.png"))`