### 2023-05-23 Misc/HTW_ToDo.qmd
more image changes and misc folder

---
title: "HTW To-do and Notes"
title-block-style: default
author: "Thomas Gorman"
date: last-modified
format:
  html:
    code-fold: true
filters:
  - collapse-callout
collapse-callout:
  todo: false
  tip: true
  note: true
highlight-style: gruvbox
---

## To do List

{{< fa chess-pawn >}} {{< fa chess-bishop >}} {{< fa chess-knight >}} {{< fa chess-rook >}} {{< fa chess-king >}}

::: {.grid}
::: {.g-col-8}

::: {.callout-note .callout-note2 collapse="false"}
## **Model To-do**
- Fit to Train then Predict Transfer vs. Fitting to all stages
- Separate ALM and EXAM Fits
- ALM + Prior Knowledge (initial anchor at 0)
- Empirical Learning Model
- Individual vs. Group fits
- Using Cognitive Model parameters to predict testing Vx vs. Deviation vs. Discrimination
- Model Recovery?
- Approximate Bayes?
:::
:::

::: {.g-col-4}
::: {.callout-note .callout-note2 icon-scale="10%"}
## **Analysis To-do**
- Discrimination
- Mixed Models?
:::
::: {.callout-note .callout-note2 icon-scale="10%"}
## **Site to-do**
- Configure react tables
:::
:::
:::

## Notes

::: {.callout-note .cTool}
## Human Regression or Rule

:::

::: {.callout-note .cBook}
## ALM & EXAM Implementations

:::

::: {.callout-note .cBook}
## ALM Likelihood

:::

::: {.callout-note .cBook}
## HTW DP
<iframe width="1100" height="655" src="https://tegorman13.github.io/DP/htw_n.html"></iframe>
:::