### 2023-05-23 Misc/Task.qmd

---
title: "HTW Task"
date: last-modified
page-layout: full
categories: [Task, js]
---

## HTW Task
* need to create a demo version without consent form. And maybe separate windows for the different versions.

Experimental Task for the HTW Project. Programmed in Javascript, and making use of phaser.js.

```{r out.width = "70%", echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
knitr::include_graphics(here::here("images/2023-04-19-16-20-55.png"))
```

<!-- ![](images/2023-04-19-16-20-55.png) -->
<!-- `r knitr::include_graphics(here::here("images/2023-04-19-16-20-55.png"))` -->

## Live Task Demo
![HTW_Task](https://pcl.sitehost.iu.edu/tg/HTW/HTW_Index_Of.html){width="900" height="800"}

<script src="https://cdn.jsdelivr.net/npm/p5@1.4.0/lib/p5.js"></script> <!-- load p5.js from CDN--> 
<!-- <script src = "/Shiny_HTW/testJs.js"></script> <!-- this will pick our script  -->
<!-- <script src = "/Shiny_HTW/testp5.js"></script> this will pick our script  --> -->

<!-- <script src = "/Shiny_HTW/p5_2.js"></script> <!-- this will pick our script  -->
<!-- <div id="idCanvas3"> this sets the position of the script --></div> -->