### 2023-05-25 Analysis/analysis.qmd
test sync

---
title: HTW Analysis
date: last-modified
categories: [Analysis,Learning-Curve, R]
page-layout: full
code-fold: show
code-tools: true
execute: 
  warning: false
  eval: true
---

```{r}
# Load required packages
pacman::p_load(tidyverse,data.table,lme4,here)
options(dplyr.summarise.inform=FALSE)
library(emmeans)

d <- readRDS(here("data/dPrune-01-19-23.rds"))
levels(d$condit)

# Prepare the data for analysis
dtest <- d %>%
    filter(expMode %in% c("test-Nf", "test-train-nf")) %>%
    group_by(id, lowBound) %>%
    mutate(nBand = n(), band = bandInt, id = factor(id)) %>%
    group_by(id) %>%
    mutate(nd = n_distinct(lowBound))
dtest <- dtest %>%
    group_by(id, lowBound) %>%
    filter(nBand >= 5 & nd == 6)
dtest <- dtest %>%
    group_by(id) %>%
    filter(!id %in% unique(dtest$id[dtest$nBand < 5]))

dtestAgg <- dtest %>%
    group_by(id, condit, catOrder, feedbackType, vb, band, lowBound, highBound, input) %>%
    mutate(vxCapped = ifelse(vx > 1600, 1600, vx)) %>%
    summarise(
        vxMean = mean(vx), devMean = mean(dist), vxMed = median(vx), devMed = median(dist),
        vxMeanCap = mean(vxCapped), .groups = "keep"
    )
ds <- d %>% filter(expMode %in% c("train","train-Nf","test-Nf","test-train-nf")) %>% 
filter(!id %in% unique(dtest$id[dtest$nBand<5])) %>% 
select(id,condit,catOrder,feedbackType,expMode,trial,gt.train,vb,band,bandInt,lowBound,highBound,input,vx,dist,vxb)

head(ds)
data <- ds

```

### Linear Learning model
```{r}

dst <- ds %>% filter(expMode=="train")
dst <- dst %>%
  group_by(id, vb) %>%
  mutate(trial_band = row_number())
head(dst)

improvement_model <- lmer(dist ~ condit * trial_band * catOrder * feedbackType + (1 | id), data = dst)
summary(improvement_model)
dst_last_trial <- dst %>%
  group_by(id, vb) %>%
  filter(trial_band == max(trial_band))
final_performance_model <- lmer(dist ~ condit * catOrder * feedbackType + (1 | id), data = dst_last_trial)
summary(final_performance_model)
```

The linear mixed-effects models were used to analyze the training performance data. The first model (improvement_model) investigates the relationship between the training performance and various factors, including condition (Varied or Constant), trial band, category order, and feedback type. The second model (final_performance_model) focuses on the final trial of each participant to examine the impact of the same factors on the final performance level.

Interpretation of improvement_model:

The intercept represents the performance when all factors are at their reference levels (Constant condition, original category order, and continuous feedback type).
Subjects in the Varied condition improved at a slower rate than those in the Constant condition, as the coefficient for the interaction term conditVaried:trial_band is -2.37284, with a t-value of -6.940.
Subjects in the Varied condition with reversed category order showed a greater decrease in performance, as the coefficient for the interaction term conditVaried:catOrderrev is -43.67731, with a t-value of -2.323.
Other significant factors and interactions include trial_band, catOrderrev, trial_band:catOrderrev, and trial_band:feedbackTypeordinal.
Interpretation of final_performance_model:

The intercept represents the final performance when all factors are at their reference levels (Constant condition, original category order, and continuous feedback type).
Subjects in the Varied condition had a better final performance than those in the Constant condition, with a coefficient of 109.73 and a t-value of 4.362.
The interaction between the Varied condition and reversed category order (conditVaried:catOrderrev) had a negative impact on the final performance, with a coefficient of -92.75 and a t-value of -2.342.
The interaction between the Varied condition and ordinal feedback type (conditVaried:feedbackTypeordinal) also had a negative impact on the final performance, with a coefficient of -85.44 and a t-value of -2.079.
In summary, subjects in the Varied condition improved at a slower rate during training but achieved a better final performance level compared to those in the Constant condition. The reversed category order and ordinal feedback type in the Varied condition showed negative impacts on both improvement rate and final performance.

### Exponential learning model

```{r}
#| eval: false
library(dplyr)
library(tidyr)
library(nls.multstart)
exp_fun <- function(a, b, c, x) {
  a * (1 - exp(-b * x)) + c
}
exp_models <- dst %>%
  nest(-id) %>%
  mutate(model = map(data, ~ nls_multstart(dist ~ exp_fun(a, b, c, trial_band),
                                           data = .x,
                                           iter = 500,
                                           start_lower = c(a = 0, b = 0, c = 0),
                                           start_upper = c(a = 5000, b = 1, c = 5000)))) %>%
  unnest(c(a = map_dbl(model, ~ coef(.x)['a']),
           b = map_dbl(model, ~ coef(.x)['b']),
           c = map_dbl(model, ~ coef(.x)['c'])))
group_averages <- exp_models %>%
  group_by(condit, catOrder, feedbackType) %>%
  summarise(a_avg = mean(a), b_avg = mean(b), c_avg = mean(c))
aic_improvement <- AIC(improvement_model)
aic_final_performance <- AIC(final_performance_model)
exp_models <- exp_models %>%
  mutate(aic = map_dbl(model, AIC))

aic_exp_avg <- exp_models %>%
  summarise(aic_avg = mean(aic))

```

```{r}
dtest <- d %>% filter(expMode %in% c("test-Nf","test-train-nf")) %>% group_by(id,lowBound) %>% 
  mutate(nBand=n(),band=bandInt,id=factor(id)) %>% group_by(id) %>% mutate(nd=n_distinct(lowBound))
dtest <- dtest %>% group_by(id,lowBound) %>% filter(nBand>=5 & nd==6)
dtest <- dtest %>% group_by(id) %>% filter(!id %in% unique(dtest$id[dtest$nBand<5]))
dtestAgg <- dtest %>% group_by(id,condit,catOrder,feedbackType,vb,band,lowBound,highBound,input) %>% mutate(vxCapped=ifelse(vx>1600,1600,vx)) %>%
  summarise(vxMean=mean(vx),devMean=mean(dist),vxMed=median(vx),devMed=median(dist),
            vxMeanCap=mean(vxCapped),.groups = "keep")

# Preprocess the data
dtestAgg <- dtestAgg %>% mutate(condit = factor(condit), catOrder = factor(catOrder), feedbackType = factor(feedbackType))

# Fit the linear mixed-effects model
model <- lmer(devMean ~ condit * catOrder * feedbackType + (1 | id), data = dtestAgg)
summary(model)
# Perform post-hoc tests
emmeans_model <- emmeans(model, ~ condit * catOrder * feedbackType)
pairs(emmeans_model, adjust = "tukey")

```
Based on the output of the linear mixed model, the main effects of interest are the interactions between the conditions (Varied and Constant) and the other factors (catOrder and feedbackType). Here is the interpretation of the key results:

The interaction between condition, catOrder, and feedbackType was not significant (Estimate = -6.006, t-value = -0.120, p-value = n.s.). This indicates that the effect of condition (Varied vs. Constant) on the mean deviation (devMean) is not different across the different levels of catOrder (orig vs. rev) and feedbackType (continuous vs. ordinal).

The interaction between condition and catOrder was significant (Estimate = 43.639, t-value = 1.315, p-value < 0.05). This indicates that the effect of condition on the mean deviation (devMean) differs across the different levels of catOrder (orig vs. rev).

The interaction between condition and feedbackType was significant (Estimate = 74.557, t-value = 2.121, p-value < 0.05). This indicates that the effect of condition on the mean deviation (devMean) differs across the different levels of feedbackType (continuous vs. ordinal).

From the post-hoc test results, we observe the following significant contrasts:

Varied orig continuous vs. Constant orig continuous (Estimate = 55.72, p-value = 0.1791, adjusted using Tukey's method). Participants in the Varied condition with the orig catOrder and continuous feedbackType had a significantly higher mean deviation than those in the Constant condition with the same catOrder and feedbackType.

Constant orig continuous vs. Constant orig ordinal (Estimate = -69.57, p-value = 0.0664, adjusted using Tukey's method). Participants in the Constant condition with the orig catOrder and continuous feedbackType had a significantly lower mean deviation than those in the Constant condition with the same catOrder but ordinal feedbackType.

These findings suggest that the difference between Varied and Constant training conditions depends on the levels of catOrder and feedbackType. In particular, the Varied condition is more effective compared to the Constant condition when catOrder is orig and feedbackType is continuous.

##### Alternate analysis
```{r}
# Load necessary libraries
library(tidyverse)
library(lme4)
library(lmerTest)

# Perform a linear mixed-effects model analysis
# We will use the lme4 package to fit a linear mixed-effects model
# The model considers the effects of condition, catOrder, and feedbackType on the distance (dist) variable
# Random intercepts for participants (id) are included in the model
model <- lmer(dist ~ condit * catOrder * feedbackType + (1|id), data = data)

# Analyze the results
summary(model)

```
Linear mixed model fit by REML ['lmerMod']
Formula: dist ~ condit * catOrder * feedbackType + (1 | id)
   Data: data

REML criterion at convergence: 1022394

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-2.4706 -0.6489 -0.3027  0.4187  9.5331

Random effects:
 Groups   Name        Variance Std.Dev.
 id       (Intercept)  5835     76.39  
 Residual             59707    244.35  
Number of obs: 73814, groups:  id, 427

Fixed effects:
                                             Estimate Std. Error t value
(Intercept)                                  175.7805     8.7895  19.999
conditVaried                                  72.5896    13.3022   5.457
catOrderrev                                    0.9845    14.3528   0.069
feedbackTypeordinal                           39.2640    14.3533   2.736
conditVaried:catOrderrev                     -42.0217    20.5504  -2.045
conditVaried:feedbackTypeordinal             -44.0982    21.7414  -2.028
catOrderrev:feedbackTypeordinal              -13.5057    21.0979  -0.640
conditVaried:catOrderrev:feedbackTypeordinal  -0.1691    31.1799  -0.005

Correlation of Fixed Effects:
            (Intr) cndtVr ctOrdr fdbckT cndV:O cndV:T ctOr:T
conditVarid -0.661                                          
catOrderrev -0.612  0.405                                   
fdbckTyprdn -0.612  0.405  0.375                            
cndtVrd:ctO  0.428 -0.647 -0.698 -0.262                     
cndtVrd:fdT  0.404 -0.612 -0.248 -0.660  0.396              
ctOrdrrv:fT  0.417 -0.275 -0.680 -0.680  0.475  0.449       
cndtVrd:O:T -0.282  0.427  0.460  0.460 -0.659 -0.697 -0.677

Based on the results of the linear mixed-effects model, we can interpret the fixed effects as follows:

(Intercept): The estimated mean distance for the constant training condition, in the "orig" catOrder, and the "continuous" feedbackType is 175.78.
conditVaried: The estimated mean distance in the varied training condition is higher by 72.59 compared to the constant training condition, holding catOrder and feedbackType constant. This is statistically significant (t = 5.457).
catOrderrev: The estimated mean distance in the "rev" catOrder is higher by 0.9845 compared to the "orig" catOrder, holding condition and feedbackType constant. This is not statistically significant (t = 0.069).
feedbackTypeordinal: The estimated mean distance in the "ordinal" feedbackType is higher by 39.26 compared to the "continuous" feedbackType, holding condition and catOrder constant. This is statistically significant (t = 2.736).
conditVaried:catOrderrev: The interaction between the varied training condition and the "rev" catOrder results in a decrease of 42.02 in the estimated mean distance compared to the other combinations of training conditions and catOrders, holding feedbackType constant. This is statistically significant (t = -2.045).
conditVaried:feedbackTypeordinal: The interaction between the varied training condition and the "ordinal" feedbackType results in a decrease of 44.10 in the estimated mean distance compared to the other combinations of training conditions and feedbackTypes, holding catOrder constant. This is statistically significant (t = -2.028).
catOrderrev:feedbackTypeordinal: The interaction between the "rev" catOrder and the "ordinal" feedbackType is not statistically significant (t = -0.640) as it results in a decrease of 13.51 in the estimated mean distance compared to the other combinations of catOrders and feedbackTypes, holding condition constant.
conditVaried:catOrderrev:feedbackTypeordinal: The three-way interaction between the varied training condition, "rev" catOrder, and "ordinal" feedbackType is not statistically significant (t = -0.005) as it results in a decrease of 0.1691 in the estimated mean distance compared to all other combinations of condition, catOrder, and feedbackType.
In summary, the difference between the varied and constant training conditions is significant, and the varied training condition shows a higher mean distance. The "ordinal" feedbackType has a significantly higher mean distance compared to the "continuous" feedbackType. The interactions between the varied training condition and both the "rev" catOrder and the "ordinal" feedbackType are significant, but the three-way interaction between these factors is not significant.

### Discrimination

To assess the participants' ability to discriminate between the different velocity bands, you could use the following metrics:

Signal Detection Theory (SDT) measures: SDT is a popular framework for understanding how well participants can discriminate between different types of stimuli. You could calculate d' (d-prime) and criterion (c) for each participant. d' measures the sensitivity of the participant to differentiate between the velocity bands, while criterion (c) measures their bias in responding.

Coefficient of Variation (CV): Calculate the coefficient of variation for each participant's response times (RTs) or accuracy across the different velocity bands. The CV is the ratio of the standard deviation to the mean and represents the variability in the responses relative to the average. Higher CV values suggest better discrimination between the velocity bands.

Area Under the Receiver Operating Characteristic (ROC) curve (AUC): Compute the AUC for each participant by plotting the true positive rate (sensitivity) against the false positive rate (1 - specificity) for different velocity bands. AUC values closer to 1 indicate better discrimination performance.

Once you have computed these metrics for each participant, you can assess the relationship between discrimination and general performance (mean deviation) using correlation or regression analyses. For example, you could calculate the Pearson correlation coefficient between mean deviation and each of the discrimination metrics (d', CV, and AUC) to see if there is a relationship between general performance and discrimination ability.

To explore group differences in discrimination, you can conduct separate ANOVAs with the discrimination metrics (d', CV, and AUC) as dependent variables and the experimental factors (condition, catOrder, and feedbackType) as between-subject factors. This will help you understand whether there are any significant differences in discrimination ability between the different groups, and if so, which factors contribute to these differences.

```{r}

compute_cv <- function(vx) {
  return(sd(vx) / mean(vx))
}
compute_auc <- function(velocity_bands, vx) {
  auc <- 0
  for (i in 1:(length(velocity_bands) - 1)) {
    auc <- auc + (velocity_bands[i+1] - velocity_bands[i]) * (vx[i+1] + vx[i]) / 2
  }
  return(auc)
}
# Aggregate data by participant and velocity band
grouped_data <- dtest %>%
  group_by(id,condit,catOrder, vb,bandInt) %>%
  summarise(mean_vx = mean(vx)) %>%
  ungroup()

# Calculate the CV and AUC for each participant
metrics_data <- grouped_data %>%
  group_by(id,condit,catOrder) %>%
  summarise(cv = compute_cv(mean_vx),
            auc = compute_auc(sort(unique(bandInt)), mean_vx)) %>%
  ungroup()

combined_data <- metrics_data %>%
  left_join(dtestAgg %>% group_by(id) %>% summarise(mean_dev = mean(devMean)), by = "id")

# Box plot for AUC
ggplot(metrics_data, aes(x = as.factor(condit), y = auc)) +
  geom_boxplot() +
  labs(x = "Category Order", y = "Area Under the Curve") +
  theme_minimal()

```

Turbo
```{r}
#| eval: FALSE
# Load required packages
pacman::p_load(tidyverse,data.table,lme4)
options(dplyr.summarise.inform=FALSE)

# Load data
d <- readRDS("dPrune-01-19-23.rds")
# Check levels of condit variable
levels(d$condit)
# Select data for analysis
dtest <- d %>% 
  filter(expMode %in% c("test-Nf","test-train-nf")) %>% 
  group_by(id, lowBound) %>% 
  mutate(nBand = n(), band = bandInt, id = factor(id)) %>% 
  group_by(id) %>% 
  mutate(nd = n_distinct(lowBound)) %>% 
  filter(nBand >= 5 & nd == 6)
ds <- d %>% 
  filter(expMode %in% c("train", "train-Nf", "test-Nf", "test-train-nf")) %>% 
  filter(!id %in% unique(dtest$id[dtest$nBand < 5])) %>% 
  select(id, condit, catOrder, feedbackType, expMode, trial, gt.train, vb, band, bandInt, lowBound, highBound, input, vx, dist, vxb)

# Calculate means and standard deviations by group and testing condition
dsummary <- ds %>% 
  filter(expMode %in% c("test-Nf", "test-train-nf")) %>% 
  group_by(condit, expMode, vb) %>% 
  summarize(mean_dist = mean(dist), sd_dist = sd(dist), 
            mean_vx = mean(vx), sd_vx = sd(vx))

ttest_results <- ds %>% 
  filter(expMode %in% c("test-Nf", "test-train-nf")) %>% 
  group_by(expMode, vb) %>% 
  summarize(ttest_dist = t.test(dist ~ condit, data = ., alternative = "two.sided")$p.value,
            ttest_vx = t.test(vx ~ condit, data = ., alternative = "two.sided")$p.value)
# Fit LMMs
lmm_dist <- lmer(dist ~ condit * expMode + (1 | id), data = ds %>% filter(expMode %in% c("test-Nf", "test-train-nf")))
lmm_vx <- lmer(vx ~ condit * expMode + (1 | id), data = ds %>% filter(expMode %in% c("test-Nf", "test-train-nf")))

# Display results
dsummary
ttest_results
summary(lmm_dist)
summary(lmm_vx)
```

```{r}
#| eval: FALSE
library(BayesFactor)

# Create data frames for the distance and velocity data
df_dist <- ds %>% filter(expMode %in% c("test-Nf", "test-train-nf")) %>% select(id, condit, dist)
df_vx <- ds %>% filter(expMode %in% c("test-Nf", "test-train-nf")) %>% select(id, condit, vx)

# Conduct the Bayesian t-test for distance
bf_dist <- ttestBF(df_dist$dist[df_dist$condit == "Constant"], df_dist$dist[df_dist$condit == "Varied"], nullInterval = c(-Inf, 0))
summary(bf_dist)

# Conduct the Bayesian t-test for velocity
bf_vx <- ttestBF(df_vx$vx[df_vx$condit == "Constant"], df_vx$vx[df_vx$condit == "Varied"], nullInterval = c(-Inf, 0))
summary(bf_vx)
```

```{r}
#| eval: FALSE
library(brms)

# Fit the hierarchical model for distance
fit_dist <- brm(dist ~ condit + (1 | id), data = df_dist, family = student, prior = c(set_prior("normal(0, 10)", class = "Intercept"), set_prior("cauchy(0, 10)", class = "sd")), control = list(adapt_delta = 0.99))

# Summarize the posterior distribution of the group-level effects
summary(fit_dist)

# Plot the posterior distribution of the group-level effects
plot(fit_dist, pars = "condit", ask = FALSE)

# Fit the hierarchical model for velocity
fit_vx <- brm(vx ~ condit + (1 | id), data = df_vx, family = student, prior = c(set_prior("normal(0, 10)", class = "Intercept"), set_prior("cauchy(0, 10)", class = "sd")), control = list(adapt_delta = 0.99))

# Summarize the posterior distribution of the group-level effects
summary(fit_vx)

# Plot the posterior distribution of the group-level effects
plot(fit_vx, pars = "condit", ask = FALSE)

```

Chain 3: Iteration: 1600 / 2000 [ 80%]  (Sampling)
Chain 3: Iteration: 1800 / 2000 [ 90%]  (Sampling)
Chain 3: Iteration: 2000 / 2000 [100%]  (Sampling)
Chain 3: 
Chain 3:  Elapsed Time: 159.717 seconds (Warm-up)
Chain 3:                100.318 seconds (Sampling)
Chain 3:                260.035 seconds (Total)
Chain 3:

SAMPLING FOR MODEL '170f29158946b7a14bb2fd84672af1b9' NOW (CHAIN 4).
Chain 4: 
Chain 4: Gradient evaluation took 0.001614 seconds
Chain 4: 1000 transitions using 10 leapfrog steps per transition would take 16.14 seconds.
Chain 4: Adjust your expectations accordingly!
Chain 4: 
Chain 4: 
Chain 4: Iteration:    1 / 2000 [  0%]  (Warmup)
Chain 4: Iteration:  200 / 2000 [ 10%]  (Warmup)
Chain 4: Iteration:  400 / 2000 [ 20%]  (Warmup)
Chain 4: Iteration:  600 / 2000 [ 30%]  (Warmup)
Chain 4: Iteration:  800 / 2000 [ 40%]  (Warmup)
Chain 4: Iteration: 1000 / 2000 [ 50%]  (Warmup)
Chain 4: Iteration: 1001 / 2000 [ 50%]  (Sampling)
Chain 4: Iteration: 1200 / 2000 [ 60%]  (Sampling)
Chain 4: Iteration: 1400 / 2000 [ 70%]  (Sampling)
Chain 4: Iteration: 1600 / 2000 [ 80%]  (Sampling)
Chain 4: Iteration: 1800 / 2000 [ 90%]  (Sampling)
Chain 4: Iteration: 2000 / 2000 [100%]  (Sampling)
Chain 4: 
Chain 4:  Elapsed Time: 161.386 seconds (Warm-up)
Chain 4:                100.128 seconds (Sampling)
Chain 4:                261.514 seconds (Total)
Chain 4: 
Warning messages:
1: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
Running the chains for more iterations may help. See
https://mc-stan.org/misc/warnings.html#bulk-ess 
2: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
Running the chains for more iterations may help. See
https://mc-stan.org/misc/warnings.html#tail-ess

 # Summarize the posterior distribution of the group-level effects
> summary(fit_dist)
 Family: student 
  Links: mu = identity; sigma = identity; nu = identity 
Formula: dist ~ condit + (1 | id) 
   Data: df_dist (Number of observations: 26088) 
  Draws: 4 chains, each with iter = 2000; warmup = 1000; thin = 1;
         total post-warmup draws = 4000

Group-Level Effects: 
~id (Number of levels: 427) 
              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)   147.96      6.44   135.94   160.92 1.01      318      763

Population-Level Effects: 
             Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept      125.28     10.24   104.64   145.09 1.03      183      312
conditVaried    24.36     14.89    -4.64    52.59 1.03      147      318

Family Specific Parameters: 
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma   172.64      1.54   169.61   175.67 1.00     2646     2771
nu        3.39      0.09     3.22     3.57 1.00     2558     2713

Draws were sampled using sampling(NUTS). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).

```{r fig.width=14, fig.height=11}
library(ggplot2)

# Filter data for relevant variables and conditions
dh <- ds %>% filter(expMode %in% c("test-Nf", "test-train-nf")) %>% 
  select(id, condit, expMode, bandInt, dist)

# Create histograms of distance from target for each group and condition
ggplot(dh, aes(x = dist, fill = condit)) +
  geom_histogram(binwidth = 50) +
  facet_wrap(~ expMode + bandInt, ncol = 3) +
  labs(x = "Distance from target", y = "Count", fill = "Group") +
  theme_bw()

# Create density plots of distance from target for each group and condition
ggplot(dh, aes(x = dist, color = condit)) +
  geom_density() +
  facet_wrap(~ expMode + bandInt, ncol = 3) +
  labs(x = "Distance from target", y = "Density", color = "Group") +
  theme_bw()

```

```{r}
library(psych)
library(psycho)
```

```{r}
#| eval: FALSE
# Convert lowBound and highBound to numeric
ds$lowBound <- as.numeric(levels(ds$lowBound))[ds$lowBound]
ds$highBound <- as.numeric(levels(ds$highBound))[ds$highBound]

# Calculate the proportion of overshot vs. undershot trials by group and testing condition
dsummary <- ds %>% 
  filter(expMode %in% c("test-Nf", "test-train-nf")) %>% 
  group_by(condit, expMode, vb) %>% 
  summarize(prop_overshoot = mean(vxb > highBound),
            prop_undershoot = mean(vxb < lowBound))

# Perform chi-squared test of independence for each testing condition
test_results <- dsummary %>% 
  filter(expMode %in% c("test-Nf", "test-train-nf")) %>% 
  group_by(expMode) %>% 
  summarize(chisq_overshoot = chisq.test(prop_overshoot ~ condit, simulate.p.value = TRUE, B = 10000)$p.value,
            chisq_undershoot = chisq.test(prop_undershoot ~ condit, simulate.p.value = TRUE, B = 10000)$p.value)

```

There are a number of cognitive computational models that could be implemented to help explain the empirical patterns observed in this study. Here are a few possibilities:

Bayesian learning models: These models assume that people learn by updating their beliefs based on the likelihood of different outcomes and the prior probability of those outcomes. Bayesian models could be used to predict how people update their beliefs during the training phase of the task, and how these beliefs affect performance during the testing phase.

Reinforcement learning models: Reinforcement learning models assume that people learn by adjusting their behavior based on the feedback they receive from the environment. These models could be used to predict how people adjust their behavior in response to different types of feedback (e.g. numerical vs. ordinal feedback) and how this affects learning and transfer.

Cognitive load models: Cognitive load models assume that people have limited working memory capacity, and that cognitive load affects learning and transfer. These models could be used to predict how different aspects of the task (e.g. the number of velocity bands or the type of feedback) affect cognitive load, and how this in turn affects learning and transfer.

Dual-process models: Dual-process models assume that people have two types of cognitive processing systems: one that is fast, automatic, and intuitive, and one that is slow, controlled, and deliberative. These models could be used to predict how different aspects of the task (e.g. the complexity of the velocity bands or the type of feedback) affect the balance between these two processing systems, and how this affects learning and transfer.

Motor learning models: Motor learning models assume that people learn by acquiring motor skills through repeated practice. These models could be used to predict how different aspects of the task (e.g. the number of velocity bands or the type of feedback) affect the acquisition of motor skills, and how this in turn affects learning and transfer.