### 2023-05-28 Visuals_Interactives/model_viz.qmd
mermaid visuals and rbf attempts

---
title: Model Visualization
date: last-modified
categories: [Visualization, R, OJS]
page-layout: full
code-fold: true
code-tools: true
execute: 
  warning: false
---

```{r}
#lapply(c('tidyverse','data.table','igraph','ggraph','kableExtra'),library,character.only=TRUE))
pacman::p_load(tidyverse,data.table,igraph,ggraph,kableExtra,DiagrammeR,png) 
```

## ggplot model visualization
```{r}
nInput=6
nOutput=10

inNodes <- seq(1,nInput,1) %>% as.integer()
outNodes <- seq(300,1000,length.out=nOutput)%>% as.integer()
weight.mat <<- matrix(0.001,nrow=nOutput,ncol=nInput) # weights initialized to 0 (as in Delosh 1997)

stim <- 1.5
c=.1
inAct <- round(exp(-c*((inNodes-stim)^2)),2)
inActLab <- paste0("x",inNodes,"=",inAct)
outAct <- weight.mat %*% inAct
output.probability <<- outAct/sum(outAct)
outLab=paste0("y",outNodes,"=",round(output.probability,2))
mean.response <<- round(sum(outNodes * output.probability),0)

resp <- mean.response
inFlow <- tibble(expand.grid(from=stim,to=inActLab)) %>% mutate_all(as.character)
outFlow <- tibble(expand.grid(from=outLab,to=mean.response)) %>% mutate_all(as.character)

gd <- tibble(expand.grid(from=inActLab,to=outLab)) %>% mutate_all(as.character) %>%
  rbind(inFlow,.) %>% rbind(.,outFlow)

xInc <- .3
yInc=.5

g = graph_from_data_frame(gd,directed=TRUE)
coords2=layout_as_tree(g)
colnames(coords2)=c("y","x")

odf <- as_tibble(coords2) %>% 
  mutate(label=vertex_attr(g,"name"),
         type=c("stim",rep("Input",nInput),rep("Output",nOutput),"Resp"),
         x=x*-1) %>%
  mutate(y=ifelse(type=="Resp",0,y),xmin=x-xInc,xmax=x+xInc,ymin=y-yInc,ymax=y+yInc)

plot_edges = gd %>% mutate(id=row_number()) %>%
  pivot_longer(cols=c("from","to"),names_to="s_e",values_to=("label")) %>%
                 mutate(label=as.character(label)) %>% 
  group_by(id) %>%
  mutate(weight=sqrt(rnorm(1,mean=0,sd=10)^2)/10) %>%
  left_join(odf,by="label") %>%
  mutate(xmin=xmin+.02,xmax=xmax-.02)

ggplot() + geom_rect(data = odf,
            mapping = aes(xmin = xmin, ymin = ymin, 
                          xmax = xmax, ymax = ymax, 
                          fill = type, colour = type),alpha = 0.01) +
  geom_text(data=odf,aes(x=x,y=y,label=label,size=3)) +
  geom_path(data=plot_edges,mapping=aes(x=x,y=y,group=id,alpha=weight)) +
  # geom_rect(aes(xmin=-1.05,xmax=-.95,ymin=-10,ymax=5),color="red",alpha=.1)+
  # geom_rect(aes(xmin=-0.05,xmax=.05,ymin=-10,ymax=5),color="blue",alpha=.1) +
  theme_void()

```

## ggraph method
```{r }

inNodes <- seq(1,6,1) %>% as.integer()
outNodes <- seq(300,1000,50)%>% as.integer()

da <- data.frame(expand.grid(inNodes,outNodes))  %>% magrittr::set_colnames(c("input","output"))
da <- da %>% mutate_all(as.character)
m = graph_from_data_frame(da, directed = TRUE)

coords = layout_with_sugiyama(m)
colnames(coords$layout) = c("y", "x")
coords$layout=coords$layout[,c("x","y")]
plot(m,layout=coords)

ggraph(m,layout=coords$layout)+
geom_edge_link0(width=0.2,colour="grey")+
  geom_node_point(col="white",size=6)+scale_x_reverse()+
  geom_node_text(aes(label=name)) +
  # draw rectangle that covers input layer at x=1, min y is min of coords$y and max y is max of coords$y
  annotate("rect",xmin=0,xmax=.1,ymin=min(coords$layout[,2]),ymax=max(coords$layout[,2]),fill="grey",alpha=0.7)
 geom_rect(xmin=0,xmax=1.1,ymin=min(coords$layout[,2]),ymax=max(coords$layout[,2]),fill="grey",alpha=0.7)

```

```{r fig.width=11, fig.height=12}
# Load the required libraries
library(ggplot2)
library(ggpp)
library(tidyverse)
library(patchwork)

# Create the Gaussian curve plots for each input node
input_node_1 <- ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 1, sd = .4)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0))

input_node_2 <- ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 2, sd = .4)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0))

input_node_3 <- ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 3, sd = .4)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0))

# Create the main plot
main_plot <- ggplot() +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  annotate("plot_npc", npcx = 0.2, npcy = 0.8, label = input_node_1) +
  annotate("plot_npc", npcx = 0.5, npcy = 0.8, label = input_node_2) +
  annotate("plot_npc", npcx = 0.8, npcy = 0.8, label = input_node_3) +
  annotate("text", x = 0.2, y = 0.7, label = "Input Node 1") +
  annotate("text", x = 0.5, y = 0.7, label = "Input Node 2") +
  annotate("text", x = 0.8, y = 0.7, label = "Input Node 3") +
  annotate("text", x = 0.5, y = 0.6, label = "Input Layer", fontface = "bold") +
  annotate("text", x = 0.5, y = 0.3, label = "Output Layer", fontface = "bold") +
  annotate("text", x = 0.5, y = 0.1, label = "ALM + EXAM Response", fontface = "bold")

# Print the main plot
print(main_plot)

# Create the Gaussian curve plots for each input node
input_node_1 <- ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0))

input_node_2 <- ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 2.2, sd = 1)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0))

input_node_3 <- ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 4.3, sd = .1)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0))

# Create the main plot
main_plot <- ggplot() +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  annotate("plot_npc", npcx = 0.2, npcy = 0.8, label = input_node_1, width = 0.1, height = 0.1) +
  annotate("plot_npc", npcx = 0.5, npcy = 0.8, label = input_node_2, width = 0.1, height = 0.1) +
  annotate("plot_npc", npcx = 0.8, npcy = 0.8, label = input_node_3, width = 0.1, height = 0.1) +
  annotate("text", x = 0.2, y = 0.9, label = "Input Node 1") +
  annotate("text", x = 0.5, y = 0.9, label = "Input Node 2") +
  annotate("text", x = 0.8, y = 0.9, label = "Input Node 3") +
  annotate("text", x = 0.5, y = 0.6, label = "Input Layer", fontface = "bold") +
  annotate("rect", xmin = 0.1, xmax = 0.3, ymin = 0.4, ymax = 0.5, fill = "grey80") +
  annotate("rect", xmin = 0.4, xmax = 0.6, ymin = 0.4, ymax = 0.5, fill = "grey80") +
  annotate("rect", xmin = 0.7, xmax = 0.9, ymin = 0.4, ymax = 0.5, fill = "grey80") +
  annotate("text", x = 0.5, y = 0.3, label = "Output Layer", fontface = "bold") +
  annotate("text", x = 0.5, y = 0.1, label = "ALM + EXAM Response", fontface ="bold")
print(main_plot)

# Create the Gaussian curve plots for each input node
input_node_1 <- ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = .41)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  ggtitle("Input Node 1")

input_node_2 <- ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 2, sd = .41)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  ggtitle("Input Node 2")

input_node_3 <- ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
  stat_function(fun = dnorm, args = list(mean = 3, sd = .41)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  ggtitle("Input Node 3")

# Create the output nodes
output_node_1 <- ggplot() +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  annotate("rect", xmin = -1, xmax = 1, ymin = -1, ymax = 1, fill = "blue") +
  ggtitle("Output Node 1")

output_node_2 <- ggplot() +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  annotate("rect", xmin = -1, xmax = 1, ymin = -1, ymax = 1, fill = "blue") +
  ggtitle("Output Node 2")

# Combine the input and output nodes using patchwork
input_layer <- input_node_1 + input_node_2 + input_node_3
output_layer <- output_node_1 + output_node_2

# Create the connection matrix
connection_matrix <- ggplot() +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  annotate("segment", x = 0.2, xend = 0.8, y = 0.8, yend = 0.2, arrow = arrow()) +
  annotate("segment", x = 0.5, xend = 0.8, y = 0.8, yend = 0.2, arrow = arrow()) +
  annotate("segment", x = 0.8, xend = 0.8, y = 0.8, yend = 0.2, arrow = arrow())

# Combine the input layer, connection matrix, and output layer
main_plot <- input_layer / connection_matrix / output_layer

# Print the main plot
print(main_plot)

# Function to create an input node
create_input_node <- function(mean, position) {
  ggplot(data.frame(x = seq(-3, 3, length.out = 100)), aes(x)) +
    stat_function(fun = dnorm, args = list(mean = mean, sd = .3)) +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0)) +
    coord_cartesian(xlim = c(0, 3), ylim = c(0, 0.5)) + coord_flip()+
    labs(title = paste("Input Node", position)) +
    theme(plot.title = element_text(hjust = 0.5))
}

# Function to create an output node
create_output_node <- function(position) {
  ggplot() +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0)) +
    annotate("rect", xmin = -1, xmax = 1, ymin = -1, ymax = 1, fill = "grey80") +
    labs(title = paste("Output Node", position)) +
    theme(plot.title = element_text(hjust = 0.5))
}

create_input_layer <- function(n) {
  input_plots <- map(1:n, ~create_gaussian_plot(.x))
  input_layer <- wrap_plots(input_plots, ncol = 1) + coord_flip()
  return(input_layer)
}

# Create the input nodes
input_nodes <- map(1:3, ~create_input_node(.x, .x))

# Create the output nodes
output_nodes <- map(1:4, ~create_output_node(.x))

## Create the connection matrix
connection_matrix <- ggplot() +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  annotate("segment", x = 1, xend = 2, y = rep(1:3, each = 4), yend = rep(1:4, times = 3), arrow = arrow())
# Combine the plots
input_layer <- wrap_plots(input_nodes,byrow=FALSE,ncol=1)
output_layer <- wrap_plots(output_nodes, ncol = 1)
main_plot <- input_layer + connection_matrix + output_layer + plot_layout(ncol = 3)

# Print the main plot
print(main_plot)

combined_plot <- input_layer + connection_matrix + output_layer +
  plot_layout(ncol = 3)

print(combined_plot)

print(input_layer)

```

```{r}
# Load the required libraries
library(ggplot2)
library(ggpp)
library(patchwork)
library(tidyverse)

# Function to create the input layer
create_input_layer <- function(n_nodes, stimulus = 0, gamma = 5) {
  # Create a data frame with the x values and node means
  df <- expand.grid(x = seq(-3, 3, length.out = 100),
                    mean = seq(-2, 1.5, length.out = n_nodes)) %>% 
    mutate(noisyMean=map(mean,~dnorm(.x,mean,sd=.5)))

  df <- df %>%
    mutate(distance = abs(x - stimulus),
           noise = rnorm(n(), sd = 0.05),
           max_activation = exp(-gamma * distance^2) + noise,
           y = max_activation * exp(-gamma * (x - mean)^2))

  # Create the plot
  ggplot(df, aes(x, y, group = mean)) +
    geom_line() +
    geom_smooth()+
    coord_cartesian(xlim = c(-2, 2), ylim = c(0, 1)) +
    labs(title = "Input Layer") +
    theme(plot.title = element_text(hjust = 0.5))
}

# Function to create an output node
create_output_node <- function(position) {
  ggplot() +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0)) +
    annotate("rect", xmin = -1, xmax = 1, ymin = -1, ymax = 1, fill = "grey80") +
    labs(title = paste("Output Node", position)) +
    theme(plot.title = element_text(hjust = 0.5))
}

# Create the input layer
input_layer <- create_input_layer(3)

# Create the output nodes
output_nodes <- map(1:4, ~create_output_node(.x))

# Create the connection matrix
connection_matrix <- ggplot() +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0)) +
  annotate("segment", x = 1, xend = 2, y = 1:3, yend = rep(1:4, each = 3), arrow = arrow())

# Combine the plots
output_layer <- wrap_plots(output_nodes, ncol = 1)
main_plot <- input_layer + connection_matrix + output_layer + plot_layout(ncol = 3)

# Print the main plot
print(main_plot)

```

```{r}
# Load the visNetwork package
library(visNetwork)

# Define the nodes
nodes <- data.frame(
  id = c("InputLayer", "I1", "I2", "I3", "OutputLayer", "O1", "O2", "O3", "O4", "ALM", "EXAM"),
  label = c("Input Layer", "Input Node 1", "Input Node 2", "Input Node 3", "Output Layer", "Output Node 1", "Output Node 2", "Output Node 3", "Output Node 4", "ALM Response", "EXAM Response"),
  shape = c("box", "image", "image", "image", "box", "circle", "circle", "circle", "circle", "box", "box"),
  image = c(NA, "gaussian_curve_input_node_1.png", "gaussian_curve_input_node_2.png", "gaussian_curve_input_node_3.png", NA, NA, NA, NA, NA, NA, NA)
)

# Define the edges
edges <- data.frame(
  from = c("InputLayer", "I1", "I1", "I1", "I1", "I2", "I2", "I2", "I2", "I3", "I3", "I3", "I3", "OutputLayer", "OutputLayer", "ALM", "EXAM"),
  to = c("I1", "O1", "O2", "O3", "O4", "O1", "O2", "O3", "O4", "O1", "O2", "O3", "O4", "ALM", "EXAM", "EXAM", "ALM"),
  label = c("", "w1", "w2", "w3", "w4", "w5", "w6", "w7", "w8", "w9", "w10", "w11", "w12", "", "", "Pure ALM Model", "ALM with EXAM Response Component")
)

# Create the network diagram
visNetwork(nodes, edges, width = "100%") %>%
  visNodes(shapeProperties = list(useBorderWithImage = TRUE)) %>%
  visLayout(randomSeed = 2)

# Define the nodes
nodes <- data.frame(
  id = c("InputLayer", "I1", "I2", "I3", "OutputLayer", "O1", "O2", "O3", "O4", "ALM", "EXAM"),
  label = c("Input Layer", "<img src='gaussian_curve_input_node_1.png' />", "<img src='gaussian_curve_input_node_2.png' />", "<img src='gaussian_curve_input_node_3.png' />", "Output Layer", "Output Node 1", "Output Node 2", "Output Node 3", "Output Node 4", "ALM Response", "EXAM Response"),
  shape = c("box", "circle", "circle", "circle", "box", "circle", "circle", "circle", "circle", "box", "box")
)

# Define the edges
edges <- data.frame(
  from = c("InputLayer", "I1", "I1", "I1", "I1", "I2", "I2", "I2", "I2", "I3", "I3", "I3", "I3", "OutputLayer", "OutputLayer", "ALM", "EXAM"),
  to = c("I1", "O1", "O2", "O3", "O4", "O1", "O2", "O3", "O4", "O1", "O2", "O3", "O4", "ALM", "EXAM", "EXAM", "ALM"),
  label = c(NA, "w1", "w2", "w3", "w4", "w5", "w6", "w7", "w8", "w9", "w10", "w11", "w12", NA, NA, "Pure ALM Model", "ALM with EXAM Response Component")
)

# Create the network diagram
visNetwork(nodes, edges, width = "100%") %>%
  visNodes(shapeProperties = list(useBorderWithImage = TRUE)) %>%
  visLayout(randomSeed = 2)

```

```{r}

# Load the DiagrammeR package
library(DiagrammeR)
# Define the graph
grViz("
  digraph ALM_EXAM {

    # Graph attributes
    graph [overlap = true, fontsize = 10]

    # Node definitions
    node [shape = box, fontname = Helvetica]
    InputLayer [label = 'Input Layer']
    OutputLayer [label = 'Output Layer']
    ALM [label = 'ALM Response']
    EXAM [label = 'EXAM Response']

    node [shape = circle, fixedsize = true, width = 0.9]
    I1 [label = 'Input Node 1']
    I2 [label = 'Input Node 2']
    I3 [label = 'Input Node 3']
    O1 [label = 'Output Node 1']
    O2 [label = 'Output Node 2']
    O3 [label = 'Output Node 3']
    O4 [label = 'Output Node 4']

    # Edge definitions
    InputLayer -> I1
    InputLayer -> I2
    InputLayer -> I3
    I1 -> O1 [label = 'w1']
    I1 -> O2 [label = 'w2']
    I1 -> O3 [label = 'w3']
    I1 -> O4 [label = 'w4']
    I2 -> O1 [label = 'w5']
    I2 -> O2 [label = 'w6']
    I2 -> O3 [label = 'w7']
    I2 -> O4 [label = 'w8']
    I3 -> O1 [label = 'w9']
    I3 -> O2 [label = 'w10']
    I3 -> O3 [label = 'w11']
    I3 -> O4 [label = 'w12']
    OutputLayer -> O1
    OutputLayer -> O2
    OutputLayer -> O3
    OutputLayer -> O4
    OutputLayer -> ALM
    OutputLayer -> EXAM
    ALM -> EXAM [label = 'Pure ALM Model']
    EXAM -> ALM [label = 'ALM with EXAM Response Component']
  }
")

```

```{r}

# Load the required packages
library(ggplot2)
library(png)

# Define the Gaussian function
gaussian <- function(x, mean, sd) {
  1/(sd*sqrt(2*pi)) * exp(-1/2 * ((x - mean)/sd)^2)
}

# Define the central values for the input nodes
central_values <- c(1, 2, 3)

# Define the input stimulus value
input_stimulus <- 2

# Generate a Gaussian curve image for each input node
for (i in seq_along(central_values)) {
  # Define the mean and standard deviation for the Gaussian curve
  mean <- central_values[i]
  sd <- 1
  # Generate the x values
  x <- seq(mean - 3*sd, mean + 3*sd, length.out = 100)
  # Generate the y values
  y <- gaussian(x, mean, sd)
  # Create the plot
  p <- ggplot(data.frame(x, y), aes(x, y)) +
    geom_line() +
    theme_minimal() +
    labs(x = "X", y = "Activation", title = paste("Input Node", i))
  # Save the plot as a PNG image
  ggsave(paste0("gaussian_curve_input_node_", i, ".png"), plot = p, width = 4, height = 3)
}

```

`r knitr::include_graphics(("gaussian_curve_input_node_1.png"))`

```{r}
grViz("digraph causal {
                      	# Nodes
                      	node [imagescale=true,shape = reactangle, fontname = Arial, style = filled]
                      	iv   [label = 'TRT', fillcolor = '#7FC97F']
                      	me   [label = 'Mediator', shape = ellipse]
                      	dv   [label = 'DLQI', fillcolor = '#7FC97F']

                      	# Edges
                      	edge [color = black, arrowhead = normal]
                      	rankdir = LR
                      	iv -> me
                      	iv -> dv [label = 'DIRECT', fontcolor = '#7FC97F', color = '#7FC97F']
                      	me -> dv
                      	# Graph
                      	graph [overlap = true, fontsize = 10]
                      }")

grViz("digraph causal {
                      	# Nodes
                      	node [shape = reactangle, fontname = Arial, style = filled]
                      	iv   [image='gaussian_curve_input_node_1.png',label = '', fillcolor = '#7FC97F']
                      	me   [label = 'Mediator', shape = ellipse]
                      	dv   [label = 'DLQI', fillcolor = '#7FC97F']

                      	# Edges
                      	edge [color = black, arrowhead = normal]
                      	rankdir = LR
                      	iv -> me
                      	iv -> dv [label = 'DIRECT', fontcolor = '#7FC97F', color = '#7FC97F']
                      	me -> dv
                      	# Graph
                      	graph [overlap = true, fontsize = 10]
                      }")

```

```{r}

# grViz('digraph structs {
#     node [shape=plaintext];
# 
#     struct1 [label=<<TABLE>
#       <TR><TD><IMG SRC="gaussian_curve_input_node_1.png"/></TD></TR>
#       <TR><TD>caption</TD></TR>
#     </TABLE>>];
# }')

svgDataUrl = "data:image/svg+xml;base64,"
grViz('
digraph {
  node[imagescale=true, shape=circle, width=4, height=4,fontname="sans-serif",penwidth=0]
  a[label=<<FONT POINT-SIZE="10">Transaction</FONT><BR/><FONT POINT-SIZE="36"><B>TRNA</B></FONT>>,image="gaussian_curve_input_node_1.png"]
  b[label=<<FONT POINT-SIZE="10">Transaction</FONT><BR/><FONT POINT-SIZE="18"><B>TRNB</B></FONT>>,image="${svgDataUrl}"]
  c[label=<<FONT POINT-SIZE="10">Transaction</FONT><BR/><FONT POINT-SIZE="24"><B>TRNC</B></FONT>>,image="${svgDataUrl}"]
  a -> b
  a -> c
}')

grViz('digraph { a[image=\"gaussian_curve_input_node_1\"]; }", {
  images: [{ href: "gaussian_curve_input_node_1.png", width: "400px", height: "300px" }]
}');

grViz('
digraph g{
  node[imagescale=true, shape=circle, width=14, height=14,fontname="sans-serif",penwidth=0]
  I1[image="gaussian_curve_input_node_1.png", label=""];
}')

grViz('digraph {
    ratio="fill";
    size="10,10!";
    margin="0,0";

    node [shape=plain];
    root [label=<<TABLE border="0"><TR><TD><IMG SRC="gaussian_curve_input_node_1"/></TD></TR>
                                   <TR><TD>text under</TD></TR></TABLE>>];
}')
```

```{r}

grViz("
  digraph ALM_EXAM {

    # Graph attributes
    graph [overlap = true, fontsize = 10, rankdir = LR]

    # Node definitions
    node [shape = box, fontname = Helvetica]
    InputLayer [label = 'Input Layer']
    OutputLayer [label = 'Output Layer']
    ALM [label = 'ALM Response']
    EXAM [label = 'EXAM Response']

    node [shape = none, label = '']
    I1 [image = 'gaussian_curve_input_node_1.png', label='']
    I2 [image = 'gaussian_curve_input_node_2.png', label='']
    I3 [image = 'gaussian_curve_input_node_3.png', label='']

    node [shape = circle, fixedsize = true, width = 0.9]
    O1 [label = 'Output Node 1']
    O2 [label = 'Output Node 2']
    O3 [label = 'Output Node 3']
    O4 [label = 'Output Node 4']

    # Edge definitions
    InputLayer -> I1
    InputLayer -> I2
    InputLayer -> I3
    I1 -> O1 [label = 'w1']
    I1 -> O2 [label = 'w2']
    I1 -> O3 [label = 'w3']
    I1 -> O4 [label = 'w4']
    I2 -> O1 [label = 'w5']
    I2 -> O2 [label = 'w6']
    I2 -> O3 [label = 'w7']
    I2 -> O4 [label = 'w8']
    I3 -> O1 [label = 'w9']
    I3 -> O2 [label = 'w10']
    I3 -> O3 [label = 'w11']
    I3 -> O4 [label = 'w12']
    OutputLayer -> O1
    OutputLayer -> O2
    OutputLayer -> O3
    OutputLayer -> O4
    OutputLayer -> ALM
    OutputLayer -> EXAM
    ALM -> EXAM [label = 'Pure ALM Model']
    EXAM -> ALM [label = 'ALM with EXAM Response Component']
  }
")

```

```{r}

DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

data1 [label = 'Dataset 1', shape = folder, fillcolor = Beige]
data2 [label = 'Dataset 2', shape = folder, fillcolor = Beige]
process [label =  'Process \n Data']
statistical [label = 'Statistical \n Analysis']
results [label= 'Results']

# edge definitions with the node IDs
{data1 data2}  -> process -> statistical -> results
}")

```

```{mermaid}
graph LR
  subgraph InputLayer
    I1["Input Node 1"]
    I2["Input Node 2"]
    I3["Input Node 3"]
  end
  subgraph OutputLayer
    O1["Output Node 1"]
    O2["Output Node 2"]
    O3["Output Node 3"]
    O4["Output Node 4"]
  end
  I1 -->|"w1"| O1
  I1 -->|"w2"| O2
  I1 -->|"w3"| O3
  I1 -->|"w4"| O4
  I2 -->|"w5"| O1
  I2 -->|"w6"| O2
  I2 -->|"w7"| O3
  I2 -->|"w8"| O4
  I3 -->|"w9"| O1
  I3 -->|"w10"| O2
  I3 -->|"w11"| O3
  I3 -->|"w12"| O4
  style InputLayer fill:#99cc99,stroke:#333,stroke-width:2px
  style OutputLayer fill:#cc99cc,stroke:#333,stroke-width:2px
  IS["Input Stimulus (X)"] --> InputLayer
  InputLayer -->|"Gaussian Activation (a_i)"| OutputLayer
  OutputLayer --> ALM["ALM Response (m(X))"]
  OutputLayer --> EXAM["EXAM Response (E[Y | X_i])"]
  ALM -->|"Pure ALM Model"| EXAM
  EXAM -->|"ALM with EXAM Response Component"| ALM
  linkStyle 0 stroke:#2ecd71,stroke-width:2px;
  linkStyle 1 stroke:#2ecd71,stroke-width:2px;
  linkStyle 2 stroke:#2ecd71,stroke-width:2px;
  linkStyle 3 stroke:#2ecd71,stroke-width:2px;
  linkStyle 4 stroke:#2ecd71,stroke-width:2px;
  linkStyle 5 stroke:#2ecd71,stroke-width:2px;
  linkStyle 6 stroke:#2ecd71,stroke-width:2px;
  linkStyle 7 stroke:#2ecd71,stroke-width:2px;
  linkStyle 8 stroke:#2ecd71,stroke-width:2px;
  linkStyle 9 stroke:#2ecd71,stroke-width:2px;
  linkStyle 10 stroke:#2ecd71,stroke-width:2px;
  linkStyle 11 stroke:#2ecd71,stroke-width:2px;
  linkStyle 12 stroke:#2ecd71,stroke-width:2px;
  linkStyle 13 stroke:#2ecd71,stroke-width:2px;
  linkStyle 14 stroke:#2ecd71,stroke-width:2px;
  linkStyle 15 stroke:#2ecd71,stroke-width:2px;
  linkStyle 16 stroke:#2ecd71,stroke-width:2px;
  linkStyle 17 stroke:#2ecd71,stroke-width:2px;
  linkStyle 0 stroke:#2ecd71,stroke-width:2px;
  linkStyle 1 stroke:#2ecd71,stroke-width:2px;
  linkStyle 2 stroke:#2ecd71,stroke-width:2px;
  linkStyle 3 stroke:#2ecd71,stroke-width:2px;
  linkStyle 4 stroke:#2ecd71,stroke-width:2px;
  linkStyle 5 stroke:#2ecd71,stroke-width:2px;
  linkStyle 6 stroke:#2ecd71,stroke-width:2px;
  linkStyle 7 stroke:#2ecd71,stroke-width:2px;
  linkStyle 8 stroke:#2ecd71,stroke-width:2px;
  linkStyle 9 stroke:#2ecd71,stroke-width:2px;
  linkStyle 10 stroke:#2ecd71,stroke-width:2px;
  linkStyle 11 stroke:#2ecd71,stroke-width:2px;
  linkStyle 12 stroke:#2ecd71,stroke-width:2px;
  linkStyle 13 stroke:#2ecd71,stroke-width:2px;
  linkStyle 14 stroke:#2ecd71,stroke-width:2px;
  linkStyle 15 stroke:#2ecd71,stroke-width:2px;
  linkStyle 16 stroke:#2ecd71,stroke-width:2px;
  linkStyle 17 stroke:#2ecd71,stroke-width:2px;

```

## P5 sketching
```{ojs}

P5 = require("p5")
function* createSketch(sketch) {
  const element = DOM.element('div');
  yield element;
  const instance = new P5(sketch, element, true);
  try {
    while (true) {
      yield element;
    }
  } finally {
    instance.remove();
  }
}
createSketch(s => {

    s.setup = function() {
      s.createCanvas(746, 300);
      s.textFont('Courgette');
      s.textStyle(s.BOLD);
      s.textAlign(s.CENTER, s.CENTER)

      s.button = s.createButton('clear');
      s.button.mousePressed(s.clearCanvas);
        s.text('Click and drag to draw', s.width/2, s.height/10);

    };
    s.draw = function() {
    if (s.mouseIsPressed) {
    s.fill(0);
    s.ellipse(s.mouseX, s.mouseY, 10, 10);
    } else {
   //s.fill(255);
    }
  // add text input

    };

  // add button to clear canvas
  s.clearCanvas = function() {
    s.clear();
  };
  // add text
  // add slider
  }
)

```

```{r}
```