### 2023-05-28 Simulations/alm_hiddenLayer.qmd
mermaid visuals etc.

---
title: alm hidden
---

```{r}

#Here's a new version of ALM with 7 input nodes, 32 output nodes, and a hidden layer with 10 nodes:
input.activation <- function(x.target, association.parameter) {
  return(exp(-1 * association.parameter * (x.target - x.plotting)^2))
}

hidden.activation <- function(input.activations, weights, association.parameter) {
  return(exp(-1 * association.parameter * (weights %*% input.activations)))
}

output.activation <- function(hidden.activations, weights, association.parameter) {
  return(weights %*% hidden.activations)
}

mean.prediction <- function(x.target, weights, association.parameter) {
  hidden.activations <- hidden.activation(input.activation(x.target, association.parameter), weights, association.parameter)
  probability <- output.activation(hidden.activations, weights, association.parameter) / sum(output.activation(hidden.activations, weights, association.parameter))
  return(y.plotting %*% probability) # integer prediction
}

update.weights <- function(x.new, y.new, weights, association.parameter, update.parameter) {
  hidden.activations <- hidden.activation(input.activation(x.new, association.parameter), weights, association.parameter)
  y.feedback.activation <- exp(-1 * association.parameter * (y.new - y.plotting)^2)
  x.feedback.activation <- output.activation(hidden.activations, weights, association.parameter)
  delta.weights <- update.parameter * (y.feedback.activation - x.feedback.activation) %*% t(hidden.activations %*% input.activation(x.new, association.parameter))
  return(weights + delta.weights)
}

learn.alm <- function(y.learning, association.parameter = 0.05, update.parameter = 0.5) {
  weights1 <- matrix(rnorm(7 * 10, mean = 0, sd = 0.1), nrow = 7, ncol = 10)
  weights2 <- matrix(rnorm(10 * 32, mean = 0, sd = 0.1), nrow = 32, ncol = 10)
  for (i in 1:length(y.learning)) {
    hidden.activations <- hidden.activation(input.activation(x.learning[i], association.parameter), weights1, association.parameter)
    weights2 <- update.weights(x.learning[i], y.learning[i], weights2, association.parameter, update.parameter, hidden.activations)
    resp = mean.prediction(x.learning[i], weights2, association.parameter)
    weights2[weights2 < 0] = 0
  }
  alm.predictions <- sapply(x.plotting, mean.prediction, weights = weights2, association.parameter = association.parameter)
  exam.predictions <- sapply(x.plotting, exam.prediction, weights = weights2, association.parameter = association.parameter)
  return(list(alm.predictions = alm.predictions, exam.predictions = exam.predictions))
}

```

```{r}
# Define input and output variables
x.learning <- seq(0, 100, length.out = 100)
y.learning <- sin(x.learning) + rnorm(length(x.learning), mean = 0, sd = 0.1)
x.plotting <- seq(0, 100, length.out = 10)
y.plotting <- sin(x.plotting)

# Train ALM model
results <- learn.alm(y.learning)

# Plot predictions
plot(x.learning, y.learning, type = "l", col = "black")
lines(x.learning, results$alm.predictions, col = "red")
lines(x.learning, results$exam.predictions, col = "blue")
legend("topright", legend = c("ALM", "Exemplar"), col = c("red", "blue"), lty = 1)
```