### 2023-06-09 Analysis/test_analysis.qmd
general commit

---
title: HTW Test Analysis
date: last-modified
categories: [Analysis, R]
page-layout: full
code-fold: show
code-tools: true
execute: 
  warning: false
  eval: true
---

```{r}
pacman::p_load(tidyverse,data.table,lme4,emmeans,here)
d <- readRDS(here("data/dPrune-05-21-23.rds"))

dtest <- d %>% filter(expMode %in% c("test-Nf","test-train-nf")) %>% group_by(id,lowBound) %>% 
  mutate(nBand=n(),band=bandInt,id=factor(id)) %>% group_by(id) %>% mutate(nd=n_distinct(lowBound))
dtest <- dtest %>% group_by(id,lowBound) %>% filter(nBand>=5 & nd==6)
dtest <- dtest %>% group_by(id) %>% filter(!id %in% unique(dtest$id[dtest$nBand<5]))

dtestAgg <- dtest %>% group_by(id,condit,catOrder,fb,vb,band,lowBound,highBound,input) %>% mutate(vxCapped=ifelse(vx>1600,1600,vx)) %>%
  summarise(vxMean=mean(vx),devMean=mean(dist),vxMed=median(vx),devMed=median(dist),
            vxMeanCap=mean(vxCapped),.groups = "keep")

```