### 2023-06-12 Visuals_Interactives/mermaid_consort.qmd
tikz plots and slight reorganization.

---
title: Mermaid with R + consort
date: last-modified
categories: [Visualization, R, OJS]
page-layout: full
code-fold: true
code-tools: true
execute: 
  warning: false
---

```{r}
#lapply(c('tidyverse','data.table','igraph','ggraph','kableExtra'),library,character.only=TRUE))
pacman::p_load(tidyverse,data.table,igraph,ggraph,kableExtra,DiagrammeR,png,consort,data.table) 
```

```{dot}
digraph ALM {
    rankdir = LR
    splines=false // straight lines, not curves

    // define input nodes
    Input1 [label="Input1"]
    Input2 [label="Input2"]

    // define output nodes
    Output1 [label="Output1"]
    Output2 [label="Output2"]
    Output3 [label="Output3"]

    // define equations
    equation1 [shape=none, label="a_i(X) = e^{-γ(X-X_i)^2}, normalized: a_i(X) / Σa_i(X)", fontsize=10]
    equation2 [shape=none, label="O_j(X) = Σ w_{ji} ⋅ a_i(X), P[Y_j | X] = O_j(X) / Σ O_k(X)", fontsize=10]

    // define invisible nodes for alignment
    invis1 [style=invis]
    invis2 [style=invis]

    // edges between input nodes and output nodes
    Input1 -> Output1 [label="w_{11}"]
    Input1 -> Output2 [label="w_{12}"]
    Input1 -> Output3 [label="w_{13}"]

    Input2 -> Output1 [label="w_{21}"]
    Input2 -> Output2 [label="w_{22}"]
    Input2 -> Output3 [label="w_{23}"]

    // align input nodes and equations on same rank
    {rank=same Input1 Input2 equation1 invis1}
    {rank=same Output1 Output2 Output3 equation2 invis2}

    // invisible edges to order nodes
    edge [style=invis]
    Input1 -> Input2 -> equation1 -> invis1
    Output1 -> Output2 -> Output3 -> equation2 -> invis2
}

```

```{dot}
digraph ALM {
    rankdir = LR
    splines=false // straight lines, not curves

    // Define input layer with rectangle
    subgraph cluster_input {
        label="Input Layer";
        style=filled;
        color=lightgrey;
        Input1 [label="Input1"]
        Input2 [label="Input2"]
    }

    // Define output layer with rectangle
    subgraph cluster_output {
        label="Output Layer";
        style=filled;
        color=lightgrey;
        Output1 [label="Output1"]
        Output2 [label="Output2"]
        Output3 [label="Output3"]
    }

    // Define equations (using simplified text as Graphviz has limited text formatting options)
    equation1 [shape=plaintext, label="Input activation ai(X) = exp(-gamma*(X-Xi)^2), normalized by dividing by sum", fontsize=10]
    equation2 [shape=plaintext, label="Output Oj(X) = sum(w_ji * ai(X)), Probability P[Yj | X] = Oj(X) / sum(Ok(X))", fontsize=10]

    // Edges between input nodes and output nodes
    Input1 -> Output1 [label="w_11"]
    Input1 -> Output2 [label="w_12"]
    Input1 -> Output3 [label="w_13"]

    Input2 -> Output1 [label="w_21"]
    Input2 -> Output2 [label="w_22"]
    Input2 -> Output3 [label="w_23"]

    // Invisible nodes for alignment
    invis1 [style=invis]
    invis2 [style=invis]

    // Align input nodes and equations on same rank
    {rank=same equation1 invis1}
    {rank=same equation2 invis2}

    // Invisible edges for ordering
    edge [style=invis]
    equation1 -> invis1
    equation2 -> invis2
}

```

```{dot}

digraph G {
rankdir = LR
splines=false  // straight lines, not curves
// define ellipsis nodes
E1[shape=none  label="&#8942;" fontsize=30]
E2[shape=none  label="&#8942;" fontsize=30]
{rank=same a b E1 z}       // align on same rank
{rank=same x1 x2 E2 x100}  // align on same rank

{a b z} -> {x1, x2, x100} -> Y

// invisible edges to order nodes
edge[style=invis]
a->b->E1->z 
x1->x2->E2->x100 
}

```

```{dot}
digraph G {
 graph [rankdir=LR]
 node [shape=none label=""]
 node [imagescale=true]
 n1  [ image="Visuals_Interactives/pic1.svg" ]
 n2  [ image="image_dir/longeqns2.png" ]
 n3  [ image="image_dir/Eggequation.jpg" ]
 n4  [ shape=rect image="image_dir/LaTeX-arrow.png" ]
 n5  [ shape=circle image="image_dir/main-qimg.png"]
 n6  [ shape=rect image="image_dir/LaTeX-arrow.png" ]

 n1 -> { n2 n3}
 n2-> n4 -> n5 -> n6

subgraph clusterX {
 node [imagescale=true]
 J1  [ image="image_dir/LaTeX-arrow.png" ]
 J2  [ shape=rect image="image_dir/longeqns2.png" ]
 J4  [ image="image_dir/LaTeX-arrow.png" ]
 J5  [ shape=rect image="image_dir/main-qimg.png"]
 J6  [ shape=circle image="image_dir/LaTeX-arrow.png" ]
}
 J1 -> J2
 J1-> J6 -> J5 -> J4
}

```

```{dot}

digraph G {
rankdir = LR
splines=false  // straight lines, not curves
// define ellipsis nodes
E1[image="Visuals_Interactives/gaussian_curve_input_node_1.png"]
E2[shape=none  label="&#8942;" fontsize=30]
{rank=same a b E1 z}       // align on same rank
{rank=same x1 x2 E2 x100}  // align on same rank

{a b z} -> {x1, x2, x100} -> Y

// invisible edges to order nodes
edge[style=invis]
a->b->E1->z 
x1->x2->E2->x100 
}

```

```{dot}

digraph G {
 graph [rankdir=LR]
 node [shape=none label=""]
 n1  [ image="images/svg1.svg" ]
 n2  [ image="images/acclaim.svg" ]
 n4  [ image="images/acm.svg" ]

 n1 -> n2 -> n4 
}

```

```{dot}
digraph ALM {
    rankdir = LR;
    splines=false;
    edge[style=invis];
    ranksep= 1.4;

    {
        node [shape=circle, color=chartreuse, style=filled, fillcolor=chartreuse];
        Input1 [label=<X<sub>1</sub>>];
        Input2 [label=<X<sub>2</sub>>];
    }

    {
        node [shape=circle, color=coral1, style=filled, fillcolor=coral1];
        Output1 [label=<Y<sub>1</sub>>];
        Output2 [label=<Y<sub>2</sub>>]; 
        Output3 [label=<Y<sub>3</sub>>];
    }

    {
        rank=same;
        Input1 -> Input2 [style=invis];
    }

    {
        rank=same;
        Output1 -> Output2 -> Output3 [style=invis];
    }

    label_input [shape=plaintext, label="Input Layer"];
    label_input -> Input1;
    {rank=same; label_input; Input1};

    label_output [shape=plaintext, label="Output Layer"];
    label_output -> Output1;
    {rank=same; label_output; Output1};

    edge[style=solid, tailport=e, headport=w, labelangle=45, labeldistance=2.0];
    Input1 -> Output1 [label=<w<sub>11</sub>>];
    Input1 -> Output2 [label=<w<sub>12</sub>>];
    Input1 -> Output3 [label=<w<sub>13</sub>>];

    Input2 -> Output1 [label=<w<sub>21</sub>>];
    Input2 -> Output2 [label=<w<sub>22</sub>>];
    Input2 -> Output3 [label=<w<sub>23</sub>>];
}

```

```{dot}
digraph G {

    // ===============================
    // rankdir: directed graph drawn from left to right 
    // details: http://www.graphviz.org/doc/info/attrs.html
    // ===============================
    rankdir=LR;  
    edge[style=solid, tailport=e, headport=w];

    // ===============================
    // splines=line:  draw straight lines to connect nodes
    // ===============================
    splines=line;
    nodesep=".05";
    node [label=""];
    subgraph cluster_0 {
        color=white;
        label="layer 1 (input layer)";
        node [color=chartreuse, 
              style=filled, 
              shape=circle];

       // ==============================
       // Below label text format specifies subscript and superscript in graphviz
       // ==============================
        x0 [color=yellow, 
            fillcolor=yellow, 
            label=<x<sub>0</sub>>];
        x1 [fillcolor=chartreuse, 
            label=<x<sub>1</sub>>];
        x2 [fillcolor=chartreuse, 
            label=<x<sub>2</sub>>];
        x3 [fillcolor=chartreuse, 
            label=<x<sub>3</sub>>];
    }

    subgraph cluster_1 {
        color=white;
        label="layer 2 (hidden layer)";
        node [color=dodgerblue, 
              style=filled, 
              shape=circle];
        a02 [color=yellow, 
             fillcolor=yellow, 
             label=<a<sub>0</sub><sup>(2)</sup>>];

       // ==============================
       // Below label text doesn't match with the node naming (i.e. a12 should have label "a12" instead of "a52")
       // This is a dirty trick that to keep the nodes in the same layer with label sorted 
       // (i.e. a12, followed by a22, followed by a32, followed by a42, followed by a52)
       // ==============================
        a12 [fillcolor=dodgerblue, 
             label=<a<sub>5</sub><sup>(2)</sup>>];
        a22 [fillcolor=dodgerblue, 
             label=<a<sub>1</sub><sup>(2)</sup>>];
        a32 [fillcolor=dodgerblue, 
             label=<a<sub>2</sub><sup>(2)</sup>>];
        a42 [fillcolor=dodgerblue, 
             label=<a<sub>3</sub><sup>(2)</sup>>];
        a52 [fillcolor=dodgerblue, 
             label=<a<sub>4</sub><sup>(2)</sup>>];
    }

    subgraph cluster_2 {
        color=white;
        label="layer 3 (hidden layer)";
        node [color=dodgerblue, 
              style=filled, 
              shape=circle];
        a03 [color=yellow, 
             fillcolor=yellow, 
             label=<a<sub>0</sub><sup>(3)</sup>>];
        a13 [fillcolor=dodgerblue, 
             label=<a<sub>3</sub><sup>(3)</sup>>];
        a23 [fillcolor=dodgerblue, 
             label=<a<sub>4</sub><sup>(3)</sup>>];
        a33 [fillcolor=dodgerblue, 
             label=<a<sub>5</sub><sup>(3)</sup>>];
        a43 [fillcolor=dodgerblue, 
             label=<a<sub>1</sub><sup>(3)</sup>>];
        a53 [fillcolor=dodgerblue, 
             label=<a<sub>2</sub><sup>(3)</sup>>];
    }

    subgraph cluster_3 {
        color=white;
        label="layer 4 (output layer)";
        node [color=coral1, 
              style=filled, 
              shape=circle];
        O1 [fillcolor=coral1, 
            label=<a<sub>1</sub><sup>(4)</sup>>];
        O2 [fillcolor=coral1, 
            label=<a<sub>2</sub><sup>(4)</sup>>];
        O3 [fillcolor=coral1, 
            label=<a<sub>3</sub><sup>(4)</sup>>];
        O4 [fillcolor=coral1, 
            label=<a<sub>4</sub><sup>(4)</sup>>];
    }

   // ===============================
   // This is the trick to enforce the bias node stays at the top of 
   // vertical array of nodes in each layer
   // style=invisible: makes the edge connection invisible
   // dir=none: hide the arrow 
   // ===============================
    x0 -> a02 [style=invisible, dir=none];
    x0 -> a12;
    x0 -> a22;
    x0 -> a32;
    x0 -> a42;
    x0 -> a52;
    x1 -> a12;
    x1 -> a22;
    x1 -> a32;
    x1 -> a42;
    x1 -> a52;
    x2 -> a12;
    x2 -> a22;
    x2 -> a32;
    x2 -> a42;
    x2 -> a52;
    x3 -> a12;
    x3 -> a22;
    x3 -> a32;
    x3 -> a42;
    x3 -> a52;
    a02 -> a03 [style=invisible, dir=none];
    a02 -> a13;
    a02 -> a23;
    a02 -> a33;
    a02 -> a43;
    a02 -> a53;
    a12 -> a13;
    a12 -> a23;
    a12 -> a33;
    a12 -> a43;
    a12 -> a53;
    a22 -> a13;
    a22 -> a23;
    a22 -> a33;
    a22 -> a43;
    a22 -> a53;
    a32 -> a13;
    a32 -> a23;
    a32 -> a33;
    a32 -> a43;
    a32 -> a53;
    a42 -> a13;
    a42 -> a23;
    a42 -> a33;
    a42 -> a43;
    a42 -> a53;
    a52 -> a13;
    a52 -> a23;
    a52 -> a33;
    a52 -> a43;
    a52 -> a53;
    a03 -> O1;
    a13 -> O1;
    a23 -> O1;
    a33 -> O1;
    a43 -> O1;
    a53 -> O1;
    a03 -> O2;
    a13 -> O2;
    a23 -> O2;
    a33 -> O2;
    a43 -> O2;
    a53 -> O2;
    a03 -> O3;
    a13 -> O3;
    a23 -> O3;
    a33 -> O3;
    a43 -> O3;
    a53 -> O3;
    a03 -> O4;
    a13 -> O4;
    a23 -> O4;
    a33 -> O4;
    a43 -> O4;
    a53 -> O4;
}

```

```{dot}
digraph G {

        rankdir=LR
	splines=line

        node [fixedsize=true, label=""];

        subgraph cluster_0 {
		color=white;
		node [style=solid,color=blue4, shape=circle];
		x1 x2 x3;
		label = "layer 1 (Input layer)";
	}

	subgraph cluster_1 {
		color=white;
		node [style=solid,color=red2, shape=circle];
		a12 a22 a32;
		label = "layer 2 (hidden layer)";
	}

	subgraph cluster_2 {
		color=white;
		node [style=solid,color=seagreen2, shape=circle];
		O;
		label="layer 3 (output layer)";
	}

        x1 -> a12;
        x1 -> a22;
        x1 -> a32;
        x2 -> a12;
        x2 -> a22;
        x2 -> a32;
        x3 -> a12;
        x3 -> a22;
        x3 -> a32;

        a12 -> O
        a22 -> O
        a32 -> O

}
```

```{dot}
digraph G {
    rankdir = LR;
    splines=false;
    edge[style=invis];
    ranksep= 1.4;
    {
    node [shape=circle, color=yellow, style=filled, fillcolor=yellow];
    x0 [label=<x<sub>0</sub>>]; 
    a02 [label=<a<sub>0</sub><sup>(2)</sup>>]; 
    a03 [label=<a<sub>0</sub><sup>(3)</sup>>];
    }
    {
    node [shape=circle, color=chartreuse, style=filled, fillcolor=chartreuse];
    x1 [label=<x<sub>1</sub>>];
    x2 [label=<x<sub>2</sub>>]; 
    x3 [label=<x<sub>3</sub>>];
}
{
    node [shape=circle, color=dodgerblue, style=filled, fillcolor=dodgerblue];
    a12 [label=<a<sub>1</sub><sup>(2)</sup>>];
    a22 [label=<a<sub>2</sub><sup>(2)</sup>>];
    a32 [label=<a<sub>3</sub><sup>(2)</sup>>];
    a42 [label=<a<sub>4</sub><sup>(2)</sup>>];
    a52 [label=<a<sub>5</sub><sup>(2)</sup>>];
    a13 [label=<a<sub>1</sub><sup>(3)</sup>>];
    a23 [label=<a<sub>2</sub><sup>(3)</sup>>];
    a33 [label=<a<sub>3</sub><sup>(3)</sup>>];
    a43 [label=<a<sub>4</sub><sup>(3)</sup>>];
    a53 [label=<a<sub>5</sub><sup>(3)</sup>>];
}
{
    node [shape=circle, color=coral1, style=filled, fillcolor=coral1];
    O1 [label=<a<sub>1</sub><sup>(4)</sup>>];
    O2 [label=<a<sub>2</sub><sup>(4)</sup>>]; 
    O3 [label=<a<sub>3</sub><sup>(4)</sup>>]; 
    O4 [label=<a<sub>4</sub><sup>(4)</sup>>];
}
    {
        rank=same;
        x0->x1->x2->x3;
    }
    {
        rank=same;
        a02->a12->a22->a32->a42->a52;
    }
    {
        rank=same;
        a03->a13->a23->a33->a43->a53;
    }
    {
        rank=same;
        O1->O2->O3->O4;
    }
    a02->a03;  // prevent tilting
    l0 [shape=plaintext, label="layer 1 (input layer)"];
    l0->x0;
    {rank=same; l0;x0};
    l1 [shape=plaintext, label="layer 2 (hidden layer)"];
    l1->a02;
    {rank=same; l1;a02};
    l2 [shape=plaintext, label="layer 3 (hidden layer)"];
    l2->a03;
    {rank=same; l2;a03};
    l3 [shape=plaintext, label="layer 4 (output layer)"];
    l3->O1;
    {rank=same; l3;O1};
    edge[style=solid, tailport=e, headport=w];
    {x0; x1; x2; x3} -> {a12;a22;a32;a42;a52};
    {a02;a12;a22;a32;a42;a52} -> {a13;a23;a33;a43;a53};
    {a03;a13;a23;a33;a43;a53} -> {O1,O2,O3,O4};
}
```

```{dot}
 digraph threevar {
      rankdir=LR;
      size="8,4";
      node [fontsize=14 shape=box];
      edge [fontsize=10];
      center=1;
      {rank=min k }
      {rank=same X1 X2 X3 }
      {rank=max z1 z2 z3 }
      z1 [shape=circle label="d1"];
      z2 [shape=circle label="d2"];
      z3 [shape=circle label="d3"];
      k [label="?" shape="ellipse"];
      k -> X1 [label="n@_{S}"];
      k -> X2 [label="&theta;@^{(h)}"];
      k -> X3 [label="?3"];
      z1 -> X1;
      z2 -> X2;
      z3 -> X3;
    }

```

```{r}
grViz('digraph threevar {
      rankdir=LR;
      size="8,4";
      node [fontsize=14 shape=box];
      edge [fontsize=10];
      center=1;
      {rank=min k }
      {rank=same X1 X2 X3 }
      {rank=max z1 z2 z3 }
      z1 [shape=circle label="d1"];
      z2 [shape=circle label="d2"];
      z3 [shape=circle label="d3"];
      k [label="?" shape="ellipse"];
      k -> X1 [label="n@_{S}"];
      k -> X2 [label="&theta;@^{(h)}"];
      k -> X3 [label="?3"];
      z1 -> X1;
      z2 -> X2;
      z3 -> X3;
    }
      ')

```

```{r}
grViz("digraph dot {
      graph [style = filled, fillcolor = white]

      node [shape = circle,
            style = filled, fillcolor = white,
            fixedsize = true, width = 0.5, height = 0.5,
            fontname = 'Times-italic']
        c
        d
      node [shape = doublecircle,
            style = filled, fillcolor = white,
            fixedsize = true, width = 0.5, height = 0.5,
            fontname = 'Times-italic']
        thetah [label = '&theta;@^{(h)}']
        thetaf [label = '&theta;@^{(f)}']
      node [shape = square,
            style = filled, fillcolor = grey,
            fixedsize = true, width=0.5, height=0.5,
            fontname = 'Times-italic']
        h
        f
        ns [label = 'n@_{S}']
        nn [label = 'n@_{N}']

      edge [color = black]
        c -> thetah -> h
        d -> thetaf -> f
        c -> thetaf
        d -> thetah
        ns -> h
        nn -> f
      {rank = max; ns; nn}
      }")

grViz("digraph dot {
      graph [style = filled, fillcolor = white,
             rankdir = LR,
             newrank = true]

      node [shape = circle,
            style = filled, fillcolor = white,
            fixedsize = true, width = 0.5, height = 0.5,
            fontname = 'Times-italic']
        p [fillcolor = gray]
        gamma [label = '&gamma;']
        omega [label = '&omega;', shape = doublecircle]
        beta [label = '&beta;']
      subgraph cluster_out{
        fontsize = 8
        label = <<I>j</I>:試行>
        labelloc = b
        labeljust = r
        subgraph cluster_in{
          fontsize = 8
          label = <<I>k</I>:回数>
          labelloc = b
          labeljust = r
          thetajk [label = '&theta;@_{jk}', shape = doublecircle]
          djk [label = 'd@_{jk}', shape = square, fillcolor = gray]
        }}
      edge [color=black]
        p -> omega -> thetajk -> djk
        gamma -> omega
        beta -> thetajk 
      {rank = same; gamma; omega}
      {rank = same; beta; thetajk}
      }")

grViz('
digraph G {
	fontname="Helvetica,Arial,sans-serif"
	node [fontname="Helvetica,Arial,sans-serif"]
	edge [fontname="Helvetica,Arial,sans-serif"]

	subgraph cluster_0 {
		style=filled;
		color=lightgrey;
		node [style=filled,color=white];
		a0 -> a1 -> a2 -> a3;
		label = "process #1";
	}

	subgraph cluster_1 {
		node [style=filled];
		b0 -> b1 -> b2 -> b3;
		label = "process #2";
		color=blue
	}
	start -> a0;
	start -> b0;
	a1 -> b3;
	b2 -> a3;
	a3 -> a0;
	a3 -> end;
	b3 -> end;

	start [shape=Mdiamond];
	end [shape=Msquare];
}')

#https://graphviz.org/Gallery/directed/neural-network.html

grViz('digraph G {
  fontname="Helvetica,Arial,sans-serif"
  node [fontname="Helvetica,Arial,sans-serif"]
  edge [fontname="Helvetica,Arial,sans-serif"]
  concentrate=True;
  rankdir=TB;
  node [shape=record];
  140087530674552 [label="title: InputLayer\n|{input:|output:}|{{[(?, ?)]}|{[(?, ?)]}}"];
  140087537895856 [label="body: InputLayer\n|{input:|output:}|{{[(?, ?)]}|{[(?, ?)]}}"];
  140087531105640 [label="embedding_2: Embedding\n|{input:|output:}|{{(?, ?)}|{(?, ?, 64)}}"];
  140087530711024 [label="embedding_3: Embedding\n|{input:|output:}|{{(?, ?)}|{(?, ?, 64)}}"];
  140087537980360 [label="lstm_2: LSTM\n|{input:|output:}|{{(?, ?, 64)}|{(?, 128)}}"];
  140087531256464 [label="lstm_3: LSTM\n|{input:|output:}|{{(?, ?, 64)}|{(?, 32)}}"];
  140087531106200 [label="tags: InputLayer\n|{input:|output:}|{{[(?, 12)]}|{[(?, 12)]}}"];
  140087530348048 [label="concatenate_1: Concatenate\n|{input:|output:}|{{[(?, 128), (?, 32), (?, 12)]}|{(?, 172)}}"];
  140087530347992 [label="priority: Dense\n|{input:|output:}|{{(?, 172)}|{(?, 1)}}"];
  140087530711304 [label="department: Dense\n|{input:|output:}|{{(?, 172)}|{(?, 4)}}"];
  140087530674552 -> 140087531105640;
  140087537895856 -> 140087530711024;
  140087531105640 -> 140087537980360;
  140087530711024 -> 140087531256464;
  140087537980360 -> 140087530348048;
  140087531256464 -> 140087530348048;
  140087531106200 -> 140087530348048;
  140087530348048 -> 140087530347992;
  140087530348048 -> 140087530711304;
}')

```

```{r}
require(data.table)
require(qreport)
require(consort)

# Load the necessary libraries
require(data.table)
require(qreport)

# Define the mermaid diagram
x <- '
graph LR
  InputLayer[Input Layer] --> I1[{{I1}}]
  InputLayer --> I2[{{I2}}]
  InputLayer --> I3[{{I3}}]
  I1 --> O1[w1]
  I1 --> O2[w2]
  I1 --> O3[w3]
  I1 --> O4[w4]
  I2 --> O1[w5]
  I2 --> O2[w6]
  I2 --> O3[w7]
  I2 --> O4[w8]
  I3 --> O1[w9]
  I3 --> O2[w10]
  I3 --> O3[w11]
  I3 --> O4[w12]
  OutputLayer[Output Layer] --> O1
  OutputLayer --> O2
  OutputLayer --> O3
  OutputLayer --> O4
  OutputLayer --> ALM[ALM Response]
  OutputLayer --> EXAM[EXAM Response]
  ALM --> EXAM[Pure ALM Model]
  EXAM --> ALM[ALM with EXAM Response Component]
'

# Generate the mermaid diagram
makemermaid(x,
            I1 = '![Gaussian Curve](gaussian_curve_input_node_1.png)',
            I2 = '![Gaussian Curve](gaussian_curve_input_node_2.png)',
            I3 = '![Gaussian Curve](gaussian_curve_input_node_3.png)',
            file = 'alm_exam.mer'
            )

# Output the mermaid diagram
# cat('```{mermaid}\n')
# cat(readLines('alm_exam.mer'), sep = '\n')
# cat('```\n')

```

```{mermaid}
%%| fig-cap: "ALM diagram produced with `mermaid` with individual exclusions linked to the overall exclusions node, and with a tooltip to show more detail"
%%| label: fig-doverview-mermaid2
%%| file: alm_exam.mer
```

```{r}
library(consort)
set.seed(1001)
N <- 300

trialno <- sample(c(1000:2000), N)
exc <- rep(NA, N)
exc[sample(1:N, 15)] <- sample(c("Sample not collected", "MRI not collected", "Other"),
                                15, replace = T, prob = c(0.4, 0.4, 0.2))

arm <- rep(NA, N)
arm[is.na(exc)] <- sample(c("Conc", "Seq"), sum(is.na(exc)), replace = T)

fow1 <- rep(NA, N)
fow1[!is.na(arm)] <- sample(c("Withdraw", "Discontinued", "Death", "Other", NA),
                            sum(!is.na(arm)), replace = T, 
                            prob = c(0.05, 0.05, 0.05, 0.05, 0.8))
fow2 <- rep(NA, N)
fow2[!is.na(arm) & is.na(fow1)] <- sample(c("Protocol deviation", "Outcome missing", NA),
                                          sum(!is.na(arm) & is.na(fow1)), replace = T, 
                                          prob = c(0.05, 0.05, 0.9))
df <- data.frame(trialno, exc, arm, fow1, fow2)
out <- consort_plot(data = df,
             order = c(trialno = "Population",
                          exc    = "Excluded",
                          arm     = "Randomized patient",
                          fow1    = "Lost of Follow-up",
                          trialno = "Finished Followup",
                          fow2    = "Not evaluable",
                          trialno = "Final Analysis"),
             side_box = c("exc", "fow1", "fow2"),
             allocation = "arm",
             labels = c("1" = "Screening", "2" = "Randomization",
                        "5" = "Final"),
             cex = 0.6)

plot(out)

```

```{r}
plot(out, grViz = TRUE)
```

```{r}
require(Hmisc)
require(data.table)
require(qreport)
hookaddcap()
N <- 1000
set.seed(1)
r <- data.table(
  id    = 1 : N,
  age   = round(rnorm(N, 60, 15)),
  pain  = sample(0 : 5, N, replace=TRUE),
  hxmed = sample(0 : 1, N, replace=TRUE, prob=c(0.95, 0.05))   )
# Set consent status to those not excluded at screening
r[age >= 40 & pain > 0 & hxmed == 0,
  consent := sample(0 : 1, .N, replace=TRUE, prob=c(0.1, 0.9))]
# Set randomization status for those consenting
r[consent == 1,
  randomized := sample(0 : 1, .N, replace=TRUE, prob=c(0.15, 0.85))]
# Add treatment and follow-up time to randomized subjects
r[randomized == 1, tx     := sample(c('A', 'B'), .N, replace=TRUE)]
r[randomized == 1, futime := pmin(runif(.N, 0, 10), 3)]
# Add outcome status for those followed 3 years
# Make a few of those followed 3 years missing
r[futime == 3,
  y := sample(c(0, 1, NA), .N, replace=TRUE, prob=c(0.75, 0.2, 0.05))]
# Print first 15 subjects
kabl(r[1 : 15, ])

```

```{r}
r[, exc := seqFreq('pain-free'  = pain  == 0,
                    'Hx med'    = hxmed == 1,
                                  age < 40,
                    noneNA=TRUE)]
eo  <- attr(r[, exc], 'obs.per.numcond')
mult <- paste0('1, 2, ≥3 exclusions: n=',
                eo[2], ', ',
                eo[3], ', ',
                eo[-(1:3)]  )

r[, .q(qual, consent, fin) :=
    .(is.na(exc),
      ifelse(consent == 1, 1, NA),
      ifelse(futime  >= 3, 1, NA))]

require(consort)
# consort_plot used to take a coords=c(0.4, 0.6) argument that prevented
# the collision you see here
consort_plot(r,
             orders = c(id      = 'Screened',
                        exc     = 'Excluded',
                        qual    = 'Qualified for Randomization',
                        consent = 'Consented',
                        tx      = 'Randomized',
                        fin     = 'Finished',
                        y       = 'Outcome\nassessed'),
             side_box = 'exc',
             allocation = 'tx',
             labels=c('1'='Screening', '3'='Consent', '4'='Randomization', '6'='Follow-up'))

```

```{r}

h <- function(n, label) paste0(label, ' (n=', n, ')')
htab <- function(x, label=NULL, split=! length(label), br='\n') {
  tab <- table(x)
  w <- if(length(label)) paste0(h(sum(tab), label), ':', br)
  f <- if(split) h(tab, names(tab)) 
  else
    paste(paste0('   ', h(tab, names(tab))), collapse=br)
  if(split) return(f)
  paste(w, f, sep=if(length(label))'' else br)
}  
count <- function(x, by=rep(1, length(x)))
  tapply(x, by, sum, na.rm=TRUE)

w <- r[, {
 g <-
   add_box(txt=h(nrow(r),       'Screened'))                    |>
   add_side_box(htab(exc,       'Excluded'))                    |>
   add_box(h(count(is.na(exc)), 'Qualified for Randomization')) |>
   add_box(h(count(consent),    'Consented'))                   |>
   add_box(h(count(randomized), 'Randomized'))                  |>
   add_split(htab(tx))                                          |>
   add_box(h(count(fin, tx),    'Finished'))                    |>
   add_box(h(count(! is.na(y), tx), 'Outcome\nassessed'))       |>
   add_label_box(c('1'='Screening',     '3'='Consent',
                   '4'='Randomization', '6'='Follow-up'))
 plot(g)
}
]

```

```{r}

addCap('fig-doverview-mermaid1', 'Consort diagram produced by `mermaid`')
x <- 'flowchart TD
  S["Screened (n={{N0}})"] --> E["{{excl}}"]
  S   --> Q["Qualified for Randomization (n={{Nq}})"]
  Q   --> C["Consented (n={{Nc}})"]
  C   --> R["Randomized (n={{Nr}})"]
  R   --> TxA["A (n={{Ntxa}})"]
  R   --> TxB["B (n={{Ntxb}})"]
  TxA --> FA["Finished (n={{Ntxaf}})"]
  TxB --> FB["Finished (n={{Ntxbf}})"]
  FA  --> OA["Outcome assessed (n={{Ntxao}})"]
  FB  --> OB["Outcome assessed (n={{Ntxbo}})"]
classDef largert fill:lightgray,width:1.5in,height:10em,text-align:right,font-size:0.8em;
class E largert;
'

w <- r[, 
makemermaid(x,
            N0   = nrow(r),
            excl = htab(exc, 'Excluded', br='<br>'),
            Nq   = count(is.na(exc)),
            Nc   = count(consent),
            Nr   = count(randomized),
            Ntxa = count(tx == 'A'),
            Ntxb = count(tx == 'B'),
            Ntxaf= count(tx == 'A' & fin),
            Ntxbf= count(tx == 'B' & fin),
            Ntxao= count(tx == 'A' & ! is.na(y)),
            Ntxbo= count(tx == 'B' & ! is.na(y)),
            file = 'mermaid1.mer'
            )
]

```

```{mermaid}
%%| fig-cap: "Consort diagram produced by `mermaid`"
%%| label: fig-doverview-mermaid1
%%| file: mermaid1.mer
```

```{r}
# Create some service functions so later it will be easy to change from
# mermaid to graphviz
makenode       <- function(name, label) paste0(name, '["', label, '"]')
makeconnection <- function(from, to)    paste0(from, ' --> ', to)

exclnodes <- function(x, from='E', root='E', seq=FALSE, remain=FALSE) {
  # Create complete node specifications for individual exclusions, each
  # linking to overall exclusion count assumed to be in node root.
  # Set seq=TRUE to make use of the fact that the exclusions were
  # done in frequency priority order so that each exclusion is in
  # addition to the previous one.  Leave seq=FALSE to make all exclusions
  # subservient to root.  Use remain=TRUE to include # obs remaining
    # remain=TRUE assumes noneNA specified to seqFreq
  tab <- table(x)
  i <- 1 : length(tab)
    rem <- if(remain) paste0(', ', length(x) - cumsum(tab), ' remain')
  labels <- paste0(names(tab), ' (n=', tab, rem, ')')
  nodes  <- if(seq) makenode(ifelse(i == 1, paste0(root, '1'), paste0(root, i)),
                             labels)
            else    makenode(paste0(root, i), labels)
  connects <- if(seq) makeconnection(ifelse(i == 1, from, paste0(root, i - 1)),
                                     paste0(root, i))
              else makeconnection(from, paste0(root, i))
   paste(c(nodes, connects), collapse='\n')
}

# Create parallel treatment nodes
# Treatments are assumed to be in order by the tx variable
# and will appear left to right in the diagram
# Treatment node names correspond to that and are Tx1, Tx2, ...
# root: root of new nodes, from: single node name to connect from
# fromparallel: root of connected-from node name which is to be
# expanded by adding the integers 1, 2, ... number of treatments.

Txs <- r[, if(is.factor(tx)) levels(tx) else sort(unique(tx))]

parNodes <- function(counts, root, from=NULL, fromparallel=NULL,
                      label=Txs) {
  if(! identical(names(counts), Txs)) stop('Txs not consistent')
  k <- length(Txs)
  ns <- paste0(' (n=', counts, ')')
   nodenames <- paste0(root, 1 : k)
  nodes <- makenode(nodenames, paste0(label, ns))
  connects <- if(length(fromparallel)) makeconnection(paste0(fromparallel, 1 : k), nodenames)
              else                     makeconnection(from,                        nodenames)
  paste(c(nodes, connects), collapse='\n')
    }

# Create tooltip text from tabulation created by seqFreq earlier
efreq <- data.frame('# Exclusions'= (1 : length(eo)) - 1,
                    '# Subjects'  = eo, check.names=FALSE)
efreq <- subset(efreq, `# Subjects` > 0)
# Convert to text which will be wrapped by the html
excltab <- paste(capture.output(print(efreq, row.names=FALSE)),
                 collapse='\n')

addCap('fig-doverview-mermaid2', 'Consort diagram produced with `mermaid` with individual exclusions linked to the overall exclusions node, and with a tooltip to show more detail')

x <- '
flowchart TD
  S["Screened (n={{N0}})"] --> E["Excluded (n={{Ne}})"]
  {{exclsep}}
  E1 & E2 & E3 --> M["{{mult}}"]
  S   --> Q["Qualified for Randomization (n={{Nq}})"]
  Q   --> C["Consented (n={{Nc}})"]
  C   --> R["Randomized (n={{Nr}})"]
  {{txcounts}}
  {{finished}}
  {{outcome}}
click E callback "{{excltab}}"
'

w <- r[, 
makemermaid(x,
  N0       = nrow(r),
  Ne       = count(! is.na(exc)),
  exclsep  = exclnodes(exc),  # add seq=TRUE to put exclusions vertical
  excltab  = excltab,         # tooltip text
  mult     = mult,  # separate node: count multiple exclusions
  Nq       = count(is.na(exc)),
  Nc       = count(consent),
  Nr       = count(randomized),
  txcounts = parNodes(table(tx),         'Tx', from='R'),
  finished = parNodes(count(fin, by=tx), 'F',  fromparallel='Tx',
                      label='Finished'),
  outcome  = parNodes(count(! is.na(y), by=tx), 'O',
                      fromparallel='F', label='Outcome assessed'),
  file='mermaid2.mer'  # save generated code for another use
)
]

makenode       <- function(name, label) paste0(name, ' [label="', label, '"];')
makeconnection <- function(from, to)    paste0(from, ' -> ', to, ';')

# Create data frame from tabulation created by seqFreq earlier
efreq <- data.frame('# Exclusions'= (1 : length(eo)) - 1,
                    '# Subjects'  = eo, check.names=FALSE)
efreq <- subset(efreq, `# Subjects` > 0)

x <- 'digraph {
  graph [pad="0.5", nodesep="0.5", ranksep="2", splines=ortho]
  //  splines=ortho for square connections
  node  [shape=box, fontsize="30"]
  rankdir=TD;
  S [label="Screened (n={{N0}})"];
  E [label="Excluded (n={{Ne}})"];
  S -> E;
  {{exclsep}}
  M [label="{{mult}}"];
  E1 -> M;
  E2 -> M;
  E3 -> M;
  Q [label="Qualified for Randomization (n={{Nq}})"];
  C [label="Consented (n={{Nc}})"];
  R [label="Randomized (n={{Nr}})"];
  S -> Q;
  Q -> C;
  C -> R;
  {{txcounts}}
  {{finished}}
  {{outcome}}
  efreq [label=<{{efreq}}>];
  M -> efreq [dir=none, style=dotted];
}
'

w <- r[, 
makegraphviz(x,
  N0       = nrow(r),
  Ne       = count(! is.na(exc)),
  exclsep  = exclnodes(exc),  # add seq=TRUE to put exclusions vertical
  efreq    = efreq,
  mult     = mult,  # separate node: count multiple exclusions
  Nq       = count(is.na(exc)),
  Nc       = count(consent),
  Nr       = count(randomized),
  txcounts = parNodes(table(tx),         'Tx', from='R'),
  finished = parNodes(count(fin, by=tx), 'F',  fromparallel='Tx',
                      label='Finished'),
  outcome  = parNodes(count(! is.na(y), by=tx), 'O',
                      fromparallel='F', label='Outcome assessed'),
  file='graphviz.dot'
)
]
#  addCap('fig-doverview-graphviza', 'Consort diagram produced with `graphviz` with detailed exclusion frequencies in a separate node', scap='Consort diagram produced with `graphviz`')

```

```{mermaid}
%%| fig-cap: "Consort diagram produced with `mermaid` with individual exclusions linked to the overall exclusions node, and with a tooltip to show more detail"
%%| label: fig-doverview-mermaid2
%%| file: mermaid2.mer
```

```{r}

getHdata(support)
setDT(support)
# addCap('fig-doverview-missflow', 'Flowchart of sequential exclusion of observations due to missing values')
vars <-  .q(age, sex, dzgroup, edu, income, meanbp, wblc,
            alb, bili, crea, glucose, bun, urine)
ex <- missChk(support, use=vars, type='seq') # seq: don't make report

# Create tooltip text from tabulation created by seqFreq
oc   <- attr(ex, 'obs.per.numcond')
freq <- data.frame('# Exclusions'= (1 : length(oc)) - 1,
                   '# Subjects'  = oc, check.names=FALSE)
freq <- subset(freq, `# Subjects` > 0)

x <- '
digraph {
  graph [pad="0.5", nodesep="0.5", ranksep="2", splines=ortho]
  //  splines=ortho for square connections
  node  [shape=box, fontsize="30"]
  rankdir=TD;
  Enr [label="Enrolled (n={{N0}})"];
  Enr;
  {{exclsep}}
    Extab [label=<{{excltab}}>];
  Enr:e -> Extab [dir=none];
}
'
makegraphviz(x,
  N0        = nrow(support),
  exclsep   = exclnodes(ex, from='Enr', seq=TRUE, remain=TRUE),
  excltab   = freq,
  file      = 'support.dot'
)

```

```{r}
grViz('
graph G {
fontname="Helvetica,Arial,sans-serif"
node [fontname="Helvetica,Arial,sans-serif"]
edge [fontname="Helvetica,Arial,sans-serif"]
I5 [shape=ellipse,color=red,style=bold,label="Caroline Bouvier Kennedy\nb. 27.11.1957 New York",image="images/165px-Caroline_Kennedy.jpg",labelloc=b];
I1 [shape=box,color=blue,style=bold,label="John Fitzgerald Kennedy\nb. 29.5.1917 Brookline\nd. 22.11.1963 Dallas",image="images/kennedyface.jpg",labelloc=b];
I6 [shape=box,color=blue,style=bold,label="John Fitzgerald Kennedy\nb. 25.11.1960 Washington\nd. 16.7.1999 over the Atlantic Ocean, near Aquinnah, MA, USA",image="images/180px-JFKJr2.jpg",labelloc=b];
I7 [shape=box,color=blue,style=bold,label="Patrick Bouvier Kennedy\nb. 7.8.1963\nd. 9.8.1963"];
I2 [shape=ellipse,color=red,style=bold,label="Jaqueline Lee Bouvier\nb. 28.7.1929 Southampton\nd. 19.5.1994 New York City",image="images/jacqueline-kennedy-onassis.jpg",labelloc=b];
I8 [shape=box,color=blue,style=bold,label="Joseph Patrick Kennedy\nb. 6.9.1888 East Boston\nd. 16.11.1969 Hyannis Port",image="images/1025901671.jpg",labelloc=b];
I10 [shape=box,color=blue,style=bold,label="Joseph Patrick Kennedy Jr\nb. 1915\nd. 1944"];
I11 [shape=ellipse,color=red,style=bold,label="Rosemary Kennedy\nb. 13.9.1918\nd. 7.1.2005",image="images/rosemary.jpg",labelloc=b];
I12 [shape=ellipse,color=red,style=bold,label="Kathleen Kennedy\nb. 1920\nd. 1948"];
I13 [shape=ellipse,color=red,style=bold,label="Eunice Mary Kennedy\nb. 10.7.1921 Brookline"];
I9 [shape=ellipse,color=red,style=bold,label="Rose Elizabeth Fitzgerald\nb. 22.7.1890 Boston\nd. 22.1.1995 Hyannis Port",image="images/Rose_kennedy.JPG",labelloc=b];
I15 [shape=box,color=blue,style=bold,label="Aristotle Onassis"];
I3 [shape=box,color=blue,style=bold,label="John Vernou Bouvier III\nb. 1891\nd. 1957",image="images/BE037819.jpg",labelloc=b];
I4 [shape=ellipse,color=red,style=bold,label="Janet Norton Lee\nb. 2.10.1877\nd. 3.1.1968",image="images/n48862003257_1275276_1366.jpg",labelloc=b];
 I1 -- I5  [style=bold,color=blue]; 
 I1 -- I6  [style=bold,color=orange]; 
 I2 -- I6  [style=bold,color=orange]; 
 I1 -- I7  [style=bold,color=orange]; 
 I2 -- I7  [style=bold,color=orange]; 
 I1 -- I2  [style=bold,color=violet]; 
 I8 -- I1  [style=bold,color=blue]; 
 I8 -- I10  [style=bold,color=orange]; 
 I9 -- I10  [style=bold,color=orange]; 
 I8 -- I11  [style=bold,color=orange]; 
 I9 -- I11  [style=bold,color=orange]; 
 I8 -- I12  [style=bold,color=orange]; 
 I9 -- I12  [style=bold,color=orange]; 
 I8 -- I13  [style=bold,color=orange]; 
 I9 -- I13  [style=bold,color=orange]; 
 I8 -- I9  [style=bold,color=violet]; 
 I9 -- I1  [style=bold,color=red]; 
 I2 -- I5  [style=bold,color=red]; 
 I2 -- I15  [style=bold,color=violet]; 
 I3 -- I2  [style=bold,color=blue]; 
 I3 -- I4  [style=bold,color=violet]; 
 I4 -- I2  [style=bold,color=red]; 
}')

```