### 2023-06-12 Visuals_Interactives/tikz_net.qmd

---
title: tikz
date: last-modified
categories: [Visualization, R, OJS]
page-layout: full
code-fold: true
code-tools: true
execute: 
  warning: true
  echo: fenced
# header-includes:
#    - \usepackage{tikz}
#    - \usepackage[usenames,dvipsnames]{xcolor}
#    - \usepackage{pgfplots}
#    - \usepackage{neuralnetwork}
#   - \usepackage{pgf,tikz,pgfplots}
#   - \usepackage{ifthen}
#   - \usepackage{asmath,contour,listofitems,xcolors}
#   - \usepackage{neuralnetwork}
---

https://www.andrewheiss.com/blog/2021/08/27/tikz-knitr-html-svg-fun/
https://gist.github.com/andrewheiss/4ece621813a27dfdcaef7f1c2d773237

```{r}
#lapply(c('tidyverse','data.table','igraph','ggraph','kableExtra'),library,character.only=TRUE))
pacman::p_load(tikzDevice, knitr) 
```

```{r point-to-ghostscript, include=FALSE}
Sys.setenv(LIBGS = "/usr/local/share/ghostscript/10.00/lib/libgs.dylib.10.00")
embed_svg_fonts <- function(before, options, envir) {
  # !before means after the chunk is run
  if (!before) {
    # Get a list of all the plot files this chunk makes
    paths <- knitr:::get_plot_files()
    # Loop through all the plot files. When dealing with SVG, knitr actually
    # creates two different files: an initial DVI and the converted SVG. We'll
    # use the original DVI and pass it through dvisvgm again, this time with
    # --font-format=woff enabled. This will overwrite the SVG that knitr makes
    knitr:::in_base_dir(
      lapply(paths, function(x) {
        message("Embedding fonts in ", x)
        path_svg <- xfun::with_ext(x, "svg")
        path_dvi <- xfun::with_ext(x, "dvi")

        if (system2('dvisvgm', c('--font-format=woff', '-o', shQuote(path_svg), shQuote(path_dvi))) != 0)
          stop('Failed to convert ', path_dvi, ' to ', path_svg)
      })
    )
  }
}
# Register the function as a possible hook
# Use this in a chunk by setting embed_svg_fonts=TRUE in the chunk options
knitr::knit_hooks$set(embed_svg_fonts = embed_svg_fonts)

# Conditional tikz output types; use PDF if this is LaTeX, otherwise use SVG
if (knitr::is_latex_output()) {
  knitr::opts_template$set(
    tikz_settings = list(fig.ext = "pdf", fig.align = "center")
  )
} else {
  knitr::opts_template$set(
    tikz_settings = list(fig.ext = "svg", fig.align = "center", embed_svg_fonts = TRUE)
  )
}
```

```{r setup, include=FALSE}

font_opts <- list(extra.preamble = c("\\usepackage{libertine}", #"\\usepackage{libertinust1math}",
                                     "\\usepackage{xstring}",
                                     "\\usepackage{neuralnetwork}"),
                  dvisvgm.opts = "--font-format=woff")

font_opts2 <- list(extra.preamble = c("\\usepackage{amsmath}", #"alignment
                                     "\\usepackage{listofitems}", # create arrays
                                     "\\usepackage{xcolor}", 
                                     "\\usepackage[outline]{contour}"),  # glow around text
                  dvisvgm.opts = "--font-format=woff")

```

```{tikz rectangle-text, fig.ext="svg", echo=FALSE, fig.align="center"}
#| engine.opts=font_opts
#| 
\begin{tikzpicture}
\draw (0,0) rectangle (6, 2) node[midway, align=center] {I am a rectangle with math: \\ $\hat{y} = \beta_0 + \beta_1 x_1 + \text{Some text} + \varepsilon$};
\end{tikzpicture}
```

## Intro
Here is a TikZ picture

```{r, engine = 'tikz'}
#| eval: true
\tikzset{
    declare function={
        sig = 0.1;
        mu = 0;
        g(\x) = 1/(sig*sqrt(2*pi)) * exp(-1/2 * ((\x-mu)/sig)^2);
    }
}

\begin{tikzpicture}[
    shorten >=1pt,
    ->,
    draw=black!50,
    node distance=2.5cm,
    scale=1.5,
    every pin edge/.style={<-,shorten <=1pt},
    neuron/.style={
        circle,fill=black!25,minimum size=17pt,inner sep=0pt,
        path picture={
            \draw[red,thick,-] plot[domain=-0.3:0.3,samples=11,smooth] ({\x},{0.05*g(\x)});
        },
    },
    input neuron/.style={neuron, fill=green!50},
    output neuron/.style={neuron, fill=red!50},
    hidden neuron/.style={neuron, fill=blue!50},
    annot/.style={text width=4em, text centered},
]

	% Draw the input layer nodes
	\foreach \name / \y in {1,...,4}
	% This is the same as writing \foreach \name / \y in {1/1,2/2,3/3,4/4}
	    \node[input neuron, pin=left:Input \y] (I-\name) at (0,-\y) {};

	% Draw the hidden layer nodes
	\foreach \name / \y in {1,...,5}
	    \path[yshift=0.5cm]
	        node[hidden neuron] (H-\name) at (2.5cm,-\y cm) {};

	% Draw the output layer node
	\node[output neuron,pin={[pin edge={->}]right:Output}, right of=H-3] (O) {};

	% Connect every node in the input layer with every node in the
	% hidden layer.
	\foreach \source in {1,...,4}
	    \foreach \dest in {1,...,5}
	        \path (I-\source) edge (H-\dest);

	% Connect every node in the hidden layer with the output layer
	\foreach \source in {1,...,5}
	    \path (H-\source) edge (O);

	% Annotate the layers
	\node[annot,above of=H-1, node distance=1cm] (hl) {Hidden layer};
	\node[annot,above of=I-1, node distance=1cm] {Input layer};
	\node[annot,above of=O] {Output layer};

\end{tikzpicture}

```

```{r, engine = 'tikz'}
#| eval: true
#| cache: true

\usetikzlibrary{matrix,chains,positioning,decorations.pathreplacing,arrows}

\begin{tikzpicture}[
	init/.style={
	  draw,
	  circle,
	  inner sep=2pt,
	  font=\Huge,
	  join = by -latex
	},
	squa/.style={
	  draw,
	  inner sep=2pt,
	  font=\Large,
	  join = by -latex
	},
	start chain=2,node distance=13mm
	]
	\node[on chain=2] 
	  (x2) {$x_2$};
	\node[on chain=2,join=by o-latex] 
	  {$w_2$};
	\node[on chain=2,init] (sigma) 
	  {$\displaystyle\Sigma$};
	\node[on chain=2,squa,label=above:{\parbox{2cm}{\centering Activate \\ function}}]   
	  {$f$};
	\node[on chain=2,label=above:Output,join=by -latex] 
	  {$y$};
	\begin{scope}[start chain=1]
	\node[on chain=1] at (0,1.5cm) 
	  (x1) {$x_1$};
	\node[on chain=1,join=by o-latex] 
	  (w1) {$w_1$};
	\end{scope}
	\begin{scope}[start chain=3]
	\node[on chain=3] at (0,-1.5cm) 
	  (x3) {$x_3$};
	\node[on chain=3,label=below:Weights,join=by o-latex] 
	  (w3) {$w_3$};
	\end{scope}
	\node[label=above:\parbox{2cm}{\centering Bias \\ $b$}] at (sigma|-w1) (b) {};

	\draw[-latex] (w1) -- (sigma);
	\draw[-latex] (w3) -- (sigma);
	\draw[o-latex] (b) -- (sigma);

	\draw[decorate,decoration={brace,mirror}] (x1.north west) -- node[left=10pt] {Inputs} (x3.south west);
\end{tikzpicture}

```

```{r, engine = 'tikz'}
#| eval: true
#| cache: true

\usetikzlibrary{positioning}

\tikzset{basic/.style={draw,fill=blue!20,text width=1em,text badly centered}}
\tikzset{input/.style={basic,circle}}
\tikzset{weights/.style={basic,rectangle}}
\tikzset{functions/.style={basic,circle,fill=blue!10}}

\begin{tikzpicture}
    \node[functions] (center) {};
    \node[below of=center,font=\scriptsize,text width=4em] {Activation function};
    \draw[thick] (0.5em,0.5em) -- (0,0.5em) -- (0,-0.5em) -- (-0.5em,-0.5em);
    \draw (0em,0.75em) -- (0em,-0.75em);
    \draw (0.75em,0em) -- (-0.75em,0em);
    \node[right of=center] (right) {};
        \path[draw,->] (center) -- (right);
    \node[functions,left=3em of center] (left) {$\sum$};
        \path[draw,->] (left) -- (center);
    \node[weights,left=3em of left] (2) {$w_2$} -- (2) node[input,left of=2] (l2) {$x_2$};
        \path[draw,->] (l2) -- (2);
        \path[draw,->] (2) -- (left);
    \node[below of=2] (dots) {$\vdots$} -- (dots) node[left of=dots] (ldots) {$\vdots$};
    \node[weights,below of=dots] (n) {$w_n$} -- (n) node[input,left of=n] (ln) {$x_n$};
        \path[draw,->] (ln) -- (n);
        \path[draw,->] (n) -- (left);
    \node[weights,above of=2] (1) {$w_1$} -- (1) node[input,left of=1] (l1) {$x_1$};
        \path[draw,->] (l1) -- (1);
        \path[draw,->] (1) -- (left);
    \node[weights,above of=1] (0) {$w_0$} -- (0) node[input,left of=0] (l0) {$1$};
        \path[draw,->] (l0) -- (0);
        \path[draw,->] (0) -- (left);
    \node[below of=ln,font=\scriptsize] {inputs};
    \node[below of=n,font=\scriptsize] {weights};
\end{tikzpicture}

```

```{r, engine = 'tikz'}
#| eval: true
#| cache: true

\usetikzlibrary{%
  calc,
  fit,
  shapes,
  backgrounds
}
% the next macro is useful to create a table
\newcommand\tabins[3]{%
 \tikz[baseline=(Tab.base)] 
           \node  [rectangle split, 
                   rectangle split parts=3, 
                   draw, 
                   align=right,
                   inner sep=.5em,
                   rectangle split horizontal] (Tab)
                           {\hbox to 4ex{#1}
           \nodepart{two}  {\hbox to 8ex{\hfill #2\$}}  
           \nodepart{three}{\hbox to 3ex{#3}}}; 
}

\parindent=0pt

\begin{tikzpicture}[%
    %every node/.style={transform shape},% now is not necessary but good for a poster
    x=1.25cm,y=2cm,  
    font=\footnotesize,
    % every group of nodes have a style except for main, the style is named by a letter
    main/.style={draw,fill=yellow,inner sep=.5em},
    R/.style={draw,fill=purple!40!blue!30,inner sep=.5em},
    M/.style={draw,fill=green!80!yellow,inner sep=.5em},
    S/.style={anchor=east},
    V/.style={anchor=west},
    P/.style={anchor=center},
    F/.style={anchor=west}
    ]

  % main node the reference Shuffle 
  \node[main] (shuffle) {Group};
  %group R reducer
  \node[R] at ($(shuffle)+(8,1)$)    (R1+) {Reduce};
  \node[R] at ($(shuffle)+(8, 0)$)   (R0)  {Reduce};
  \node[R] at ($(shuffle)+(8,-1)$)   (R1-) {Reduce};
  % group M Mapper
  \node[M] at ($(shuffle)+(-6,+2.5)$)   (M3+)  {Map};
  \node[M] at ($(shuffle)+(-6,+ 1.5)$)  (M2+)  {Map};
  \node[M] at ($(shuffle)+(-6,+ .5)$)   (M1+)  {Map};
  \node[M] at ($(shuffle)+(-6,- .5)$)   (M1-)  {Map};
  \node[M] at ($(shuffle)+(-6,- 1.5)$)  (M2-)  {Map};
  \node[M] at ($(shuffle)+(-6,-2.5)$)   (M3-)  {Map};
  % group S Start the first nodes
  \node[S] at ($(M3+)+(-1.5,0)$)  (S3+) {\Big($k_1$,\tabins{4711}{59.90}{NY}\Big)};
  \node[S] at ($(M2+)+(-1.5,0)$)  (S2+) {\Big($k_2$,\tabins{4713}{142.99}{CA}\Big)};
  \node[S] at ($(M1+)+(-1.5,0)$)  (S1+) {\Big($k_3$,\tabins{4714}{72.00}{NY}\Big)}; 
  \node[S] at ($(M1-)+(-1.5,0)$)  (S1-) {\Big($k_4$,\tabins{4715}{108.75}{NY}\Big)}; 
  \node[S] at ($(M2-)+(-1.5,0)$)  (S2-) {\Big($k_5$,\tabins{4718}{19.89}{WA}\Big)};  
  \node[S] at ($(M3-)+(-1.5,0)$)  (S3-) {\Big($k_6$,\tabins{4719}{36.60}{CA}\Big)};  
  % group V  why not
  \node[V] at ($(M3+)+(1.5,0)$)  (V3+) {\Big(NY,59.90\$\Big)};
  \node[V] at ($(M2+)+(1.5,0)$)  (V2+) {\Big(CA,142.99\$\Big)};
  \node[V] at ($(M1+)+(1.5,0)$)  (V1+) {\Big(NY,72.00\$\Big)}; 
  \node[V] at ($(M1-)+(1.5,0)$)  (V1-) {\Big(NY,108.75\$\Big)}; 
  \node[V] at ($(M2-)+(1.5,0)$)  (V2-) {\Big(WA,19.89\$\Big)};  
  \node[V] at ($(M3-)+(1.5,0)$)  (V3-) {\Big(CA,36.60\$\Big)};

  \node[P] at ($(R1+)+(-4,0)$) (P1+) {\Big(CA,\big[142.99\$,36.60\$\big]\Big)};
  \node[P] at ($(R0) +(-4,0)$) (P0)  {\Big(NY,\big[59.90\$,72.00\$,108.75\big]\Big)};
  \node[P] at ($(R1-)+(-4,0)$) (P1-) {\Big(WA,\big[19.89\$\big]\Big)};

  \node[F] (F1+) at ($(R1+)+(1.5,0)$) {(CA,89.80\$)};
  \node[F] (F0)  at ($(R0) +(1.5,0)$) {(NY,80.22\$)}; 
  \node[F] (F1-) at ($(R1-)+(1.5,0)$) {(WA,72.00\$)};

  % wrappers
  \begin{scope}[on background layer]
      \node[fill=lightgray!50,inner sep = 4mm,fit=(shuffle),label=above:Shuffle] {}; 
  \end{scope} 
  \begin{scope}[on background layer]
      \node[fill=lightgray!50,inner sep = 4mm,fit=(R1+)(R1-),label=above:Reducer] {}; 
  \end{scope}  
  \begin{scope}[on background layer]
      \node[fill=lightgray!50,inner sep = 4mm,fit=(M3+)(M3-),label=above:Mapper] {}; 
  \end{scope}

  %edges

  \foreach \indice in {3+,2+,1+,1-,2-,3-} \draw[->] (S\indice.east) -- (M\indice.west); 
  \foreach \indice in {3+,2+,1+,1-,2-,3-} \draw[->] (M\indice.east) -- (V\indice.west);
  \foreach \indice in {3+,2+,1+,1-,2-,3-} \draw[->] (V\indice.east) to [out=0,in=180] (shuffle.west); 
  \foreach \indice in {1+,0,1-} \draw[->] (shuffle.east) to [out=0,in=180] (P\indice.west);  
  \foreach \indice in {1+,0,1-} \draw[->] (P\indice.east) -- (R\indice.west);
  \foreach \indice in {1+,0,1-} \draw[->] (R\indice.east) -- (F\indice.west);   
\end{tikzpicture}

```

```{r, engine = 'tikz'}
#| eval: true
#| cache: true
#| 
\def\layersep{3cm}
\def\nodeinlayersep{1.5cm}

\begin{tikzpicture}
  [
    shorten >=1pt,->,
    draw=black!50,
    node distance=\layersep,
    every pin edge/.style={<-,shorten <=1pt},
    neuron/.style={circle,fill=black!25,minimum size=17pt,inner sep=0pt},
    input neuron/.style={neuron, fill=green!50,},
    output neuron/.style={neuron, fill=red!50},
    hidden neuron/.style={neuron, fill=blue!50},
    annot/.style={text width=4em, text centered},
    bias/.style={neuron, fill=yellow!50,minimum size=4em},%<-- added %%%
  ]

  % Draw the input layer nodes
  \foreach \name / \y in {1,...,3}
    \node[input neuron, pin=left:Input \#\y] (I-\name) at (0,-\y-2.5) {};

  % set number of hidden layers
  \newcommand\Nhidden{2}

  % Draw the hidden layer nodes
  \foreach \N in {0,...,\Nhidden} {
      \foreach \y in {0,...,5} { % <-- added 0 instead of 1 %%%%%
      \ifnum \y=4
      \ifnum \N>0 %<-- added %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \node at (\N*\layersep,-\y*\nodeinlayersep) {$\vdots$};  % add dots
        \else\fi %<-- added %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      \else
          \ifnum \y=0 %<-- added %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          \ifnum \N<3 %<-- added %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
            \node[bias] (H\N-\y) at (\N*\layersep,-\y*\nodeinlayersep ) {Bias}; %<-- added
            \else\fi %<-- added %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
          \else %<-- added %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
            \ifnum \N>0 %<-- added %%%%%%%%%%%%%%%%%%%%%%%%%
            % print function
            \node[hidden neuron] (H\N-\y) at (\N*\layersep,-\y*\nodeinlayersep ) {$\frac{1}{1+e^{-x}}$}; %<-- added %%%%%%%%%%%
                \else\fi %<-- added %%%%%%%%%%%%
          \fi %<-- added %%%%%%%
      \fi
    }
      \ifnum \N>0 %<-- added %%%%%%
      % print hidden layer labels at the top
    \node[annot,above of=H\N-1, node distance=1cm,yshift=2cm] (hl\N) {Hidden layer \N}; % <- added yshift=2cm %%%%%%%%%%%%
    \else\fi %<-- added %%%%%
  }

  % Draw the output layer node and label
  \node[output neuron,pin={[pin edge={->}]right:Output}, right of=H\Nhidden-3] (O) {};

  % Connect bias every node in the input layer with every node in the
  % hidden layer.
  \foreach \source in {1,...,3}
      \foreach \dest in {1,...,3,5} {
        % \path[yellow] (H-0) edge (H1-\dest);
        \path[dashed,orange] (H0-0) edge (H1-\dest); %<-- added %%%%%
          \path[green!50] (I-\source) edge (H1-\dest);  % change to green, yellow gets blended
    };

  % connect all hidden stuff
  \foreach [remember=\N as \lastN (initially 1)] \N in {2,...,\Nhidden}
      \foreach \source in {0,...,3,5} 
          \foreach \dest in {1,...,3,5}{
              \ifnum \source=0 %<-- added %%%%%%%%%%%%%%%%%%%%%%%
          \path[dashed,red](H\lastN-\source) edge (H\N-\dest);%<-- added 
            \else %<-- added %%%
            \path[blue!50] (H\lastN-\source) edge (H\N-\dest);%<-- added 
            \fi %<-- added %%%
            }; %<-- added %%%%

  % Connect every node in the hidden layer with the output layer
  \foreach \source in {1,...,3,5}
    \path[green!50] (H\Nhidden-\source) edge (O);
    \path[dashed,red] (H2-0) edge (O); %<-- added %%%%

% Annotate the input and output layers
  \node[annot,left of=hl1] {Input layer};
  \node[annot,right of=hl\Nhidden] {Output layer};  
\end{tikzpicture}

```

```{r, engine = 'tikz'}
#| eval: true
#| cache: true

\usetikzlibrary{positioning}

\tikzstyle{inputNode}=[draw,circle,minimum size=10pt,inner sep=0pt]
\tikzstyle{stateTransition}=[-stealth, thick]

\begin{tikzpicture}
	\node[draw,circle,minimum size=25pt,inner sep=0pt] (x) at (0,0) {$\Sigma$ $\sigma$};

	\node[inputNode] (x0) at (-2, 1.5) {$\tiny +1$};
	\node[inputNode] (x1) at (-2, 0.75) {$\tiny x_1$};
	\node[inputNode] (x2) at (-2, 0) {$\tiny x_2$};
	\node[inputNode] (x3) at (-2, -0.75) {$\tiny x_3$};
	\node[inputNode] (xn) at (-2, -1.75) {$\tiny x_n$};

	\draw[stateTransition] (x0) to[out=0,in=120] node [midway, sloped, above] {$w_0$} (x);
	\draw[stateTransition] (x1) to[out=0,in=150] node [midway, sloped, above] {$w_1$} (x);
	\draw[stateTransition] (x2) to[out=0,in=180] node [midway, sloped, above] {$w_2$} (x);
	\draw[stateTransition] (x3) to[out=0,in=210] node [midway, sloped, above] {$w_3$} (x);
	\draw[stateTransition] (xn) to[out=0,in=240] node [midway, sloped, above] {$w_n$} (x);
	\draw[stateTransition] (x) -- (4,0) node [midway,above] {$\sigma\left(w_0 + \sum\limits_{i=1}^{n}{w_ix_i}\right)$};
	\draw[dashed] (0,-0.43) -- (0,0.43);
	\node (dots) at (-2, -1.15) {$\vdots$};
	\node[inputNode, thick] (i1) at (6, 0.75) {};
	\node[inputNode, thick] (i2) at (6, 0) {};
	\node[inputNode, thick] (i3) at (6, -0.75) {};

	\node[inputNode, thick] (h1) at (8, 1.5) {};
	\node[inputNode, thick] (h2) at (8, 0.75) {};
	\node[inputNode, thick] (h3) at (8, 0) {};
	\node[inputNode, thick] (h4) at (8, -0.75) {};
	\node[inputNode, thick] (h5) at (8, -1.5) {};

	\node[inputNode, thick] (o1) at (10, 0.75) {};
	\node[inputNode, thick] (o2) at (10, -0.75) {};

	\draw[stateTransition] (5, 0.75) -- node[above] {$I_1$} (i1);
	\draw[stateTransition] (5, 0) -- node[above] {$I_2$} (i2);
	\draw[stateTransition] (5, -0.75) -- node[above] {$I_3$} (i3);

	\draw[stateTransition] (i1) -- (h1);
	\draw[stateTransition] (i1) -- (h2);
	\draw[stateTransition] (i1) -- (h3);
	\draw[stateTransition] (i1) -- (h4);
	\draw[stateTransition] (i1) -- (h5);
	\draw[stateTransition] (i2) -- (h1);
	\draw[stateTransition] (i2) -- (h2);
	\draw[stateTransition] (i2) -- (h3);
	\draw[stateTransition] (i2) -- (h4);
	\draw[stateTransition] (i2) -- (h5);
	\draw[stateTransition] (i3) -- (h1);
	\draw[stateTransition] (i3) -- (h2);
	\draw[stateTransition] (i3) -- (h3);
	\draw[stateTransition] (i3) -- (h4);
	\draw[stateTransition] (i3) -- (h5);

	\draw[stateTransition] (h1) -- (o1);
	\draw[stateTransition] (h1) -- (o2);
	\draw[stateTransition] (h2) -- (o1);
	\draw[stateTransition] (h2) -- (o2);
	\draw[stateTransition] (h3) -- (o1);
	\draw[stateTransition] (h3) -- (o2);
	\draw[stateTransition] (h4) -- (o1);
	\draw[stateTransition] (h4) -- (o2);
	\draw[stateTransition] (h5) -- (o1);
	\draw[stateTransition] (h5) -- (o2);

	\node[above=of i1, align=center] (l1) {Input \\ layer};
	\node[right=2.3em of l1, align=center] (l2) {Hidden \\ layer};
	\node[right=2.3em of l2, align=center] (l3) {Output \\ layer};

	\draw[stateTransition] (o1) -- node[above] {$O_1$} (11, 0.75);
	\draw[stateTransition] (o2) -- node[above] {$O_2$} (11, -0.75);

	\path[dashed, double, ultra thick, gray] (x.north) edge[bend left=0] (h5.north);
	\path[dashed, double, ultra thick, gray] (x.south) edge[bend right=0] (h5.south);
\end{tikzpicture}

```

```{r, engine = 'tikz'}
#| eval: false

\usetikzlibrary{positioning,decorations.pathreplacing,shapes}

\newcommand*{\cancer}{\text{cancer}}
\newcommand*{\testp}{\text{test}+}

\begin{tikzpicture}[%
   % common options for blocks:
   block/.style = {draw, fill=blue!30, align=center, anchor=west,
               minimum height=0.65cm, inner sep=0},
   % common options for the circles:
   ball/.style = {circle, draw, align=center, anchor=north, inner sep=0}]

   % circle illustrating all women
   \node[ball,text width=3cm,fill=purple!20] (all) at (6,0) {All women};

   % two circles showing split of p{cancer} and p{~cancer}
   \node[ball,fill=red!70,text width=0.1cm,anchor=base] (pcan) at (3.5,-5.5) {};
   \node[ball,fill=blue!40,text width=2.9cm,anchor=base] (pncan) at (8.5,-6)
      {Women without cancer\\
      $\p({\sim}\cancer) = 99\%$};

   % arrows showing split from all women to cancer and ~cancer
   \draw[->,thick,draw=red!50] (all.south) to [out=270,in=90] (pcan.north);
   \draw[->,thick,draw=blue!80] (all.south) to [out=270,in=110] (pncan.100);

   % transition from all women to actual cancer rates
   \node[anchor=north,text width=10cm,inner sep=.05cm,align=center,fill=white]
   (why1) at (6,-3.7) {In measuring, we find:};

   % note illustration the p{cancer} circle (text wont fit inside)
   \node[inner sep=0,anchor=east,text width=3.3cm] (note1) at (3.2,-5.5) {
      Women with cancer $\p(\cancer) = 1\%$};

   % draw the sieves
   \node[block,anchor=north,text width=4.4cm,fill=green!50] (tray1) at
      (3.5,-8.8) {\small{$\p(\testp\mid\cancer)=0.8$}};

   \node[block,anchor=north,text width=4.4cm,fill=green!50] (tray2) at
      (8.5,-8.8) {$\p(\testp\mid{\sim}\cancer)=0.096$};

   % text explaining how p{cancer} and p{~cancer} behave as they
   % pass through the sieves
   \node[anchor=west,text width=6cm] (note1) at (-6,-9.1) {
      Now we pass both groups through the sieve; note that both
      sieves are \emph{the same}; they just behave differently
      depending on which group is passing through. \\ 
      Let $\testp=$ a positve mammography.};

   % arrows showing the circles passing through the seives
   \draw[->,thick,draw=red!80] (3.5,-5.9) -- (3.5,-8.6);
   \draw[->,thick,draw=blue!50] (8.5,-8.1) -- (8.5,-8.6);

   % numerator
   \node[ball,text width=0.05cm,fill=red!70] (can) at (6,-10.5) {};

   % dividing line
   \draw[thick] (5,-11) -- (7,-11);

   % demoniator
   \node[ball,text width=0.39cm,fill=blue!40,anchor=base] (ncan) at (6.5,-11.5) {};
   \node[ball,text width=0.05cm,fill=red!70,anchor=base] (can2) at (5.5,-11.5) {};

   % plus sign in denominator
   \draw[thick] (5.9,-11.4) -- (5.9,-11.6);
   \draw[thick] (5.8,-11.5) -- (6,-11.5);

   % arrows showing the output of the sieves formed the fraction
   \draw[->,thick,draw=red!80] (tray1.south) to [out=280,in=180] (can);
   \draw[->,thick,draw=red!80] (tray1.south) to [out=280,in=180] (can2);
   \node[anchor=north,inner sep=.1cm,align=center,fill=white] (why2) at
      (3.8,-9.8) {$1\% * 80\%$};

   \draw[->,thick,draw=blue!50] (tray2.south) to [out=265,in=0] (ncan);
   \node[anchor=north,inner sep=.1cm,align=center,fill=white] (why2) at
      (8.4,-9.8) {$99\% * 9.6\%$};

   % explanation of final formula
   \node[anchor=north west,text width=6.5cm] (note2) at (-6,-12.5)
      {Finally, to find the probability that a positive test
         \emph{actually means cancer}, we look at those who passed
         through the sieve \emph{with cancer}, and divide by all who
         received a positive test, cancer or not.};

   % illustrated fraction turned into math
   \node[anchor=north,text width=10cm] (solution) at (6,-12.5) {
   \begin{align*}
         \frac{\p(\testp\mid\cancer)}{\p(\testp\mid\cancer)
         + \p(\testp\mid{\sim}\cancer)} &= \\
         \frac{1\% * 80\%}{(1\% * 80\%) + (99\% * 9.6\%)} &= 7.8\%
         = \p(\cancer\mid\testp)
      \end{align*}};
\end{tikzpicture}

```

```{r, engine = 'tikz'}
#| eval: false

# \usetikzlibrary{arrows}
# \usetikzlibrary{positioning}
# \usetikzlibrary{calc}
# \usetikzlibrary{arrows.meta}
# \usetikzlibrary{decorations.pathreplacing}

# \begin{tikzpicture}
# \draw (0,0)node(a){} -- (10,0) node (b) {} ;
# \foreach \x in  {0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10} % edit here for the vertical lines
# \draw[shift={(\x,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
# \foreach \x in {0, 0.2, 0.4, 0.6, 0.8, 1} % edit here for the numbers
# \draw[shift={(\x*10,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below]
# {$\x$};
# \node at (8, 0.5) (eq1) {$\textcolor{red}{\boldsymbol{SQ}}$};
# \node at (4, 0.5) (eq2) {$\textcolor{purple}{\boldsymbol{G_i(0)}}$}; 
# \node at (7, 0.5) (eq2) {$\textcolor{purple}{\boldsymbol{G_i(1)}}$}; 
# \node at (3, 0.5) (eq3) {$\textcolor{blue}{\boldsymbol{P}}$};
# \node at (0, 0.5) (eq4) {$\textcolor{black}{\boldsymbol{x_i}}$};
# \draw[decorate, decoration={brace, amplitude=6pt, mirror},] ([yshift=0.5cm]4,0.5)-- node[above=0.25cm]
# {\shortstack{Text}}([yshift=0.5cm]3,0.5);
# \draw[decorate, decoration={brace, amplitude=6pt},] ([yshift=-1cm]7,0)-- node[below=0.25cm]
# {\shortstack{Text}}([yshift=-1cm]3,0);
# \end{tikzpicture}
```

### Tikz Neural Networks
https://tikz.net/neural_networks/

```{r, engine = 'tikz', engine.opts=font_opts2, cache=TRUE}
#| cache: true

\usetikzlibrary{arrows.meta} % for arrow size

\tikzset{>=latex} % for LaTeX arrow head
\colorlet{myred}{red!80!black}
\colorlet{myblue}{blue!80!black}
\colorlet{mygreen}{green!60!black}
\colorlet{myorange}{orange!70!red!60!black}
\colorlet{mydarkred}{red!30!black}
\colorlet{mydarkblue}{blue!40!black}
\colorlet{mydarkgreen}{green!30!black}
\tikzstyle{node}=[thick,circle,draw=myblue,minimum size=22,inner sep=0.5,outer sep=0.6]
\tikzstyle{node in}=[node,green!20!black,draw=mygreen!30!black,fill=mygreen!25]
\tikzstyle{node hidden}=[node,blue!20!black,draw=myblue!30!black,fill=myblue!20]
\tikzstyle{node convol}=[node,orange!20!black,draw=myorange!30!black,fill=myorange!20]
\tikzstyle{node out}=[node,red!20!black,draw=myred!30!black,fill=myred!20]
\tikzstyle{connect}=[thick,mydarkblue] %,line cap=round
\tikzstyle{connect arrow}=[-{Latex[length=4,width=3.5]},thick,mydarkblue,shorten <=0.5,shorten >=1]
\tikzset{ % node styles, numbered for easy mapping with \nstyle
  node 1/.style={node in},
  node 2/.style={node hidden},
  node 3/.style={node out},
}
\def\nstyle{int(\lay<\Nnodlen?min(2,\lay):3)} % map layer number onto 1, 2, or 3

\begin{tikzpicture}[x=2.7cm,y=1.6cm]
  \message{^^JNeural network activation}
  \def\NI{5} % number of nodes in input layers
  \def\NO{4} % number of nodes in output layers
  \def\yshift{0.4} % shift last node for dots

  % INPUT LAYER
  \foreach \i [evaluate={\c=int(\i==\NI); \y=\NI/2-\i-\c*\yshift; \index=(\i<\NI?int(\i):"n");}]
              in {1,...,\NI}{ % loop over nodes
    \node[node in,outer sep=0.6] (NI-\i) at (0,\y) {$a_{\index}^{(0)}$};
  }

  % OUTPUT LAYER
  \foreach \i [evaluate={\c=int(\i==\NO); \y=\NO/2-\i-\c*\yshift; \index=(\i<\NO?int(\i):"m");}]
    in {\NO,...,1}{ % loop over nodes
    \ifnum\i=1 % high-lighted node
      \node[node hidden]
        (NO-\i) at (1,\y) {$a_{\index}^{(1)}$};
      \foreach \j [evaluate={\index=(\j<\NI?int(\j):"n");}] in {1,...,\NI}{ % loop over nodes in previous layer
        \draw[connect,white,line width=1.2] (NI-\j) -- (NO-\i);
        \draw[connect] (NI-\j) -- (NO-\i)
          node[pos=0.50] {\contour{white}{$w_{1,\index}$}};
      }
    \else % other light-colored nodes
      \node[node,blue!20!black!80,draw=myblue!20,fill=myblue!5]
        (NO-\i) at (1,\y) {$a_{\index}^{(1)}$};
      \foreach \j in {1,...,\NI}{ % loop over nodes in previous layer
        %\draw[connect,white,line width=1.2] (NI-\j) -- (NO-\i);
        \draw[connect,myblue!20] (NI-\j) -- (NO-\i);
      }
    \fi
  }

  % DOTS
  \path (NI-\NI) --++ (0,1+\yshift) node[midway,scale=1.2] {$\vdots$};
  \path (NO-\NO) --++ (0,1+\yshift) node[midway,scale=1.2] {$\vdots$};

  % EQUATIONS
  \def\agr#1{{\color{mydarkgreen}a_{#1}^{(0)}}}
  \node[below=17,right=11,mydarkblue,scale=0.95] at (NO-1)
    {$\begin{aligned} %\underset{\text{bias}}{b_1}
       &= \color{mydarkred}\sigma\left( \color{black}
            w_{1,0}\agr{0} + w_{1,1}\agr{1} + \ldots + w_{1,n}\agr{n} + b_1^{(0)}
          \color{mydarkred}\right)\\
       &= \color{mydarkred}\sigma\left( \color{black}
            \sum_{i=1}^{n} w_{1,i}\agr{i} + b_1^{(0)}
           \color{mydarkred}\right)
     \end{aligned}$};
  \node[right,scale=0.9] at (1.3,-1.3)
    {$\begin{aligned}
      {\color{mydarkblue}
      \begin{pmatrix}
        a_{1}^{(1)} \\[0.3em]
        a_{2}^{(1)} \\
        \vdots \\
        a_{m}^{(1)}
      \end{pmatrix}}
      &=
      \color{mydarkred}\sigma\left[ \color{black}
      \begin{pmatrix}
        w_{1,0} & w_{1,1} & \ldots & w_{1,n} \\
        w_{2,0} & w_{2,1} & \ldots & w_{2,n} \\
        \vdots  & \vdots  & \ddots & \vdots  \\
        w_{m,0} & w_{m,1} & \ldots & w_{m,n}
      \end{pmatrix}
      {\color{mydarkgreen}
      \begin{pmatrix}
        a_{1}^{(0)} \\[0.3em]
        a_{2}^{(0)} \\
        \vdots \\
        a_{n}^{(0)}
      \end{pmatrix}}
      +
      \begin{pmatrix}
        b_{1}^{(0)} \\[0.3em]
        b_{2}^{(0)} \\
        \vdots \\
        b_{m}^{(0)}
      \end{pmatrix}
      \color{mydarkred}\right]\\[0.5em]
      {\color{mydarkblue}a^{(1)}}
      &= \color{mydarkred}\sigma\left( \color{black}
           \mathbf{W}^{(0)} {\color{mydarkgreen}a^{(0)}}+\mathbf{b}^{(0)}
         \color{mydarkred}\right)
         %\color{black},\quad \mathbf{W}^{(0)} \in \mathbb{R}^{m\times n}
    \end{aligned}$};

\end{tikzpicture}

```

### Autoencoder
https://tikz.net/autoencoder/
```{r, engine = 'tikz', engine.opts=font_opts}
#| eval: true
#| cache: true

\newcommand{\xin}[2]{$x_#2$}
\newcommand{\xout}[2]{$\hat x_#2$}

\begin{neuralnetwork}[height=8]
  \tikzstyle{input neuron}=[neuron, fill=orange!70];
  \tikzstyle{output neuron}=[neuron, fill=blue!60!black, text=white];

  \inputlayer[count=8, bias=false, title=Input Layer, text=\xin]

  \hiddenlayer[count=5, bias=false]
  \linklayers

  \hiddenlayer[count=3, bias=false, title=Latent\\Representation]
  \linklayers

  \hiddenlayer[count=5, bias=false]
  \linklayers

  \outputlayer[count=8, title=Output Layer, text=\xout]
  \linklayers

\end{neuralnetwork}

```

### VAE
https://tikz.net/vae/
```{r, engine = 'tikz', engine.opts=font_opts, cache=TRUE}
#| eval: true

\usetikzlibrary{fit,positioning}

\newcommand\drawNodes[2]{
  % #1 (str): namespace
  % #2 (list[list[str]]): list of labels to print in the node of each neuron
  \foreach \neurons [count=\lyrIdx] in #2 {
    \StrCount{\neurons}{,}[\lyrLength] % use xstring package to save each layer size into \lyrLength macro
    \foreach \n [count=\nIdx] in \neurons
      \node[neuron] (#1-\lyrIdx-\nIdx) at (2*\lyrIdx, \lyrLength/2-1.4*\nIdx) {\n};
  }
}

\newcommand\denselyConnectNodes[2]{
  % #1 (str): namespace
  % #2 (list[int]): number of nodes in each layer
  \foreach \n [count=\lyrIdx, remember=\lyrIdx as \previdx, remember=\n as \prevn] in #2 {
    \foreach \y in {1,...,\n} {
      \ifnum \lyrIdx > 1
        \foreach \x in {1,...,\prevn}
          \draw[->] (#1-\previdx-\x) -- (#1-\lyrIdx-\y);
      \fi
    }
  }
}

\begin{tikzpicture}[
    shorten >=1pt, shorten <=1pt,
    neuron/.style={circle, draw, minimum size=4ex, thick},
    legend/.style={font=\large\bfseries},
  ]

  % encoder
  \drawNodes{encoder}{{{,,,,}, {,,,}, {,,}}}
  \denselyConnectNodes{encoder}{{5, 4, 3}}

  % decoder
  \begin{scope}[xshift=11cm]
    \drawNodes{decoder}{{{,,}, {,,,}, {,,,,}}}
    \denselyConnectNodes{decoder}{{3, 4, 5}}
  \end{scope}

  % mu, sigma, sample nodes
  \foreach \idx in {1,...,3} {
      \coordinate[neuron, right=2 of encoder-3-2, yshift=\idx cm,, fill=yellow, fill opacity=0.2] (mu-\idx);
      \coordinate[neuron, right=2 of encoder-3-2, yshift=-\idx cm, fill=blue, fill opacity=0.1] (sigma-\idx);
      \coordinate[neuron, right=4 of encoder-3-2, yshift=\idx cm-2cm, fill=green, fill opacity=0.1] (sample-\idx);
    }

  % mu, sigma, sample boxes
  \node [label=$\mu$, fit=(mu-1) (mu-3), draw, fill=yellow, opacity=0.45] (mu) {};
  \node [label=$\sigma$, fit=(sigma-1) (sigma-3), draw, fill=blue, opacity=0.3] (sigma) {};
  \node [label=sample, fit=(sample-1) (sample-3), draw, fill=green, opacity=0.3] (sample) {};

  % mu, sigma, sample connections
  \draw[->] (mu.east) edge (sample.west) (sigma.east) -- (sample.west);
  \foreach \a in {1,2,3}
  \foreach \b in {1,2,3} {
      \draw[->] (encoder-3-\a) -- (mu-\b);
      \draw[->] (encoder-3-\a) -- (sigma-\b);
      \draw[->] (sample-\a) -- (decoder-1-\b);
    }

  % input + output labels
  \foreach \idx in {1,...,5} {
      \node[left=0 of encoder-1-\idx] {$x_\idx$};
      \node[right=0 of decoder-3-\idx] {$\hat x_\idx$};
    }
  \node[above=0.1 of encoder-1-1] {input};
  \node[above=0.1 of decoder-3-1] {output};

\end{tikzpicture}

```

## bayes vs regular nn
https://tikz.net/regular-vs-bayes-nn/
```{r, engine = 'tikz', engine.opts=font_opts}
#| cache: true
#| fig-align: "center"

\usetikzlibrary{calc}
\def\layersep{3cm}

\newcommand\nn[1]{
    % Input layer
    \foreach \y in {1,...,2}
        \node[neuron, fill=green!40] (i\y-#1) at (0,\y+1) {$i\y$};

    % Hidden layer
    \foreach \y in {1,...,4}
        \path node[neuron, fill=blue!40] (h\y-#1) at (\layersep,\y) {$h\y$};

    % Output node
    \node[neuron, fill=red!40] (o-#1) at (2*\layersep,2.5) {$o$};

    % Connect every node in the input layer with every node in the hidden layer.
    \foreach \source in {1,...,2}
        \foreach \dest in {1,...,4}
            \path (i\source-#1) edge (h\dest-#1);

    % Connect every node in the hidden layer with the output layer
    \foreach \source in {1,...,4}
        \path (h\source-#1) edge (o-#1);
}

\begin{tikzpicture}[
    scale=1.2,
    shorten >=1pt,->,draw=black!70, node distance=\layersep,
    neuron/.style={circle,fill=black!25,minimum size=20,inner sep=0},
    edge/.style 2 args={pos={(mod(#1+#2,2)+1)*0.33}, font=\tiny},
    distro/.style 2 args={
        edge={#1}{#2}, node contents={}, minimum size=0.6cm, path picture={\draw[double=orange,white,thick,double distance=1pt,shorten >=0pt] plot[variable=\t,domain=-1:1,samples=51] ({\t},{0.2*exp(-100*(\t-0.05*(#1-1))^2 - 3*\t*#2))});}
      },
    weight/.style 2 args={
        edge={#1}{#2}, node contents={\pgfmathparse{0.35*#1-#2*0.15}\pgfmathprintnumber[fixed]{\pgfmathresult}}, fill=white, inner sep=2pt
      }
  ]
  \nn{regular}

  \begin{scope}[xshift=8cm]
    \nn{bayes}
  \end{scope}

  % Draw weights for all regular edges.
  \foreach \i in {1,...,2}
  \foreach \j in {1,...,4}
  \path (i\i-regular) -- (h\j-regular) node[weight={\i}{\j}];
  \foreach \i in {1,...,4}
  \path (h\i-regular) -- (o-regular) node[weight={\i}{1}];

  % Draw distros for all Bayesian edges.
  \foreach \i in {1,...,2}
  \foreach \j in {1,...,4}
  \path (i\i-bayes) -- (h\j-bayes) node[distro={\i}{\j}];
  \foreach \i in {1,...,4}
  \path (h\i-bayes) -- (o-bayes) node[distro={\i}{1}];
\end{tikzpicture}

```

## nndiagram
https://github.com/ccfang2/nndiagram
```{r}
#| results: 'asis'
#| eval: false

library(nndiagram)
nnd <- nndiagram(input=3, hidden=c(4,4,4))
cat(paste(nnd,"\n"))
```

```{r, engine = 'tikz'}
#| eval: true
#| cache: true

\def\layersep{2.5cm} 
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{ 
  \node[shape=rectangle,inner sep=3pt, draw=black!100, fill= black!25] (char) {#1};}}

\centering 
\begin{tikzpicture}[shorten >=1pt,->,draw=black!100, node distance=\layersep, scale=1] 
  \tikzstyle{every pin edge}=[<-,shorten <=1pt]; 
  \tikzstyle{neuron}=[circle, draw=black!100, minimum size=17pt,inner sep=0pt]; 
  \tikzstyle{input neuron}=[neuron]; 
  \tikzstyle{output neuron}=[neuron]; 
  \tikzstyle{hidden n