### 2023-06-22 Analysis/data_organize.qmd
small org updates

---
title: Organization
---

```{r}
pacman::p_load(tidyverse,lme4,emmeans,here)
d <- readRDS(here("data/dPrune-05-21-23.rds"))
d$stage <- factor(d$stage, levels = c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23"))
d$tOrder <- factor(d$tOrder, levels=c("testFirst","trainFirst"))

```

```{r}
head(d)
colnames(d)

# identify which columns are factors
d %>% select_if(is.factor) %>% colnames()

# show levels of each factor variable
d %>% select_if(is.factor) %>% map(levels)

```

```{r}

```