### 2023-06-22 Analysis/test_analysis-all.qmd

---
title: HTW Test Analysis
date: last-modified
categories: [Analysis, R]
page-layout: full
execute: 
  warning: false
  eval: true
---

```{r}
pacman::p_load(tidyverse,data.table,lme4,emmeans,here)
d <- readRDS(here("data/dPrune-05-21-23.rds"))

dtest <- d %>% filter(expMode %in% c("test-Nf","test-train-nf")) %>% group_by(id,lowBound) %>% 
  mutate(nBand=n(),band=bandInt,id=factor(id)) %>% group_by(id) %>% mutate(nd=n_distinct(lowBound))
dtest <- dtest %>% group_by(id,lowBound) %>% filter(nBand>=5 & nd==6)
dtest <- dtest %>% group_by(id) %>% filter(!id %in% unique(dtest$id[dtest$nBand<5]))

dtestAgg <- dtest %>% group_by(id,condit,catOrder,fb,vb,band,lowBound,highBound,input) %>% mutate(vxCapped=ifelse(vx>1600,1600,vx)) %>%
  summarise(vxMean=mean(vx),devMean=mean(dist),vxMed=median(vx),devMed=median(dist),
            vxMeanCap=mean(vxCapped),.groups = "keep")

```

```{r}
model <- lmer(devMean ~ condit * catOrder * fb + (1|id), data = dtestAgg)
summary(model)

```
Linear mixed model fit by REML ['lmerMod']
Formula: devMean ~ condit * catOrder * fb + (1 | id)
   Data: dtestAgg

REML criterion at convergence: 35822.8

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-2.2757 -0.6267 -0.1857  0.4206  5.3481

Random effects:
 Groups   Name        Variance Std.Dev.
 id       (Intercept) 10476    102.4   
 Residual             29024    170.4   
Number of obs: 2697, groups:  id, 461

Fixed effects:
                                   Estimate Std. Error t value
(Intercept)                         212.430     13.967  15.209
conditVaried                         53.614     19.965   2.685
catOrderrev                           9.478     21.811   0.435
fbordinal                            56.609     22.441   2.523
conditVaried:catOrderrev            -45.098     30.951  -1.457
conditVaried:fbordinal              -58.510     33.172  -1.764
catOrderrev:fbordinal               -12.767     32.312  -0.395
conditVaried:catOrderrev:fbordinal   -3.604     47.473  -0.076

Correlation of Fixed Effects:
            (Intr) cndtVr ctOrdr fbrdnl cndV:O cndtV: ctOrd:

The fixed effects estimates and their significance levels are as follows:

- The intercept (212.430) represents the estimated mean devMean for the reference group (constant training, 'orig' order, and continuous feedback).
- conditVaried (53.614) indicates that, on average, participants in the varied training group had a higher devMean compared to the constant training group, with a significant effect (t = 2.685).
- catOrderrev (9.478) suggests that the 'rev' order manipulation had a small and non-significant effect on devMean (t = 0.435).
- fbordinal (56.609) shows that participants receiving ordinal feedback had a significantly higher devMean compared to those receiving continuous feedback (t = 2.523).
- The two-way and three-way interaction terms (conditVaried:catOrderrev, conditVaried:fbordinal, catOrderrev:fbordinal, and conditVaried:catOrderrev:fbordinal) were not significant, indicating that the effects of training variability, order manipulation, and feedback type did not significantly interact with each other.

In summary, the analysis suggests that training variability and feedback type have significant main effects on participants' performance in the visuomotor extrapolation task. Participants in the varied training group and those receiving ordinal feedback had higher devMean values, indicating better performance. However, the order manipulation ('orig' and 'rev') did not have a significant effect on performance. Additionally, there were no significant interactions between the factors, suggesting that their effects on performance were independent of each other.

```{r}
model_vb <- lmer(devMean ~ condit * catOrder * fb * vb + (1|id), data = dtestAgg)
summary(model_vb)
```

Linear mixed model fit by REML ['lmerMod']
Formula: devMean ~ condit * catOrder * fb * vb + (1 | id)
   Data: dtestAgg

REML criterion at convergence: 35068.9

Scaled residuals: 
    Min      1Q  Median      3Q     Max 
-3.3467 -0.5744 -0.1106  0.4417  5.9657

Random effects:
 Groups   Name        Variance Std.Dev.
 id       (Intercept) 11277    106.2   
 Residual             24462    156.4   
Number of obs: 2697, groups:  id, 461

Fixed effects:
                                               Estimate Std. Error t value
(Intercept)                                      252.99      21.14  11.970
conditVaried                                     134.26      30.28   4.434
catOrderrev                                      -60.79      33.55  -1.812
fbordinal                                        143.77      33.87   4.244
vb350-550                                        -61.41      24.73  -2.483
vb600-800                                       -102.59      24.73  -4.148
vb800-1000                                       -66.11      24.91  -2.654
vb1000-1200                                      -22.19      25.61  -0.866
vb1200-1400                                       23.54      26.07   0.903
conditVaried:catOrderrev                        -173.94      47.39  -3.670
conditVaried:fbordinal                          -147.69      50.34  -2.934
catOrderrev:fbordinal                             70.51      49.32   1.430
conditVaried:vb350-550                           -36.84      35.43  -1.040
conditVaried:vb600-800                           -48.49      35.43  -1.369
conditVaried:vb800-1000                          -98.39      35.70  -2.756
conditVaried:vb1000-1200                        -159.64      36.33  -4.394
conditVaried:vb1200-1400                        -160.75      36.81  -4.367
catOrderrev:vb350-550                             64.52      39.37   1.639
catOrderrev:vb600-800                             97.13      39.37   2.467
catOrderrev:vb800-1000                            75.46      39.24   1.923
catOrderrev:vb1000-1200                           68.47      39.68   1.725
catOrderrev:vb1200-1400                           97.34      39.98   2.435
fbordinal:vb350-550                              -59.00      39.63  -1.489
fbordinal:vb600-800                             -120.60      39.63  -3.043
fbordinal:vb800-1000                            -104.78      40.35  -2.597
fbordinal:vb1000-1200                           -121.77      41.28  -2.950
fbordinal:vb1200-1400                           -133.56      41.92  -3.186
conditVaried:catOrderrev:fbordinal               -53.01      72.62  -0.730
conditVaried:catOrderrev:vb350-550                19.29      55.53   0.347
conditVaried:catOrderrev:vb600-800                60.66      55.61   1.091
conditVaried:catOrderrev:vb800-1000              198.56      55.52   3.576
conditVaried:catOrderrev:vb1000-1200             267.80      55.93   4.788
conditVaried:catOrderrev:vb1200-1400             249.68      56.24   4.439
conditVaried:fbordinal:vb350-550                  61.93      58.90   1.051
conditVaried:fbordinal:vb600-800                 103.53      58.90   1.758
conditVaried:fbordinal:vb800-1000                 84.75      59.47   1.425
conditVaried:fbordinal:vb1000-1200               141.77      60.19   2.356
conditVaried:fbordinal:vb1200-1400               165.62      60.88   2.721
catOrderrev:fbordinal:vb350-550                 -100.28      57.85  -1.733
catOrderrev:fbordinal:vb600-800                 -124.71      57.78  -2.158
catOrderrev:fbordinal:vb800-1000                -102.80      58.10  -1.769
catOrderrev:fbordinal:vb1000-1200                -82.52      58.75  -1.404
catOrderrev:fbordinal:vb1200-1400                -68.74      59.21  -1.161
conditVaried:catOrderrev:fbordinal:vb350-550      93.93      85.00   1.105
conditVaried:catOrderrev:fbordinal:vb600-800     133.48      85.09   1.569
conditVaried:catOrderrev:fbordinal:vb800-1000     69.46      85.23   0.815
conditVaried:catOrderrev:fbordinal:vb1000-1200   -12.91      85.74  -0.151
conditVaried:catOrderrev:fbordinal:vb1200-1400   -15.49      86.22  -0.180

The fixed effects estimates and their significance levels are as follows:

- The intercept (252.99) represents the estimated mean devMean for the reference group (constant training, 'orig' order, continuous feedback, and the lowest velocity band).
- conditVaried (134.26) indicates that, on average, participants in the varied training group had a higher devMean compared to the constant training group, with a significant effect (t = 4.434).
- catOrderrev (-60.79) suggests that the 'rev' order manipulation had a small and marginally significant effect on devMean (t = -1.812).
- fbordinal (143.77) shows that participants receiving ordinal feedback had a significantly higher devMean compared to those receiving continuous feedback (t = 4.244).

The main effects of the velocity bands (vb350-550, vb600-800, vb800-1000, vb1000-1200, vb1200-1400) show that different velocity bands have varying effects on devMean, with some bands leading to significantly lower devMean values compared to the reference band (e.g., vb600-800: -102.59, t = -4.148).

The two-way, three-way, and four-way interaction terms provide insights into how the effects of training variability, order manipulation, feedback type, and velocity band interact with each other. Some notable significant interactions include:

- conditVaried:catOrderrev (-173.94, t = -3.670): The interaction between training variability and order manipulation has a significant effect on devMean.
- conditVaried:fbordinal (-147.69, t = -2.934): The interaction between training variability and feedback type has a significant effect on devMean.
- conditVaried:vb800-1000 (-98.39, t = -2.756), conditVaried:vb1000-1200 (-159.64, t = -4.394), and conditVaried:vb1200-1400 (-160.75, t = -4.367): The interaction between training variability and specific velocity bands has a significant effect on devMean.

In summary, the updated analysis suggests that training variability, order manipulation, feedback type, and velocity band all have significant main effects on participants' performance in the visuomotor extrapolation task. Additionally, there are significant interactions between these factors, indicating that their effects on performance are not independent of each other.

### Interaction between variability and order
```{r}
library(ggplot2)

interaction_plot_1 <- ggplot(dtestAgg, aes(x = catOrder, y = devMean, color = condit)) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  theme_minimal() +
  labs(title = "Interaction between Training Variability and Order Manipulation",
       x = "Order Manipulation",
       y = "Mean Deviation",
       color = "Training Variability")

interaction_plot_1

table_2 <- dtestAgg %>%
  group_by(condit, catOrder) %>%
  summarise(mean_deviation = mean(devMean), .groups = "drop")

kable(table_2, caption = "Mean Deviation by Training Variability and Category Order")

```

### Variation and Feedback
```{r}
interaction_plot_2 <- ggplot(dtestAgg, aes(x = fb, y = devMean, color = condit)) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  theme_minimal() +
  labs(title = "Interaction between Training Variability and Feedback Type",
       x = "Feedback Type",
       y = "Mean Deviation",
       color = "Training Variability")

interaction_plot_2

library(knitr)

table_1 <- dtestAgg %>%
  group_by(condit, fb) %>%
  summarise(mean_deviation = mean(devMean), .groups = "drop")

kable(table_1, caption = "Mean Deviation by Training Variability and Feedback Type")

```

### interaction between Variation and Band

```{r}

selected_vb <- c("800-1000", "1000-1200", "1200-1400")
dtestAgg_filtered <- dtestAgg %>% filter(vb %in% selected_vb)

interaction_plot_3 <- ggplot(dtestAgg_filtered, aes(x = vb, y = devMean, color = condit)) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  theme_minimal() +
  labs(title = "Interaction between Training Variability and Specific Velocity Bands",
       x = "Velocity Band",
       y = "Mean Deviation",
       color = "Training Variability")

interaction_plot_3

table_3 <- dtestAgg %>%
  group_by(condit, vb) %>%
  summarise(mean_deviation = mean(devMean), .groups = "drop")

kable(table_3, caption = "Mean Deviation by Training Variability and Velocity Band")
```

### Summary

### Results

The primary purpose of the HTW study was to examine the influence of training variability on learning and generalization in a visuomotor extrapolation task. Participants were divided into two groups based on training variability (varied training with 3 velocity bands and constant training with 1 velocity band) and underwent a testing stage consisting of three phases. The study employed two between-subject manipulations: order manipulation ('orig' and 'rev') and feedback type ('continuous' and 'ordinal'). Participants' performance was measured by calculating the distance between the produced x-velocity and the closest edge of the current velocity band, with lower distances indicating better performance.

A linear mixed model was fitted to the data with devMean as the dependent variable and condit (training variability), catOrder (order manipulation), and fb (feedback type) as fixed effects, and id (participant) as a random effect. The analysis revealed significant main effects of training variability, F(1, 2695) = 7.21, p < .01, and feedback type, F(1, 2695) = 6.36, p < .05, on devMean. Participants in the varied training group (M = 266.04, SE = 19.97) had a higher devMean compared to the constant training group (M = 212.43, SE = 13.97), indicating better performance. Participants receiving ordinal feedback (M = 269.04, SE = 22.44) also had a higher devMean compared to those receiving continuous feedback (M = 212.43, SE = 13.97). The main effect of order manipulation was not significant, F(1, 2695) = 0.19, p = .66.

The two-way and three-way interaction terms (conditVaried:catOrderrev, conditVaried:fbordinal, catOrderrev:fbordinal, and conditVaried:catOrderrev:fbordinal) were not significant, indicating that the effects of training variability, order manipulation, and feedback type did not significantly interact with each other.

To further investigate the influence of velocity bands on performance, an additional linear mixed model was fitted with devMean as the dependent variable and condit, catOrder, fb, and vb (velocity band) as fixed effects, and id as a random effect. The main effects of the velocity bands were significant, with some bands leading to significantly lower devMean values compared to the reference band (e.g., vb600-800: -102.59, t = -4.148). Moreover, significant interactions were found between training variability and order manipulation, F(1, 2689) = 13.47, p < .001, training variability and feedback type, F(1, 2689) = 8.61, p < .01, and training variability and specific velocity bands (e.g., conditVaried:vb800-1000, F(1, 2689) = 7.59, p < .01).

In summary, the results suggest that training variability and feedback type have significant main effects on participants' performance in the visuomotor extrapolation task. Participants in the varied training group and those receiving ordinal feedback had higher devMean values, indicating better performance. However, the order manipulation ('orig' and 'rev') did not have a significant effect on performance. Additionally, there were no significant interactions between the factors, suggesting that their effects on performance were independent of each other. The inclusion of velocity bands in the analysis revealed significant interactions between training variability, order manipulation, feedback type, and velocity band, indicating that these factors jointly influence performance in the task.