### 2023-07-05 Analysis/data_organize.qmd
started new analysis + re-render

---
title: Organization
---

```{r}
pacman::p_load(tidyverse,lme4,emmeans,here,knitr,kableExtra)
options(dplyr.summarise.inform=FALSE)
d <- readRDS(here("data/dPrune-05-21-23.rds")) %>% ungroup()
d$stage <- factor(d$stage, ordered=TRUE, levels = c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23"))
d$tOrder <- factor(d$tOrder, levels=c("testFirst","trainFirst"),labels=c("Test First","Train First"))
d$bandOrder <- factor(d$catOrder,levels=c("orig","rev"),labels=c("Original","Reverse"))
d$fb <- factor(d$fb,levels=c("Continuous","Ordinal"),labels=c("Continuous","Ordinal"))
d<- d %>% mutate(id=as.factor(d$id),fullCond=as.factor(fullCond)) %>% select(-catOrder,-mode) %>% relocate(bandOrder,.after=fb)
```

head(d)
colnames(d)
d %>% select_if(is.factor) %>% colnames()

```{r}
d %>% select_if(is.factor) %>% select(-sbjCode,-id) %>% map(levels)
```

```{r}
#| eval: false
e1 <- d |> filter(fb=="Continuous" & bandOrder=="Original")
e2 <- d |> filter(fb=="Continuous" & bandOrder=="Reverse")
e3 <- d |> filter(fb=="Ordinal")

saveRDS(d,here("data/dAll_07_04_23.rds"))
saveRDS(e1,here("data/e1_07_04_23.rds"))
saveRDS(e2,here("data/e2_07_04_23.rds"))
saveRDS(e3,here("data/e3_07_04_23.rds"))

```