### 2023-07-27 Analysis/ME_Slopes.qmd
test plots and gam code

---
title: Experiment 1 Testing
date: last-modified
categories: [Analysis, R]
page-layout: full
# fig-width: 15
# fig-height: 8
code-fold: true
code-tools: true
toc: false
execute: 
  warning: false
  eval: false
---

```{r setup}
pacman::p_load(tidyverse,tidybayes, brms, broom, broom.mixed, lme4,emmeans,here,knitr,kableExtra,gt,ggh4x)
e1 <- readRDS(here("data/e1_07_04_23.rds"))
source(here("Functions/Display_Functions.R"))

options(mc.cores = 4,  # Use 4 cores
        brms.backend = "cmdstanr")
bayes_seed <- 1234

test <- e1 |> filter(expMode %in% c("test-Nf","test-train-nf"))

nested_settings <- strip_nested(
  text_x = list(element_text(family = "serif", 
                             face = "plain"), NULL),
  background_x = list(element_rect(fill = "grey92"), NULL),
  by_layer_x = TRUE)

testAvg <- test %>% group_by(id, condit, vb, bandInt) %>%
  summarise(vx=mean(vx),dist=mean(dist))

```

```{r}
model_super_boring <- e1 |> filter(expMode %in% c("test-Nf","test-train-nf")) |>  with(lm(vx ~ vb))
tidy(model_super_boring)

```

```{r}
model_boring <- brm(bf(vx ~ vb),
                    data=filter(e1,expMode %in% c("test-Nf","test-train-nf")),
                    chains=2,seed=bayes_seed)

tidy(model_boring)
```

```{r}
model_fixed <- brm(
  bf(vx ~ vb + (1 | id)),
  data = filter(e1,expMode %in% c("test-Nf","test-train-nf")),
  control = list(adapt_delta = 0.95),
  chains = 4, seed = bayes_seed
)
tidy(model_fixed)
summary(model_fixed)

tidy(model_fixed,effects="fixed")
tidy(model_fixed,effects="ran_pars")

indvEst1 <- ranef(model_fixed)$id %>% as_tibble(rownames="id")
indvEst1

coef(model_fixed)$id %>% as_tibble(rownames="id") %>% select(id,starts_with("Estimate"))

model_fixed %>% 
  emmeans(~ id + vb,
          at = list(vb = 0),  # Look at predicted values for 1952
          epred = TRUE,  # Use expected predictions from the posterior
          re_formula = NULL)
```
```{r}
model_cont <- brm(
  bf(vx ~ bandInt + (1 | id)),
  data = filter(e1,expMode %in% c("test-Nf","test-train-nf")),
  control = list(adapt_delta = 0.95),
  chains = 4, seed = bayes_seed
)
tidy(model_cont)
summary(model_cont)

tidy(model_cont,effects="fixed")
tidy(model_cont,effects="ran_pars")

indvEst1 <- ranef(model_cont)$id %>% as_tibble(rownames="id")
indvEst1

coef(model_cont)$id %>% as_tibble(rownames="id") %>% select(id,starts_with("Estimate"))

model_cont %>% 
  emmeans(~ id + bandInt,
          at = list(bandInt = 100),  # Look at predicted values for 1952
          epred = TRUE,  # Use expected predictions from the posterior
          re_formula = NULL)

```

```{r fig.width=14, fig.height=16}

pmcy = test %>% select(id,bandInt) %>% mutate(id=factor(id,levels=unique(id)))
pmcy = expand_grid(id=unique(test$id),bandInt=unique(test$bandInt)) %>% mutate(id=factor(id,levels=unique(id)))

preds1 <- model_cont %>% epred_draws(pmcy,re_formula = NULL)

nd <- preds1 %>% left_join(test,by=c("id","bandInt"))

preds1 <- model_cont %>% epred_draws(pmcy,re_formula = NULL,ndraws=1) 
nd <- preds1 %>% left_join(test,by=c("id","bandInt"))
nd <- test %>% mutate(pred=preds1$.epred) %>% relocate(bandInt,pred,vx,.after=nGoodTrial)

ggplot(nd[1:5000,],aes(x=bandInt,y=pred))+geom_point(aes(y=vx))+
  stat_lineribbon(alpha=.5)+scale_fill_brewer(palette="Reds") +
  labs(title = "Intercepts and slopes for year trend vary by country",
       subtitle = "lifeExp ~ year + (1 + year | country)",
       x = NULL, y = "Predicted life expectancy") +
  guides(fill = "none") +
  facet_nested_wrap(vars(condit, id), nrow = 4, strip = nested_settings) +
  theme(legend.position = "bottom",
        plot.subtitle = element_text(family = "serif"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

```

```{r}

model_slope <- brm(
  bf(vx ~ bandInt + (1 + bandInt | id)),
  data = filter(e1,expMode %in% c("test-Nf","test-train-nf")),
  control = list(adapt_delta = 0.95),
  chains = 4, seed = bayes_seed
)
tidy(model_slope)
summary(model_slope)

tidy(model_slope,effects="fixed")
tidy(model_slope,effects="ran_pars")

indvEst2 <- ranef(model_slope)$id %>% as_tibble(rownames="id")
indvEst2

ms_coef <- coef(model_slope)$id %>% as_tibble(rownames="id") %>% select(id,starts_with("Estimate"))

model_slope %>% 
  emmeans(~ id + bandInt,
          at = list(bandInt = 100),  # Look at predicted values for 1952
          epred = TRUE,  # Use expected predictions from the posterior
          re_formula = NULL)

testAvg <- testAvg %>% left_join(ms_coef,by="id")

testAvg %>% ggplot(aes(x=condit,y=Estimate.bandInt)) + 
  stat_summary(geom="bar",fun=mean)+stat_summary(geom="errorbar",fun.data=mean_se,width=.5)

testAvg %>% ggplot(aes(x=Estimate.bandInt,y=dist,col=condit))+geom_point()+geom_line(alpha=.5)+
  facet_wrap(~vb)

testAvg %>% ggplot(aes(x=Estimate.Intercept.y,y=dist,col=condit))+geom_point()+geom_line(alpha=.5)+
  facet_wrap(~vb)
testAvg %>% ggplot(aes(x=Estimate.bandInt,y=vx,col=condit))+geom_point()+geom_line(alpha=.5)+
  facet_wrap(~vb)

```
```{r}
# center model on bandInt 800, and fit slopes again with brms, intercept and slope for each id

# center around bandInt 800
e1$bandInt2 <- e1$bandInt - 800

model_slope2 <- brm(
  bf(vx ~ bandInt2 + (1 + bandInt2 | id)),
  data = filter(e1,expMode %in% c("test-Nf","test-train-nf")),
  control = list(adapt_delta = 0.95),
  chains = 4, seed = bayes_seed
)

tidy(model_slope2,effects="fixed")
ms_coef2 <- coef(model_slope2)$id %>% as_tibble(rownames="id") %>% select(id,starts_with("Estimate"))

testAvg <- testAvg %>% left_join(ms_coef2,by="id")

testAvg %>% ggplot(aes(x=condit,y=Estimate.bandInt2)) + 
  stat_summary(geom="bar",fun=mean)+stat_summary(geom="errorbar",fun.data=mean_se,width=.5)

testAvg %>% ggplot(aes(x=Estimate.bandInt2,y=dist,col=condit))+geom_point()+geom_line(alpha=.5)+
  facet_wrap(~vb)

testAvg %>% ggplot(aes(x=Estimate.Intercept.y,y=dist,col=condit))+geom_point()+geom_line(alpha=.5)+
  facet_wrap(~vb)

testAvg %>% ggplot(aes(x=Estimate.bandInt2,y=vx,col=condit))+geom_point()+geom_line(alpha=.5)+
  facet_wrap(~vb)
testAvg %>% ggplot(aes(x=Estimate.Intercept.y,y=vx,col=condit))+geom_point()+geom_line(alpha=.5)+
  facet_wrap(~vb)

```

```{r}
head(testAvg)

model_slope3 <- brm(
  bf(vx ~ bandInt + (1 + bandInt | id)),
  data = testAvg,
  control = list(adapt_delta = 0.95),
  chains = 4, seed = bayes_seed
)
tidy(model_slope3)
ms_coef3 <- coef(model_slope3)$id %>% as_tibble(rownames="id") %>% select(id,starts_with("Estimate"))

left_join(ms_coef,ms_coef3,by="id")

```

```{r}

```

```{r}

```