### 2023-07-27 Analysis/gam_discrim.qmd

---
title: GAM
date: last-modified
categories: [Analysis, R]
page-layout: full
# fig-width: 15
# fig-height: 8
code-fold: true
code-tools: true
toc: false
execute: 
  warning: false
  eval: false
---

```{r}
pacman::p_load(tidyverse,lme4,emmeans,here,knitr,kableExtra,gt,mgcv,data.table)
e1 <- readRDS(here("data/e1_07_04_23.rds"))
source(here("Functions/Display_Functions.R"))

test <- e1 |> filter(expMode %in% c("test-Nf","test-train-nf"))
testAvg <- test %>% group_by(id, condit, vb, bandInt) %>%
  summarise(vx=mean(vx),dist=mean(dist)) %>% data.table()

```

```{r}

testAvg %>% ggplot(aes(x=bandInt,y=vx))+geom_point(color="grey")+geom_line()+theme_bw()

tm <- gam(vx ~s(bandInt,k=3),sp=0,data=testAvg)
summary(tm)

results <- tibble(k = integer(), r_sq = numeric(), edf = numeric(), p_value = numeric(), scale = numeric(), deviance = numeric(), null_deviance = numeric())

# Iterate through different k values
for (k in 1:6) {
  temp_result <- tryCatch({
    # Fit the GAM model
    tm <- gam(vx ~ s(bandInt, k = k), sp = 0, data = testAvg)
    model_sum <- summary(tm)

    # Extract additional metrics from the model summary and the model itself
    edf <- model_sum$s.table[1]
    p_value <- model_sum$s.pv[1]
    scale <- tm$scale
    deviance <- tm$deviance
    null_deviance <- tm$null.deviance

    # Return data frame with results
    tibble(k = k, r_sq = model_sum$r.sq, edf = edf, p_value = p_value, scale = scale, deviance = deviance, null_deviance = null_deviance)
  }, error = function(e) {
    # Print error message
    cat(paste("Skipping k =", k, "due to error:", conditionMessage(e)), "\n")

    # Return data frame with NA values
    tibble(k = k, r_sq = NA_real_, edf = NA_real_, p_value = NA_real_, scale = NA_real_, deviance = NA_real_, null_deviance = NA_real_)
  })

  # Add results to data frame
  results <- bind_rows(results, temp_result)
}

print(results)

# Plot R-squared
results %>%
  ggplot(aes(x = k, y = r_sq)) +
  geom_line(color = "blue") +
  labs(x = "k", y = "R-squared") +
  theme_minimal()

# Plot deviance
results %>%
  ggplot(aes(x = k, y = deviance)) +
  geom_line(color = "red") +
  labs(x = "k", y = "Deviance") +
  theme_minimal()

```

```{r}

# Iterate through different k values
for (k in 1:6) {
  for (group in unique(testAvg$condit)) {

    temp_result <- tryCatch({
      # Filter data by group
      group_data <- testAvg %>% filter(condit == group)

      # Fit the GAM model with interaction and more flexible splines
      tm <- gam(vx ~ s(bandInt, k = k, bs = "cr"), data = group_data)
      model_sum <- summary(tm)

      # Extract additional metrics from the model summary and the model itself
      edf <- model_sum$s.table[1]
      p_value <- model_sum$s.pv[1]
      scale <- tm$scale
      deviance <- tm$deviance
      null_deviance <- tm$null.deviance
      threshold <- NA_real_

      # Attempt to calculate discrimination threshold
      tryCatch({
        # Value of bandInt where the derivative of the fitted spline is highest
        bandInt_values <- seq(min(group_data$bandInt), max(group_data$bandInt), length.out = 1000)
        fitted_spline <- predict(tm, newdata = data.frame(bandInt = bandInt_values, condit_num = ifelse(group == "Varied", 1, 0)), type = "terms")
        fitted_derivative <- diff(fitted_spline[,1]) / diff(bandInt_values)
        threshold <- bandInt_values[which.max(fitted_derivative)]
      }, error = function(e) {
        cat(paste("Failed to calculate threshold for k =", k, "in group", group, "due to error:", conditionMessage(e)), "\n")
      })

      # Return data frame with results
      tibble(group = group, k = k, r_sq = model_sum$r.sq, edf = edf, p_value = p_value, scale = scale, deviance = deviance, null_deviance = null_deviance, threshold = threshold)
    }, error = function(e) {
      # Print error message
      cat(paste("Skipping k =", k, "for group", group, "due to error:", conditionMessage(e)), "\n")

      # Return data frame with NA values
      tibble(group = group, k = k, r_sq = NA_real_, edf = NA_real_, p_value = NA_real_, scale = NA_real_, deviance = NA_real_, null_deviance = NA_real_, threshold = NA_real_)
    })

    # Add results to data frame
    results <- bind_rows(results, temp_result)
  }
}
print(results)

```

```{r}

est_gam <- function(k, train_data, test_data, group) 
{
  # Filter by group
  train_data_group <- train_data %>% filter(condit == group)
  test_data_group <- test_data %>% filter(condit == group)

  # Train a model  
  trained_model <- gam(vx ~ s(bandInt, k = k, bs = "tp"), data = train_data_group)

  # Print out the number of rows in train_data_group
  print(paste("Number of rows in train_data_group: ", nrow(train_data_group)))

  # Predict y for the train datasets
  predicted_values_train <- predict(trained_model, newdata = train_data_group)
  print(paste("Number of predicted values for train data: ", length(predicted_values_train)))

  # Add the predicted values as a new column
  train_data_group <- train_data_group %>% mutate(vx_hat = predicted_values_train, type = "Train")

  # Repeat the same for the test datasets
  predicted_values_test <- predict(trained_model, newdata = test_data_group)
  print(paste("Number of predicted values for test data: ", length(predicted_values_test)))

  test_data_group <- test_data_group %>% mutate(vx_hat = predicted_values_test, type = "Test")

  # Combine before returning
  return_data <- bind_rows(train_data_group, test_data_group) %>% 
                 mutate(num_knots = k, group = group)

  return(return_data)
}

# define k-values
k_values <- 1:5

# Split the data into training and testing datasets
set.seed(123)
data_split <- testAvg %>% 
  resample_partition(c(Train = 0.75, Test = 0.25))

# Extract the data frames from the resample objects
train_data <- as.data.frame(data_split$Train)
test_data <- as.data.frame(data_split$Test)

# Use a nested loop to apply the est_gam function over all combinations of groups and k_values
results <- map_df(unique(testAvg$condit), ~{
  group <- .
  map_df(k_values, ~{
    est_gam(.x, train_data %>% filter(condit == group), 
            test_data %>% filter(condit == group), group)
  })
})

```

```{r}

# Create an empty data frame to store results
results <- tibble(group = character(), k = integer(), r_sq = numeric(), edf = numeric(), p_value = numeric(), scale = numeric(), deviance = numeric(), null_deviance = numeric())

# Iterate through different k values
for (k in 1:6) {
  # Iterate through different conditions
  for (condit in unique(testAvg$condit)) {
    temp_result <- tryCatch({
      # Subset the data for the current condition
      data_subset <- testAvg[testAvg$condit == condit, ]

      # Fit the GAM model
      tm <- gam(vx ~ s(bandInt, k = k), sp = 0, data = data_subset)
      model_sum <- summary(tm)

      # Extract additional metrics from the model summary and the model itself
      edf <- model_sum$s.table[1]
      p_value <- model_sum$s.pv[1]
      scale <- tm$scale
      deviance <- tm$deviance
      null_deviance <- tm$null.deviance

      # Return data frame with results
      tibble(group = condit, k = k, r_sq = model_sum$r.sq, edf = edf, p_value = p_value, scale = scale, deviance = deviance, null_deviance = null_deviance)
    }, error = function(e) {
      # Print error message
      cat(paste("Skipping k =", k, "in group", condit, "due to error:", conditionMessage(e)), "\n")

      # Return data frame with NA values
      tibble(group = condit, k = k, r_sq = NA_real_, edf = NA_real_, p_value = NA_real_, scale = NA_real_, deviance = NA_real_, null_deviance = NA_real_)
    })

    # Add results to data frame
    results <- bind_rows(results, temp_result)
  }
}

print(results)

# Plot R-squared
results %>%
  ggplot(aes(x = k, y = r_sq, color = group)) +
  geom_line() +
  labs(x = "k", y = "R-squared") +
  theme_minimal()

# Plot deviance
results %>%
  ggplot(aes(x = k, y = deviance, color = group)) +
  geom_line() +
  labs(x = "k", y = "Deviance") +
  theme_minimal()

# Add additional plots as necessary

```

```{r}

# Create an empty data frame to store results
results <- data.frame(k = numeric(), r_sq = numeric(), deviance_explained = numeric())

# Iterate through different k values
for (k in 1:6) {
  temp_result <- tryCatch({
    # Fit the GAM model
    tm <- gam(vx ~ s(bandInt, k = k), sp = 0, data = testAvg)
    model_sum <- summary(tm)

    # Return data frame with results
    data.frame(k = k, r_sq = model_sum$r.sq.adj, deviance_explained = model_sum$dev.expl)
  }, error = function(e) {
    # Print error message
    cat(paste("Skipping k =", k, "due to error:", conditionMessage(e)), "\n")

    # Return data frame with NA values
    data.frame(k = k, r_sq = NA, deviance_explained = NA)
  })

  # Add results to data frame
  results <- rbind(results, temp_result)
}

# Check if we have successfully fitted models and results have more than zero rows 
if(nrow(results) > 0){
  # No need to set column names here, you've already set them when creating the data frames inside the loop

  # Plotting
  plot(results$k, results$r_sq, type = "l", xlab = "k", ylab = "R-squared", col = "blue")
  lines(results$k, results$deviance_explained, type = "l", col = "red") 
  legend("topright", legend = c("R-squared", "Deviance Explained"), col = c("blue", "red"), lty = 1)
}

```

```{r}

gam_fit <- gam(vx ~ s(bandInt, k = 6, bs = "cr"), data = testAvg)
gam_fit$coefficient

basis_data <- gam_fit %>%  predict(., type = "lpmatrix") %>%
  data.table() %>% 
  .[, bandInt := testAvg[, bandInt]] %>% 
  melt(id.var = "bandInt")

data.table(
  variable = unique(basis_data$variable)[-1],
  coef = gam_fit$coefficient[-1]
) %>% 
.[basis_data, on = "variable"] %>% 
.[, .(y_no_int = sum(coef * value)), by = bandInt] %>% 
.[, y_hat := gam_fit$coefficient[1] + vx] %>% 
ggplot(data = .) +
  geom_line(aes(y = vx, x = bandInt, color = "gam-fitted")) +
  geom_line(data = data, aes(y = vx, x = bandInt, color = "True")) +
  scale_color_manual(
    name = "",
    values = c("gam-fitted" = "red", "True" = "blue")
  ) +
  ylab("y") +
  xlab("x") +
  theme_bw()
```

```{r}

```

```{r}

```