### 2023-08-05 Analysis/clust_gmm.qmd
extremely minor organization

https://mhahsler.github.io/Introduction_to_Data_Mining_R_Examples/book/clustering-analysis.html#gaussian-mixture-models

https://cran.r-project.org/web/packages/mclust/vignettes/mclust.html

https://joshuamrosenberg.com/posts/lpa-in-r-using-mclust/

```{r}
pacman::p_load(tidyverse,here, mclust,furrr,future,broom)
select <- dplyr::select
mutate <- dplyr::mutate
filter <- dplyr::filter
map <- purrr::map

e1 <- readRDS(here("data/e1_08-04-23.rds"))

test <- e1 |> filter(expMode2 == "Test")
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vxMean=mean(vx),vxMed=median(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

get_mclust_param <- function(model, param) {
  if (!is.null(model$parameters[[param]])) {
    return(model$parameters[[param]])
  } else {
    return(NA_real_)
  }
}

plan(multisession, workers = 8)

mcFree <- test |> select(id,condit,bandInt,vb,vx) %>% ungroup()

mcFree2 <- get_mcFree(mcFree, "id", "condit", c("bandInt", "vb", "vx", "vy"))

mcFit <- function(df, id_var, condit_var, clustvars,G=1:9) {
  df %>%
    select({{ id_var }}, {{ condit_var }}, !!!clustvars) %>%
    ungroup() %>%
    group_by({{ id_var }}, {{ condit_var }}) %>%
    nest(data = clustvars) %>%
    mutate(mc = future_map(data, ~Mclust(.x[c(clustvars)],G=G))) %>%
    mutate(
      bic= as.numeric(map(mc,"bic")),
      Gfit = as.numeric(map(mc, "G")),
      means = map(mc, get_mclust_param, param = "mean"),
      proportions = map(mc, get_mclust_param, param = "pro")
    )
}

mcFree1 <- mcFit(test, id, condit, c("vx"))
mcFreeVxb <- mcFit(test, id, condit, c("vxb"))
mcFree2 <- mcFit(test, id, condit, c("vx","vy"))

mcFreeVb <- mcFit(test, id, condit, c("vx","bandInt"))

df <- test |> select(id,condit,bandInt,vx) |> group_by(id,bandInt) |> mutate(m=mean(vx),sd=sd(vx),n=n())

# test1 <- dfs |> filter(id==1 || id==2) |> mutate(samps = map2(m,sd, ~rnorm(n=10,mean=.x,sd=.y)))
# test1 <- test1 %>% unnest(samps)

dfs <- test |> select(id,condit,bandInt,vx,vb) |> group_by(id,vb,condit) |> summarise(m=mean(vx),sd=sd(vx),med=median(vx),n=n(),se=sd/sqrt(n))

bdf <- dfs |> mutate(samps = map2(m,se, ~rnorm(n=100,mean=.x,sd=.y))) |> unnest(samps) %>% ungroup()
mcBoot2 <- mcFit(bdf, id, condit, c("samps"))

mcBoot2 |> group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

```

Sample Stat bootstrap
```{r}

dfs <- test |> select(id,condit,bandInt,vx,vb) |> group_by(id,vb,condit) |> summarise(m=mean(vx),sd=sd(vx),med=median(vx),n=n(),se=sd/sqrt(n))

bdf <- dfs |> mutate(samps = map2(m,se, ~rnorm(n=100,mean=.x,sd=.y))) |> unnest(samps) %>% ungroup()
mcBoot2 <- mcFit(bdf, id, condit, c("samps"))

mcBoot2 |> group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

dfs2 <- test |> select(id,condit,bandInt,vx,vb,vxb) |> group_by(id,vb,condit) |> summarise(m=mean(vxb),sd=sd(vxb),med=median(vxb),n=n(),se=sd/sqrt(n))

bdf2 <- dfs2 |> mutate(samps = map2(m,se, ~rnorm(n=100,mean=.x,sd=.y))) |> unnest(samps) %>% ungroup()
mcBoot2b <- mcFit(bdf2, id, condit, c("samps"))

mcBoot2b |> group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

```

```{r}
#| fig.width=12
#| 
ta <- test %>% group_by(id,condit, vb) %>% summarise(mean(dist),mean(vx),mad(vx),sd(vx),IQR(vx),moments::skewness(vx))

plots <- map(names(ta)[-(1:3)], ~ {
  var_name <- .x
  ggplot(ta, aes(x = vb, y = !!sym(var_name), color = condit)) +
    gghalves::geom_half_violin()+
    gghalves::geom_half_boxplot(alpha=.5)+
    labs(x = "Velocity Band", y = var_name,
         title = paste("Plot of", var_name, "by Velocity Band and Condition"),
         color = "Condition") +
    theme_minimal()
})

# Combine the plots using patchwork
combined_plot <- wrap_plots(plots, ncol = 2)

# Print the combined plot
print(combined_plot)

dispersion_vars <- names(ta)[-(1:5)]
correlation_results <- map(dispersion_vars, ~ {
  var_name <- .x
  cor_result <- cor.test(ta[['mean(dist)']], ta[[var_name]], method = "pearson")
  list(var_name = var_name, cor_result = cor_result)
})

# Print the correlation results
correlation_results

dispersion_vars <- names(ta)[-(1:5)]
dispersion_vars <- c('mean(dist)', dispersion_vars)
ta_subset <- ta[dispersion_vars]
correlation_matrix <- cor(ta_subset, method = "pearson")
correlation_matrix

library(corrplot)

correlation_matrices <- ta %>% 
  group_by(condit) %>% 
  nest() %>% 
  mutate(correlation_matrix = map(data, ~ cor(.x[dispersion_vars], method = "pearson")))

# Create a correlation plot for each condition
correlation_plots <- correlation_matrices %>% 
  mutate(plot = map(correlation_matrix, ~ corrplot(.x, method = "color")))

# Print the correlation plots
correlation_plots$plot

GGally::ggcorr(ta[-2], aes(colour = condit), method = c("everything", "pearson"))

```

Empirical Bootstrap

```{r}

# Define the bootstrap function
bootstrap <- function(x, n) {
  sample(x, size = n, replace = TRUE)
}

# Apply the bootstrap function to your data
bdf2 <- test %>%
  select(id, condit, bandInt, vx, vb, vxb) %>%
  group_by(id, vb, condit) %>%
  summarise(vxb_bootstrap = list(bootstrap(vx, 20)), .groups = "drop") %>%
  unnest(vxb_bootstrap)

eb1 <- mcFit(bdf2, id, condit, c("vxb_bootstrap"))

eb1 |> group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

eb2<- test %>%
  select(id, condit, bandInt, vx, vb, vxb) %>%
  group_by(id, vb, condit) %>%
  summarise(vxb_bootstrap = list(bootstrap(vx, 100)), .groups = "drop") %>%
  unnest(vxb_bootstrap) %>% mcFit(.,id,condit,c("vxb_bootstrap"))

eb2 |> group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

```

```{r}

```

Population Clustering
```{r}

bic <- mclustBIC(test$vx)
plot(bic)
mod1 <- Mclust(test$vx,x=bic)
summary(mod1,parameters=TRUE)

bic <- mclustBIC(c(test$vx,test$vy))
plot(bic)

bic <- mclustBIC(c(test$sdist))
plot(bic)

bic <- mclustBIC(c(test$dist))
plot(bic)

bic <- mclustBIC(c(test$dist,test$vx))
plot(bic)

bic <- mclustBIC(c(test$vx,test$bandInt))
plot(bic)
summary(bic)
```

```{r}
mv4 <- mclustBIC(subset(test,select = c(vx,bandInt)),
                 G=1:7,
                 modelNames=c("EII", "VVI", "EEE", "VVV"))
 y <- mv4%>%
        as.data.frame.matrix() %>%
        rownames_to_column("n_mixtures") %>%
        rename(`Constrained variance, fixed covariance` = EII, 
               `Freed variance, fixed covariance` = VVI,
               `Constrained variance, constrained covariance` = EEE,
               `Freed variance, freed covariance` = VVV)

to_plot <- y %>%  gather(`Covariance matrix structure`, val, -n_mixtures) %>% 
    mutate(`Covariance matrix structure` = as.factor(`Covariance matrix structure`),
           val = abs(val)) # this is to make the BIC values positive (to align with more common formula / interpretation of BIC)
ggplot(to_plot, aes(x = n_mixtures, y = val, color = `Covariance matrix structure`, group = `Covariance matrix structure`)) +
    geom_line() +
    geom_point() +
    ylab("BIC (smaller value is better)")

```

```{r}
m = Mclust(test$vx)
test$vx

summary(m)

mv <- Mclust(subset(test, condit == "Varied", select = vx))
summary(mv)
plot(mv, what = "classification")

mc <- Mclust(subset(test, condit == "Constant", select = vx))
summary(mc)
plot(mc, what = "classification")

mv2 <- densityMclust(subset(test, condit == "Varied", select = c(vxb,vy)))
summary(mv2)
plot(mv2, what = "density",type="persp")
plot(mv2, what = "BIC")

mv3 <- mclustBIC(subset(test, condit == "Varied", select = vx),
                 G=1:7)
 y <- mv3%>%
        as.data.frame.matrix() %>%
        rownames_to_column("n_mixtures") %>%
        rename(`Constrained variance, constrained covariance` = E,
               `Freed variance, freed covariance` = V)

plot(mc, what = "classification")

mc <- Mclust(subset(test, condit == "Constant", select = vx),G=6)
summary(mc)
tidy(mc)
glance(mc)
```

Vx and bandInt
```{r}

mcFreeVb <- mcFit(test, id, condit, c("vx","bandInt"),G=1:6)

mcFreeVb %>% group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

mcFreeVb2 <- mcFit(test, id, condit, c("vxb","bandInt"),G=1:6)

mcFreeVb2 %>% group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

mcFreeVb3 <- mcFit(test, id, condit, c("vxb","bandInt","vy"),G=1:6)

mcFreeVb3 %>% group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

```

```{r}

mcFreeVb1 <- mcFit(test, id, condit, c("vx"),G=1:6)

mcFreeVb1 %>% group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

mcFreeVb1b <- mcFit(test, id, condit, c("vxb"),G=1:6)

mcFreeVb1b %>% group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

boot1 <- MclustBootstrap(mcFreeVb1b[[4]][[1]], nboot = 1000, type = "bs")
summary(boot1)

```

```{r}

mcFree <- test |> select(id,condit,bandInt,vb,vx) %>% ungroup()

mcFree <- mcFree %>%
  group_by(id,condit) %>%
  nest(data = c(bandInt, vb, vx)) %>%
  mutate(mc = future_map(data, ~MclustBootstrap(.x$vx)))

mcFree <- mcFree %>% 
  mutate(
    Gfit = as.numeric(map(mc, "G")),
    means = map(mc, get_mclust_param, param = "mean"),
    proportions = map(mc, get_mclust_param, param = "pro")
  )

```

Fit free to each id
```{r}

mcFree <- test |> select(id,condit,bandInt,vb,vx) %>% ungroup()

mcFree <- mcFree %>%
  group_by(id,condit) %>%
  nest(data = c(bandInt, vb, vx)) %>%
  mutate(mc = future_map(data, ~Mclust(.x$vx)))

mcFree <- mcFree %>% 
  mutate(
    Gfit = as.numeric(map(mc, "G")),
    means = map(mc, get_mclust_param, param = "mean"),
    proportions = map(mc, get_mclust_param, param = "pro")
  )

ggplot(mcFree,aes(x=condit,y=Gfit))+geom_boxplot()+geom_jitter()

ggplot(mcFree,aes(x=Gfit,fill=condit))+geom_bar(position=position_dodge())

# mcFree <- test |> select(id,condit,bandInt,vb,vx) %>% ungroup()
# mcFree <- mcFree |> group_by(id) |> nest(id) |> mutate(mc=map(vx, ~Mclust(.x$vx))) 
#

plot(mcCondit6[[2]][[1]], what = "classification")
plot(mcCondit6[[2]][[2]], what = "classification")

k=summary(mcCondit6[[2]][[2]],parameters=TRUE)

```

Fit xy free to each id
```{r}

mcFree2 <- test |> select(id,condit,bandInt,vb,vx,vy) %>% ungroup()

mcFree2 <- mcFree2 %>%
  group_by(id,condit) %>%
  nest(data = c(bandInt, vb, vx,vy)) %>%
  mutate(mc = future_map(data, ~Mclust(c(.x$vx,.x$vy))))

mcFree2 <- mcFree2 %>% 
  mutate(
    Gfit = as.numeric(map(mc, "G")),
    means = map(mc, get_mclust_param, param = "mean"),
    proportions = map(mc, get_mclust_param, param = "pro")
  )

ggplot(mcFree2,aes(x=condit,y=Gfit))+geom_boxplot()+geom_jitter()

ggplot(mcFree2,aes(x=Gfit,fill=condit))+geom_bar(position=position_dodge())

# mcFree <- test |> select(id,condit,bandInt,vb,vx) %>% ungroup()
# mcFree <- mcFree |> group_by(id) |> nest(id) |> mutate(mc=map(vx, ~Mclust(.x$vx))) 
#

plot(mcCondit6[[2]][[1]], what = "classification")
plot(mcCondit6[[2]][[2]], what = "classification")

k=summary(mcCondit6[[2]][[2]],parameters=TRUE)

```

Fit xy free to each id
```{r}

mcFree2b <- test |> select(id,condit,bandInt,vb,vx) %>% ungroup()

mcFree2b <- mcFree2b %>%
  group_by(id,condit) %>%
  nest(data = c(bandInt, vb, vx)) %>%
  mutate(mc = future_map(data, ~Mclust(c(.x$vx,.x$bandInt))))

mcFree2b <- mcFree2b %>% 
  mutate(
    Gfit = as.numeric(map(mc, "G")),
    means = map(mc, get_mclust_param, param = "mean"),
    proportions = map(mc, get_mclust_param, param = "pro")
  )

ggplot(mcFree2b,aes(x=condit,y=Gfit))+geom_boxplot()+geom_jitter()
ggplot(mcFree2b,aes(x=Gfit,fill=condit))+geom_bar(position=position_dodge())

mcFree2b %>% group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

```

```{r}

mcFree2b <- test |> select(id,condit,bandInt,vb,vx,vy) %>% ungroup()

mcFree2b <- mcFree2b %>%
  group_by(id,condit) %>%
  nest(data = c(bandInt, vb, vx,vy)) %>%
  mutate(mc = future_map(data, ~Mclust(c(.x$vx,.x$bandInt,.x$vy))))

mcFree2b <- mcFree2b %>% 
  mutate(
    Gfit = as.numeric(map(mc, "G")),
    means = map(mc, get_mclust_param, param = "mean"),
    proportions = map(mc, get_mclust_param, param = "pro")
  )

ggplot(mcFree2b,aes(x=condit,y=Gfit))+geom_boxplot()+geom_jitter()
ggplot(mcFree2b,aes(x=Gfit,fill=condit))+geom_bar(position=position_dodge())

mcFree2b %>% group_by(condit) |> mutate(n=n()) |> 
   group_by(condit,Gfit) |> mutate(cg=n(),pg=cg/n) |>
  ggplot(aes(x=Gfit,fill=condit))+geom_col(aes(y=pg),position=position_dodge())+
  scale_x_continuous(breaks=seq(1,9))

```

```{r}

mcConditFree <- test |> select(condit,bandInt,vb,vx) %>% ungroup()

mcConditFree <- mcConditFree %>%
  group_by(condit) %>%
  nest(data = c(bandInt, vb, vx)) %>%
  reframe(mc = future_map(data, ~Mclust(.x$vx)))

mcConditFree <- mcConditFree %>% 
  mutate(
    Gfit = as.numeric(map(mc, "G")),
    means = map(mc, get_mclust_param, param = "mean"),
    proportions = map(mc, get_mclust_param, param = "pro")
  )

mcCondit6 <- test |> select(condit,bandInt,vb,vx) %>% ungroup()

mcCondit6 <- mcCondit6 %>%
  group_by(condit) %>%
  nest(data = c(bandInt, vb, vx)) %>%
  reframe(mc = future_map(data, ~Mclust(.x$vx,G=6)))

mcCondit6 <- mcCondit6 %>% 
  mutate(
    Gfit = as.numeric(map(mc, "G")),
    bic=as.numeric(map(mc,"bic")),
    means = map(mc, get_mclust_param, param = "mean"),
    proportions = map(mc, get_mclust_param, param = "pro")
  )

table(class,mcCondit6[[2]][[2]]$classification)

bic <- mclustBIC(test$vx)

```

```{r}
df <- test %>%
    select(id, condit, bandInt, vb, vx) %>%
    group_by(id) %>%
    nest(data = c(condit, bandInt, vb, vx))

# Define a function to apply the Mclust operation to one group
process_group <- function(group_data, group_id) {
    print(paste("Processing group", group_id))
    if(length(unique(group_data$vx))>=12) {
        Mclust(group_data$vx, G=6)
    } else {
        Mclust(rep(0, 12), G=6)  # dummy Mclust model
    }
}

# Apply the function to each group one by one
df$mc <- map2(df$data, df$id, process_group)

df <- df %>%
  mutate(
    means = map(mc, get_mclust_param, param = "mean"),
    proportions = map(mc, get_mclust_param, param = "pro")
  )

```

```{r}

centers <- tibble(
  cluster = factor(1:3),
  # number points in each cluster
  num_points = c(100, 150, 50),
  # x1 coordinate of cluster center
  x1 = c(5, 0, -3),
  # x2 coordinate of cluster center
  x2 = c(-1, 1, -2)
)

points <- centers %>%
  mutate(
    x1 = map2(num_points, x1, rnorm),
    x2 = map2(num_points, x2, rnorm)
  ) %>%
  select(-num_points, -cluster) %>%
  unnest(c(x1, x2))

m <- Mclust(points)

df <- df %>%
  mutate(
    mclust_success = map_lgl(mc, ~ !is.null(.x$parameters))
  ) %>%
  filter(mclust_success)

```

```{r}
library(EMCluster)
library(patchwork)

set.seed(1234)

x <- list(test$vx,test$vy)
out <- init.EM(x, nclass = 1)

```

```{r}

fitK <- kmeans(test$vx,6)

ggplot(test,aes(x=vx,group=as.factor(fitK$cluster),fill=as.factor(fitK$cluster)))+geom_histogram()
ggplot(test,aes(x=vx,group=as.factor(fitK$cluster),fill=as.factor(fitK$cluster)))+geom_density()

# Decide how many clusters to look at
n_clusters <- 7

# Initialize total within sum of squares error: wss
wss <- numeric(n_clusters)

set.seed(123)

# Look over 1 to n possible clusters
for (i in 1:n_clusters) {
  # Fit the model: km.out
  km.out <- kmeans(test$vx, centers = i, nstart = 20)
  # Save the within cluster sum of squares
  wss[i] <- km.out$tot.withinss
}

# Produce a scree plot
wss_df <- tibble(clusters = 1:n_clusters, wss = wss)

scree_plot <- ggplot(wss_df, aes(x = clusters, y = wss, group = 1)) +
    geom_point(size = 4)+
    geom_line() +
    scale_x_continuous(breaks = c(2, 4, 6, 8, 10)) +
    xlab('Number of clusters')
scree_plot

```

```{r}
library(mixR)

mr1 <- mixfit(test$vx,ncomp=3)
plot(mr1)
mr1

mr2 <- mixfit(test$vx,ncomp=6,family="gamma")
plot(mr2)
mr2
```

```{r}

library(flexmix)
#https://rdrr.io/cran/flexmix/man/flexmix.html

fm1=flexmix(vx~1|id,data=test,k=6)
summary(fm1)
plot(fm1)
parameters(fm1)

test %>% group_by(vb) %>% summarise(mean(vx),mad(vx),sd(vx),IQR(vx))

```

```{r}
#http://www.sthda.com/english/articles/29-cluster-validation-essentials/96-determining-the-optimal-number-of-clusters-3-must-known-methods/

e1 <- readRDS(here("data/e1_08-04-23.rds"))
test <- e1 |> filter(expMode2 == "Test") |> ungroup()
library(factoextra)
library(NbClust)

fviz_nbclust(test |> select(vx), kmeans,method = "gap_stat")

NbClust(testS |> select(vx),distance="euclidean",min.nc=2,max.nc=6,method="average")

```

```{r}
#| eval: false

library(brms)

testS <- test |> filter(id %in% 1:20)

options(mc.cores = 2, brms.backend = "cmdstanr")

mix <- mixture(gaussian, gaussian, gaussian,gaussian,gaussian,gaussian)
prior <- c(
  prior(normal(200, 100), Intercept, dpar = mu1),
  prior(normal(450, 100), Intercept, dpar = mu2),
  prior(normal(700, 100), Intercept, dpar = mu3),
  prior(normal(900, 100), Intercept, dpar = mu4),
  prior(normal(1100, 100), Intercept, dpar = mu5),
  prior(normal(1300, 100), Intercept, dpar = mu6)
)
fit1 <- brm(bf(vxC ~ 1), testS, family = mix,
            prior = prior, chains = 2)

summary(fit1)
pp_check(fit1,type="stat_grouped",ndraws=500,group="bandInt",stat="mean")

# saveRDS(object = fit1,
#         file = here("data/model_cache/mix6_vxC_test6.rda"), 
#         compress = "xz")

```

7x speedup for scaled vx
```{r}

library(brms)

test$scaleVx <- scale(test$vx)
test %>% group_by(vb) %>% summarise(mean(scaleVx))
# vb        `mean(scaleVx)`
#   <fct>               <dbl>
# 1 100-300           -0.544 
# 2 350-550           -0.273 
# 3 600-800           -0.0194
# 4 800-1000           0.449 
# 5 1000-1200          0.774 
# 6 1200-1400          1.00

testS <- test |> filter(id %in% 1:20)

options(mc.cores = 2, brms.backend = "cmdstanr")

mix <- mixture(gaussian, gaussian, gaussian,gaussian,gaussian,gaussian)
prior <- c(
  prior(normal(-.5, 10), Intercept, dpar = mu1),
  prior(normal(-.25, 10), Intercept, dpar = mu2),
  prior(normal(-.02, 10), Intercept, dpar = mu3),
  prior(normal(.4, 10), Intercept, dpar = mu4),
  prior(normal(.7, 10), Intercept, dpar = mu5),
  prior(normal(1, 10), Intercept, dpar = mu6)
)
fitScale <- brm(bf(scaleVx ~ 1), testS, family = mix,
            prior = prior, chains = 2)

summary(fitScale)
pp_check(fit1,type="stat_grouped",ndraws=500,group="bandInt",stat="mean")

# saveRDS(object = fit1,
#         file = here("data/model_cache/mix6_vxC_test6.rda"), 
#         compress = "xz")

```

```{r}

options(mc.cores = 4, brms.backend = "cmdstanr")

mix <- mixture(gaussian, gaussian, gaussian,gaussian,gaussian,gaussian)
prior <- c(
  prior(normal(-.5, 10), Intercept, dpar = mu1),
  prior(normal(-.25, 10), Intercept, dpar = mu2),
  prior(normal(-.02, 10), Intercept, dpar = mu3),
  prior(normal(.4, 10), Intercept, dpar = mu4),
  prior(normal(.7, 10), Intercept, dpar = mu5),
  prior(normal(1, 10), Intercept, dpar = mu6)
)

fitScaleInt <- brm(bf(scaleVx ~ 1 + (1|id)), test, family = mix,
            prior = prior, chains = 4)

summary(fitScaleInt)
pp_check(fitScaleInt,type="stat_grouped",ndraws=500,stat="mean")

# saveRDS(object = fit1,
#         file = here("data/model_cache/mix6_vxC_test6.rda"), 
#         compress = "xz")

```