### 2023-08-09 Analysis/e1_brms_test.qmd
sync changes

---
title: "E1 Testing - Bayesian Mixed Models"
subtitle: "Fitting mixed effects models"
date: last-modified
categories: [Analysis, R, Bayesian]
code-fold: true
---

```{r}
pacman::p_load(tidyverse,tidybayes,brms,bayesplot,broom,broom.mixed,lme4,emmeans,here,knitr,kableExtra,gt,gghalves,patchwork,ggforce,ggdist)
e1 <- readRDS(here("data/e1_08-04-23.rds"))
source(here("Functions/Display_Functions.R"))
source(here("Functions/org_functions.R"))
test <- e1 |> 
  filter(expMode2 == "Test") |> 
  group_by(vb) |> 
  mutate(distS = as.numeric(scale(dist, scale = FALSE)), 
         distS2 = custom_scale(dist))

options(brms.backend="cmdstanr",mc.cores=4)

test %>% group_by(condit) |> summarise(mean=mean(dist),sd=sd(dist),sk=moments::skewness(dist)) 
test %>% group_by(vb) |> summarise(mean=mean(dist),sd=sd(dist),sk=moments::skewness(dist))

```

```{r fig.width=12}

test |> ggplot(aes(x=dist))+geom_histogram() + facet_wrap(~vb) + ggtitle("empirical_dist")  +
test |> ggplot(aes(x=distS))+geom_histogram() + facet_wrap(~vb) + ggtitle("centered dist")  +
test |> ggplot(aes(x=distS2))+geom_histogram() + facet_wrap(~vb) + ggtitle("scaled dist")

test |> filter(id %in% 1:5) |> ggplot(aes(x=dist))+geom_histogram() + facet_wrap(id~vb) + ggtitle("empirical_dist")

```

```{r}

sk1 <- brm(dist ~ vb,family=skew_normal(),data=test,iter=800,chains=2)
g1 <- brm(dist ~ vb,family=gaussian(),data=test,iter=800,chains=2)
ga2 <- brm(dist+.01 ~ vb,family=Gamma(),data=test,iter=800,chains=2)
ln1 <- brm(dist+.01 ~ vb,family=lognormal(),data=test,iter=800,chains=2)
ln2 <- brm(dist+.0001 ~ vb,family=lognormal(link="inverse"),data=test,iter=800,chains=2)

g2 <- brm(bf(dist+.001|trunc(lb=0) ~ vb),data=testS,iter=800,chains=2,family=gaussian())
bayesplot::ppc_dens_overlay_grouped(testS$dist,yrep=posterior_predict(g2,ndraws=200),group=testS$vb)
pp_check(g2,type="stat_grouped",ndraws=200, group="vb",stat="mean")

bayesplot::ppc_dens_overlay_grouped(test$dist,yrep=posterior_predict(sk1,ndraws=200),group=test$vb)
bayesplot::ppc_dens_overlay_grouped(test$dist,yrep=posterior_predict(g1,ndraws=200),group=test$vb)
bayesplot::ppc_dens_overlay_grouped(test$dist,yrep=posterior_predict(ga2,ndraws=200),group=test$vb)
bayesplot::ppc_dens_overlay_grouped(test$dist,yrep=posterior_predict(ln1,ndraws=200),group=test$vb)

bayes_R2(g1)
bayes_R2(ga1)
bayes_R2(sk1)
bayes_R2(sk1)
bayes_R2(g1S)

#testS <- test %>% group_by(id,vb) |> filter(id %in% c("1","2","139")) |> select(id,vb,gt.stage,trial,condit,vx,dist,distS,distS2)
testS <- test %>% group_by(id,vb) |> filter(id %in% 1:15) |> select(id,vb,gt.stage,trial,condit,vx,dist,distS,distS2)

testS |> ggplot(aes(x=trial,y=dist,col=vb))+geom_line()+facet_wrap(~id)
testS |> ggplot(aes(x=trial,y=distS2,col=vb))+geom_line()+facet_wrap(~id)

g1S <- brm(distS ~ 0+ vb + condit,family=gaussian(),data=test,iter=1000,chains=4)
bayesplot::ppc_dens_overlay_grouped(test$distS,yrep=posterior_predict(g1S,ndraws=200),group=test$vb)

g1S2 <- brm(distS2 ~ vb,family=gaussian(),data=test,iter=1000,chains=4)
bayesplot::ppc_dens_overlay_grouped(test$distS2,yrep=posterior_predict(g1S2,ndraws=200),group=test$vb)

g1S_F <- brm(distS ~ 0+ vb + condit + (0+vb|id),family=gaussian(),data=test,iter=1000,chains=4,
             file=paste0(here::here("data/model_cache","e1_test_centeredDistS")))

bayesplot::ppc_dens_overlay_grouped(test$distS,yrep=posterior_predict(g1S_F,ndraws=200),group=test$vb)
bayesplot::ppc_dens_overlay_grouped(test$distS,yrep=posterior_predict(g1S_F,ndraws=200),group=test$vb)

g1S_gamma <- brm(dist+.001 ~ 1+ vb + condit + (0+vb|id),family=Gamma(),data=test,iter=1000,chains=4,
             file=paste0(here::here("data/model_cache","e1_test_Gamma")))
bayesplot::ppc_dens_overlay_grouped(test$dist,yrep=posterior_predict(g1S_gamma,ndraws=200),group=test$vb)

pp_check(g1S_weibull)+ xlim(c(-300,300))

g1S_exG <- brm(dist ~ 0 + vb + condit + (0+vb|id),family=exgaussian(),data=test,iter=1000,chains=4,
             file=paste0(here::here("data/model_cache","e1_test_exgauss")))
bayesplot::ppc_dens_overlay_grouped(test$dist,yrep=posterior_predict(g1S_exG,ndraws=200),group=test$vb)

```

## Fixed Effects Only
```{r}

modelName <- "e1_testDist"
e1_testDist <- brm(dist ~ condit,data=test,file=paste0(here::here("data/model_cache",modelName)))
brms_eq_tidy <-tidyMCMC(e1_testDist, conf.int = TRUE, conf.level = 0.95,
           estimate.method = "median", conf.method = "HPDinterval")

modelName <- "e1_testDist0"
e1_testDist0 <- brm(dist ~ 0+condit,data=test,file=paste0(here::here("data/model_cache",modelName)))
brms_eq_tidy <-tidyMCMC(e1_testDist0, conf.int = TRUE, conf.level = 0.95,
           estimate.method = "median", conf.method = "HPDinterval")

modelName <- "e1_testDist_uneq"
e1_testDist_uneq <- brm(bf(dist ~ condit,sigma~condit),data=test,file=paste0(here::here("data/model_cache",modelName)))
brms_eq_tidy_uneq <-tidyMCMC(e1_testDist_uneq, conf.int = TRUE, conf.level = 0.95,
           estimate.method = "median", conf.method = "HPDinterval")
brms_eq_tidy_uneq %>%
  mutate_at(vars(estimate, std.error, conf.low, conf.high),
            funs(ifelse(str_detect(term, "sigma"), exp(.), .)))

modelName <- "e1_testDist_uneq_robust"
brms_uneq_robust <- brm(
  bf(dist ~ condit, sigma ~ condit), data=test,
  family = student,file=paste0(here::here("data/model_cache",modelName)))

brms_uneq_robust_tidy <- tidyMCMC(brms_uneq_robust, conf.int = TRUE, conf.level = 0.95,estimate.method = "median", conf.method = "HPDinterval") %>% mutate_at(vars(estimate, std.error, conf.low, conf.high),
            funs(ifelse(str_detect(term, "sigma"), exp(.), .)))
brms_uneq_robust_tidy

brm(dist ~ (0+vb)+(1+condit),data=test)

bayes_R2(e1_testDist)
sjPlot::tab_model(e1_testDist)

```

## Random Effects
```{r}

modelName <- "e1_testDistRF"
e1_testDistRF <- brm(dist ~ condit + vb + (1 + vb|id),data=test,file=paste0(here::here("data/model_cache",modelName)))
brms_eq_tidy <-tidyMCMC(e1_testDistRF, conf.int = TRUE, conf.level = 0.95,
           estimate.method = "median", conf.method = "HPDinterval")

bayes_R2(e1_testDistRF)
sjPlot::tab_model(e1_testDistRF)

```

## Random Effects - Interaction condit x vb
```{r}

modelName <- "e1_testDistRF2"
e1_testDistRF2 <- brm(dist ~ condit * vb + (1 + vb|id),data=test,file=paste0(here::here("data/model_cache",modelName)))
brms_eq_tidy <-tidyMCMC(e1_testDistRF2, conf.int = TRUE, conf.level = 0.95,
           estimate.method = "median", conf.method = "HPDinterval")

bayes_R2(e1_testDistRF2)
sjPlot::tab_model(e1_testDistRF2)

fixef(e1_testDistRF2)
newdat <-data.frame(crossing(condit=c("Constant","Varied"), vb = unique(test$vb)))
preds <- fitted(e1_testDistRF2, re_formula = NA, newdata = newdat, probs = c(0.025, 0.975))

#shinystan::launch_shinystan(e1_testDistRF2)

```

## 0 intercept;  Random Effects - Interaction condit x vb
```{r}

modelName <- "e1_testDistRF2_0"
e1_testDistRF2_0 <- brm(dist ~ 0 + (condit * vb) + (0 + vb|id),data=test,file=paste0(here::here("data/model_cache",modelName)), 
                        iter=1000,chains=2)
brms_eq_tidy <-tidyMCMC(e1_testDistRF2_0, conf.int = TRUE, conf.level = 0.95,
           estimate.method = "median", conf.method = "HPDinterval")

bayes_R2(e1_testDistRF2_0)
sjPlot::tab_model(e1_testDistRF2_0)

fixef(e1_testDistRF2_0)
newdat <-data.frame(crossing(condit=c("Constant","Varied"), vb = unique(test$vb)))
preds <- fitted(e1_testDistRF2_0, re_formula = NA, newdata = newdat, probs = c(0.025, 0.975))

shinystan::launch_shinystan(e1_testDistRF2_0)

```

## 0 intercept;  Random Effects - Interaction condit x vb
```{r}

modelName <- "e1_testDistRF2_02"
e1_testDistRF2_02 <- brm(dist ~ 0 + condit:vb + (0 + vb|id),data=test,file=paste0(here::here("data/model_cache",modelName)))
brms_eq_tidy <-tidyMCMC(e1_testDistRF2_02, conf.int = TRUE, conf.level = 0.95,
           estimate.method = "median", conf.method = "HPDinterval")

bayes_R2(e1_testDistRF2_02)
sjPlot::tab_model(e1_testDistRF2_02)

fixef(e1_testDistRF2_02)
newdat <-data.frame(crossing(condit=c("Constant","Varied"), vb = unique(test$vb)))
preds <- fitted(e1_testDistRF2_02, re_formula = NA, newdata = newdat, probs = c(0.025, 0.975))

shinystan::launch_shinystan(e1_testDistRF2_02)

```

```{r}

modelName <- "e1_test_Dist_Int"
e1_testDist_Int <- brm(dist ~ 0 + Intercept + bandInt + condit + (1|id), 
                       data=test,
                       iter=1000, chains=4, silent=0,
                       file=paste0(here::here("data/model_cache",modelName)))

e1_testDist_Int

pp_check(e1_testDist_Int,group="bandInt")
bayesplot::ppc_dens_overlay_grouped(test$dist,yrep=posterior_predict(e1_testDist_Int,ndraws=100),group=test$bandInt)
pp_check(e1_testDist_Int,type="stat_grouped",ndraws=500, group="bandInt",stat="mean")
coef(e1_testDist_Int)$id

```

```{r}

modelName <- "e1Test_conditBand_RS1"
e1Test_conditBand_RS1 <- brm(dist ~ 1 + bandInt + condit + (1+ bandInt|id), 
                       data=test,
                       iter=1000, chains=4, silent=0,
                       file=paste0(here::here("data/model_cache",modelName)))

e1Test_conditBand_RS1
fixef(e1Test_conditBand_RS1)

pp_check(e1Test_conditBand_RS1,group="bandInt")
bayesplot::ppc_dens_overlay_grouped(test$dist,yrep=posterior_predict(e1Test_conditBand_RS1,ndraws=100),group=test$bandInt)
pp_check(e1Test_conditBand_RS1,type="stat_grouped",ndraws=500, group="bandInt",stat="mean")
coef(e1Test_conditBand_RS1)$id

e1Test_conditBand_RS1 |> 
  tidy_draws() |> 
  select(starts_with("b_"),.chain,.iteration,.draw)

e1Test_conditBand_RS1 |> 
  spread_draws()

```

```{r}

modelName <- "e1Test_conditVb_RS1"
e1Test_conditVb_RS1 <- brm(dist ~ 1 + vb + condit + (1+ vb|id), 
                       data=test,
                       iter=1000, chains=4, silent=0,
                       file=paste0(here::here("data/model_cache",modelName)))

e1Test_conditVb_RS1

pp_check(e1Test_conditVb_RS1,group="bandInt")
bayesplot::ppc_dens_overlay_grouped(test$dist,yrep=posterior_predict(e1Test_conditVb_RS1,ndraws=100),group=test$bandInt)
pp_check(e1Test_conditVb_RS1,type="stat_grouped",ndraws=500, group="vb",stat="mean")
coef(e1Test_conditVb_RS1)$id

```

```{r}

```

```{r}

```

```{r}

```