### 2023-08-12 Analysis/e1_discrim.qmd
google sync, new posterior function

---
title: "E1 Discrimination Analysis"
subtitle: "Fitting mixed effects models"
date: last-modified
categories: [Analysis, R]
code-fold: true
---

```{r setup}
pacman::p_load(tidyverse,tidybayes,brms,broom,broom.mixed,lme4,emmeans,here,knitr,kableExtra,gt,gghalves,patchwork,ggforce,ggdist,equatiomatic)
e1 <- readRDS(here("data/e1_08-04-23.rds"))
source(here("Functions/Display_Functions.R"))

test <- e1 |> filter(expMode2 == "Test") 
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vxMean=mean(vx),vxMed=median(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

#test %>% filter(id==1, band==1) %>% select(vx) %>% unlist() %>% as.vector()

```

```{r}

lmerFit <- function(mform,df) {
m=lmer(mform, df)
  fe <- fixef(m)
  dfm <- coef(m)$id %>% as.tibble(rownames="id") %>%rename("Intercept"=`(Intercept)`)
  for (effect in names(fe)) {
      dfm <- dfm %>% mutate(!!str_c("group_", effect) := fe[effect])
  }
  dfm
}

# Input formulas and corresponding dataframes
formulas <- c("vx ~ 1 + bandInt + (1 + bandInt | id)", 
              "vxMean ~ 1 + bandInt + (1 + bandInt | id)", 
              "vxMed ~ 1 + bandInt + (1 + bandInt | id)")
dfs <- list(test, testAvg, testAvg)

# Run lmerFit function over all combinations of formulas and dataframes
m_all <- map2_dfr(formulas, dfs, ~ lmerFit(.x, .y)) %>%
  mutate(mformula = rep(formulas, each = length(unique(test$id))), 
         dframe = rep(c('test', 'testAvg', 'testAvg'), each=length(unique(test$id))))

head(m_all)
m_all |> filter(id==1)

# m1 <- lmerFit("vx ~ 1 + bandInt + (1 + bandInt | id)",test)
# m2 <- lmerFit("vxMean ~ 1 + bandInt + (1 + bandInt | id)",testAvg)
# m3 <- lmerFit("vxMed ~ 1 + bandInt + (1 + bandInt | id)",testAvg)

m1 <- lmer(vx ~ 1 + bandInt + condit + (1 + bandInt | id),data=test)
coef(m1)$id %>% as.tibble(rownames="id") %>%rename("Intercept"=`(Intercept)`)
summary(m1)

fitted(m1)
new_data <- testAvg  |> ungroup() |> select(id,condit,bandInt)

new_data=expand_grid(unique(test[, c("id", "condit")]), bandInt=unique(test$bandInt))
new_data <- new_data[rep(row.names(new_data), each = 10), ]

# Compute predictions
new_data$p <- predict(m1, newdata = new_data)

options(mc.cores = 4, brms.backend = "cmdstanr")

m1_bayes <- brm(vx ~ 1 + bandInt + condit + (1 + bandInt | id), data = test, family = gaussian())
summary(m1_bayes)

new_data=map_dfr(1, ~data.frame(unique(test[,c("id","condit","bandInt")]))) |> dplyr::arrange(id,bandInt)
predictions <- t(posterior_predict(m1_bayes, newdata = new_data,ndraws=10))
new_data$prediction <- rowMeans(predictions)

predicted_draws <- new_data %>%
  tidybayes::add_predicted_draws(m1_bayes,ndraws=5)

test$censored_vx <- ifelse(test$vx < 50, 50, test$vx)
test$censored_vx <- ifelse(test$censored_vx > 1600, 1600, test$censored_vx)
test$censored <- ifelse((test$vx <50 | test$vx>1600), TRUE, FALSE)

# fit model
m1_bayes <- brm(vx~ 1 + bandInt + condit + (1 + bandInt | id), 
                data = test, family = gaussian())

m1_bayesSN <- brm(vx ~ 1 + bandInt + condit + (1 + bandInt | id), data = test, family = skew_normal(),chains = 4)
summary(m1_bayesSN)

pst <- posterior_samples(m1_bayes, "b")
ggplot(testAvg, aes(x = bandInt, y = vxMean)) +
    geom_point(shape = 1) +
    geom_abline(
        data = pst, alpha = .01, size = .1,
        aes(intercept = b_Intercept, slope = b_bandInt)
    )

X <- cbind(test[,c("id","condit","trial","vb","bandInt","vx")], fitted(m1_bayes)[,-2]) %>% as_tibble()

```

$$
\begin{aligned}
  \operatorname{vx}_{i}  &\sim N \left(\alpha_{j[i]} + \beta_{1j[i]}(\operatorname{bandInt}), \sigma^2 \right) \\    
\left(
  \begin{array}{c} 
    \begin{aligned}
      &\alpha_{j} \\
      &\beta_{1j}
    \end{aligned}
  \end{array}
\right)
  &\sim N \left(
\left(
  \begin{array}{c} 
    \begin{aligned}
      &\gamma_{0}^{\alpha} + \gamma_{1}^{\alpha}(\operatorname{condit}_{\operatorname{Varied}}) \\
      &\mu_{\beta_{1j}}
    \end{aligned}
  \end{array}
\right)
, 
\left(
  \begin{array}{cc}
     \sigma^2_{\alpha_{j}} & \rho_{\alpha_{j}\beta_{1j}} \\ 
     \rho_{\beta_{1j}\alpha_{j}} & \sigma^2_{\beta_{1j}}
  \end{array}
\right)
 \right)
    \text{, for id j = 1,} \dots \text{,J}
\end{aligned}
$$

`r cat(extract_eq(m1, use_coefs = TRUE))`

$$
\begin{aligned}
  \operatorname{\widehat{vx}}_{i}  &\sim N \left(480.37_{\alpha_{j[i]}} + 0.63_{\beta_{1j[i]}}(\operatorname{bandInt}), \sigma^2 \right) \\    
\left(
  \begin{array}{c} 
    \begin{aligned}
      &\alpha_{j} \\
      &\beta_{1j}
    \end{aligned}
  \end{array}
\right)
  &\sim N \left(
\left(
  \begin{array}{c} 
    \begin{aligned}
      &56.83_{\gamma_{1}^{\alpha}}(\operatorname{condit}_{\operatorname{Varied}}) \\
      &0
    \end{aligned}
  \end{array}
\right)
, 
\left(
  \begin{array}{cc}
     351.8 & -0.8 \\ 
     -0.8 & 0.38
  \end{array}
\right)
 \right)
    \text{, for id j = 1,} \dots \text{,J}
\end{aligned}
$$

```{r}
m1_bayes <- brm(vx ~ 1 + bandInt + condit + (1 + bandInt | id), 
data = test, family = gaussian(), file="band_gauss_vx")

m1_bayesb <- brm(vxb ~ 1 + bandInt + condit + (1 + bandInt | id), 
data = test, family = gaussian(), file="band_gauss_vxb")

m1_bayes0 <- brm(vx ~ 0 + bandInt + condit + (0 + bandInt | id), 
data = test, family = gaussian(), file="band0_gauss_vx")

m1_bayes0b <- brm(vxb ~ 0 + bandInt + condit + (0 + bandInt | id), 
data = test, family = gaussian(), file="band0_gauss_vxb")

summary(m1_bayes)

```

```{r}
# Calculate the standard deviation of the estimates for each level of id
m_all %>%
  summarise(across(starts_with("group_"), sd, na.rm = TRUE))

# Calculate the correlation between the estimates for each level of id
m_all %>%
  group_by(id) %>%
  summarise(correlation = cor(`bandInt`, `Intercept`, use = "pairwise.complete.obs"))

m_all |> filter(id==10)

```

```{r}
#| eval: false

options(mc.cores = 8,  # Use 4 cores
        brms.backend = "cmdstanr")

model_fixed <- brm(
  bf(vxMean ~ bandInt + (1 +bandInt | id)),
  data = testAvg,
  control = list(adapt_delta = 0.95),
  chains = 4, seed = 1234
)

fixef(model_fixed)

mm_Mean_b <- coef(model_fixed)$id |> as.tibble(rownames="id") |> select(id,starts_with("Estimate"))
head(mm_Mean_b)

X <- cbind(testAvg[,1:8], fitted(model_fixed)[,-2]) %>% as_tibble()

ggplot(X[1:90,],aes(y=Estimate,x=vb,group=id))+geom_point(aes(y=vxMean),shape=1)+
  geom_line(aes(y=Q2.5),lty=2)+
  geom_line(aes(y=Q97.5),lty=2)+
  geom_smooth(aes(y=vxMean),method = "lm", fill = "dodgerblue", level = .95)+
  facet_wrap(~id)

conditional_effects(model_fixed)
#conditional_effects(model_fixed,conditions=distinct(testAvg,id))

predict(model_fixed,data.frame(id=1,bandInt=800))

epred_draws(model_fixed,newdata=expand_grid(bandInt=unique(e1$bandInt),id=factor(e1$id,levels=unique(e1$id))),ndraws = 1)

```

```{r}
#| eval: false

log_m1 <- brm(
  bf(vx ~ bandInt + (1 +bandInt | id)),
  data = test,
  family=shifted_lognormal(),
  control = list(adapt_delta = 0.95),
  chains = 2, seed = 1234
)

summary(log_m1)

skew_m1 <- brm(
  bf(vxMean ~ bandInt + (1 +bandInt | id)),
  data = testAvg,
  family=skew_normal(),
  control = list(adapt_delta = 0.95),
  chains = 2, seed = 1234
)

summary(skew_m1)

```

```{r}
#| eval: false
library(brms)

bmerFit <- function(mform,df) {
  m <- brm(mform, data = df, control = list(adapt_delta = 0.95), chains = 1, seed = 1234)
  fe <- fixef(m)
  dfm <- coef(m)$id %>% as.tibble(rownames="id") %>% select(id, starts_with("Estimate"))

  for (effect in names(fe)) {
    dfm <- dfm %>% mutate(!!str_c("group_", effect) := fe[effect])
  }
  dfm
}

# Input formulas and corresponding dataframes
formulas <- c("vx ~ 1 + bandInt + (1 + bandInt | id)", 
              "vxMean ~ 1 + bandInt + (1 + bandInt | id)", 
              "vxMed ~ 1 + bandInt + (1 + bandInt | id)")
dfs <- list(test, testAvg, testAvg)

# Run bmerFit function over all combinations of formulas and dataframes
m_all_bayes <- map2_dfr(formulas, dfs, ~ bmerFit(.x, .y)) %>%
  mutate(mformula = rep(formulas, each = length(unique(test$id))), 
         dframe = rep(c('test', 'testAvg', 'testAvg'), each=length(unique(test$id))))

m_all_bayes(m_all)
m_all_bayes |> filter(id==10)

```

```{r}

```