### 2023-08-18 Misc/indv_diff.qmd

---
title: Individual Differenes
date: last-modified
categories: [Analysis, R]
page-layout: full
toc: false
execute: 
  warning: false
  eval: false
---

```{r setup, include=FALSE}
source(here::here("Functions", "packages.R"))
test <- readRDS(here("data/e1_08-04-23.rds")) |> filter(expMode2 == "Test") 
options(brms.backend="cmdstanr",mc.cores=4)
e1Sbjs <- test |> group_by(id,condit) |> summarise(n=n())

testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k_Ml1")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
mt2 <-GetModelStats(e1_vxBMM ) |> kable(escape=F,booktabs=T)

new_data_grid=map_dfr(1, ~data.frame(unique(test[,c("id","condit","bandInt")]))) |> 
  dplyr::arrange(id,bandInt) |> 
  mutate(condit_dummy = ifelse(condit == "Varied", 1, 0))

nTop <- 5

indv_coefs <- coef(e1_vxBMM)$id |> 
  as_tibble(rownames="id") |> 
  select(id, starts_with("Est")) |>
  left_join(e1Sbjs, by=join_by(id) ) |> 
  group_by(condit) |> 
  mutate(rank = rank(desc(Estimate.bandInt)),
         intErrorRank=rank((Est.Error.Intercept)),
         bandErrorRank=rank((Est.Error.bandInt)),
         nCond = n()) |> arrange(intErrorRank) |> ungroup()

top_indv_coefs <- indv_coefs %>%
  filter(rank <=nTop) |>
  select(id,condit,rank,Est.BandInt=Estimate.bandInt, Est.Intercept=Estimate.Intercept)

small_slope <- indv_coefs %>%
  filter(rank >=nCond - nTop) |>
  select(id,condit,rank,Est.BandInt=Estimate.bandInt, Est.Intercept=Estimate.Intercept)

well_fit <- indv_coefs %>%
  filter(intErrorRank<=nTop)

```

```{r}

coef(e1_vxBMM)$id %>% as_tibble(rownames="id") %>% select(id, starts_with("Est")) |> print(n=15)

n_lines <- 20
f <- fitted(e1_vxBMM, newdata = new_data_grid, re_formula = NA, summary=F, ndraws=n_lines) |> 
  as_tibble() %>% 
  mutate(draw = 1:n()) %>% 
  pivot_longer(-draw)

random_effects <- e1_vxBMM |> 
  gather_draws(`^r_id.*$`, regex = TRUE, ndraws = 1000) |> 
  separate(.variable, into = c("effect", "id", "term"), sep = "\\[|,|\\]") |> 
  mutate(id = factor(id,levels=levels(test$id))) |> 
  pivot_wider(names_from = term, values_from = .value) |> arrange(id,.chain,.draw,.iteration)

fixed_effects <- e1_vxBMM |> 
  spread_draws(`^b_.*`,regex=TRUE) |> arrange(.chain,.draw,.iteration)

combined_df <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt) |>
  right_join(new_data_grid, by = join_by("id")) |> 
  mutate(
    fixed_effects = b_Intercept +
      (bandInt * b_bandInt) +
      condit_dummy * b_conditVaried + 
      (bandInt * condit_dummy) * `b_conditVaried:bandInt` ,  # Note: Replace : with correct operator
    random_effects = Intercept + bandInt_RF,  # Assuming random effects for intercept and bandInt
    estimate = fixed_effects + random_effects
  )

combined_df |> 
  filter(id %in% top_indv_coefs$id) |> 
  group_by(id, bandInt) |>
  sample_n(50) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  geom_abline(data=indv_coefs |> filter(id %in% top_indv_coefs$id),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt,color="red")) +
  stat_halfeye() +
  stat_halfeye(data=testAvg |> filter(id %in% top_indv_coefs$id), aes(x=bandInt,y=vx),color="blue") +
  facet_wrap(~id)

combined_df |> 
  filter(id %in% well_fit$id) |> 
  group_by(id, bandInt) |>
  sample_n(50) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  geom_abline(data=indv_coefs |> filter(id %in% well_fit$id),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt,color="red")) +
  stat_halfeye() +
  stat_halfeye(data=testAvg |> filter(id %in% well_fit$id), aes(x=bandInt,y=vx),color="blue") +
  ggh4x::facet_nested_wrap(vars(condit,id))

combined_df |> 
  filter(id %in% (indv_coefs |> filter(rank<=2) |> pull(id)) )  |> 
  group_by(id, bandInt) |>
  sample_n(50) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  geom_abline(data=indv_coefs |> filter(id %in% (indv_coefs |> filter(rank<=2) |> pull(id))),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt,color="red")) +
  stat_halfeye() +
  stat_halfeye(data=testAvg |> filter(id %in% (indv_coefs |> filter(rank<=2) |> pull(id))), aes(x=bandInt,y=vx),color="blue") +
  ggh4x::facet_nested_wrap(vars(condit,id))

```

```{r}

indv_coefs |> 
  filter(id %in% top_indv_coefs$id) |> ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  stat_halfeye()

all_effects <- e1_vxBMM |> 
  gather_draws(b_Intercept, b_conditVaried, b_bandInt, `b_conditVaried:bandInt`, `^r_id.*$`, regex = TRUE, ndraws = 1)

head(all_effects)

 nd <- e1_vxBMM |>  
   spread_draws(b_Intercept, b_bandInt, `b_conditVaried`,`b_conditVaried:bandInt`,  r_id[id,term], ndraws=10)

```

```{r}

indv_coefs |> 
  filter(id %in% top_indv_coefs$id) |> ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  stat_halfeye()

all_effects <- e1_vxBMM |> 
  gather_draws(b_Intercept, b_conditVaried, b_bandInt, `b_conditVaried:bandInt`, `^r_id.*$`, regex = TRUE, ndraws = 1)

head(all_effects)

 nd <- e1_vxBMM |>  
   spread_draws(b_Intercept, b_bandInt, `b_conditVaried`,`b_conditVaried:bandInt`,  r_id[id,term], ndraws=10)

```

```{r}

df_posterior <- e1_vxBMM |> as_tibble()

ggplot(df_posterior) + 
  aes(x = b_Intercept, y = b_bandInt) + 
  stat_density_2d( geom = "polygon")

df_posterior |> 
  sample_n(1000) |>
  ggplot() + 
  aes(x = b_Intercept, y = b_bandInt) + 
  stat_dist_dotsinterval()

```