### 2023-08-24 Analysis/Appendix/E1_Appendix.qmd
re-render + update plot functions

```{r setupAe1, include=FALSE}
source(here::here("Functions", "packages.R"))
test <- readRDS(here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Test") 
options(brms.backend="cmdstanr",mc.cores=4)
e1Sbjs <- test |> group_by(id,condit) |> summarise(n=n())
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

modelName <- "e1_testDistBand_RF_5K"
e1_distBMM <- brm(dist ~ condit * bandInt + (1 + bandInt|id),
                      data=test,file=paste0(here::here("data/model_cache",modelName)),
                      iter=5000,chains=4)

e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))

```

## E1 Appendix

### Posterior Predictive Distributions

```{r}
#| label: fig-post-pred
#| fig-cap: Posterior Predictive distributions for Deviation and Vx. Posterior Draws in Blue, colored lines are empirical data. 
#| fig-subcap: ["Posterior Predictive Distribution - Deviation", "Posterior Predictive Distribution - X Velocity"]
#| fig-width: 13
#| fig-height: 12

vx_pred <- 
  posterior_predict(e1_vxBMM, ndraws = 500) |> 
  array_branch(margin=1) |> 
   map_dfr( 
    function(yrep_iter) {
      test  |>
        mutate(vx_pred = yrep_iter)
    },
    .id = 'iter'
  ) |>
  mutate(iter = as.numeric(iter))

dist_pred <- 
  posterior_predict(e1_distBMM, ndraws = 500) |> 
  array_branch(margin=1) |> 
   map_dfr( 
    function(yrep_iter) {
      test  |>
        mutate(dist_pred = yrep_iter)
    },
    .id = 'iter'
  ) |>
  mutate(iter = as.numeric(iter))

dist_pred  |>
  filter(iter < 100) %>%
  ggplot(aes(dist_pred, group = iter)) +
  geom_line(alpha = .03, stat = 'density', color = 'blue') +
  geom_density(data = test,
               aes(dist,col=vb),
               inherit.aes = FALSE,
               size = 0.7) + # 1
  facet_grid(condit ~ vb) +
  xlab('Deviation')

vx_pred  |>
  filter(iter < 100) %>%
  ggplot(aes(vx_pred, group = iter)) +
  geom_line(alpha = .03, stat = 'density', color = 'blue') +
  geom_density(data = test,
               aes(vx,col=vb),
               inherit.aes = FALSE,
               size = 0.7) + # 1
  facet_grid(condit ~ vb) +
  xlab('Vx')

```

### Empirical vs. Predicted

```{r}
#| label: fig-empVsPred
#| fig-cap: Bayesian Mixed Model predictions vs. Empirical Predictions - X velocity
#| fig-height: 15
#| fig-width: 12

{
vx_pred  |>
  filter(iter < 100)  |> group_by(id,condit,vb,iter) |>
  summarise(vx_pred=mean(vx_pred)) %>%
  ggplot(aes(x=vb,y=vx_pred,fill=condit)) + 
  geom_flat_violin( position = position_nudge(x = 0.1, y = 0),
                   adjust = 1.5,
                   trim = FALSE, alpha = .5, colour = NA) +
  # geom_point(aes(x = as.numeric(vb) - 0.15, y = vx_pred, colour = vb),
  #            position = position_jitter(width = 0.05, height = 0),
  #            size = 1, shape = 20) +
  geom_boxplot(aes(x = vb, y = vx_pred, fill = condit),
               outlier.shape = NA,
               alpha = 0.5,
               width = 0.1,
               colour = "black") +
  geom_hline(yintercept = 0,
             linetype = 'dashed',
             color = 'red',
             size = 0.4) + 
  coord_flip() + ggtitle("Predicted Vx")  } /
{
vx_pred  |>
  filter(iter < 2)  |> group_by(id,condit,vb) |>
  summarise(vx=mean(vx)) %>%
  ggplot(aes(x=vb,y=vx,fill=condit)) + 
  geom_flat_violin( position = position_nudge(x = 0.1, y = 0),
                   adjust = 1.5,
                   trim = FALSE,
                   alpha = .5,
                   colour = NA) +
  geom_point(aes(x = as.numeric(vb) - 0.15,col=condit),
             # position = position_jitter(width = 0.05),
             position = position_jitter(width = 0.05, height = 0),
             size = 1,
             shape = 20) +
  geom_boxplot(
               outlier.shape = NA,
               alpha = 0.5,
               width = 0.1,
               colour = "black") +
  geom_hline(yintercept = 0,
             linetype = 'dashed',
             color = 'red',
             size = 0.4) + 
  coord_flip() + ggtitle("Empirical Vx") }
```

### Different Aggregations

```{r fig.width=14, fig.height=13}
#| label: fig-empirical-vxGrp
#| fig-cap: E1. Distribution of vx at Participant and Trial level
#| fig-width: 13
#| fig-height: 14
epId <- vx_pred  |>
  filter(iter < 2)  |> group_by(id,condit,vb) |>
  summarise(vx=median(vx)) |>
  ggplot(aes(x=vb,y=vx,fill=condit)) + 
  geom_flat_violin(aes(fill=condit), position = position_nudge(x = 0.1, y = 0),
                   adjust = 1.5,trim = FALSE, alpha = .5, colour = NA) +
  geom_point(aes(x = as.numeric(vb) - 0.15, col=condit),
             position = position_jitter(width = 0.05, height = 0),
             size = 1, shape = 20, alpha=.7) +
  geom_boxplot(aes(x=vb,y=vx,fill=condit),
               outlier.shape = NA,
               alpha = 0.5, width = 0.1) +
  geom_hline(yintercept = 0,
             linetype = 'dashed',
             color = 'red',
             size = 0.4) + 
  coord_flip() + ggtitle("Empirical Vx - Subject level averaging")

epTrial <- vx_pred  |>
  filter(iter < 2)  |> group_by(id,condit,vb) |>
  ggplot(aes(x=vb,y=vx,fill=condit)) + 
  geom_flat_violin(aes(fill=condit), position = position_nudge(x = 0.1, y = 0),
                   adjust = 1.5,trim = FALSE, alpha = .5, colour = NA) +
  geom_point(aes(x = as.numeric(vb) - 0.15, col=condit),
             position = position_jitter(width = 0.05, height = 0),
             size = .5, shape = 20, alpha=.7) +
  geom_boxplot(aes(x=vb,y=vx,fill=condit),
               outlier.shape = NA,
               alpha = 0.5, width = 0.1) +
  geom_hline(yintercept = 0,
             linetype = 'dashed',
             color = 'red',
             size = 0.4) + 
  coord_flip() + ggtitle("Empirical Vx - Raw Trial") +
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank())

epId + epTrial
```

```{r fig.width=14, fig.height=13}
#| label: fig-empirical-distGrp
#| fig-cap: E1. Distribution of Vx  at Participant and Trial level
#| fig-width: 13
#| fig-height: 14
epId <- dist_pred  |>
  filter(iter < 2)  |> group_by(id,condit,vb) |>
  summarise(dist=median(dist)) |>
  ggplot(aes(x=vb,y=dist,fill=condit)) + 
  geom_flat_violin(aes(fill=condit), position = position_nudge(x = 0.1, y = 0),
                   adjust = 1.5,trim = FALSE, alpha = .5, colour = NA) +
  geom_point(aes(x = as.numeric(vb) - 0.15, col=condit),
             position = position_jitter(width = 0.05, height = 0),
             size = 1, shape = 20, alpha=.7) +
  geom_boxplot(aes(x=vb,y=dist,fill=condit),
               outlier.shape = NA,
               alpha = 0.5, width = 0.1) +
  geom_hline(yintercept = 0,
             linetype = 'dashed',
             color = 'red',
             size = 0.4) + 
  coord_flip() + ggtitle("Empirical Deviation - Subject level averaging")

epTrial <- dist_pred  |>
  filter(iter < 2)  |> group_by(id,condit,vb) |>
  ggplot(aes(x=vb,y=dist,fill=condit)) + 
  geom_flat_violin(aes(fill=condit), position = position_nudge(x = 0.1, y = 0),
                   adjust = 1.5,trim = FALSE, alpha = .5, colour = NA) +
  geom_point(aes(x = as.numeric(vb) - 0.15, col=condit),
             position = position_jitter(width = 0.05, height = 0),
             size = .5, shape = 20, alpha=.7) +
  geom_boxplot(aes(x=vb,y=dist,fill=condit),
               outlier.shape = NA,
               alpha = 0.5, width = 0.1) +
  geom_hline(yintercept = 0,
             linetype = 'dashed',
             color = 'red',
             size = 0.4) + 
  coord_flip() + ggtitle("Empirical Deviation - Raw Trial") +
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank())

epId + epTrial
```