### 2023-08-30 Misc/indv_diff.qmd

---
title: Individual Differenes
date: last-modified
categories: [Analysis, R]
page-layout: full
toc: false
execute: 
  warning: false
  eval: true
---

```{r setup, include=FALSE}
source(here::here("Functions", "packages.R"))
test <- readRDS(here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Test") 
options(brms.backend="cmdstanr",mc.cores=4)
e1Sbjs <- test |> group_by(id,condit) |> summarise(n=n()) |>
  mutate(condit_dummy = ifelse(condit == "Varied", 1, 0))

testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n) |> ungroup()

e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
mt2 <-GetModelStats(e1_vxBMM ) |> kable(escape=F,booktabs=T)

# nTop <- 5
# top_indv_coefs <- indv_coefs %>%
#   filter(rank <=nTop) |>
#   select(id,condit,rank,Est.BandInt=Estimate.bandInt, Est.Intercept=Estimate.Intercept) 
```

```{r}

new_data_grid=map_dfr(1, ~data.frame(unique(test[,c("id","condit","bandInt")]))) |> 
  dplyr::arrange(id,bandInt) |> 
  mutate(condit_dummy = ifelse(condit == "Varied", 1, 0))

indv_coefs <- coef(e1_vxBMM)$id |> 
  as_tibble(rownames="id") |> 
  select(id, starts_with("Esti")) |>
  left_join(e1Sbjs, by=join_by(id))

indv_coefs_preds <- new_data_grid |> left_join(indv_coefs,by=join_by(id,condit,condit_dummy)) |>
  mutate(
    estimate = (Estimate.Intercept) + 
      (bandInt*(Estimate.bandInt)) + 
      (condit_dummy*Estimate.conditVaried) +
      (bandInt * condit_dummy) * `Estimate.conditVaried:bandInt`)

indv_linpred <- e1_vxBMM |> linpred_draws(new_data_grid) |> 
  summarize(across(.linpred, lst(mean, sd, median), .names = "{.fn}"))

indv_epred <- e1_vxBMM |> epred_draws(new_data_grid) |>
  summarize(across(.epred, lst(mean, sd, median), .names = "{.fn}"))

indv_pred <- e1_vxBMM |> predicted_draws(new_data_grid) |>
   summarize(across(.prediction, lst(mean, sd, median), .names = "{.fn}"))

# fitted(e1_vxBMM,re_formula=NA)
# fitted(e1_vxBMM,re_formula=NULL) # matches epred

# |> 
#   group_by(condit) |> 
#   mutate(rank = rank(desc(Estimate.bandInt)),
#          intErrorRank=rank((Est.Error.Intercept)),
#          bandErrorRank=rank((Est.Error.bandInt)),
#          nCond = n()) |> arrange(id) |> ungroup()

fixed_effects <- e1_vxBMM |> 
  spread_draws(`^b_.*`,regex=TRUE) |> arrange(.chain,.draw,.iteration)

random_effects <- e1_vxBMM |> 
  gather_draws(`^r_id.*$`, regex = TRUE, ndraws = 1000) |> 
  separate(.variable, into = c("effect", "id", "term"), sep = "\\[|,|\\]") |> 
  mutate(id = factor(id,levels=levels(test$id))) |> 
  pivot_wider(names_from = term, values_from = .value) |> arrange(id,.chain,.draw,.iteration)

 indvDraws <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt,RF_Intercept=Intercept) |>
  right_join(new_data_grid, by = join_by("id")) |> 
  mutate(
    Slope = bandInt_RF+b_bandInt,
    Intercept= RF_Intercept + b_Intercept,
    estimate = (b_Intercept + RF_Intercept) + 
      (bandInt*(b_bandInt+bandInt_RF)) + 
      (condit_dummy*b_conditVaried)+
      (bandInt * condit_dummy) * `b_conditVaried:bandInt`,
    SlopeInt=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt` + bandInt_RF)*condit_dummy),
    SlopeInt2=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt`)*condit_dummy)
  )

  indvSlopes <- indvDraws |> group_by(id) |> median_qi(Slope,SlopeInt,SlopeInt2, Intercept,b_Intercept,b_bandInt) |>
  left_join(e1Sbjs, by=join_by(id)) |> group_by(condit) |>
    select(id,condit,condit_dummy,Intercept,b_Intercept,Slope,SlopeInt,SlopeInt2,b_bandInt, n) |>
  mutate(rankSlope=rank(Slope)) #|> arrange(rankSlope)   |> ungroup()

```

```{r}

full_tab <- testAvg |> select(id,condit,vx,vb,bandInt) |> 
  left_join(indvDraws |> group_by(id,bandInt) |> median_qi(estimate,Intercept,Slope), by=join_by(id,bandInt)) |>
  mutate(resid=vx-estimate) |> relocate(vx,resid,.before=estimate) |>
  mutate(across(where(is.numeric), \(x) round(x,2)))

indvResid <- full_tab |> group_by(id,condit,bandInt) |> summarize(across(where(is.numeric), \(x) mean(x))) |>
  group_by(id,condit) |> mutate(meanResid=mean(resid)) |>
  arrange(meanResid)

full_tab |> group_by(condit) |> summarise(mean(resid),sd(resid))
full_tab |> group_by(condit,id) |> summarise(m=mean(resid),s=sd(resid)) |> summarise(m=mean(m),s=sd(s))

full_tab |> DT::datatable(options = list(pageLength=10))
```

```{r}
  indvSlopes |> group_by(condit) |> reframe(enframe(quantile(Slope, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "Slope")) |> 
  pivot_wider(names_from=quantile,values_from=Slope,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

  indvSlopes |> group_by(condit)  |> reframe(enframe(quantile(SlopeInt, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "SlopeInt")) |> 
  pivot_wider(names_from=quantile,values_from=SlopeInt,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

  indvSlopes |> group_by(condit)  |> reframe(enframe(quantile(SlopeInt2, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "SlopeInt2")) |> 
  pivot_wider(names_from=quantile,values_from=SlopeInt2,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

  ggplot(indvSlopes, aes(x = Slope, color = condit)) + 
  geom_density() 
    ggplot(indvSlopes, aes(x = SlopeInt2, color = condit)) + 
  geom_density() 
```

```{r fig.width=15, fig.height=18}

e1_vxBMM %>% as_draws_df(variable = c('^b_', 'sigma'), regex = TRUE) |>
  select(starts_with("b_"),sigma) |> 
  pivot_longer(
    everything(),
    names_to = 'key',
    values_to = 'value') %>% 
  ggplot(aes(x = value, fill = key)) +
  stat_halfeye(
    # plot the mode
    point_interval = mode_hdi,
    # plot HDIs be at both the 95% levels
    .width = c(.95),
    outline_bars = TRUE,
    normalize = 'panels',
    show.legend = FALSE
  ) +
  facet_wrap(~key, scales = 'free', ncol = 4) +
  ggtitle('Posterior distributions')

 indvDraws |> filter(id %in% 1:10) |> group_by(id,condit,.draw) |> summarise(Slope=mean(Slope)) |> 
   ggplot(aes(x=Slope,col=condit))+geom_density() + facet_wrap(~id)

  indvDraws  |> group_by(id,condit,.draw) |> summarise(Slope=mean(Slope)) |> 
   ggplot(aes(x=Slope,col=condit,fill=condit))+stat_halfeye()

indvDraws |> group_by(condit,.draw) |> summarise(Slope=mean(Slope)) |> 
   ggplot(aes(x=Slope,col=condit,fill=condit))+stat_halfeye()

 indvDraws |> group_by(id,condit) |> summarise(Slope=mean(Slope)) |> 
   ggplot(aes(x=Slope,col=condit,fill=condit))+geom_density()

  indvDraws |> group_by(id,condit,.draw) |> summarise(SlopeInt2=mean(SlopeInt2)) |> 
   ggplot(aes(x=SlopeInt2,col=condit,fill=condit))+geom_density()

  # pp_check(e1_vxBMM,
  #        type = 'stat_grouped',
  #        ndraws = 200,
  #        group = 'id',
  #        stat = 'mean')

```

```{r fig.width=13, fig.height=12.5}

  n_sbj = 3

{indvDraws |> filter(condit=="Varied")  |> 
   indv_model_plot(indvSlopes |> filter(condit=="Varied"), 
                   testAvg |> filter(condit=="Varied"),slopeVar=Slope, rank_variable=Slope,n_sbj=n_sbj,"max") + 
     ggtitle("Slope = bandInt_RF+b_bandInt") } /

{indvDraws |> filter(condit=="Varied")  |> 
  indv_model_plot(indvSlopes |> filter(condit=="Varied"), 
                  testAvg |> filter(condit=="Varied"),slopeVar=SlopeInt, rank_variable=Slope,n_sbj=n_sbj,"max")  + 
     ggtitle("SlopeInt=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt` + bandInt_RF)*condit_dummy)") } /

indvDraws |> filter(condit=="Varied")  |> 
  indv_model_plot(indvSlopes |> filter(condit=="Varied"), 
                  testAvg |> filter(condit=="Varied"),slopeVar=SlopeInt2, rank_variable=Slope,n_sbj=n_sbj,"max") + 
     ggtitle("SlopeInt2=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt` + b_bandInt)*condit_dummy)")

{indvDraws |> filter(condit=="Varied")  |> 
   indv_model_plot(indvSlopes |> filter(condit=="Varied"), 
                   test |> filter(condit=="Varied"),slopeVar=Slope, rank_variable=Slope,n_sbj=n_sbj,"max") + 
     ggtitle("Slope = bandInt_RF+b_bandInt") } /

{indvDraws |> filter(condit=="Varied")  |> 
  indv_model_plot(indvSlopes |> filter(condit=="Varied"), 
                  test |> filter(condit=="Varied"),slopeVar=SlopeInt, rank_variable=Slope,n_sbj=n_sbj,"max")  + 
     ggtitle("SlopeInt=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt` + bandInt_RF)*condit_dummy)") } /

indvDraws |> filter(condit=="Varied")  |> 
  indv_model_plot(indvSlopes |> filter(condit=="Varied"), 
                  test |> filter(condit=="Varied"),slopeVar=SlopeInt2, rank_variable=Slope,n_sbj=n_sbj,"max") + 
     ggtitle("SlopeInt2=bandInt_RF+b_bandInt + ((`b_conditVaried:bandInt` + b_bandInt)*condit_dummy)")

```

```{r fig.height=15, fig.width=12}
#| fig-height: 16
#| fig-width: 12
{indvDraws  |> indv_model_plot(indvSlopes, test,slopeVar=Slope, rank_variable=Slope,n_sbj=n_sbj,"max") + 
     ggtitle("Slope = bandInt_RF+b_bandInt") } /
 {ggplot(indvSlopes, aes(x = Slope, color = condit)) + geom_density() } /
 gridExtra::tableGrob( indvSlopes |> group_by(condit) |> summarise(Mean=mean(Slope),Sd=sd(Slope)) ) + 
  #gt_temp(indvSlopes |> group_by(condit) |> summarise(Mean=mean(Slope),Sd=sd(Slope)) ) +
  plot_layout(heights=unit(c(16,6,3),c('cm'))) + plot_annotation("vx ~ condit * bandInt + (1 + bandInt | id) ")
```

Band Only Model
```{r fig.width=11, fig.height=15}
#| fig-height: 16
#| fig-width: 12
e1_testVx_grRF8 <- readRDS(here::here("data/model_cache/e1_testVx_grRF8.rds"))
# e1_testVx_grRF8$formula
# vx ~ bandInt + (1 + bandInt || gr(id, by = condit))

new_data_grid=map_dfr(1, ~data.frame(unique(test[,c("id","bandInt")]))) |> 
  dplyr::arrange(id,bandInt)

fixed_effects <- e1_testVx_grRF8 |> 
  spread_draws(`^b_.*`,regex=TRUE) |> arrange(.chain,.draw,.iteration)

random_effects <- e1_testVx_grRF8 |> 
  gather_draws(`^r_id.*$`, regex = TRUE, ndraws = 1000) |> 
  separate(.variable, into = c("effect", "id", "term"), sep = "\\[|,|\\]") |> 
  mutate(id = factor(id,levels=levels(test$id))) |> 
  pivot_wider(names_from = term, values_from = .value) |> arrange(id,.chain,.draw,.iteration)

indvDraws <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt,RF_Intercept=Intercept) %>%
  right_join(e1Sbjs,., by=join_by(id)) |>
  right_join(new_data_grid, by = join_by("id")) |> 
  mutate(
    Slope = bandInt_RF+b_bandInt,
    Intercept= RF_Intercept + b_Intercept,
    estimate = (b_Intercept + RF_Intercept) + (bandInt*(b_bandInt+bandInt_RF)),
  )

indvSlopes <- indvDraws |> group_by(id) |> median_qi(RF_Intercept,b_Intercept,Intercept,bandInt_RF,b_bandInt,Slope) |>
  left_join(e1Sbjs, by=join_by(id)) |> group_by(condit) |>
    select(id,condit,RF_Intercept,b_Intercept,Intercept,bandInt_RF,b_bandInt,Slope, n) |>
  ungroup() |>
  mutate(rankSlope=rank(Slope), 
         Slope2 = ((coef(e1_testVx_grRF8)$id |> as_tibble(rownames="id") ) |> pull(Estimate.bandInt)) )

{ indvDraws |> indv_model_plot(indvSlopes ,test, slopeVar=Slope, rank_variable=Slope,n_sbj=n_sbj,"max") + 
     ggtitle("Slope = bandInt_RF+b_bandInt") } /
 {ggplot(indvSlopes, aes(x = Slope, color = condit)) + geom_density() } /
 gridExtra::tableGrob( indvSlopes |> group_by(condit) |> summarise(Mean=mean(Slope),Sd=sd(Slope)) ) + 
 # gt_temp(indvSlopes |> group_by(condit) |> summarise(Mean=mean(Slope),Sd=sd(Slope)) ) +
  plot_layout(heights=unit(c(16,6,2),c('cm'))) + plot_annotation("vx ~ bandInt + (1 + bandInt || gr(id, by = condit))  ")

```

```{r}
noInt_tab <- testAvg |> select(id,condit,vx,vb,bandInt) |> 
  left_join(indvDraws |> group_by(id,bandInt) |> median_qi(estimate,Intercept,Slope), by=join_by(id,bandInt)) |>
  mutate(resid=vx-estimate) |> relocate(vx,resid,.before=estimate) |>
  mutate(across(where(is.numeric), \(x) round(x,0)))

noInt_tab |> group_by(condit) |> summarise(mean(resid),sd(resid))
noInt_tab |> group_by(condit,id) |> summarise(m=mean(resid),s=sd(resid)) |> summarise(m=mean(m),s=sd(s))

noInt_tab |> DT::datatable(options = list(pageLength=15))
```

```{r fig.height=9,fig.width=10}
#| eval: false
random_effects |> group_by(id) |> median_hdi()

indv_coefs |> select(id,condit,Estimate.bandInt)
indv_coefs |> select(id,condit,Estimate.bandInt) |> group_by(condit) |> summarise(mean(Estimate.bandInt))

fixef(e1_vxBMM)
ranef(e1_vxBMM)$id[, ,"bandInt"] |> as_tibble(rownames="id") |> 
  left_join(e1Sbjs,by=join_by(id)) |>
  group_by(condit) |> summarise(mean(Estimate))

ranef(e1_vxBMM)$id[, ,"bandInt"] |> as_tibble(rownames="id") |> 
  left_join(e1Sbjs,by=join_by(id)) |> ggplot(aes(Estimate,color=condit)) + geom_density()

popSlope <- tidy(e1_vxBMM, effects = "fixed") |> filter(term=="bandInt") |> pull(estimate)
popInteraction <- tidy(e1_vxBMM, effects = "fixed") |> filter(term=="conditVaried:bandInt") |> pull(estimate)

e1Vx <- ranef(e1_vxBMM)$id[, ,"bandInt"] |> as_tibble(rownames="id") %>%
  left_join(e1Sbjs,., by=join_by(id)) |> select(id,condit,condit_dummy,Estimate) |> 
  mutate(pop=popSlope,Slope1=Estimate + pop) |> 
  left_join(indv_coefs |> select(id,Slope2=Estimate.bandInt), join_by(id)) |>
  mutate(Slope3=Estimate + pop + ((popInteraction + Estimate)*condit_dummy))

e1Vx |> group_by(condit) |> summarise(Slope_with_int=mean(Slope3), sd_Slope=sd(Slope3))

e1Vx |> group_by(condit) |> reframe(enframe(quantile(Slope3, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "Slope3")) |> 
  pivot_wider(names_from=quantile,values_from=Slope3,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

{ e1Vx |> ggplot(aes(Slope1,color=condit)) + geom_density() +ggtitle("RF_Estimate + Population_Est") } /
e1Vx |> ggplot(aes(Slope3,color=condit)) + geom_density() + 
  ggtitle("Slope3=Estimate + pop + ((popInteraction + Estimate)*condit_dummy)")

```

Examine slopes from different models

```{r fig.height=10,fig.width=9}
#| eval: false 
#| 
# vx ~ bandInt + (1 + bandInt || gr(id, by = condit)) 
bandOnly <- ranef(e1_testVx_grRF8)$id[, ,"bandInt"] |> as_tibble(rownames="id") |> 
  left_join(e1Sbjs,by=join_by(id)) |> select(id,condit,condit_dummy,starts_with("Estimate")) |>
 # group_by(condit) |> 
  #summarise(Estimate=mean(Estimate)) |>
  mutate(pop = tidy(e1_testVx_grRF8, effects = "fixed") |> filter(term=="bandInt") |> pull(estimate),
         Slope1=pop+Estimate,
         Slope2 = (coef(e1_testVx_grRF8)$id |> as_tibble(rownames="id") |> pull(Estimate.bandInt)))

# vx ~ condit + bandInt + (1 + bandInt | id) 
noInt <- ranef(e1_conditPlusBand_RF)$id[, ,"bandInt"] |> as_tibble(rownames="id") |> 
  left_join(e1Sbjs,by=join_by(id)) |> select(id,condit,condit_dummy,starts_with("Estimate")) |>
 # group_by(condit) |> 
  #summarise(Estimate=mean(Estimate)) |>
  mutate(pop = tidy(e1_conditPlusBand_RF, effects = "fixed") |> filter(term=="bandInt") |> pull(estimate),
         Slope1=pop+Estimate,
         Slope2 = (coef(e1_conditPlusBand_RF)$id |> as_tibble(rownames="id") |> pull(Estimate.bandInt)))

# vx ~ condit * bandInt + (1 + bandInt | gr(id, by = condit)) 
popSlope <- tidy(e1_testVxBand_grRF, effects = "fixed") |> filter(term=="bandInt") |> pull(estimate)
popInteraction <- tidy(e1_testVxBand_grRF, effects = "fixed") |> filter(term=="conditVaried:bandInt") |> pull(estimate)
vx_grRF <- ranef(e1_testVxBand_grRF)$id[, ,"bandInt"] |> as_tibble(rownames="id") |> 
  left_join(e1Sbjs,by=join_by(id)) |> select(id,condit,condit_dummy,starts_with("Estimate")) |>
 # group_by(condit) |> 
  #summarise(Estimate=mean(Estimate)) |>
  mutate(pop = tidy(e1_testVxBand_grRF, effects = "fixed") |> filter(term=="bandInt") |> pull(estimate),
         Slope1=pop+Estimate,
         Slope2 = (coef(e1_testVxBand_grRF)$id |> as_tibble(rownames="id") |> pull(Estimate.bandInt)), 
         Slope3=Estimate + pop + ((popInteraction + Estimate)*condit_dummy))

bandOnly |> group_by(condit) |> summarise(Estimate=mean(Slope1))
noInt |> group_by(condit) |> summarise(Estimate=mean(Slope1))
vx_grRF |> group_by(condit) |> summarise(Estimate=mean(Slope3))

modelsummary(list("Band Only" = e1_testVx_grRF8,
                  "No Interaction" = e1_conditPlusBand_RF,
                  "Full Model - Gr" = e1_testVxBand_grRF,
                  "Full Model" = e1_vxBMM),
             coef_map = c("b_Intercept"="Intercept","b_bandInt"="Band Slope", "b_conditVaried"="Condition",
                          'b_conditVaried:bandInt'="Interaction","sigma"="Sigma"),
             output="markdown",gof_map=NA,fmt=2)

{ bandOnly |> ggplot(aes(Slope1,color=condit)) + geom_density() + 
  ggtitle("vx ~ bandInt + (1 + bandInt || gr(id, by = condit))") } /
  { noInt |> ggplot(aes(Slope1,color=condit)) + geom_density() + 
  ggtitle("vx ~ condit + bandInt + (1 + bandInt | id)  ") } / 
{ vx_grRF |> ggplot(aes(Slope3,color=condit)) + geom_density() + 
  ggtitle("vx ~ condit * bandInt + (1 + bandInt | gr(id, by = condit)) ") } /
{ e1Vx |> ggplot(aes(Slope3,color=condit)) + geom_density() + 
  ggtitle("vx ~ condit * bandInt + (1 + bandInt | id) ") }

bandOnly |> group_by(condit) |> reframe(enframe(quantile(Slope1, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "Slope1")) |> 
  pivot_wider(names_from=quantile,values_from=Slope1,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> 
  pandoc.table(caption="vx ~ bandInt + (1 + bandInt || gr(id, by = condit))")

noInt |> group_by(condit) |> reframe(enframe(quantile(Slope1, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "Slope1")) |> 
  pivot_wider(names_from=quantile,values_from=Slope1,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> 
  pandoc.table(caption="vx ~ condit + bandInt + (1 + bandInt | id)  ")

vx_grRF |> group_by(condit) |> reframe(enframe(quantile(Slope3, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "Slope1")) |> 
  pivot_wider(names_from=quantile,values_from=Slope1,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> 
  pandoc.table(caption="vx ~ condit * bandInt + (1 + bandInt | gr(id, by = condit))  ")

e1Vx |> group_by(condit) |> reframe(enframe(quantile(Slope3, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "Slope3")) |> 
  pivot_wider(names_from=quantile,values_from=Slope3,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |>
  pandoc.table(caption="vx ~ condit * bandInt + (1 + bandInt | id)  ")

```

```{r}

```

```{r fig.width=11, fig.height=10}

# 
# indv_model_plot <- function(combined_df, indv_coefs, testAvg, rank_variable = "Estimate.Intercept", n_sbj = 5, order = "min") {
#   slice_fn <- if (order == "min") slice_min else slice_max
#     rank_variable_name <- deparse(substitute(rank_variable))
#   selected_ids <- indv_coefs %>%
#     slice_fn({{ rank_variable }}, n = n_sbj, by = condit) %>%
#     select(id, rank_var = {{ rank_variable }})
#   print(head(selected_ids))
#    cdf <- combined_df %>%
#     inner_join(selected_ids, by = join_by(id)) %>%
#     mutate(label = paste0("Subject ", id, "; ", rank_variable_name, "=", round(rank_var, 2)))
#     #print(head(cdf))
# 
#     cdf |>
#     group_by(id, bandInt)  |>
#     sample_n(50) %>%
#     ggplot(aes(x = bandInt, y = estimate)) +
#     geom_abline(aes(intercept = Intercept + b_Intercept, slope = bandInt_RF + b_bandInt), color = "grey50") +
#     geom_abline(data = indv_coefs %>%
#                   slice_fn({{ rank_variable }}, n = n_sbj, by = condit),
#                 aes(intercept = {{ rank_variable }}, slope = {{ rank_variable }}), color = "red") +
#     stat_halfeye() +
#     stat_halfeye(data = testAvg %>%
#                    filter(id %in% selected_ids$id),
#                  aes(x = bandInt, y = vx), color = "blue") +
#     ggh4x::facet_nested_wrap(vars(condit, label)) +
#        ggtitle(glue("Order {n_sbj} Individuals for {rank_variable_name}"))
# 
# }

```

```{r fig.width=11, fig.height=10}
#| eval: false

combined_df |> indv_model_plot(indv_coefs, testAvg, rank_variable=Estimate.Intercept,n_sbj=6,"min")
combined_df |> indv_model_plot(indv_coefs, testAvg, rank_variable=Estimate.Intercept,n_sbj=6,"max")
combined_df |> indv_model_plot(indv_coefs, testAvg, rank_variable=Estimate.bandInt,n_sbj=6,"min")
combined_df |> indv_model_plot(indv_coefs, testAvg, rank_variable=Estimate.bandInt,n_sbj=6,"max")

combined_df |> indv_model_plot(indv_coefs, testAvg, rank_variable=Est.Error.Intercept,n_sbj=6,"min")
combined_df |> indv_model_plot(indv_coefs, testAvg, rank_variable=Est.Error.Intercept,n_sbj=6,"max")

## Sort by intercept 
combined_df |> 
  filter(id %in% (indv_coefs |> slice_min(Estimate.Intercept,n=5,by=condit) |> pull(id))) |> 
  group_by(id, bandInt) |>
  sample_n(50) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  geom_abline(data=indv_coefs |> slice_min(Estimate.Intercept,n=5,by=condit),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt,color="red")) +
  stat_halfeye() +
  stat_halfeye(data=testAvg |> filter(id %in% (indv_coefs |> slice_min(Estimate.Intercept,n=5,by=condit) |> pull(id))), aes(x=bandInt,y=vx),color="blue") +
  ggh4x::facet_nested_wrap(vars(condit,id))

indv_coefs |> ggplot(aes(y=id, x=Estimate.bandInt)) + geom_pointrange()

cd <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt) |>
  mutate(Slope=bandInt_RF+b_bandInt) |> group_by(id)

cdMed <- cd |> group_by(id) |> median_qi(Slope) |> mutate(rankSlope=rank(Slope)) |> arrange(rankSlope) |> left_join(e1Sbjs, by=join_by(id))

cdMed %>% ggplot(aes(y=rankSlope, x=Slope,fill=condit,color=condit)) + 
  geom_pointrange(aes(xmin=.lower , xmax=.upper)) + 
  labs(x="Estimated Slope", y="Participant") # + facet_wrap(~condit) +

cd |>  ggplot(aes(y=id, x=Slope)) + geom_pointrange()
```

```{r fig.width=12, fig.height=11}
#| eval: false
coef(e1_vxBMM)$id %>% as_tibble(rownames="id") %>% select(id, starts_with("Est")) |> print(n=15)

n_lines <- 20
f <- fitted(e1_vxBMM, newdata = new_data_grid, re_formula = NA, summary=F, ndraws=n_lines) |> 
  as_tibble() %>% 
  mutate(draw = 1:n()) %>% 
  pivot_longer(-draw)

combined_df |> 
  filter(id %in% top_indv_coefs$id) |> 
  group_by(id, bandInt) |>
  sample_n(50) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  geom_abline(data=indv_coefs |> filter(id %in% top_indv_coefs$id),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt,color="red")) +
  stat_halfeye() +
  stat_halfeye(data=testAvg |> filter(id %in% top_indv_coefs$id), aes(x=bandInt,y=vx),color="blue") +
  facet_wrap(~id)

## Sort by intercept 
combined_df |> 
  filter(id %in% (indv_coefs |> slice_min(Estimate.Intercept,n=5,by=condit) |> pull(id))) |> 
  group_by(id, bandInt) |>
  sample_n(50) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  geom_abline(data=indv_coefs |> slice_min(Estimate.Intercept,n=5,by=condit),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt,color="red")) +
  stat_halfeye() +
  stat_halfeye(data=testAvg |> filter(id %in% (indv_coefs |> slice_min(Estimate.Intercept,n=5,by=condit) |> pull(id))), aes(x=bandInt,y=vx),color="blue") +
  ggh4x::facet_nested_wrap(vars(condit,id))

## Sort by intercept fit error
combined_df |> 
  filter(id %in% (indv_coefs |> slice_min(Est.Error.Intercept,n=5,by=condit) |> pull(id))) |> 
  group_by(id, bandInt) |>
  sample_n(50) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  geom_abline(data=indv_coefs |> slice_min(Est.Error.Intercept,n=5,by=condit),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt,color="red")) +
  stat_halfeye() +
  stat_halfeye(data=testAvg |> filter(id %in% well_fit$id), aes(x=bandInt,y=vx),color="blue") +
  ggh4x::facet_nested_wrap(vars(condit,id))

combined_df |> 
  filter(id %in% (indv_coefs |> filter(rank<=2) |> pull(id)) )  |> 
  group_by(id, bandInt) |>
  sample_n(50) |>
  ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  geom_abline(data=indv_coefs |> filter(id %in% (indv_coefs |> filter(rank<=2) |> pull(id))),aes(intercept=Estimate.Intercept,slope=Estimate.bandInt,color="red")) +
  stat_halfeye() +
  stat_halfeye(data=testAvg |> filter(id %in% (indv_coefs |> filter(rank<=2) |> pull(id))), aes(x=bandInt,y=vx),color="blue") +
  ggh4x::facet_nested_wrap(vars(condit,id))

```

```{r}
#| eval: false

indv_coefs |> 
  filter(id %in% top_indv_coefs$id) |> ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  stat_halfeye()

all_effects <- e1_vxBMM |> 
  gather_draws(b_Intercept, b_conditVaried, b_bandInt, `b_conditVaried:bandInt`, `^r_id.*$`, regex = TRUE, ndraws = 1)

head(all_effects)

 nd <- e1_vxBMM |>  
   spread_draws(b_Intercept, b_bandInt, `b_conditVaried`,`b_conditVaried:bandInt`,  r_id[id,term], ndraws=10)

```

```{r}
#| eval: false
df_pred <- 
  posterior_predict(e1_distBMM, ndraws = 500) |>
  array_branch(margin=1) |> 
   map_dfr( 
    function(yrep_iter) {
      test %>%
        mutate(dist_pred = yrep_iter)
    },
    .id = 'iter'
  ) |>
  mutate(iter = as.numeric(iter))

df_pred %>% filter(id %in% 1:162 ) |>
  filter(iter < 100) %>%
  ggplot(aes(dist_pred, group = iter)) +
  geom_line(alpha = .05, stat = 'density', color = 'blue') +
  geom_density(data = test |> filter(id %in% 1:162),
               aes(dist,col=vb),
               inherit.aes = FALSE,
               size = 0.8) + # 1
  facet_grid(condit ~ vb) +
  xlab('vx')

df_pred %>% filter(id %in% 1:10 ) |>
  filter(iter < 100) %>%
  ggplot(aes(dist_pred, group = iter)) +
  geom_line(alpha = .05, stat = 'density', color = 'blue') +
  geom_density(data = test |> filter(id %in% 1:10),
               aes(dist),
               inherit.aes = FALSE,
               size = 0.5) + # 1
  facet_wrap(id ~ .) +
  xlab('vx')
```

```{r}
#| eval: false
vx_pred <- 
  posterior_predict(e1_vxBMM, ndraws = 500) |> 
  array_branch(margin=1) |> 
   map_dfr( 
    function(yrep_iter) {
      test %>%
        mutate(vx_pred = yrep_iter)
    },
    .id = 'iter'
  ) %>%
  mutate(iter = as.numeric(iter))

vx_pred %>% filter(id %in% 1:2 ) |>
  filter(iter < 100) %>%
  ggplot(aes(vx_pred, group = iter)) +
  geom_line(alpha = .05, stat = 'density', color = 'blue') +
  geom_density(data = test |> filter(id %in% 1:2),
               aes(vx,col=vb),
               inherit.aes = FALSE,
               size = 0.5) + # 1
  facet_grid(id ~ vb) +
  xlab('vx')

vx_pred  |>
  filter(iter < 100) %>%
  ggplot(aes(vx_pred, group = iter)) +
  geom_line(alpha = .05, stat = 'density', color = 'blue') +
  geom_density(data = test,
               aes(vx,col=vb),
               inherit.aes = FALSE,
               size = 0.5) + # 1
  facet_grid(condit ~ vb) +
  xlab('vx')

vx_pred %>% filter(id %in% 1:9 ) |>
  filter(iter < 100) %>%
  ggplot(aes(vx_pred, group = iter)) +
  geom_line(alpha = .05, stat = 'density', color = 'blue') +
  geom_density(data = test |> filter(id %in% 1:9),
               aes(vx),
               inherit.aes = FALSE,
               size = 0.5) + # 1
  facet_wrap(id ~ .,ncol=3) +
  xlab('vx')

```
```

```{r}
#| eval: false

indv_coefs |> 
  filter(id %in% top_indv_coefs$id) |> ggplot(aes(x=bandInt,y=estimate)) + 
  geom_abline(aes(intercept=Intercept+b_Intercept, slope=bandInt_RF+b_bandInt), color="grey50") +
  stat_halfeye()

all_effects <- e1_vxBMM |> 
  gather_draws(b_Intercept, b_conditVaried, b_bandInt, `b_conditVaried:bandInt`, `^r_id.*$`, regex = TRUE, ndraws = 1)

head(all_effects)

 nd <- e1_vxBMM |>  
   spread_draws(b_Intercept, b_bandInt, `b_conditVaried`,`b_conditVaried:bandInt`,  r_id[id,term], ndraws=10)

```

```{r}
#| eval: false
df_posterior <- e1_vxBMM |> as_tibble()

ggplot(df_posterior) + 
  aes(x = b_Intercept, y = b_bandInt) + 
  stat_density_2d( geom = "polygon")

df_posterior |> sample_n(100) |> ggplot() + 
  aes(x = b_Intercept, y = b_bandInt) + 
  geom_raster(interpolate = T)

df_posterior |> sample_n(100) |> ggplot(aes(x = b_Intercept, y = b_bandInt)) + 
  stat_ellipse(geom = "polygon", level = 0.1, alpha = 1/2)

df_posterior |> 
  sample_n(1000) |>
  ggplot() + 
  aes(x = b_Intercept, y = b_bandInt) + 
  stat_dist_dotsinterval()

```