### 2023-08-31 Analysis/e1_test.qmd
ordinal models

---
title: "HTW E1 Testing"
categories: [Analysis, R, Bayesian]
---

```{r setup, include=FALSE}
source(here::here("Functions", "packages.R"))
test <- readRDS(here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Test") 
options(brms.backend="cmdstanr",mc.cores=4)
e1Sbjs <- test |> group_by(id,condit) |> summarise(n=n())
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)
```

## Analyses Strategy

To assess differences between groups, we used Bayesian Mixed Effects
Regression. Model fitting was performed with the brms package in R
@burknerBrmsPackageBayesian2017, and descriptive stats and tables were
extracted with the BayestestR package
@makowskiBayestestRDescribingEffects2019a. Mixed effects regression
enables us to take advantage of partial pooling, simultaneously
estimating parameters at the individual and group level. Our use of
Bayesian, rather than frequentist methods allows us to directly quantify
the uncertainty in our parameter estimates, as well as circumventing
convergence issues common to the frequentist analogues of our mixed
models. For each model, we report the median values of the posterior
distribution, and 95% credible intervals.

Each model was set to run with 4 chains, 5000 iterations per chain, with
the first 2500 of which were discarded as warmup chains. Rhat values
were generally within an acceptable range, with values \<=1.02 (see
appendix for diagnostic plots). We used uninformative priors for the
fixed effects of the model (condition and velocity band), and weakly
informative Student T distributions for for the random effects.

We compared varied and constant performance across two measures,
deviation and discrimination. Deviation was quantified as the absolute
deviation from the nearest boundary of the velocity band, or set to 0 if
the throw velocity fell anywhere inside the target band. Thus, when the
target band was 600-800, throws of 400, 650, and 1100 would result in
deviation values of 200, 0, and 300, respectively. Discrimination was
measured by fitting a linear model to the testing throws of each
subjects, with the lower end of the target velocity band as the
predicted variable, and the x velocity produced by the participants as
the predictor variable. Participants who reliably discriminated between
velocity bands tended to have positive slopes with values \~1, while
participants who made throws irrespective of the current target band
would have slopes \~0.

{{< include ../Misc/bmm_table.qmd >}}

## Results

### Testing Phase - No feedback.

In the first part of the testing phase, participants are tested from
each of the velocity bands, and receive no feedback after each throw.

#### Deviation From Target Band

Descriptive summaries testing deviation data are provided in
@tbl-e1-test-nf-deviation and @fig-e1-test-dev. To model differences in
accuracy between groups, we used Bayesian mixed effects regression
models to the trial level data from the testing phase. The primary model
predicted the absolute deviation from the target velocity band (dist) as
a function of training condition (condit), target velocity band (band),
and their interaction, with random intercepts and slopes for each
participant (id).

```{=tex}
\begin{equation}
dist_{ij} = \beta_0 + \beta_1 \cdot condit_{ij} + \beta_2 \cdot band_{ij} + \beta_3 \cdot condit_{ij} \cdot band_{ij} + b_{0i} + b_{1i} \cdot band_{ij} + \epsilon_{ij}
\end{equation}
```
```{r}
#| label: tbl-e1-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]
#| layout-ncol: 2
#| echo: fenced

result <- test_summary_table(test, "dist", mfun = list(mean = mean, median = median, sd = sd))
result$constant
result$varied
```

```{r}
#| label: fig-e1-test-dev
#| fig-cap: E1. Deviations from target band during testing without feedback stage. 
test |>  ggplot(aes(x = vb, y = dist,fill=condit)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
  labs(x="Band", y="Deviation From Target")
```

The model predicting absolute deviation (dist) showed clear effects of
both training condition and target velocity band (Table X). Overall, the
varied training group showed a larger deviation relative to the constant
training group (β = 135, 95% CI \[43, 231\]). Deviation also depended on
target velocity band, with lower bands showing less deviation. See
@tbl-e1-bmm-dist for full model output.

```{r}
#| label: tbl-e1-bmm-dist
#| tbl-cap: "Experiment 1. Bayesian Mixed Model predicting absolute deviation as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Model Coefficients", "Contrasts"]
#| layout-ncol: 2
#| echo: fenced
#| column: page-inset-right

#contrasts(test$condit) 
#contrasts(test$vb)
modelName <- "e1_testDistBand_RF_5K"
e1_distBMM <- brm(dist ~ condit * bandInt + (1 + bandInt|id),
                      data=test,file=paste0(here::here("data/model_cache",modelName)),
                      iter=5000,chains=4)
GetModelStats(e1_distBMM) |> kable(escape=F,booktabs=T)

e1_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |> 
   summarize(median_qi(.value),pd=sum(.value>0)/n()) |>
   select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
   mutate(across(where(is.numeric), \(x) round(x, 2)),
          pd=ifelse(value<0,1-pd,pd)) |>
   kbl()

```

```{r}
#| label: fig-e1-bmm-dist
#| fig-cap: E1. Conditioinal Effect of Training Condition and Band. Ribbon indicated 95% Credible Intervals.

condEffects <- function(m){
  m |> ggplot(aes(x = bandInt, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() + stat_halfeye(alpha=.2) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  ylab("Predicted X Velocity") + xlab("Band")
}

e1_distBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |>
 condEffects()+
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(test$vb), 
                     limits = c(0, 1400)) 
```

#### Discrimination between bands

```{r}
#| label: fig-e1-test-vx
#| fig-cap: E1 testing x velocities. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 11
#| fig-height: 9
test %>% group_by(id,vb,condit) |> plot_distByCondit()

```

```{r}
#| label: tbl-e1-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 2
#| echo: fenced

result <- test_summary_table(test, "vx", mfun = list(mean = mean, median = median, sd = sd))
result$constant 
result$varied

```

```{r}
#| label: tbl-e1-bmm-vx
#| tbl-cap: "Experiment 1. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Model fit to all 6 bands", "Model fit to 3 extrapolation bands"]
#| layout-ncol: 2
#| echo: fenced
e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
GetModelStats(e1_vxBMM ) |> kable(escape=F,booktabs=T)

modelName <- "e1_extrap_testVxBand"
e1_extrap_VxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                  data=test |>
                    filter(expMode=="test-Nf"),file=paste0(here::here("data/model_cache",modelName)),
                  iter=5000,chains=4)
GetModelStats(e1_extrap_VxBMM ) |> kable(escape=F,booktabs=T)
```

See @tbl-e1-bmm-vx for the full model results. The estimated coefficient
for training condition (β = 164.67, 95% CrI [45.50, 278.85]) suggests
that the varied group tends to produce harder throws than the constant
group, but is not in and of itself useful for assessing discrimination.
Most relevant to the issue of discrimination is the slope on Velocity
Band (β = 0.71, 95% CrI [.62, .80]). Although the median slope does fall
underneath the ideal of value of 1, the fact that the 95% credible
interval does not contain 0 provides strong evidence that participants
exhibited some discrimination between bands. The slope coefficient (β =
0.71 ) also seems in accordance with an examination of
@tbl-e1-test-nf-vx, which shows that average x-velocity fell within the
range of the target band for four out of six of the bands ($4/6$=.67).
The estimate for the interaction between slope and condition (β = -.12,
95% CI [-.25, 0.00]), suggests that the varied condition had
significantly lower slopes than did the constant group. This difference
is depicted visually in @fig-e1-bmm-vx.

```{r}
#| label: tbl-vx-emmeans
#| tbl-cap: Conditional Means and Contrasts
#| tbl-subcap: ["6 Band Model Contrasts", "3 Band Model Contrasts"]
#| layout-ncol: 2
#| echo: fenced
#| column: page-inset-right

e1_vxBMM  |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> 
  gather_emmeans_draws()  |> 
  summarize(pd=sum(.value>0)/n(), median_hdi(.value) ) |> 
  select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
  mutate(pd=ifelse(value<0,1-pd,pd),across(where(is.numeric), \(x) round(x, 2))) |>
  kbl()

e1_extrap_VxBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> 
  gather_emmeans_draws()  |> 
  summarize(pd=sum(.value>0)/n(), median_hdi(.value) ) |> 
  select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
  mutate(pd=ifelse(value<0,1-pd,pd),across(where(is.numeric), \(x) round(x, 2))) |>
  kbl()

```

```{r}
#| label: fig-e1-bmm-vx
#| fig-cap: Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands. 
#| fig-subcap: ["Model fit to all 6 bands", "Model fit to only 3 extrapolation bands"]
#| layout-ncol: 2

e1_vxBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects() +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(test$vb), 
                     limits = c(0, 1400))

e1_extrap_VxBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600))) |>
  gather_emmeans_draws() |>
  condEffects() +
  scale_x_continuous(breaks = c(100, 350, 600), 
                     labels = levels(test$vb)[1:3], 
                     limits = c(0, 1000))

```

```{r}
#| label: tbl-e1-slope-quartile
#| tbl-cap: "Mean slopes for each quartile, by condition"

new_data_grid=map_dfr(1, ~data.frame(unique(test[,c("id","condit","bandInt")]))) |> 
  dplyr::arrange(id,bandInt) |> 
  mutate(condit_dummy = ifelse(condit == "Varied", 1, 0))

indv_coefs <- coef(e1_vxBMM)$id |> 
  as_tibble(rownames="id") |> 
  select(id, starts_with("Est")) |>
  left_join(e1Sbjs, by=join_by(id) )

fixed_effects <- e1_vxBMM |> 
  spread_draws(`^b_.*`,regex=TRUE) |> arrange(.chain,.draw,.iteration)

random_effects <- e1_vxBMM |> 
  gather_draws(`^r_id.*$`, regex = TRUE, ndraws = 1500) |> 
  separate(.variable, into = c("effect", "id", "term"), sep = "\\[|,|\\]") |> 
  mutate(id = factor(id,levels=levels(test$id))) |> 
  pivot_wider(names_from = term, values_from = .value) |> arrange(id,.chain,.draw,.iteration)

 indvDraws <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt,RF_Intercept=Intercept) |>
  right_join(new_data_grid, by = join_by("id")) |> 
  mutate(
    Slope = bandInt_RF+b_bandInt,
    Intercept= RF_Intercept + b_Intercept,
    estimate = (b_Intercept + RF_Intercept) + (bandInt*(b_bandInt+bandInt_RF)) + (bandInt * condit_dummy) * `b_conditVaried:bandInt`,
  )

  indvSlopes <- indvDraws |> group_by(id) |> median_qi(Slope,SlopeInt,SlopeInt2, Intercept,b_Intercept,b_bandInt) |>
  left_join(e1Sbjs, by=join_by(id)) |> group_by(condit) |>
    select(id,condit,condit_dummy,Intercept,b_Intercept,Slope,SlopeInt,SlopeInt2,b_bandInt, n) |>
  mutate(rankSlope=rank(Slope)) |> arrange(rankSlope)   |> ungroup()

  indvSlopes |> group_by(condit) |> reframe(enframe(quantile(Slope, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "Slope")) |> 
  pivot_wider(names_from=quantile,values_from=Slope,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

  indvSlopes |> reframe(enframe(quantile(SlopeInt, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "SlopeInt")) |> 
  pivot_wider(names_from=quantile,values_from=SlopeInt,names_prefix="Q_") |>
  group_by(condit) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

 ggplot(indvSlopes, aes(x = Slope, color = condit)) + 
  geom_density()

```

```{r}
#| label: fig-e1-bmm-bx2
#| fig-cap: Slope estimates by participant - ordered from lowest to highest within each condition. 
#| fig-height: 9
#| fig-width: 10

  indvSlopes |> ggplot(aes(y=rankSlope, x=Slope,fill=condit,color=condit)) + 
  geom_pointrange(aes(xmin=.lower , xmax=.upper)) + 
  labs(x="Estimated Slope", y="Participant")  + facet_wrap(~condit)

```

```{r}
#| label: fig-e1-indv-slopes
#| fig-cap: Subset of Varied and Constant Participants with the smallest and largest estimated slope values. Red lines represent the best fitting line for each participant, gray lines are 200 random samples from the posterior distribution. Colored points and intervals at each band represent the empirical median and 95% HDI. 
#| fig-subcap: ["subset with largest slopes", "subset with smallest slopes"]
#| fig-height: 11
#| fig-width: 12
nSbj <- 3
indvDraws  |> indv_model_plot(indv_coefs, testAvg, Slope,rank_variable=Estimate.bandInt,n_sbj=nSbj,"max")
indvDraws |> indv_model_plot(indv_coefs, testAvg,Slope, rank_variable=Estimate.bandInt,n_sbj=nSbj,"min")

```

#### E1 Results Discussion

NEEDS TO BE WRITTEN

```{r}
#| label: Band Deviation
#| tbl-cap: "Deviations from target"
#| layout: [[70, 30],[-5]]

```

## References

::: {#refs}
:::

## Appendix

{{< include Appendix/E1_Appendix.qmd >}}