### 2023-10-09 Analysis/e3_test.qmd

---
title: "HTW E3 Testing"
categories: [Analysis, R, Bayesian]
---

```{r setup, include=FALSE}
source(here::here("Functions", "packages.R"))
testE3 <- readRDS(here("data/e3_08-04-23.rds")) |> filter(expMode2 == "Test") 
options(brms.backend="cmdstanr",mc.cores=4)
e3Sbjs <- testE3 |> group_by(id,condit) |> summarise(n=n())
testE3Avg <- testE3 %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)
```

The major manipulation adjustment of experiment 3 is for participants to receive ordinal feedback, in contrast to the continuous feedback of the earlier experiments. Experiment 3 participants were randomly assigned to both a training condition (Constant vs. Varied) and a Band Order condition (original order used in Experiment 1, or the Reverse order of Experiment 2).

adlkdj

## Results

### Testing Phase - No feedback.

In the first part of the testing phase, participants are tested from each of the velocity bands, and receive no feedback after each throw. Note that these no-feedback testing trials are identical to those of Experiment 1 and 2, as the ordinal feedback only occurs during the training phase, and final testing phase, of Experiment 3.

#### Deviation From Target Band

Descriptive summaries testing deviation data are provided in @tbl-e3-test-nf-deviation and @fig-e3-test-dev. 
To model differences in accuracy between groups, we used Bayesian mixed effects regression models to the trial level data from the testing phase. The primary model predicted the absolute deviation from the target velocity band (dist) as a function of training condition (condit), target velocity band (band), and their interaction, with random intercepts and slopes for each participant (id).

\begin{equation}
dist_{ij} = \beta_0 + \beta_1 \cdot condit_{ij} + \beta_2 \cdot band_{ij} + \beta_3 \cdot condit_{ij} \cdot band_{ij} + b_{0i} + b_{1i} \cdot band_{ij} + \epsilon_{ij}
\end{equation}

**need to make E3 tables differently since ppts do both original and reverse order**

```{r}
#| label: tbl-e3-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]
#| layout-ncol: 2
#| echo: fenced

resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant 
resultOrig$varied

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant 
resultRev$varied

```

```{r}
#| label: fig-e3-test-dev
#| fig-cap: e3. Deviations from target band during testing without feedback stage. 
testE3 |>  ggplot(aes(x = vb, y = dist,fill=condit)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
  labs(x="Band", y="Deviation From Target") + facet_wrap(~bandOrder)
```

```{r}
#| label: tbl-e3-bmm-dist
#| tbl-cap: "Experiment 3. Bayesian Mixed Model predicting absolute deviation as a function of condition (Constant vs. Varied) and Velocity Band"
#contrasts(test$condit)

# contrasts(testE3$vb)

modelName <- "e3_testDistBand_RF_5K"
e3_distBMM <- brm(dist ~ condit * bandInt + (1 + bandInt|id),
                      data=testE3,file=paste0(here::here("data/model_cache",modelName)),
                      iter=5000,chains=4)
mp3 <- GetModelStats(e3_distBMM) |> kable(escape=F,booktabs=T)
mp3

cd1 <- get_coef_details(e3_distBMM, "conditVaried")
sc1 <- get_coef_details(e3_distBMM, "bandInt")
intCoef1 <- get_coef_details(e3_distBMM, "conditVaried:bandInt")
```

In experiment 3, the overall effect of training condition shows a similar pattern to that of Experiment 2, with a  moderately strong effect tendency for the varied group to show lower deviation (β = `r cd1$estimate`, 95% CrI \[`r cd1$conf.low`, `r cd1$conf.high`\]), with 97% of the posterior distribution falling under 0. (NEED TO CONTROL FOR BAND ORDER HERE)

```{r}
#| label: fig-e3-bmm-dist
#| fig-cap: e3. Conditioinal Effect of Training Condition and Band. Ribbon indicated 95% Credible Intervals.

e3_distBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |>
  ggplot(aes(x = bandInt, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
    ylab("Predicted Deviation") + xlab("Velocity Band")+
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE3$vb), 
                     limits = c(0, 1400)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) 
```

#### Discrimination between Velocity Bands

In addition to accuracy/deviation. We also assessed the ability of participants to reliably discriminate between the velocity bands (i.e. responding differently when prompted for band 600-800 than when prompted for band 150-350). @tbl-e3-test-nf-vx shows descriptive statistics of this measure, and Figure 1 visualizes the full distributions of throws for each combination of condition and velocity band. To quantify discrimination, we again fit Bayesian Mixed Models as above, but this time the dependent variable was the raw x velocity generated by participants.

\begin{equation}
vx_{ij} = \beta_0 + \beta_1 \cdot condit_{ij} + \beta_2 \cdot bandInt_{ij} + \beta_3 \cdot condit_{ij} \cdot bandInt_{ij} + b_{0i} + b_{1i} \cdot bandInt_{ij} + \epsilon_{ij}
\end{equation}

```{r}
#| label: fig-e3-test-vx
#| fig-cap: e3 testing x velocities. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 11
#| fig-height: 9
# testE3 |> filter(bandOrder=="Original")|> group_by(id,vb,condit) |> plot_distByCondit()
# testE3 |> filter(bandOrder=="Reverse")|> group_by(id,vb,condit) |> plot_distByCondit() +ggtitle("test")

testE3 |> group_by(id,vb,condit,bandOrder) |> plot_distByCondit() + facet_wrap(bandOrder~condit,scale="free_x")

```

**need to make E3 tables differently since ppts do both original and reverse order**

```{r}
#| label: tbl-e3-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 2
#| echo: fenced

resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant 
resultOrig$varied

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant 
resultRev$varied 
```

```{r}
#| label: tbl-e3-bmm-vx
#| tbl-cap: "Experiment 3. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
e3_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=testE3,file=paste0(here::here("data/model_cache", "e3_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
mt4 <-GetModelStats(e3_vxBMM ) |> kable(escape=F,booktabs=T)
mt4

cd1 <- get_coef_details(e3_vxBMM, "conditVaried")
sc1 <- get_coef_details(e3_vxBMM, "bandInt")
intCoef1 <- get_coef_details(e3_vxBMM, "conditVaried:bandInt")
```

See @tbl-e3-bmm-vx for the full model results. The estimated coefficient for training condition (β = `r cd1$estimate`, 95% CrI \[`r cd1$conf.low`, `r cd1$conf.high`\]) suggests that the varied group tends to produce softer throws than the constant group. More relevant to the issue of discrimination is the slope on Velocity Band (β = `r sc1$estimate`, 95% CrI \[`r sc1$conf.low`, `r sc1$conf.high`\]).
Interaction: (β = `r intCoef1$estimate`, 95% CrI \[`r intCoef1$conf.low`, `r intCoef1$conf.high`\])

It's also noteworthy that the difference between the E3 slope and E1 & E2 slope is quite strong, with the upper end of the 95% CI for E3 falling underneath the lower end of the 95% CI for both of the earlier experiments.

```{r}
#| label: fig-e3-bmm-vx
#| fig-cap: Conditional effect of training condition and Band. Ribbons indicate 95% HDI.

e3_vxBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |>
  ggplot(aes(x = bandInt, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  ylab("Predicted X Velocity") + xlab("Band")+
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE3$vb), 
                     limits = c(0, 1400)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) 
```

```{r}
#| label: tbl-e3-slope-quartile
#| tbl-cap: "Slope coefficients by quartile, per condition"

new_data_grid=map_dfr(1, ~data.frame(unique(testE3[,c("id","condit","bandInt")]))) |> 
  dplyr::arrange(id,bandInt) |> 
  mutate(condit_dummy = ifelse(condit == "Varied", 1, 0))

indv_coefs <- as_tibble(coef(e3_vxBMM)$id, rownames="id")|> 
  select(id, starts_with("Est")) |>
  left_join(e3Sbjs, by=join_by(id) )

fixed_effects <- e3_vxBMM |> 
  spread_draws(`^b_.*`,regex=TRUE) |> arrange(.chain,.draw,.iteration)

random_effects <- e3_vxBMM |> 
  gather_draws(`^r_id.*$`, regex = TRUE, ndraws = 1500) |> 
  separate(.variable, into = c("effect", "id", "term"), sep = "\\[|,|\\]") |> 
  mutate(id = factor(id,levels=levels(testE3$id))) |> 
  pivot_wider(names_from = term, values_from = .value) |> arrange(id,.chain,.draw,.iteration)

 indvDraws <- left_join(random_effects, fixed_effects, by = join_by(".chain", ".iteration", ".draw")) |> 
  rename(bandInt_RF = bandInt,RF_Intercept=Intercept) |>
  right_join(new_data_grid, by = join_by("id")) |> 
  mutate(
    Slope = bandInt_RF+b_bandInt,
    Intercept= RF_Intercept + b_Intercept,
    estimate = (b_Intercept + RF_Intercept) + (bandInt*(b_bandInt+bandInt_RF)) + (bandInt * condit_dummy) * `b_conditVaried:bandInt`,
    SlopeInt = Slope + (`b_conditVaried:bandInt`*condit_dummy)
  )

  indvSlopes <- indvDraws |> group_by(id) |> median_qi(Slope,SlopeInt, Intercept,b_Intercept,b_bandInt) |>
  left_join(e3Sbjs, by=join_by(id)) |> group_by(condit) |>
    select(id,condit,Intercept,b_Intercept,starts_with("Slope"),b_bandInt, n) |>
  mutate(rankSlope=rank(Slope)) |> arrange(rankSlope)   |> ungroup()

  indvSlopes |> mutate(Condition=condit) |>  group_by(Condition) |> 
    reframe(enframe(quantile(SlopeInt, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "SlopeInt")) |> 
  pivot_wider(names_from=quantile,values_from=SlopeInt,names_prefix="Q_") |>
  group_by(Condition) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

```

@fig-e3-bmm-bx2 shows the distributions of estimated slopes relating velocity band to x velocity for each participant, ordered from lowest to highest within condition. Slope values are lower overall for varied training compared to constant training. Figure Xb plots the density of these slopes for each condition. The distribution for varied training has more mass at lower values than the constant training distribution. Both figures illustrate the model's estimate that varied training resulted in less discrimination between velocity bands, evidenced by lower slopes on average.

```{r}
#| label: fig-e3-bmm-bx2
#| fig-cap: Slope distributions between condition
#| fig-subcap: ["Slope estimates by participant - ordered from lowest to highest within each condition. ", "Destiny of slope coefficients by training group"]
#| fig-height: 11
#| fig-width: 12

indvSlopes |> ggplot(aes(y=rankSlope, x=SlopeInt,fill=condit,color=condit)) + 
  geom_pointrange(aes(xmin=SlopeInt.lower , xmax=SlopeInt.upper)) + 
  labs(x="Estimated Slope", y="Participant")  + facet_wrap(~condit) +
ggplot(indvSlopes, aes(x = SlopeInt, color = condit)) + 
geom_density() + labs(x="Slope Coefficient",y="Density")

```

```{r}
#| label: fig-e3-indv-slopes
#| fig-cap: Subset of Varied and Constant Participants with the smallest and largest estimated slope values. Red lines represent the best fitting line for each participant, gray lines are 200 random samples from the posterior distribution. Colored points and intervals at each band represent the empirical median and 95% HDI. 
#| fig-subcap: ["subset with largest slopes", "subset with smallest slopes"]
#| fig-height: 11
#| fig-width: 12
nSbj <- 3
indvDraws  |> indv_model_plot(indvSlopes, testE3Avg, SlopeInt,rank_variable=Slope,n_sbj=nSbj,"max")
indvDraws |> indv_model_plot(indvSlopes, testE3Avg,SlopeInt, rank_variable=Slope,n_sbj=nSbj,"min")

```