### 2023-10-23 Misc/Intro.qmd
rearrangements

# Introduction

In project 1, I applied model-based techniques to quantify and control for the similarity between training and testing experience, which in turn enabled us to account for the difference between varied and constant training via an extended version of a similarity based generalization model. In project 2, we will go a step further, implementing a full process model capable of both 1) producing novel responses and 2) modeling behavior in both the learning and testing stages of the experiment. Project 2 also places a greater emphasis on extrapolation performance following training - as varied training has often been purported to be particularly beneficial in such situations. Extrapolation has long been a focus of the literature on function learning [@brehmerHypothesesRelationsScaled1974; @carrollFunctionalLearningLearning1963]. Central questions of the function learning literature have included the relative difficulties of learning various functional forms (e.g. linear vs.bilinear vs. quadratic), and the relative effectiveness of rule-based vs. association-based exemplar models vs. various hybrid models [@bottNonmonotonicExtrapolationFunction2004; @deloshExtrapolationSineQua1997; @jonesActiveFunctionLearning2018; @kalishPopulationLinearExperts2004; @mcdanielConceptualBasisFunction2005; @mcdanielPredictingTransferPerformance2009]. However the issue of training variation has received surprisingly little attention in this area.

The study of human function learning investigates how people learn to map continuous input values to continuous output values based on examples. Early pioneering work by Carroll (1963) proposed that individuals abstract functional rules from examples to guide generalization to novel inputs. In his experiments, participants associated stimulus magnitudes to response magnitudes that followed linear, quadratic, or random functions. Results showed that people could extrapolate responses to novel inputs by applying the underlying rule, rather than simply matching responses to the most similar input as proposed by associative models.

Building on Carroll's ideas, Brehmer (1974) compared how well people could learn different types of functions, finding that linear functions were easier to acquire than nonlinear functions. Brehmer proposed a hypothesis testing model whereby learners begin with an expectation of a positive linear trend and adjust their functional hypothesis given sufficient evidence.

More recently, Koh and Meyer (1991) developed an adaptive regression model that incorporated a simplicity bias to account for patterns in early learning. Their model assumes learners treat stimuli and responses as logarithmically scaled and acquire polynomial rules relating the subjective stimulus and response scales.

Together, these early rule-based accounts proposed that people do not simply associate inputs to outputs but infer abstract rules to guide generalization. While subsequent work revealed limits in pure rule-based models, the pioneering work of Carroll, Brehmer, and Koh established the importance of rules and abstraction in theories of human function learning.

The study of human function learning delves into the process of learning by which individuals navigate the mapping of continuous input values to continuous output values, based on an array of examples. Foundational studies, such as those carried out by Busemeyer, Byun, DeLosh, and McDaniel in 1997, have revealed a hierarchy in function learning difficulty. Positive linear functions emerge as the most readily learned, closely followed by negative linear functions, monotonic nonlinear functions, and then non-monotonic functions. This sequence of learning difficulty is further validated by the works of Byun in 1996 and Carroll in 1963.

Moreover, a proclivity towards linear extrapolation, even in cases of nonlinear functions, has been documented by researchers like DeLosh, Busemeyer, and McDaniel in 1997. This behavioral tendency has been interpreted through two primary classes of computational models: similarity-based models and rule-based models. The Associative Learning Model (ALM), a representative of the former, suggests that individuals memorize specific input-output examples and base predictions for new inputs on the likeness to these recorded instances. These models, however, grapple with the challenge of explaining accurate extrapolation, a crucial aspect of function learning.

On the flip side, rule-based models, as proposed by Brehmer in 1974 and Carroll in 1963, posit that individuals abstract a functional rule encapsulating consistency across examples. These models exhibit a superior capability for accurate extrapolation but often fall short in predicting human performance in function learning tasks, indicating the necessity for a more comprehensive model.

In response to this need, hybrid models like the Extrapolation-Association Model (EXAM) designed by DeLosh et al., fuse elements of similarity-based learning and basic rule-based extrapolation. This model archives input-output examples like the ALM but applies a linear rule for extrapolation beyond these instances. Comparative studies, such as the one conducted by McDaniel and Busemeyer in 2005, hint that this hybrid approach best portrays overall human function learning performance.

The study by DeLosh et al. manipulated a critical factor, the 'density' of training, referring to the number of unique stimulus-response pairs presented during training. Despite this manipulation, the density did not significantly impact the participants' performance during training or the subsequent transfer phase. This outcome indicates that the ability to extrapolate and abstract functions in human learners could be less influenced by the volume of training data and more by the underlying cognitive processes involved in function learning.

More recent models, like the Population of Linear Experts (POLE) proposed by Kalish, Lewandowsky, and Kruschke in 2004, suggest that individuals break down complex functions into local linear experts. This mixture-of-experts approach accounts for phenomena such as knowledge partitioning but struggles in predicting extrapolation without fitting to the test data. Despite these newer models, hybrid strategies like EXAM continue to provide the most complete account of human function learning and generalization, particularly in relation to extrapolation.

The study by McDaniel et al. (2009) directly contrasts the POLE and EXAM models. They trained participants on piecewise linear functions, with one segment sparsely trained. This setup put the models to test on extrapolating from sparse data. Additionally, they based model comparisons on a priori predictions rather than post-hoc fits. Parameters were fit to just the training data, and models independently predicted unseen transfer stimuli. This approach offers more diagnostic tests than fitting the models simultaneously to all data. The crucial findings reveal that while POLE provides a better quantitative fit to the training data, EXAM more accurately predicts human performance on essential extrapolation trials. For instance, POLE tends to erroneously predict that extrapolation will follow the slope of the densely trained segment, rather than the slope of the sparsely trained region as exhibited by humans. In contrast, EXAM, with its exemplar representations and linear extrapolation rule, better represents the human tendency to extrapolate along the slope defined by the sparse training data.