### 2023-10-26 paper.qmd

---
title: Manuscript
toc: false
page-layout: full
---

```{=html}
<p class="text-center">
  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="manuscript">
  <iconify-icon inline="" icon="fa6-solid:file-lines"></iconify-icon> &ensp;HTML
  </a>

  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="_blank">
  <iconify-icon inline="" icon="fa6-solid:arrow-up-right-from-square"></iconify-icon> &ensp;HTML <small>(new window)</small>
  </a>

  <a class="btn btn-danger btn-sm mb-1" href="/manuscript/output/htw_manuscript.pdf" target="manuscript">
  <iconify-icon inline="" icon="fa6-solid:file-pdf"></iconify-icon> &ensp;PDF
  </a>

  <!-- <a class="btn btn-danger btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.pdf">
  <iconify-icon inline="" icon="fa6-solid:file-arrow-down"></iconify-icon> &ensp;PDF <small>(download)</small>
  </a>
   -->
  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-ms/manuscript-ms.pdf" target="manuscript">
  <iconify-icon inline="" icon="fa6-solid:file-pdf"></iconify-icon> &ensp;Manuscripty PDF
  </a>

  <!-- <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-ms/manuscript-ms.pdf">
  <iconify-icon inline="" icon="fa6-solid:file-arrow-down"></iconify-icon> &ensp;Manuscripty PDF <small>(download)</small>
  </a> -->

<!--  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-apa/manuscrip-apa.pdf" target="manuscript">
  <iconify-icon inline="" icon="fa6-solid:file-pdf"></iconify-icon> &ensp;Manuscript-apa
  </a> -->

    <!-- <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-apa/manuscript.pdf">
  <iconify-icon inline="" icon="fa6-solid:file-arrow-down"></iconify-icon> &ensp;Manuscript-apa PDF <small>(download)</small>
  </a> -->

<!--   <a class="btn btn-warning btn-sm mb-1" href="/manuscript/manuscript.pdf.md" target="manuscript">
  <iconify-icon inline="" icon="fa6-solid:file-pdf"></iconify-icon> &ensp;Manuscript-md
  </a> 
-->

</p>

<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:95%;width:100%;border:2px solid #ccc;" height="95%" width="100%" name="manuscript"></iframe>
```