### 2023-11-05 Simulations/DeLosh97_Sim.qmd
update DeLosh simulation functions

---
title: Simulating DeLosh 1997
date: last-modified
categories: [Simulation, ALM, EXAM, R]
code-fold: true
code-tools: true
execute: 
  warning: false
---

```{r}
#lapply(c('tidyverse','data.table','igraph','ggraph','kableExtra'),library,character.only=TRUE))
pacman::p_load(tidyverse,data.table,igraph,ggraph,kableExtra, patchwork) 
purrr::walk(here::here(c("Functions/Display_Functions.R")),source)
source(here::here("Functions","deLosh_data.R"))
```

```{r}
#https://nrennie.rbind.io/blog/2022-06-06-creating-flowcharts-with-ggplot2/
inNodes <- seq(1,6,1) %>% as.integer()
outNodes <- seq(300,1000,50)%>% as.integer()
stim <- "Stim"
resp <- "Response"
inFlow <- tibble(expand.grid(from=stim,to=inNodes)) %>% mutate_all(as.character)
outFlow <- tibble(expand.grid(from=outNodes,to=resp)) %>% mutate_all(as.character)

gd <- tibble(expand.grid(from=inNodes,to=outNodes)) %>% mutate_all(as.character) %>%
  rbind(inFlow,.) %>% rbind(.,outFlow)

g = graph_from_data_frame(gd,directed=TRUE)
coords2=layout_as_tree(g)
colnames(coords2)=c("y","x")
odf <- as_tibble(coords2) %>% 
  mutate(label=vertex_attr(g,"name"),
         type=c("stim",rep("Input",length(inNodes)),rep("Output",length(outNodes)),"Resp"),
         x=x*-1) %>%
  mutate(y=ifelse(type=="Resp",0,y),xmin=x-.05,xmax=x+.05,ymin=y-.35,ymax=y+.35)

plot_edges = gd %>% mutate(id=row_number()) %>%
  pivot_longer(cols=c("from","to"),names_to="s_e",values_to=("label")) %>%
                 mutate(label=as.character(label)) %>% 
  group_by(id) %>%
  mutate(weight=sqrt(rnorm(1,mean=0,sd=10)^2)/10) %>%
  left_join(odf,by="label") %>%
  mutate(xmin=xmin+.02,xmax=xmax-.02)

ggplot() + geom_rect(data = odf,
            mapping = aes(xmin = xmin, ymin = ymin, 
                          xmax = xmax, ymax = ymax, 
                          fill = type, colour = type),alpha = 0.5) +
  geom_text(data=odf,aes(x=x,y=y,label=label,size=3)) +
  geom_path(data=plot_edges,mapping=aes(x=x,y=y,group=id,alpha=weight)) +
  theme_void() + theme(legend.position = "none")

```

```{r}

linear_function <- function(x) 2.2 * x + 30
exponential_function <- function(x) 200 * (1 - exp(-x/25))
quadratic_function <- function(x) 210 - (x - 50)^2 / 12

extrapLines <- list(geom_vline(xintercept=30,color="black",alpha=.4,linetype="dashed"),
                    geom_vline(xintercept=70,color="black",alpha=.4,linetype="dashed"))

linear_plot <- ggplot(deLosh_data$human_data_linear, aes(x, y)) +
    geom_point(shape=1) + stat_function(fun = linear_function, color = "black") +
  labs(y="Response Magnitude", title="Linear Function",x="") + extrapLines

exponential_plot <- ggplot(deLosh_data$human_data_exp, aes(x, y)) +
  geom_point(aes(shape = "Observed", color = "Observed"),shape=1) + 
  stat_function(aes(color = "True Function"),fun = exponential_function, geom="line")+
  labs(x="Stimulus Magnitude", title="Exponential Function",y="")  +
  extrapLines +
  scale_shape_manual(values = c(1)) +
  scale_color_manual(values = c("Observed" = "black", "True Function" = "black")) +
  theme(legend.title = element_blank(), legend.position="top") +
  guides(color = guide_legend(override.aes = list(shape = c(1, NA), 
                                                  linetype = c(0, 1))))

quadratic_plot <- ggplot(deLosh_data$human_data_quad, aes(x = x, y = y)) +
  geom_point( shape = 1) +
  stat_function( fun = quadratic_function, geom = "line") +
  labs(title="Quadratic Function",x="",y="") + extrapLines

linear_plot + exponential_plot + quadratic_plot

```

## ALM Definition

::: column-margin
###### Input Activation

$$
a_i(X)=\exp \left|-\gamma \cdot\left[X-X_i\right]^2\right|
$$

###### Output activation

$$
o_j(X)=\Sigma_{i=1, M} w_{j i} \cdot a_i(X) 
$$

###### Output Probability

$$
P\left[Y_j \mid X\right]=o_j(X) / \Sigma_{k=1, L} o_k(X) 
$$

###### Mean Response

$$
m(X)=\Sigma_{j=1, L} Y_j \cdot P\left[Y_j \mid X\right] 
$$
:::

::: column-margin
###### Feedback Signal

$$
f_j(Z)=e^{-c\cdot(Z-Y_j)^2}
$$

###### Weight Updates

$$
w_{ji}(t+1)=w_{ji}(t)+\alpha \cdot {f_i(Z(t))-O_j(X(t))} \cdot a_i(X(t))
$$
:::

::: column-margin

###### Input node actvation

$$
P[X_i|X] = \frac{a_i(X)}{\\sum_{k=1}^Ma_k(X)}
$$

###### Slope Computation

$$
E[Y|X_i]=m(X_i) + \bigg[\frac{m(X_{i+1})-m(X_{i-1})}{X_{i+1} - X_{i-1}} \bigg]\cdot[X-X_i]
$$
:::

### Generate Response

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
alm.response <- function(input = 1, c, input.layer, output.layer,weight.mat) {
  input.activation <- exp(-c * (input.layer - input)^2) / sum(exp(-c * (input.layer - input)^2))
  output.activation <- weight.mat %*% input.activation
  output.probability <- output.activation / sum(output.activation)
  mean.response <- sum(output.layer * output.probability)
  list(mean.response = mean.response, input.activation = input.activation, output.activation = output.activation)
}
```

### Update Weights Based on Feedback

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
#| code-overflow: wrap
alm.update <- function(corResp, c, lr, output.layer, input.activation, output.activation, weight.mat) {
  fz <- exp(-c * (output.layer - corResp)^2)
  teacherSignal <- (fz - output.activation) * lr
  wChange <- teacherSignal %*% t(input.activation)
  weight.mat <- weight.mat + wChange
  weight.mat[weight.mat < 0] = 0
  return(weight.mat)
}

alm.trial <- function(input, corResp, c, lr, input.layer, output.layer, weight.mat) {
  alm_resp <- alm.response(input, c, input.layer,output.layer, weight.mat)
  updated_weight.mat <- alm.update(corResp, c, lr, output.layer, alm_resp$input.activation, alm_resp$output.activation, weight.mat)
  return(list(mean.response = alm_resp$mean.response, weight.mat = updated_weight.mat))
}
```

### Exam Generalization

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
exam.response <- function(input, c, trainVec, input.layer = INPUT_LAYER_DEFAULT,output.layer = OUTPUT_LAYER_DEFAULT, weight.mat) {
  nearestTrain <- trainVec[which.min(abs(input - trainVec))]
  aresp <- alm.response(nearestTrain, c, input.layer = input.layer,output.layer = OUTPUT_LAYER_DEFAULT,weight.mat)$mean.response

  xUnder <- ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver <- ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])

  mUnder <- alm.response(xUnder, c, input.layer = input.layer, output.layer, weight.mat)$mean.response
  mOver <- alm.response(xOver, c, input.layer = input.layer,output.layer, weight.mat)$mean.response

  exam.output <- round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (input - nearestTrain), 3)
  exam.output
}
```

### Simulation Functions

```{r}
#| code-fold: show

# simulation function
alm.sim <- function(dat, c, lr, input.layer = INPUT_LAYER_DEFAULT, output.layer = OUTPUT_LAYER_DEFAULT) {
  weight.mat <- matrix(0.00, nrow = length(output.layer), ncol = length(input.layer))
  xt <- dat$x
  n <- nrow(dat)
  st <- numeric(n) # Initialize the vector to store mean responses
  for(i in 1:n) {
    trial <- alm.trial(dat$x[i], dat$y[i], c, lr, input.layer, output.layer, weight.mat)
    weight.mat <- trial$weight.mat
    st[i] <- trial$mean.response
  }
  dat <- dat %>% mutate(almResp = st)
  return(list(d = dat, wm = weight.mat, c = c, lr = lr))
}

simOrganize <- function(simOut) {
  dat <- simOut$d
  weight.mat <- simOut$wm
  c <- simOut$c
  lr <- simOut$lr
  trainX <- unique(dat$x)

  almResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
    mutate(model = "ALM", resp = alm.response(x, c, input.layer = INPUT_LAYER_DEFAULT,output.layer = OUTPUT_LAYER_DEFAULT, weight.mat = weight.mat)$mean.response)

  examResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
    mutate(model = "EXAM", resp = exam.response(x, c, trainVec = trainX, input.layer = INPUT_LAYER_DEFAULT,output.layer = OUTPUT_LAYER_DEFAULT, weight.mat))

  organized_data <- bind_rows(almResp, examResp) %>% 
    mutate(type = first(dat$type),
           error = abs(resp - y),
           c = c,
           lr = lr,
           type = factor(type, levels = c("linear", "exponential", "quadratic")),
           test_region = ifelse(x %in% trainX, "train", 
                                ifelse(x > min(trainX) & x < max(trainX), "interpolate", "extrapolate")))
  organized_data
}

generateSimData <- function(density, envTypes, noise) {
  reps <- 200 / length(trainingBlocks[[density]])
  map_dfr(envTypes, ~ 
            generate.data(rep(trainingBlocks[[density]], reps), type = .x, noise)) |>
    group_by(type) |>
    mutate(block = rep(1:reps, each = length(trainingBlocks[[density]])),
           trial=seq(1,200))
}

simulateAll <- function(density,envTypes, noise, c = .2, lr = .2) {
  trainMat <- generateSimData(density, envTypes, noise)
  trainData <- map(envTypes, ~ alm.sim(trainMat %>% filter(type == .x), c = c, lr = lr))
  assign(paste(density),list(train=trainData, test=map_dfr(trainData, simOrganize) %>% mutate(density = density)))
}

```

### Simulate Training and Testing

```{r fig.width=10, fig.height=8}
#| column: page-inset-right

envTypes <- c("linear", "exponential", "quadratic")
densities <- c("low", "med", "high")
noise=0
INPUT_LAYER_DEFAULT <- seq(0, 100, 0.5)
OUTPUT_LAYER_DEFAULT <- seq(0, 250, 1)

c = 1.4
lr=.8

results <- map(densities, ~ simulateAll(.x, envTypes, noise, c, lr)) |>
  set_names(densities)

trainAll <- results %>%
  map_df(~ map_df(.x$train, pluck, "d"), .id = "density") |>
  mutate(stage=as.numeric(cut(trial,breaks=20,labels=seq(1,20))),
         dev=sqrt((y-almResp)^2),
         density=factor(density,levels=c("low","med","high")),
         type=factor(type,levels=c("linear","exponential","quadratic"))) |>
  dplyr::relocate(density,type,stage)

simTestAll <- results |> map("test") |> bind_rows() |>
  group_by(type,density,model) %>%
  mutate(type=factor(type,levels=c("linear","exponential","quadratic")),
         density=factor(density,levels=c("low","med","high"))) %>%
  dplyr::relocate(density,type,test_region)

```

### Training Data

```{r fig.width=10, fig.height=8}

trainAll %>% ggplot(aes(x=block,y=dev,color=type)) + stat_summary(geom="line",fun=mean,alpha=.4)+
  stat_summary(geom="point",fun=mean,alpha=.4)+
  stat_summary(geom="errorbar",fun.data=mean_cl_normal,alpha=.4)+facet_wrap(~density, scales="free_x")

```

Predictions for Generalization

```{r fig.width=11,fig.height=10}
#| column: page-inset-right
#| 
simTestAll %>% ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

```

Collpasing Across Density Levels gives us:

```{r}
#| eval: false
#| column: page-inset-right
simTestAll %>% group_by(type,model,x,y) %>% summarise(resp=mean(resp))  %>% ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  facet_grid(~type) + 
  theme_bw() + theme(legend.position="bottom")
```

### Table

| Model & Definition                    | R Code                                                                                                      |
|---------------------------------------|-------------------------------------------------------------------------------------------------------------|
| $a_i(X)=\exp \left|-\gamma \cdot\left[X-X_i\right]^2\right|$   | `exp(-c * (input.layer - input)^2)`        |
| $o_j(X)=\Sigma_{i=1, M} w_{j i} \cdot a_i(X)$                 | `weight.mat %*% input.activation`                                                     |
| $P\left[Y_j \mid X\right]=o_j(X) / \Sigma_{k=1, L} o_k(X)$   | `output.activation / sum(output.activation)`                                         |
| $m(X)=\Sigma_{j=1, L} Y_j \cdot P\left[Y_j \mid X\right]$    | `sum(output.layer * output.probability)`                                                  |
| $f_j(Z)=e^{-c\cdot(Z-Y_j)^2}$                                 | `exp(-c * (output.layer - corResp)^2)`                                                               |
| $w_{ji}(t+1)=w_{ji}(t)+\alpha \cdot {f_i(Z(t))-O_j(X(t))} \cdot a_i(X(t)$   | ` lr *(fz - output.activation) %*% t(input.activation)`  |
                                                      |
| $E[Y|X_i]=m(X_i) + \bigg[\frac{m(X_{i+1})-m(X_{i-1})}{X_{i+1} - X_{i-1}} \bigg]\cdot[X-X_i]$ | ` trainVec[which.min(abs(input - trainVec))]; xUnder <- ...; xOver <- ...; mUnder <- ...; mOver <- ...; exam.output <- round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (input - nearestTrain), 3)`|

### Master Function for full simulation

```{r}
#| eval: false
#| include: false 
# Function that goes through every step of generating data, simulating training, and simulating generalization
full.sim <- function(c,lr,noise)
{

envTypes <- c("linear", "exponential", "quadratic")
lowDensityTrainBlock <- c(30.5, 36.0, 41.0, 46.5, 53.5, 59.0, 64.0, 69.5)
medDensityTrainBlock <- c(
    30.0, 31.5, 33.0, 34.5, 36.5, 38.5, 41.0, 43.5, 46.0,
    48.5, 51.5, 54.0, 56.5, 59.0, 61.5, 63.5, 65.5, 67.0, 68.5, 70.0
)
highDensityTrainBlock <- c(
    30.0, 30.5, 31.0, 32.0, 33.0, 33.5, 34.5, 35.5,
    36.5, 37.0, 38.0, 38.5, 39.5, 40.5, 41.5, 42.0, 43.0,
    43.5, 44.5, 45.5, 46.5, 47.0, 48.0, 48.5, 49.0, 51.0, 51.5, 52.0,
    53.0, 53.5, 54.5, 55.5, 56.5, 57.0, 58.0, 58.5, 59.5, 60.5, 61.5, 
    62.0, 63.0,63.5, 64.5, 65.5, 66.5, 67.0, 68.0, 69.0, 69.5, 70.0
)
# low density has 25 training blocks, medium has 10 blocks, high has 4 blocks. 
# generate training data, for each combination of environment type and density. Use purrr map functions. Rep each dataset by its number of blocks.
a
lowSim <- map(envTypes, ~ alm.sim(lowTrain %>% filter(type == .x), c = 1.4, lr = .4))
medSim <- map(envTypes, ~ alm.sim(medTrain %>% filter(type == .x), c = 1.4, lr = .4))
highSim <- map(envTypes, ~ alm.sim(highTrain %>% filter(type == .x), c = 1.4, lr = .4))

simAll <- rbind(bind_rows(lowSim %>% map("d")) %>% mutate(density = "low"), 
                bind_rows(medSim %>% map("d")) %>% mutate(density = "med"), 
                bind_rows(highSim %>% map("d")) %>% mutate(density = "high"))

simAll <- simAll %>% mutate(stage=as.numeric(cut(trial,breaks=20,labels=seq(1,20))),
                                      dev=sqrt((y-almResp)^2),
                                  #reorder density factor levels
                            density=factor(density,levels=c("low","med","high")),
                            type=factor(type,levels=c("linear","exponential","quadratic"))) %>%
                            dplyr::relocate(density,type,stage)

lowSimTest <- map_dfr(lowSim,simOrganize) %>% mutate(density = "low")
medSimTest <- map_dfr(medSim,simOrganize) %>% mutate(density = "med")
highSimTest <- map_dfr(highSim,simOrganize) %>% mutate(density = "high")

simTestAll <- rbind(lowSimTest,medSimTest,highSimTest) %>% group_by(type,density,model) %>%
  mutate(type=factor(type,levels=c("linear","exponential","quadratic")),
         density=factor(density,levels=c("low","med","high"))) %>%
  dplyr::relocate(density,type,test_region)

return(list(simAll=list(simAll),simTestAll=list(simTestAll)))

}
```

### Simulations with noise

```{r fig.height=10,fig.width=11}
#| eval: false
#| include: false
k = full.sim(c=1.4,lr=.4,noise=2.0)
k4 = full.sim(c=1.4,lr=.4,noise=4.0)

# run simulation with noise=10, 3 times, average results together. 
k10 = map_dfr(1:3, ~ full.sim(c=1.4,lr=.4,noise=10.0)) %>% group_by(type,density,model) %>%
  mutate(type=factor(type,levels=c("linear","exponential","quadratic")),
         density=factor(density,levels=c("low","med","high"))) %>%
  dplyr::relocate(density,type,test_region)

k %>% pluck("simAll") %>% ggplot(aes(x=block,y=dev,color=type)) + stat_summary(geom="line",fun=mean,alpha=.4)+
  stat_summary(geom="point",fun=mean,alpha=.4)+
  stat_summary(geom="errorbar",fun.data=mean_cl_normal,alpha=.4)+facet_wrap(~density, scales="free_x")

k4 %>% pluck("simAll") %>% ggplot(aes(x=block,y=dev,color=type)) + stat_summary(geom="line",fun=mean,alpha=.4)+
  stat_summary(geom="point",fun=mean,alpha=.4)+
  stat_summary(geom="errorbar",fun.data=mean_cl_normal,alpha=.4)+facet_wrap(~density, scales="free_x")

k %>% pluck("simTestAll") %>%ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

k4 %>% pluck("simTestAll") %>%ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

k10 %>% pluck("simTestAll") %>%ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

```

```{r}
#| eval: false
#| include: false
# label each each simulation with its density level (low, med, high), then combine all 3
lowSimTest <- lowSimTest %>% mutate(density = "low")
medSimTest <- medSimTest %>% mutate(density = "med")
highSimTest <- highSimTest %>% mutate(density = "high")

simTest <- bind_rows(lowSimTest,medSimTest,highSimTest)

# extract element d from each list, and bind rows, remove Nan's, group by type, mutate new variable "Stage", which is set to first for block 1, last for final block, and middle for all other glocks, pipe to ggplot, plotting y and almResp in different colors, facet by type and stage (only first and last block)
ls2=bind_rows(lowSim %>% map("d")) %>% filter(!is.na(almResp)) %>% 
  group_by(type) %>% 
  mutate(stage = ifelse(block == 1, "first", ifelse(block == 25, "last", "middle")),
         stage2=cut(trial,breaks=20,labels=seq(1,20)),
         dev=sqrt((y-almResp)^2))

ls2 %>% ggplot() + geom_point(aes(x=stage2,y=dev))+facet_grid(~type)

ls2 %>% filter(stage %in% c("first","last")) %>% ggplot() + geom_point(aes(x=x,y=y),color="red",alpha=.3) + 
geom_point(aes(x=x,y=almResp),color="blue",alpha=.4)+ 
facet_grid(type~stage) + theme_bw() + theme(legend.position="bottom")

lowSimTest %>% ggplot() + geom_point(aes(x=x,y=resp,color=model)) + geom_line(aes(x=x,y=y),alpha=.3)+ facet_grid(~type) + theme_bw() + theme(legend.position="bottom")

# make grid of each combination of train dataset and envType
simGrid <- expand.grid(envTypes=envTypes,trainData=c("lowTrain","medTrain","highTrain"))

# map over the grid, for each row of simGrid, filter the training dataset by the envType, and then run the simulation function, and then bind rows of the output data frame.
simOut <- map2(c("lowTrain","medTrain","highTrain"),envTypes, ~ alm.sim(get(.x) %>% filter(type == .y), c = 1.4, lr = .4))

# simOut <- map(c("lowTrain","medTrain","highTrain"), ~ map_dfr(envTypes, ~ alm.sim(get(.) %>% filter(type == .x), c = 1.4, lr = .4) %>% a))
# 
# simOut <- map_dfr(simGrid, ~ alm.sim(get(.x$trainData) %>% filter(type == .x$envTypes), c = 1.4, lr = .4) %>% simOrganize)
# 
# for each training dataset in c(lowTrain,medTrain,highTrain), run simulation function, separately for each type, and then bind rows of the output data frame.

```

### Primary Functions

```{r}
#| include: false
#| eval: false
alm.response <- function(input=1,c) {
input.activation <- exp(-c*(input.layer - input)^2)
input.activation <<- input.activation/sum(input.activation)
#print(length(input.activation)); print(dim(weight.mat))
output.activation <<- weight.mat %*% input.activation
output.probability <<- output.activation/sum(output.activation)
mean.response <<- sum(output.layer * output.probability)
mean.response
}

alm.update <- function(corResp,c,lr){
  fz <- exp(-c*(output.layer - corResp)^2)
  teacherSignal <- (fz - output.activation)*lr
  #print(length(teacherSignal)); print(length(fz))
  wChange <- teacherSignal %*% t(input.activation)
  weight.mat <<- weight.mat + (wChange)
  weight.mat[weight.mat<0]=0
 # weight.mat[weight.mat>1]=1
  weight.mat <<- weight.mat
}

alm.trial <- function(input, corResp,c,lr){
  alm.response(input,c)
  alm.update(corResp,c,lr)
  print(paste0("input=",input,"; corResp=",corResp,"; mean.response=",mean.response))
  mean.response
}

exam.response <- function(input,c){
  # Find the index of the input node with the highest activation
  trainVec = sort(unique(xt))
  nearestTrain <- trainVec[which.min(abs(input - trainVec))]
  aresp <- alm.response(nearestTrain,c)
  #max.index <- which.max(input.activation)
  xUnder = ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver = ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])

  mUnder <- alm.response(xUnder,c)
  mOver <- alm.response(xOver,c)

   exam.output = round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (input - nearestTrain), 3)
  # Determine the input nodes and associated weights for computing the slope
  exam.output
}

```