### 2023-11-14 Misc/scratch/_alm_table.qmd
overhaul tables

### ALM Activation & Response

| Step                          | Equation                                                                            |
|--------------------|----------------------------------------------------|
| **ALM Activation & Response** |                                                                                     |
| Input Activation              | $a_i(X) = \frac{e^{-c(X-X_i)^2}}{\sum_{k=1}^M e^{-c(X-X_k)^2}}$                     |
| Output Activation             | $O_j(X) = \sum_{k=1}^M w_{ji} \cdot a_i(X)$                                         |
| Output Probability            | $P[Y_j|X] = \frac{O_j(X)}{\sum_{k=1}^M O_k(X)}$                                     |
| Mean Output                   | $m(x) = \sum_{j=1}^L Y_j \cdot \frac{O_j(x)}{\sum_{k=1}^M O_k(X)}$                  |
| **ALM Learning**              |                                                                                     |
| Feedback Activation           | $f_j(Z) = e^{-c(Z-Y_j)^2}$                                                          |
| Update Weights                | $w_{ji}^{new}=w_{ji}+\eta\alpha_{ji}$                                               |
| **EXAM**                      |                                                                                     |
| Extrapolation                 | $P[X_i|X] = \frac{a_i(X)}{\sum_{k=1}^M a_k(X)}$                                     |
|                               | $E[Y|X_i] = m(X_i) + \frac{m(X_{i+1})-m(X_{i-1})}{X_{i+1}-X_{i-1}} \cdot [X - X_i]$ |

::: column-page-inset-right
## Model Table

### ALM Activation & Response

|                    | ALM Response Generation                                            |                                                                                               |
|------------------|------------------------------|-------------------------|
| Input Activation   | $a_i(X) = \frac{e^{-c(X-X_i)^2}}{\sum_{k=1}^M e^{-c(X-X_k)^2}}$    | Input nodes activate as a function of Gaussian similarity to stimulus                         |
| Output Activation  | $O_j(X) = \sum_{k=1}^M w_{ji} \cdot a_i(X)$                        | Output unit $O_j$ activation is the weighted sum of input activations and association weights |
| Output Probability | $P[Y_j|X] = \frac{O_j(X)}{\sum_{k=1}^M O_k(X)}$                    | The response, $Y_j$ probabilites computed via Luce's choice rule                              |
| Mean Output        | $m(X) = \sum_{j=1}^L Y_j \cdot \frac{O_j(x)}{\sum_{k=1}^M O_k(X)}$ | Weighted average of probabilities determines response to X                                    |
|                    | **ALM Learning**                                                   |                                                                                               |
| Feedback           | $f_j(Z) = e^{-c(Z-Y_j)^2}$                                         | feedback signal Z computed as similarity between ideal response and observed response         |
| magnitude of error | $\Delta_{ji}=(f_{j}(Z)-o_{j}(X))a_{i}(X)$                          | Delta rule to update weights.                                                                 |
| Update Weights     | $w_{ji}^{new}=w_{ji}+\eta\Delta_{ji}$                              | Updates scaled by learning rate parameter $\eta$.                                             |
|                    | **EXAM Extrapolation**                                             |                                                                                               |
| Instance Retrieval | $P[X_i|X] = \frac{a_i(X)}{\sum_{k=1}^M a_k(X)}$                    | Novel test stimulus $X$ activates input nodes $X_i$                                           |
| Slope Computation  | $S =$ $\frac{m(X_{1})-m(X_{2})}{X_{1}-X_{2}}$                      | Slope value, $S$ computed from nearest training instances                                              |
| Response           | $E[Y|X_i] = m(X_i) + S \cdot [X - X_i]$                          | ALM response $m(X_i)$ adjusted by slope.                                                      |

: ALM & EXAM Equations {#tbl-alm-exam2}
:::

{{< pagebreak >}}