### 2024-01-04 Model/abc_group.qmd
abc group code

---
title: ABC Group Level Fits
date: last-modified
categories: [Simulation, ALM, EXAM, R]
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

1. Load the group level ABC results - i.e. the posterior distribution for the c and lr parameters. 
2. Preview the posterior distribution.

```{r}
pacman::p_load(tidyverse, data.table,dtplyr, here, patchwork, conflicted)
conflict_prefer_all("dplyr", quiet = TRUE)

group_prior=abc_2M_p001 = readRDS(here::here("data/abc_2M_rmse_p001.rds"))
names(group_prior)
str(group_prior$abc_ev[[1]])

#tetr_v <- group_prior$abc_ev$teter_results

teter <- abc_2M_p001 |> map_dfr(~tibble(pluck(.x$teter_results))) 
te <- abc_2M_p001 |> map_dfr(~tibble(pluck(.x$te_results)))
tr <- abc_2M_p001 |> map_dfr(~tibble(pluck(.x$tr_results)))
```

```{r fig.width=12, fig.height=10}

teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(expMode2=="Test") |>
  mutate(facet_label = paste0("rank: ", rank, "\n", "c: ", round(c, 4), "\n", "lr: ", round(lr, 4), "\n", "distance: ", round(distance, 1))) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(~Model)

teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(expMode2=="Test",rank<=10) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(~Model)

teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(expMode2=="Test",rank<=10) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(rank~Model)

teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(condit=="Constant",expMode2=="Test",rank<=20) |>
  mutate(facet_label = paste0("rank: ", rank, "\n", "c: ", round(c, 4), "\n", "lr: ", round(lr, 4), "\n", "distance: ", round(distance, 1))) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(~facet_label)

tr |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(condit=="Constant",expMode2=="Test",rank<=20) |>
  mutate(facet_label = paste0("rank: ", rank, "\n", "c: ", round(c, 4), "\n", "lr: ", round(lr, 4), "\n", "distance: ", round(distance, 1))) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(~facet_label)

teter |> select(sim_dat) |> unnest(sim_dat)  %>%  filter(condit=="Varied",expMode2=="Test",rank<=20) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(rank~Model)

post_data_v %>%  filter(expMode2=="Test") |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge())

post_data %>% filter(rank>1970, expMode2=="Test") |> 
  mutate(facet_label = paste0("rank: ", rank, "\n", "c: ", round(c, 4), "\n", "lr: ", round(lr, 4), "\n", "distance: ", round(distance, 1))) |>
  ggplot(aes(x = x, y = pred, fill=condit)) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge()) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  facet_wrap(~facet_label)

```

```{r}

# ggplot(tetr, aes(x = c, y = lr, color = distance)) +
#   geom_point()

```

# A tibble: 2 × 9
# Rowwise: 
#   distance        c    lr sim_index  rank Model Group  Fit_Method   sim_dat           
#      <dbl>    <dbl> <dbl>     <int> <int> <chr> <chr>  <chr>        <list>            
# 1     94.5 0.000161 0.626    774776     1 EXAM  Varied Test & Train <tibble [90 × 13]>
# 2     97.4 0.000142 0.678   1353940     2 EXAM  Varied Test & Train <tibble [90 × 13]>

# the sim_dat column contains the simulated data for each parameter set in the ABC run.
# examine posterior predictive distribution for the c and lr parameters.

post_data <- tetr_v %>% select(sim_dat) |> 
  unnest(sim_dat)

str(post_data_v)
# tibble [180,000 × 13] (S3: tbl_df/tbl/data.frame)
#  $ Model     : chr [1:180000] "EXAM" "EXAM" "EXAM" "EXAM" ...
#  $ Fit_Method: chr [1:180000] "Test & Train" "Test & Train" "Test & Train" "Test & Train" ...
#  $ c         : num [1:180000] 0.000161 0.000161 0.000161 0.000161 0.000161 ...
#  $ lr        : num [1:180000] 0.626 0.626 0.626 0.626 0.626 ...
#  $ distance  : Named num [1:180000] 94.5 94.5 94.5 94.5 94.5 ...
#   ..- attr(*, "names")= chr [1:180000] "...774776" "...774776" "...774776" "...774776" ...
#  $ rank      : int [1:180000] 1 1 1 1 1 1 1 1 1 1 ...
#  $ tr        : num [1:180000] 1 2 3 4 5 6 7 8 9 10 ...
#  $ condit    : Factor w/ 2 levels "Constant","Varied": 2 2 2 2 2 2 2 2 2 2 ...
#  $ x         : num [1:180000] 1000 800 800 1000 800 1000 800 1000 1000 1000 ...
#  $ expMode2  : Factor w/ 4 levels "Train","Train-Nf",..: 1 1 1 1 1 1 1 1 1 1 ...
#  $ y         : num [1:180000] 1136 1006 819 1037 923 ...
#  $ pred      : num [1:180000] 38.2 1081.6 1000.2 1179 857.5 ...
#  $ resid     : num [1:180000] 1097.4 -75.3 -181.6 -141.6 65.1 ...

```{r}

```