### 2024-02-12 Assets/tmp_plots/_temp_reject.qmd
new report folder

---
title: "`r params$title`"
author: Thomas Gorman
date: "`r Sys.Date()`"
format: 
  html:
    page-layout: full
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
params:
    file_name: file_name
    title: title
   #ind_fits: ind_fits

---

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, stringr,future,furrr, knitr, reactable, flextable, ggstance,htmltools)
conflict_prefer_all("dplyr", quiet = TRUE)
options(scipen = 999)
#walk(c("Display_Functions","fun_alm","fun_indv_fit"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
source("../../../Functions/fun_indv_fit.R")
source("../../../Functions/Display_Functions.R")
source("../../../Functions/fun_model.R")
source("../../../Functions/fun_alm.R")

ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
test <- readRDS(here::here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Test") 
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

input_layer <<- output_layer <<-  c(100,350,600,800,1000,1200)

ids <- c(1,2,4,5,6,7,8, 10,11,12,13)
ids2 <- c(1,66,36)
ids3 <- c(20,71,101,4,76,192)
idsBad <- c(76,192, 101)

file_name <- params$file_name
# list.files(here('data/abc_reject'))
# (grep("Train",list.files(here(paste0('data/abc_reject/',file_name)),
#                                            pattern="EXAM_Test",full.names = TRUE),
#                                 invert=TRUE, value=TRUE))

ind_fits <- map(list.files(here(paste0('data/abc_reject/'),file_name),full.names=TRUE), readRDS)
# ind_fits <- map(list.files(here('data/abc_reject/n_iter_2000_ntry_10_2918'),full.names=TRUE), readRDS)
ind_fits_df <- ind_fits |> map(~list(dat=.x[[1]], Model = .x[["Model"]], Fit_Method=.x[["Fit_Method"]]))
ind_fits_df <- ind_fits_df |> map(~rbindlist(.x$dat) |> mutate(Model = .x$Model, Fit_Method = .x$Fit_Method)) |> rbindlist()

#run_params <- tibble::lst(file_name,cMean=ind_fits[[1]]$cMean, cSig=ind_fits[[1]]$cSig,

run_params <- keep(ind_fits[[1]], ~any(class(.x) %in% c("character", "numeric")))
run_params <- tibble(!!!run_params) |> select(-Model,-Fit_Method)

extract_info <- function(raw_names, run_params) {
  fname <- tools::file_path_sans_ext(basename(raw_names))
  n_samp <- str_extract(raw_names, "(?<=_)\\d+(?=_)")
  ntry_ <- str_extract(raw_names, "(?<=ntry_)\\d+")
  type <- 'ss'
  data.frame(run_params, n_samp, ntry_, type, fname)
}

run_info <- extract_info(file_name,run_params)

```

```{r}

generate_data <- function(Model, post_samples, data, num_samples = 1, return_dat = "train_data, test_data") {
  # Filter data for the specific id without invalidating selfref
  sbj_data <- copy(data[id == post_samples$id[1]])
  simulation_function <- ifelse(Model == "EXAM", full_sim_exam, full_sim_alm)

  target_data <- switch(return_dat,
                        "test_data" = copy(sbj_data[expMode2 == "Test"]),
                        "train_data" = copy(sbj_data[expMode2 == "Train"]),
                        "train_data, test_data" = copy(sbj_data[expMode2 %in% c("Test", "Train")]))

  post_samples <- post_samples[order(mean_error)][1:num_samples, .(c, lr, mean_error, rank = .I)]

  simulated_data_list <- lapply(1:nrow(post_samples), function(i) {
    params <- post_samples[i]
    sim_data <- simulation_function(sbj_data, params$c, params$lr, input_layer = input_layer, 
                                    output_layer = output_layer, return_dat = return_dat)
    sim_data_dt <- data.table(id = sbj_data$id[1], condit = sbj_data$condit[1], 
                              expMode2 = target_data$expMode2, Model = Model,tr=target_data$tr,
                              y = target_data$y, x = target_data$x, c = params$c, 
                              lr = params$lr, mean_error = params$mean_error, rank = i,
                              pred = sim_data)
    return(sim_data_dt)
  })

  result_dt <- rbindlist(simulated_data_list)
  setcolorder(result_dt, c("id", "condit", "expMode2","tr", "c", "lr", "x", "y", "pred"))
  return(result_dt)
}

future::plan(multisession)

nestSbjModelFit <- ind_fits_df %>% nest(.by=c(id,Model,Fit_Method))

post_dat <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
   generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 50, return_dat="test_data")
   })) |> 
  select(Fit_Method,pp,-data) |>  
  unnest(pp) |> filter(expMode2=="Test") |> as.data.table()

post_dat_avg <- post_dat |> group_by(id, condit, Model, Fit_Method, x, c, lr, rank) |> 
  summarise(y = mean(y), pred = mean(pred), error = y - pred) |> as.data.table()

setorder(post_dat_avg, id, x, rank)
post_dat_l <- melt(post_dat_avg, id.vars = c("id", "condit", "Model", "Fit_Method", "x", "c", "lr", "rank","error"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val")
post_dat_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))]
setorder(post_dat_l, id, Resp)
rm(post_dat_avg)
plan(sequential)

# organize training data predictions
pd_train <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
   generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 10, return_dat="train_data")
   })) |>
  select(Fit_Method,pp,-data) |>
  unnest(pp) |> as.data.table() |> filter(expMode2=="Train")

nbins <- 3
pd_train <- pd_train |> group_by(id,condit,Model,Fit_Method) |>
  mutate(Block=cut(tr,breaks=seq(1,max(tr), length.out=nbins+1),include.lowest=TRUE,labels=FALSE))
setorder(pd_train, id, x,Block, rank)

pd_train_l <- melt(pd_train, id.vars = c("id", "condit", "Model","Block", "Fit_Method", "x", "c", "lr", "rank"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val") |> as.data.table()
pd_train_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))] 
setorder(pd_train_l, id,Block, Resp)

pd_train_l <- pd_train_l |> mutate(dist=abs(x-val))

plan(sequential)

# AIC and BIC
# mc2 <- post_dat |> filter(rank==1) |>
#   group_by(id,condit,Model,Fit_Method) |>
#   mutate(e2=abs(y-pred),n=n(),k=2) |>
#   group_by(id,condit,Model,Fit_Method,x) |>
#   summarise(y=mean(y), pred=mean(pred), e2=mean(e2), mean_error=first(mean_error),n=first(n),k=first(k)) |>
#   group_by(id,condit,Model,Fit_Method) |>
#   summarise(e2=mean(e2), mean_error=first(mean_error),n=first(n),k=first(k)) |>
#   mutate(AIC=2*k+n*mean_error, BIC=log(n)*k+n*mean_error)

# Single run extraction:
# exam_test <- ind_fits_df |> filter(Model == "EXAM", Fit_Method == "Test")
# post_dat_trial <- exam_test %>% split(f =c(.$id), drop=TRUE) |> 
#   map(~generate_data(.x$Model, .x, ds, num_samples = 15, return_dat="train_data, test_data")) |> 
#   rbindlist() |> 
#   filter(expMode2=="Test")
# 
# post_dat_avg <- post_dat_trial |> group_by(id,condit,x,c,lr,rank) |> 
#   summarise(y = mean(y),pred = mean(pred)) |> arrange(id,x,rank)

#post_dat_l |> filter(Model == "EXAM", Fit_Method == "Test",id==1) |> arrange(rank)

# head(post_dat_l)

```

#### `{r} head(run_info)`

##  Posterior Average Table: 
```{r fig.width=12, fig.height=17}
#| eval: true
#| tbl-cap: "Posterior Distribution"
#| tbl-subcap: 
#| - "Full Posterior"
#| - "Best Posterior"

# post_tabs <- abc_tables(post_dat_l)
# post_tabs$et_sum |> gt::gt()

post_tabs <- abc_tables(post_dat,post_dat_l)

# tables with fit error
# post_tabs$agg_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
#                        `ME` = as_paragraph(mean_error)) |> as_flextable() |>
#   set_caption("Mean from full_posterior")

# post_tabs$agg_best |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
#                        `ME` = as_paragraph(mean_error)) |> as_flextable() |>
#   set_caption("Mean from best parameters")

post_tabs$agg_pred_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

post_tabs$agg_pred_best |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

```

## Interactive Test Predictions Table
```{r}
tags$select(
  tags$option(value = "", "All"),
  purrr::map(unique(post_tabs$agg_x_full$condit), tags$option),
    purrr::map(unique(post_tabs$agg_x_full$Model), tags$option),
    purrr::map(unique(post_tabs$agg_x_full$Fit_Method), tags$option),
   purrr::map(unique(post_tabs$agg_x_full$x), tags$option)
)

selectFilter <- function(tableId, style = "width: 100%; height: 100%;") {
  function(values, name) {
    tags$select(
      # Set to undefined to clear the filter
      onchange = sprintf("
        const value = event.target.value
        Reactable.setFilter('%s', '%s', value === '__ALL__' ? undefined : value)
      ", tableId, name),
      # "All" has a special value to clear the filter, and is the default option
      tags$option(value = "__ALL__", "All"),
      lapply(unique(values), tags$option),
      "aria-label" = sprintf("Filter %s", name),
      style = style
    )
  }
}

post_tabs$agg_x_full |> 
  left_join(post_tabs$agg_x_best |> rename("best_error"=mean_error,"best_pred"=pred), 
            by=c("condit","Model","Fit_Method","x","y")) |> 
  arrange(desc(condit),x,Fit_Method) |>
  mutate(x=as.character(x)) |>
  reactable::reactable(
    columns = list(condit=colDef(name="Condit",filterInput=selectFilter("my-tbl")),
                   Model=colDef(name="Model",filterInput=selectFilter("my-tbl")),
                   Fit_Method=colDef(name="Fit_Method",filterInput=selectFilter("my-tbl"),
                                     filterMethod = JS("(rows, columnId, filterValue) => {
        return rows.filter(row => row.values[columnId] === filterValue)
      }")),
      x=colDef(name="x",filterInput=selectFilter("my-tbl"),
               filterMethod = JS("(rows, columnId, filterValue) => {
        return rows.filter(row => row.values[columnId] === filterValue)
      }"))),
    elementId = "my-tbl",
    defaultPageSize=15,
    filterable=TRUE) 
```

## Train Predictions
```{r}

pt_train <- abc_train_tables(pd_train,pd_train_l)
pt_train$agg_pred_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

pt_train$block_pred_full |> 
  flextable::tabulator(rows=c("Fit_Method","Model","Block"), columns=c("condit"), 
                       `Observed` = as_paragraph(y), 
                       `Predicted` = as_paragraph(pred),
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

pt_train$block_pred_full |> filter(condit=="Varied") |>
  flextable::tabulator(rows=c("Fit_Method","Block"), columns=c("Model"), 
                      # `Observed` = as_paragraph(y), 
                       `Predicted` = as_paragraph(pred),
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

best_indices <- pt_train$block_pred_full_x %>%
   filter(!(Fit_Method=="Test")) |>
  nest(data = -condit) %>%
  mutate(indices = map(data, ~{
    .x %>%
      select(-y, -pred, -mean_error) %>%
      pivot_wider(names_from = Model, values_from = best) %>%
      ungroup() |>
      select(ALM, EXAM) %>%
      pmap(.f = function(ALM, EXAM) {
        which(c(ALM, EXAM) == 1)
      })
  })) %>%
  select(condit, indices)

pt_train$block_pred_full_x |> filter(condit=="Varied",!(Fit_Method=="Test")) |>
  flextable::tabulator(rows=c("Fit_Method","Block","x"), columns=c("Model"), 
                      # `Observed` = as_paragraph(y), 
                       `Predicted` = as_paragraph(pred),
                       `ME` = as_paragraph(mean_error)) |> 
  as_flextable() %>% 
  apply_best_formatting(.,pull(filter(best_indices,condit=="Varied"),indices) %>% .[[1]]) |> 
  set_caption("Vared Train Predictions")

pt_train$block_pred_full_x |> filter(condit=="Constant",!(Fit_Method=="Test")) |>
  flextable::tabulator(rows=c("Fit_Method","Block","x"), columns=c("Model"), 
                      # `Observed` = as_paragraph(y), 
                       `Predicted` = as_paragraph(pred),
                       `ME` = as_paragraph(mean_error)) |> 
  as_flextable() %>% 
  apply_best_formatting(.,pull(filter(best_indices,condit=="Constant"),indices) %>% .[[1]]) |> 
  set_caption("Constant Train Predictions")

# pull(filter(best_indices,condit=="Varied"),indices) %>% .[[1]]

```

## Posterior Predictive: 
```{r, fig.width=12, fig.height=15}
#| eval: true
group_predictive_plots(post_dat_l) 
```

### Alt grouping of Posterior Predictive
```{r, fig.width=12, fig.height=15}
 group_predictive_plots2(post_dat_l) 
```

```{r fig.width=11, fig.height=12}
plot_indv_posterior(ind_fits_df |> mutate(Group=condit))

plot_indv_posterior(post_dat |> filter(rank==1) |> mutate(Group=condit))

# plot join density of c and lr
# post_dat |> filter(rank==1, c<.001) |> 
#   ggplot(aes(x=c, y=lr, color=condit)) + geom_point() + facet_wrap(~Model+Fit_Method)

post_dat |>  filter(c<.001) |> ggplot(aes(x=c,y=condit,fill=condit)) + geom_boxploth(width=.5) + facet_wrap(~Model+Fit_Method,scales="free")

```

## Model Comparison:
```{r, fig.width=12, fig.height=13}
#| eval: true
indv_best_plots(post_dat_l)

# post_dat_l |> group_by(condit,Model,Fit_Method,rank) |> 
#    summarise(mean_error=mean(abs(error)), n=n()) |> 
#    ggplot(aes(x=rank,y=mean_error,fill=Model))+geom_col()+facet_wrap(~Fit_Method)

post_dat_l |> group_by(condit,Model,Fit_Method,rank,x) |> 
   summarise(mean_error=mean(abs(error)), n=n()) |> 
   ggplot(aes(x=rank,y=mean_error,fill=Model))+geom_col()+facet_wrap(Fit_Method~x,ncol=6)

group_best_plots(post_dat_l)

```

## Individual Predictive Plots

```{r fig.width=11, fig.height=12}
indv_predictive_plots(post_dat_l, ids2)
#indv_predictive_plots(post_dat_l, idsBad)
```

### Subject 1
```{r fig.width=11, fig.height=12}

indv_predictive_dist((post_dat_l |> filter(rank<=200)),ind_fits_df, sbj=list(1))

abc_tables(post_dat |> filter(id==1))$agg_full |> flextable() |> set_caption("Mean from full posterior")
abc_tables(post_dat |> filter(id==1))$agg_best |> flextable() |> set_caption("Mean from best parameters")

# post_dat_l |> filter(id==1) |> group_by(condit,Model,Fit_Method,rank) |> 
#    summarise(mean_error=mean(abs(error)), n=n()) |> 
#    ggplot(aes(x=rank,y=mean_error,fill=Model))+geom_col()+facet_wrap(~Fit_Method)

# plot_indv_posterior(ind_fits_df |> filter(id==1))
# ind_fits_df |> filter(id==1, Fit_Method=="Test Only", Model=="EXAM") |> pull(c) |> unique()

```

### Subject 36
```{r fig.width=11, fig.height=12}
#| eval: false
#| 
indv_predictive_dist(post_dat_l,ind_fits_df, sbj=list(36))
#abc_tables(post_dat_l |> filter(id==36))$et_sum |> gt::gt()

abc_tables(post_dat |> filter(id==36))$agg_full |> flextable() |> set_caption("Mean from full posterior")
abc_tables(post_dat |> filter(id==36))$agg_best |> flextable() |> set_caption("Mean from best parameters")
# plot_indv_posterior(ind_fits_df |> filter(id==1))
# ind_fits_df |> filter(id==1, Fit_Method=="Test Only", Model=="EXAM") |> pull(c) |> unique()
```

## More Comparisons
```{r fig.width=11, fig.height=11}

best_id_x <- post_dat |> group_by(id,condit,Model,Fit_Method,x) |>
  mutate(e2=abs(y-pred)) |>
  summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2)) |>
  group_by(id,condit,Fit_Method,x) |> mutate(best=ifelse(mean_error==min(mean_error),1,0)) |>
  group_by(id,condit,Fit_Method,Model) |> mutate(n_best=sum(best))

best_id <- best_id_x |> group_by(id,condit,Fit_Method,Model) |>
  summarise(mean_error=mean(mean_error), n_best=first(n_best))

lowest_error_model <- best_id %>%
  group_by(id, condit,Fit_Method) %>%
  summarise(Best_Model = Model[which.min(mean_error)],
            n_best = n_best[which.min(mean_error)],
            Lowest_error = min(mean_error),
            differential = min(mean_error) - max(mean_error)) %>%
  ungroup()

error_difference <- best_id %>%
  select(id, condit, Model,Fit_Method, mean_error) %>%
  pivot_wider(names_from = Model, values_from = mean_error) %>%
  mutate(Error_difference = (ALM - EXAM))

full_comparison <- lowest_error_model |> left_join(error_difference, by=c("id","condit","Fit_Method"))  |> 
  group_by(condit,Fit_Method,Best_Model) |> mutate(nGrp=n(), model_rank = nGrp - rank(Error_difference) ) |> 
  arrange(Fit_Method,-Error_difference)

full_comparison |> round_tibble(1) |> select(-Error_difference,-nGrp) |> reactable(filterable=T,defaultPageSize=15)

ggplot(full_comparison, aes(x = Best_Model, fill = condit)) +
  geom_bar(position = "dodge") +
  facet_wrap(~Fit_Method)+
  labs(title = "Distribution of Subjects Best Fit by Model",
       x = "Model with Lowest Error",
       y = "Count of Subjects") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")

# Scatter plot showing the differential in mean error
ggplot(full_comparison, aes(x = Error_difference, y = condit, color = Best_Model)) +
  geom_point(alpha=.5) +
  stat_pointinterval()+
  facet_wrap(~Fit_Method)+
  labs(title = "Differential in Mean Error Between Models",
       x = "Error Difference (ALM - EXAM)",
       y = "Condition") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

# ggplot(full_comparison, aes(x = Error_difference)) +
#   geom_histogram(aes(fill=Best_Model),bins = 40, alpha = 0.7) +
#   geom_vline(aes(xintercept = mean(Error_difference)), color = "red", linetype = "dashed", size = 1) +
#   facet_wrap(~Fit_Method)+
#   labs(
#     title = "Distribution of Error Differential (ALM - EXAM)",
#     x = "Error Differential",
#     y = "Frequency"
#   ) +
#   theme_minimal()

best_exam <- full_comparison |> filter(Best_Model=="EXAM") 
best_alm <- full_comparison |> filter(Best_Model=="ALM")

d <- testAvg |> left_join(full_comparison, by=c("id","condit"))

d |> ggplot(aes(x=vb,y=dist,fill=condit)) + stat_bar + facet_wrap(Fit_Method~Best_Model,ncol=2)

d |> ggplot(aes(x=vb,y=vx,fill=condit)) + stat_bar + facet_wrap(Fit_Method~Best_Model,ncol=2)

d |> filter(Fit_Method=="Test") |>
  ggplot(aes(x=Error_difference,y=dist,col=condit)) + geom_point() + facet_wrap(~vb,scales="free",ncol=2)

d |> group_by(id,condit,Fit_Method) |> summarise(Error_difference=mean(Error_difference),dist=mean(dist)) |> 
  ggplot(aes(x=Error_difference,y=dist,col=condit)) + geom_point() + facet_wrap(~Fit_Method,scales="free")

# ds |> filter(expMode2=="Test", id %in% best_exam$id) |> 
#   group_by(id,condit,x) |> summarise(y=mean(y), dist=mean(abs(x-y))) |>
#   mutate(x=as.factor(x)) |>
#   ggplot(aes(x=x,y=dist,fill=condit)) + stat_bar

```

## Best Test - EXAM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Test",Best_Model=="EXAM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Test - ALM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Test",Best_Model=="ALM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Test_Train - EXAM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<1)
strong_fits |> filter(Fit_Method=="Test_Train",Best_Model=="EXAM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Test_Train - ALM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Test_Train",Best_Model=="ALM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Train - EXAM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Train",Best_Model=="EXAM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Train - ALM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Train",Best_Model=="ALM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Learning Curves
```{r fig.width=11, fig.height=16}
#| eval: true
#| fig-cap: "Learning Curves"
#| fig-width: 11
#| fig-height: 13

pd_train_l |> ggplot(aes(x=Block,y=dist,col=Resp))+
  geom_point( stat = "summary", fun = mean) + 
  stat_summary( geom = "line", fun = mean) +
  stat_summary(geom="errorbar",fun.data=mean_se,width=.1,alpha=.4) +
  #facet_wrap(condit~x,scales="free")
  ggh4x::facet_nested_wrap(Fit_Method~condit~x,scales="free") +
  #coord_cartesian(ylim = c(600, 1300)) +
  labs(title="Full Posterior")

pd_train_l |>filter(rank==1) |> ggplot(aes(x=Block,y=val,col=Resp))+
  geom_point( stat = "summary", fun = mean) + 
  stat_summary( geom = "line", fun = mean) +
  stat_summary(geom="errorbar",fun.data=mean_se,width=.1,alpha=.4) +
  #facet_wrap(condit~x,scales="free")
  ggh4x::facet_nested_wrap(Fit_Method~condit~x,scales="free") +
  coord_cartesian(ylim = c(600, 1300)) +labs(title="Best Parameters")

pd_train_l |> filter(Fit_Method=="Train") |> 
  mutate(x=as.factor(x), Resp=as.factor(Resp), Block=as.factor(Block)) |>
  ggplot(aes(x=x,y=val,fill=Block))+
  stat_bar +
  #facet_wrap(condit~x,scales="free")
  facet_wrap(condit~Resp,scales="free",ncol=3) +
  coord_cartesian(ylim = c(600, 1300))

# pd_train_l |> filter(id %in% c(1,33,66, 36,76), Fit_Method=="Train") |> 
#   ggplot(aes(x=Block,y=val,col=Resp))+
#   geom_point( stat = "summary", fun = mean) + 
#   stat_summary( geom = "line", fun = mean) +
#   stat_summary(geom="errorbar",fun.data=mean_se,width=.1,alpha=.4) +
#   #facet_wrap(condit~x,scales="free")
#   facet_wrap(id~x,scales="free",ncol=3) +
#   coord_cartesian(ylim = c(600, 1300))

# pd_train_l |> 
#   #filter(id %in% c(1,33,66, 36,76), Fit_Method=="Train") |> 
#   mutate(x=as.factor(x), Resp=as.factor(Resp), Block=as.factor(Block)) |>
#   ggplot(aes(x=x,y=dist,fill=Block))+
#   stat_bar +
#   #facet_wrap(condit~x,scales="free")
#    #coord_cartesian(ylim = c(600, 1300)) +
#   facet_wrap(id~Resp,scales="free",ncol=3)

```

```{r}
#| eval: true
# post_dat |> ggplot(aes(x=Model,y=mean_error,fill=Model))+geom_col()+facet_wrap(~Fit_Method)
# post_dat |> ggplot(aes(x=Model,y=mean_error,fill=Model))+geom_col()+facet_wrap(condit~Fit_Method)

```

```{r}
# print(file_name)
# print(getwd())
post_tabs <- map(post_tabs, ~cbind(.x,run_info) |> mutate(fn=file_name))
post_tabs$agg_pred_full_train <- pt_train$agg_pred_full
post_tabs$block_pred_full <- pt_train$block_pred_full
post_tabs$block_pred_full_x <- pt_train$block_pred_full_x
path1 = "../../../data/abc_tabs"
saveRDS(post_tabs, paste0(tools::file_path_sans_ext(basename(file_name)),"_post_tab", ".rds"))
saveRDS(post_tabs, paste0(path1, "/", tools::file_path_sans_ext(basename(file_name)),"_post_tab", ".rds"))

```

```{r}
#| include: false

rm(post_dat_l)
rm(post_dat)
rm(ind_fits_df)
gc()
```