### 2024-02-12 Model/abc_reject.qmd

---
title: Individual ABC Fits
author: Thomas Gorman
date: "`r Sys.Date()`"
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

### Approximate Bayesian Computation

Models were fit to each individual participant. To estimate parameters, we used approximate bayesian computation (ABC), enabling us to obtain an approximation of the posterior distribution of the generalization and learning rate parameters for each individual. ABC belongs to the class of simulation-based inference methods [@cranmerFrontierSimulationbasedInference2020], which have begun being used for parameter estimation in cognitive modeling relatively recently [@turnerTutorialApproximateBayesian2012; @turnerBayesianAnalysisSimulationbased2016; @kangasraasioParameterInferenceComputational2019]. Although they can be applied to any model from which data can be simulated, ABC methods are most useful for complex models that lack a well defined likelihood function (e.g. many neural network and evidence accumulation models). The general approach of ABC is to 1) define a prior distribution over model parameters; 2) sample parameter values from the prior, which are used to simulate data from the model; 3) compute a measure of discrepancy between the simulated dataset and the observed dataset. 4) Establish a minimum tolerance, or degree of discrepancy, which determines whether a set of parameters are accepted into the posterior distribution.

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools)
conflict_prefer_all("dplyr", quiet = TRUE)
options(scipen = 999)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
test <- readRDS(here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Test") 
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

input_layer <<- output_layer <<-  c(100,350,600,800,1000,1200)

ids <- c(1,2,4,5,6,7,8, 10,11,12,13)
ids2 <- c(1,66,36)
ids3 <- c(20,71,101,4,76,192)
idsBad <- c(76,192, 101)

#file_name <- "n_iter_300_ntry_3000_0800"
#file_name <- "n_iter_100_ntry_200_4509"
file_name <- "n_iter_100_ntry_400_3247"
file_name <- "n_iter_200_ntry_300_5354"

# list.files(here('data/abc_reject'))
# (grep("Train",list.files(here(paste0('data/abc_reject/',file_name)),
#                                            pattern="EXAM_Test",full.names = TRUE),
#                                 invert=TRUE, value=TRUE))

ind_fits <- map(list.files(here(paste0('data/abc_reject/'),file_name),full.names=TRUE), readRDS)
# ind_fits <- map(list.files(here('data/abc_reject/n_iter_2000_ntry_10_2918'),full.names=TRUE), readRDS)
ind_fits_df <- ind_fits |> map(~list(dat=.x[[1]], Model = .x[["Model"]], Fit_Method=.x[["Fit_Method"]]))
ind_fits_df <- ind_fits_df |> map(~rbindlist(.x$dat) |> mutate(Model = .x$Model, Fit_Method = .x$Fit_Method)) |> rbindlist()

#run_params <- tibble::lst(file_name,cMean=ind_fits[[1]]$cMean, cSig=ind_fits[[1]]$cSig,

run_params <- keep(ind_fits[[1]], ~any(class(.x) %in% c("character", "numeric")))
run_params <- tibble(!!!run_params) |> select(-Model,-Fit_Method)

extract_info <- function(raw_names, run_params) {
  fname <- tools::file_path_sans_ext(basename(raw_names))
  n_samp <- str_extract(raw_names, "(?<=_)\\d+(?=_)")
  ntry_ <- str_extract(raw_names, "(?<=ntry_)\\d+")
  type <- 'ss'
  data.frame(run_params, n_samp, ntry_, type, fname)
}

run_info <- extract_info(file_name,run_params)

```

```{r}

generate_data <- function(Model, post_samples, data, num_samples = 1, return_dat = "train_data, test_data") {
  # Filter data for the specific id without invalidating selfref
  sbj_data <- copy(data[id == post_samples$id[1]])
  simulation_function <- ifelse(Model == "EXAM", full_sim_exam, full_sim_alm)

  target_data <- switch(return_dat,
                        "test_data" = copy(sbj_data[expMode2 == "Test"]),
                        "train_data" = copy(sbj_data[expMode2 == "Train"]),
                        "train_data, test_data" = copy(sbj_data[expMode2 %in% c("Test", "Train")]))

  post_samples <- post_samples[order(mean_error)][1:num_samples, .(c, lr, mean_error, rank = .I)]

  simulated_data_list <- lapply(1:nrow(post_samples), function(i) {
    params <- post_samples[i]
    sim_data <- simulation_function(sbj_data, params$c, params$lr, input_layer = input_layer, 
                                    output_layer = output_layer, return_dat = return_dat)
    sim_data_dt <- data.table(id = sbj_data$id[1], condit = sbj_data$condit[1], 
                              expMode2 = target_data$expMode2, Model = Model,tr=target_data$tr,
                              y = target_data$y, x = target_data$x, c = params$c, 
                              lr = params$lr, mean_error = params$mean_error, rank = i,
                              pred = sim_data)
    return(sim_data_dt)
  })

  result_dt <- rbindlist(simulated_data_list)
  setcolorder(result_dt, c("id", "condit", "expMode2","tr", "c", "lr", "x", "y", "pred"))
  return(result_dt)
}

future::plan(multisession)

nestSbjModelFit <- ind_fits_df %>% nest(.by=c(id,Model,Fit_Method))

# organize test data predictions
post_dat <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
   generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 50, return_dat="test_data")
   })) |> 
  select(Fit_Method,pp,-data) |>  
  unnest(pp) |>  filter(expMode2=="Test") |> as.data.table()

post_dat_avg <- post_dat |> group_by(id, condit, Model, Fit_Method, x, c, lr, rank) |> 
  summarise(y = mean(y), pred = mean(pred), error = y - pred) |> as.data.table()

setorder(post_dat_avg, id, x, rank)
post_dat_l <- melt(post_dat_avg, id.vars = c("id", "condit", "Model", "Fit_Method", "x", "c", "lr", "rank","error"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val")
post_dat_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))]
setorder(post_dat_l, id, Resp)
rm(post_dat_avg)

# organize training data predictions
pd_train <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
   generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 20, return_dat="train_data")
   })) |>
  select(Fit_Method,pp,-data) |>
  unnest(pp) |> as.data.table() |> filter(expMode2=="Train")

nbins <- 3
pd_train <- pd_train |> group_by(id,condit,Model,Fit_Method) |>
  mutate(Block=cut(tr,breaks=seq(1,max(tr), length.out=nbins+1),include.lowest=TRUE,labels=FALSE))
setorder(pd_train, id, x,Block, rank)

pd_train_l <- melt(pd_train, id.vars = c("id", "condit", "Model","Block", "Fit_Method", "x", "c", "lr", "rank"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val") |> as.data.table()
pd_train_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))] 
setorder(pd_train_l, id,Block, Resp)

pd_train_l <- pd_train_l |> mutate(dist=abs(x-val))

plan(sequential)

# AIC and BIC
# mc2 <- post_dat |> filter(rank==1) |>
#   group_by(id,condit,Model,Fit_Method) |>
#   mutate(e2=abs(y-pred),n=n(),k=2) |>
#   group_by(id,condit,Model,Fit_Method,x) |>
#   summarise(y=mean(y), pred=mean(pred), e2=mean(e2), mean_error=first(mean_error),n=first(n),k=first(k)) |>
#   group_by(id,condit,Model,Fit_Method) |>
#   summarise(e2=mean(e2), mean_error=first(mean_error),n=first(n),k=first(k)) |>
#   mutate(AIC=2*k+n*mean_error, BIC=log(n)*k+n*mean_error)

# Single run extraction:
# exam_test <- ind_fits_df |> filter(Model == "EXAM", Fit_Method == "Test")
# post_dat_trial <- exam_test %>% split(f =c(.$id), drop=TRUE) |> 
#   map(~generate_data(.x$Model, .x, ds, num_samples = 15, return_dat="train_data, test_data")) |> 
#   rbindlist() |> 
#   filter(expMode2=="Test")
# 
# post_dat_avg <- post_dat_trial |> group_by(id,condit,x,c,lr,rank) |> 
#   summarise(y = mean(y),pred = mean(pred)) |> arrange(id,x,rank)

#post_dat_l |> filter(Model == "EXAM", Fit_Method == "Test",id==1) |> arrange(rank)
# head(post_dat_l)

```

#### `{r} kable(run_info)`

##  Posterior Average Table 
```{r fig.width=12, fig.height=17}
#| eval: true
#| tbl-cap: "Posterior Distribution"
#| tbl-subcap: 
#| - "Full Posterior"
#| - "Best Posterior"

# post_tabs <- abc_tables(post_dat_l)
# post_tabs$et_sum |> gt::gt()

post_tabs <- abc_tables(post_dat,post_dat_l)

# tables with fit error
# post_tabs$agg_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
#                        `ME` = as_paragraph(mean_error)) |> as_flextable() |>
#   set_caption("Mean from full_posterior")

# post_tabs$agg_best |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
#                        `ME` = as_paragraph(mean_error)) |> as_flextable() |>
#   set_caption("Mean from best parameters")

post_tabs$agg_pred_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

post_tabs$agg_pred_best |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

```

## Interactive Test Predictions Table
```{r}
tags$select(
  tags$option(value = "", "All"),
  purrr::map(unique(post_tabs$agg_x_full$condit), tags$option),
    purrr::map(unique(post_tabs$agg_x_full$Model), tags$option),
    purrr::map(unique(post_tabs$agg_x_full$Fit_Method), tags$option),
   purrr::map(unique(post_tabs$agg_x_full$x), tags$option)
)

selectFilter <- function(tableId, style = "width: 100%; height: 100%;") {
  function(values, name) {
    tags$select(
      # Set to undefined to clear the filter
      onchange = sprintf("
        const value = event.target.value
        Reactable.setFilter('%s', '%s', value === '__ALL__' ? undefined : value)
      ", tableId, name),
      # "All" has a special value to clear the filter, and is the default option
      tags$option(value = "__ALL__", "All"),
      lapply(unique(values), tags$option),
      "aria-label" = sprintf("Filter %s", name),
      style = style
    )
  }
}

post_tabs$agg_x_full |> 
  left_join(post_tabs$agg_x_best |> rename("best_error"=mean_error,"best_pred"=pred), 
            by=c("condit","Model","Fit_Method","x","y")) |> 
  arrange(desc(condit),x,Fit_Method) |>
  mutate(x=as.character(x)) |>
  reactable::reactable(
    columns = list(condit=colDef(name="Condit",filterInput=selectFilter("my-tbl")),
                   Model=colDef(name="Model",filterInput=selectFilter("my-tbl")),
                   Fit_Method=colDef(name="Fit_Method",filterInput=selectFilter("my-tbl"),
                                     filterMethod = JS("(rows, columnId, filterValue) => {
        return rows.filter(row => row.values[columnId] === filterValue)
      }")),
      x=colDef(name="x",filterInput=selectFilter("my-tbl"),
               filterMethod = JS("(rows, columnId, filterValue) => {
        return rows.filter(row => row.values[columnId] === filterValue)
      }"))),
    elementId = "my-tbl",
    defaultPageSize=15,
    filterable=TRUE) 
```

## Train Predictions
```{r}

pt_train <- abc_train_tables(pd_train,pd_train_l)
pt_train$agg_pred_full |> flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"), 
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

pt_train$block_pred_full |> 
  flextable::tabulator(rows=c("Fit_Method","Model","Block"), columns=c("condit"), 
                       `Observed` = as_paragraph(y), 
                       `Predicted` = as_paragraph(pred),
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

pt_train$block_pred_full |> filter(condit=="Varied") |>
  flextable::tabulator(rows=c("Fit_Method","Block"), columns=c("Model"), 
                      # `Observed` = as_paragraph(y), 
                       `Predicted` = as_paragraph(pred),
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |>
  set_caption("Mean from full_posterior")

best_indices <- pt_train$block_pred_full_x %>%
   filter(!(Fit_Method=="Test")) |>
  nest(data = -condit) %>%
  mutate(indices = map(data, ~{
    .x %>%
      select(-y, -pred, -mean_error) %>%
      pivot_wider(names_from = Model, values_from = best) %>%
      ungroup() |>
      select(ALM, EXAM) %>%
      pmap(.f = function(ALM, EXAM) {
        which(c(ALM, EXAM) == 1)
      })
  })) %>%
  select(condit, indices)

pt_train$block_pred_full_x |> filter(condit=="Varied",!(Fit_Method=="Test")) |>
  flextable::tabulator(rows=c("Fit_Method","Block","x"), columns=c("Model"), 
                      # `Observed` = as_paragraph(y), 
                       `Predicted` = as_paragraph(pred),
                       `ME` = as_paragraph(mean_error)) |> 
  as_flextable() %>% 
  apply_best_formatting(.,pull(filter(best_indices,condit=="Varied"),indices) %>% .[[1]]) |> 
  set_caption("Vared Train Predictions")

pt_train$block_pred_full_x |> filter(condit=="Constant",!(Fit_Method=="Test")) |>
  flextable::tabulator(rows=c("Fit_Method","Block","x"), columns=c("Model"), 
                      # `Observed` = as_paragraph(y), 
                       `Predicted` = as_paragraph(pred),
                       `ME` = as_paragraph(mean_error)) |> 
  as_flextable() %>% 
  apply_best_formatting(.,pull(filter(best_indices,condit=="Constant"),indices) %>% .[[1]]) |> 
  set_caption("Constant Train Predictions")

# pull(filter(best_indices,condit=="Varied"),indices) %>% .[[1]]

```

## Posterior Predictive: 
```{r, fig.width=12, fig.height=15}
#| eval: true
group_predictive_plots(post_dat_l) 
```

### Alt grouping of Posterior Predictive
```{r, fig.width=12, fig.height=15}
 group_predictive_plots2(post_dat_l) 
```

```{r fig.width=11, fig.height=12}
plot_indv_posterior(ind_fits_df |> mutate(Group=condit))

plot_indv_posterior(post_dat |> filter(rank==1) |> mutate(Group=condit))

# plot join density of c and lr
# post_dat |> filter(rank==1, c<.001) |> 
#   ggplot(aes(x=c, y=lr, color=condit)) + geom_point() + facet_wrap(~Model+Fit_Method)

post_dat |>  filter(c<.001) |> ggplot(aes(x=c,y=condit,fill=condit)) + geom_boxploth(width=.5) + facet_wrap(~Model+Fit_Method,scales="free")

```

## Model Comparison:
```{r, fig.width=12, fig.height=13}
#| eval: true
indv_best_plots(post_dat_l)

post_dat_l |> group_by(condit,Model,Fit_Method,rank) |> 
   summarise(mean_error=mean(abs(error)), n=n()) |> 
   ggplot(aes(x=rank,y=mean_error,fill=Model))+geom_col()+facet_wrap(~Fit_Method,scales="free")

post_dat_l |> group_by(condit,Model,Fit_Method,rank,x) |> 
   summarise(mean_error=mean(abs(error)), n=n()) |> 
   ggplot(aes(x=rank,y=mean_error,fill=Model))+geom_col()+facet_wrap(Fit_Method~x,ncol=6,scales="free")

group_best_plots(post_dat_l)

```

## Individual Predictive Plots

```{r fig.width=11, fig.height=12}
indv_predictive_plots(post_dat_l, ids2)
indv_predictive_plots(post_dat_l, idsBad)
```

### Subject 1
```{r fig.width=11, fig.height=12}

indv_predictive_dist((post_dat_l |> filter(rank<=200)),ind_fits_df, sbj=list(1))

abc_tables(post_dat |> filter(id==1))$agg_full |> flextable() |> set_caption("Mean from full posterior")
abc_tables(post_dat |> filter(id==1))$agg_best |> flextable() |> set_caption("Mean from best parameters")

post_dat_l |> filter(id==1) |> group_by(condit,Model,Fit_Method,rank) |> 
   summarise(mean_error=mean(abs(error)), n=n()) |> 
   ggplot(aes(x=rank,y=mean_error,fill=Model))+geom_col()+facet_wrap(~Fit_Method)

# plot_indv_posterior(ind_fits_df |> filter(id==1))
# ind_fits_df |> filter(id==1, Fit_Method=="Test Only", Model=="EXAM") |> pull(c) |> unique()

```

### Subject 36
```{r fig.width=11, fig.height=12}

indv_predictive_dist(post_dat_l,ind_fits_df, sbj=list(36))
#abc_tables(post_dat_l |> filter(id==36))$et_sum |> gt::gt()

abc_tables(post_dat |> filter(id==36))$agg_full |> flextable() |> set_caption("Mean from full posterior")
abc_tables(post_dat |> filter(id==36))$agg_best |> flextable() |> set_caption("Mean from best parameters")
# plot_indv_posterior(ind_fits_df |> filter(id==1))
# ind_fits_df |> filter(id==1, Fit_Method=="Test Only", Model=="EXAM") |> pull(c) |> unique()
```

## More Comparisons
```{r fig.width=11, fig.height=12}

best_id_x <- post_dat |> group_by(id,condit,Model,Fit_Method,x) |>
  mutate(e2=abs(y-pred)) |>
  summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2)) |>
  group_by(id,condit,Fit_Method,x) |> mutate(best=ifelse(mean_error==min(mean_error),1,0)) |>
  group_by(id,condit,Fit_Method,Model) |> mutate(n_best=sum(best))

best_id <- best_id_x |> group_by(id,condit,Fit_Method,Model) |>
  summarise(mean_error=mean(mean_error), n_best=first(n_best))

lowest_error_model <- best_id %>%
  group_by(id, condit,Fit_Method) %>%
  summarise(Best_Model = Model[which.min(mean_error)],
            n_best = n_best[which.min(mean_error)],
            Lowest_error = min(mean_error),
            differential = min(mean_error) - max(mean_error)) %>%
  ungroup()

error_difference <- best_id %>%
  select(id, condit, Model,Fit_Method, mean_error) %>%
  pivot_wider(names_from = Model, values_from = mean_error) %>%
  mutate(Error_difference = (ALM - EXAM))

full_comparison <- lowest_error_model |> left_join(error_difference, by=c("id","condit","Fit_Method"))  |> 
  group_by(condit,Fit_Method,Best_Model) |> mutate(nGrp=n(), model_rank = nGrp - rank(Error_difference) ) |> 
  arrange(Fit_Method,-Error_difference)

full_comparison |> round_tibble(1) |> reactable(filterable=T,defaultPageSize=15)

ggplot(full_comparison, aes(x = Best_Model, fill = condit)) +
  geom_bar(position = "dodge") +
  facet_wrap(~Fit_Method)+
  labs(title = "Distribution of Subjects Best Fit by Model",
       x = "Model with Lowest Error",
       y = "Count of Subjects") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")

# Scatter plot showing the differential in mean error
ggplot(full_comparison, aes(x = Error_difference, y = condit, color = Best_Model)) +
  geom_point(alpha=.5) +
  stat_pointinterval()+
  facet_wrap(~Fit_Method)+
  labs(title = "Differential in Mean Error Between Models",
       x = "Error Difference (ALM - EXAM)",
       y = "Condition") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

# ggplot(full_comparison, aes(x = Error_difference)) +
#   geom_histogram(aes(fill=Best_Model),bins = 40, alpha = 0.7) +
#   geom_vline(aes(xintercept = mean(Error_difference)), color = "red", linetype = "dashed", size = 1) +
#   facet_wrap(~Fit_Method)+
#   labs(
#     title = "Distribution of Error Differential (ALM - EXAM)",
#     x = "Error Differential",
#     y = "Frequency"
#   ) +
#   theme_minimal()

best_exam <- full_comparison |> filter(Best_Model=="EXAM") 
best_alm <- full_comparison |> filter(Best_Model=="ALM")

d <- testAvg |> left_join(full_comparison, by=c("id","condit"))

d |> ggplot(aes(x=vb,y=dist,fill=condit)) + stat_bar + facet_wrap(Fit_Method~Best_Model,ncol=2)

d |> ggplot(aes(x=vb,y=vx,fill=condit)) + stat_bar + facet_wrap(Fit_Method~Best_Model,ncol=2)

d |> filter(Fit_Method=="Test") |>
  ggplot(aes(x=Error_difference,y=dist,col=condit)) + geom_point() + facet_wrap(~vb,scales="free",ncol=2)

d |> group_by(id,condit,Fit_Method) |> summarise(Error_difference=mean(Error_difference),dist=mean(dist)) |> 
  ggplot(aes(x=Error_difference,y=dist,col=condit)) + geom_point() + facet_wrap(~Fit_Method,scales="free")

# ds |> filter(expMode2=="Test", id %in% best_exam$id) |> 
#   group_by(id,condit,x) |> summarise(y=mean(y), dist=mean(abs(x-y))) |>
#   mutate(x=as.factor(x)) |>
#   ggplot(aes(x=x,y=dist,fill=condit)) + stat_bar

```

## Best Test - EXAM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Test",Best_Model=="EXAM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Test - ALM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Test",Best_Model=="ALM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Test_Train - EXAM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<1)
strong_fits |> filter(Fit_Method=="Test_Train",Best_Model=="EXAM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Test_Train - ALM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Test_Train",Best_Model=="ALM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Train - EXAM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Train",Best_Model=="EXAM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Best Train - ALM Fits
```{r fig.width=11, fig.height=12}

strong_fits <- full_comparison |> filter(model_rank<=1)
strong_fits |> filter(Fit_Method=="Train",Best_Model=="ALM") |> pull(id) %>% indv_predictive_plots(post_dat_l, .)
```

## Learning Curves
```{r fig.width=11, fig.height=16}
#| fig-cap: "Learning Curves"
#| fig-width: 11
#| fig-height: 13

pd_train_l |> ggplot(aes(x=Block,y=dist,col=Resp))+
  geom_point( stat = "summary", fun = mean) + 
  stat_summary( geom = "line", fun = mean) +
  stat_summary(geom="errorbar",fun.data=mean_se,width=.1,alpha=.4) +
  #facet_wrap(condit~x,scales="free")
  ggh4x::facet_nested_wrap(Fit_Method~condit~x,scales="free") +
  #coord_cartesian(ylim = c(600, 1300)) +
  labs(title="Full Posterior")

pd_train_l |>filter(rank==1) |> ggplot(aes(x=Block,y=val,col=Resp))+
  geom_point( stat = "summary", fun = mean) + 
  stat_summary( geom = "line", fun = mean) +
  stat_summary(geom="errorbar",fun.data=mean_se,width=.1,alpha=.4) +
  #facet_wrap(condit~x,scales="free")
  ggh4x::facet_nested_wrap(Fit_Method~condit~x,scales="free") +
  coord_cartesian(ylim = c(600, 1300)) +labs(title="Best Parameters")

pd_train_l |> filter(Fit_Method=="Train") |> 
  mutate(x=as.factor(x), Resp=as.factor(Resp), Block=as.factor(Block)) |>
  ggplot(aes(x=x,y=val,fill=Block))+
  stat_bar +
  #facet_wrap(condit~x,scales="free")
  facet_wrap(condit~Resp,scales="free",ncol=3) +
  coord_cartesian(ylim = c(600, 1300))

# pd_train_l |> filter(id %in% c(1,33,66, 36,76), Fit_Method=="Train") |> 
#   ggplot(aes(x=Block,y=val,col=Resp))+
#   geom_point( stat = "summary", fun = mean) + 
#   stat_summary( geom = "line", fun = mean) +
#   stat_summary(geom="errorbar",fun.data=mean_se,width=.1,alpha=.4) +
#   #facet_wrap(condit~x,scales="free")
#   facet_wrap(id~x,scales="free",ncol=3) +
#   coord_cartesian(ylim = c(600, 1300))

pd_train_l |> 
  #filter(id %in% c(1,33,66, 36,76), Fit_Method=="Train") |> 
  mutate(x=as.factor(x), Resp=as.factor(Resp), Block=as.factor(Block)) |>
  ggplot(aes(x=x,y=dist,fill=Block))+
  stat_bar +
  #facet_wrap(condit~x,scales="free")
   #coord_cartesian(ylim = c(600, 1300)) +
  facet_wrap(id~Resp,scales="free",ncol=3)

```

```{r}
#| eval: false
post_dat |> ggplot(aes(x=Model,y=mean_error,fill=Model))+geom_col()+facet_wrap(~Fit_Method)
post_dat |> ggplot(aes(x=Model,y=mean_error,fill=Model))+geom_col()+facet_wrap(condit~Fit_Method)

```

```{r}

# post_tabs <- map(post_tabs, ~cbind(.x,run_info) |> mutate(fn=file_name))
post_tabs$agg_pred_full_train <- pt_train$agg_pred_full
post_tabs$block_pred_full <- pt_train$block_pred_full
post_tabs$block_pred_full_x <- pt_train$block_pred_full_x

# # path1 = "../../../data/abc_tabs"
# saveRDS(post_tabs, paste0(path1, "/", tools::file_path_sans_ext(basename(file_name)),"_post_tab", ".rds"))
# saveRDS(post_tabs, paste0(tools::file_path_sans_ext(basename(file_name)),"_post_tab", ".rds"))
```