### 2024-02-24 Model/htw_model_e2.qmd

---
title: HTW Model e2
author: Thomas Gorman
date: "`r Sys.Date()`"
page-layout: full
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools)
#conflict_prefer_all("dplyr", quiet = TRUE)
options(scipen = 999)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
```

### Modelling Results

```{r}
#| cache: false

#ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
ds <- readRDS(here::here("data/e2_md_02-23-24.rds"))  |> as.data.table()
nbins <- 3

fd <- readRDS(here("data/e2_08-21-23.rds"))
test <- fd |> filter(expMode2 == "Test") 
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

trainAvg <- fd |> filter(expMode2 == "Train") |> group_by(id) |> 
  mutate(tr=trial,x=vb,Block=case_when(expMode2=="Train" ~ cut(tr,breaks=seq(1,max(tr), length.out=nbins+1),include.lowest=TRUE,labels=FALSE),
                                         expMode2=="Test" ~ 4)) |> 
  group_by(id,condit,vb,x,Block) |> 
  summarise(dist=mean(dist),y=mean(vx))

input_layer <<- output_layer <<-  c(100,350,600,800,1000,1200)
ids2 <- c(1,66,36)

file_name <- "e2_n_iter_50_ntry_200_2506"
#file_name <- "n_iter_400_ntry_100_2944"

ind_fits <- map(list.files(here(paste0('data/abc_reject/'),file_name),full.names=TRUE), readRDS)
ind_fits_df <- ind_fits |> map(~list(dat=.x[[1]], Model = .x[["Model"]], Fit_Method=.x[["Fit_Method"]]))
ind_fits_df <- ind_fits_df |> map(~rbindlist(.x$dat) |> mutate(Model = .x$Model, Fit_Method = .x$Fit_Method)) |> rbindlist() 
```

```{r}
#| cache: false

generate_data <- function(Model, post_samples, data, num_samples = 1, return_dat = "train_data, test_data") {
  # Filter data for the specific id without invalidating selfref
  sbj_data <- copy(data[id == post_samples$id[1]])
  simulation_function <- ifelse(Model == "EXAM", full_sim_exam, full_sim_alm)

  target_data <- switch(return_dat,
                        "test_data" = copy(sbj_data[expMode2 == "Test"]),
                        "train_data" = copy(sbj_data[expMode2 == "Train"]),
                        "train_data, test_data" = copy(sbj_data[expMode2 %in% c("Test", "Train")]))

  post_samples <- post_samples[order(mean_error)][1:num_samples, .(c, lr, mean_error, rank = .I)]

  simulated_data_list <- lapply(1:nrow(post_samples), function(i) {
    params <- post_samples[i]
    sim_data <- simulation_function(sbj_data, params$c, params$lr, input_layer = input_layer, 
                                    output_layer = output_layer, return_dat = return_dat)
    sim_data_dt <- data.table(id = sbj_data$id[1], condit = sbj_data$condit[1], 
                              expMode2 = target_data$expMode2, Model = Model,tr=target_data$tr,
                              y = target_data$y, x = target_data$x, c = params$c, 
                              lr = params$lr, mean_error = params$mean_error, rank = i,
                              pred = sim_data)
    return(sim_data_dt)
  })

  result_dt <- rbindlist(simulated_data_list)
  setcolorder(result_dt, c("id", "condit", "expMode2","tr", "c", "lr", "x", "y", "pred"))
  return(result_dt)
}

#future::plan(multisession)

nestSbjModelFit <- ind_fits_df %>% nest(.by=c(id,Model,Fit_Method))

# organize test data predictions
 post_dat <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
    generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 50, return_dat="test_data")
    })) |> 
   select(Fit_Method,pp,-data) |>  
   unnest(pp) |>  filter(expMode2=="Test") |> as.data.table()
# 
# saveRDS(post_dat, here("data/model_cache/post_dat.rds"))

#post_dat <- readRDS(here("data/model_cache/post_dat.rds"))

post_dat_avg <- post_dat |> group_by(id, condit, Model, Fit_Method, x, c, lr, rank) |> 
  mutate(error2 = y - pred) |>
  summarise(y = mean(y), pred = mean(pred), error = y - pred, error2=mean(error2)) |> as.data.table()

setorder(post_dat_avg, id, x, rank)
post_dat_l <- melt(post_dat_avg, id.vars = c("id", "condit", "Model", "Fit_Method", "x", "c", "lr", "rank","error"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val")
post_dat_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))]
setorder(post_dat_l, id, Resp)
#rm(post_dat_avg)

# organize training data predictions
 pd_train <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
   generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 20, return_dat="train_data")
    })) |>
   select(Fit_Method,pp,-data) |>
  unnest(pp) |> as.data.table() |> filter(expMode2=="Train")

#saveRDS(pd_train, here("data/model_cache/pd_train.rds"))

#pd_train <- readRDS(here("data/model_cache/pd_train.rds"))

nbins <- 3
pd_train <- pd_train |> group_by(id,condit,Model,Fit_Method) |>
  mutate(Block=cut(tr,breaks=seq(1,max(tr), length.out=nbins+1),include.lowest=TRUE,labels=FALSE))
setorder(pd_train, id, x,Block, rank)

pd_train_l <- melt(pd_train, id.vars = c("id", "condit", "Model","Block", "Fit_Method", "x", "c", "lr", "rank"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val") |> as.data.table()
pd_train_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))] 
setorder(pd_train_l, id,Block, Resp)

pd_train_l <- pd_train_l  |>
  mutate(dist = case_when(
    val >= x & val <= x + 200 ~ 0,                 
    val < x ~ abs(x - val),                       
    val > x + 200 ~ abs(val - (x + 200)),           
    TRUE ~ NA_real_                                 
  ))

#plan(sequential)

```

### Group level aggregations

```{r}
#| eval: true
#| label: tbl-htw-modelError
#| tbl-cap: "Mean model errors predicting testing data, aggregated over all participants and velocity bands. Note that Fit Method refers to how model parameters were optimized, while error values reflect mean absolute error for the 6 testing bands"
post_tabs <- abc_tables(post_dat,post_dat_l)

post_tabs$agg_pred_full |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) |>
  flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"),
                       `ME` = as_paragraph(mean_error)) |> as_flextable()

#post_tabs$agg_pred_full |> pander::pandoc.table()

```

```{r fig.height=12,fig.width=11}
#| label: fig-htw-resid-pred
#| column: page-inset-right
#| fig-cap: A) Model residuals for each combination of training condition, fit method, and model. Residuals reflect the difference between observed and predicted values. Lower values indicate better model fit. Note that y axes are scaled differently between facets. B) Full posterior predictive distributions vs. observed data from participants.Points represent median values, thicker intervals represent 66% credible intervals and thin intervals represent 95% credible intervals around the median. 
#| fig-height: 19
#| fig-width: 12
##| layout: [[45,-5, 45], [100]]
##| fig-subcap: ["Model Residuals - training data", "Model Residuals - testing data","Full posterior predictive distributions vs. observed data from participants."]
train_resid <- pd_train |> group_by(id,condit,Model,Fit_Method, Block) |> 
  summarise(y = mean(y), pred = mean(pred), error = y - pred) |>
  ggplot(aes(x = Block, y = abs(error), fill=Model)) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, scales="free",ncol=2) +
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  labs(title="Model Residual Errors - Training Stage", y="RMSE", x= "Training Block") +
  theme(legend.title = element_blank(), legend.position="top")

test_resid <-  post_dat |> 
   group_by(id,condit,x,Model,Fit_Method,rank) |>
   summarize(error=mean(abs(y-pred)),n=n()) |>
   group_by(id,condit,x,Model,Fit_Method) |>
   summarize(error=mean(error)) |>
  mutate(vbLab = factor(paste0(x,"-",x+200))) |>
  ggplot(aes(x = vbLab, y = abs(error), fill=Model)) + 
  stat_bar + 
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, axes = "all",ncol=2,scale="free") +
  labs(title="Model Residual Errors - Testing Stage",y="RMSE", x="Velocity Band") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

group_pred <- post_dat_l |> 
  mutate(vbLab = factor(paste0(x,"-",x+200),levels=levels(testAvg$vb))) |>
  ggplot(aes(x=val,y=vbLab,col=Resp)) + 
  stat_pointinterval(position=position_dodge(.5), alpha=.9) + 
  scale_color_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, axes = "all",ncol=2,scale="free") +
  labs(title="Posterior Predictions - Testing Stage",y="Velocity Band (lower bound)", x="X Velocity") +
theme(legend.title=element_blank(),axis.text.y = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

((train_resid | test_resid) / group_pred) +
  plot_layout(heights=c(1,1.5)) & 
  plot_annotation(tag_levels = list(c('A1','A2','B')),tag_suffix = ') ') & 
  theme(plot.tag.position = c(0, 1))

```

```{r}
#| label: fig-htw-post-dist
#| fig-cap: Posterior Distributions of $c$ and $lr$ parameters. Points represent median values, thicker intervals represent 66% credible intervals and thin intervals represent 95% credible intervals around the median. Note that the y axes of the plots for the c parameter are scaled logarithmically.
#| fig-height: 7
#| fig-width: 11

c_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=log(c), x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.2)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="c parameter") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4))

lr_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=lr, x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.4)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="learning rate parameter") +
  theme(legend.title = element_blank(), legend.position = "none",plot.title=element_text(hjust=.5))
c_post + lr_post
```

### Accounting for individual patterns

To more accurately assess the relative abilities of ALM and EXAM to capture important empirical patterns - we will now examine the predictions of both models for the subset of individual participants shown in @fig-htw-indv-pred. Panel A presents three varied and constant participants who demonstrated a reasonable degree of discrimination between the 6 velocity bands during testing.

-   \*\* comment on the different ways ALM can completely fail to mimic discrimination patterns (sbj. 35; sbj. 137),and on how it can sometimes partially succeed (sbj. 11; 14,74)

-   \*\* comment on how EXAM can somtimes mimic non-monotonic spacing between bands due to associative stregth from training (i.e. subject 47)

-   \*\* compare c values to slope parameters from the statistical models earlier in paper

```{r  fig.width=11, fig.height=10}
#| label: fig-htw-indv-pred
#| fig-cap: Model predictions alongside observed data for a subset of individual participants. A) 3 constant and 3 varied participants fit to both the test and training data. B) 3 constant and 3 varied subjects fit to only the trainign data. 
#| fig-height: 13
#| fig-width: 12

cId_tr <- c(137, 181, 11)
vId_tr <- c(14, 193, 47)
cId_tt <- c(11, 93, 35)
vId_tt <- c(1,14,74)
# filter(id %in% (filter(bestTestEXAM,group_rank<=9, Fit_Method=="Test")

testIndv <- post_dat_l |> filter(id %in% c(cId_tt,vId_tt), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Individual Participant fits from Test & Train Fitting Method",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

trainIndv <- post_dat_l |> filter(id %in% c(cId_tr,vId_tr), Fit_Method=="Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp), flab=paste0("Subject: ",id)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Individual Participant fits from Train Only Fitting Method", y="X Velocity",
       fill="Target Velocity") +
     guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

(testIndv  / trainIndv) +
  plot_annotation(tag_levels = list(c('A','B')),tag_suffix = ') ') & 
  theme(plot.tag.position = c(0, 1))

```

```{r}
#| label: fig-htw-best-model
#| fig-cap: Difference in model errors for each participant, with models fit to both train and test data. Positive values favor EXAM, while negative values favor ALM. 
#| fig-height: 9
#| fig-width: 11

# could compute best model for each posterior parameter - examine consistency
# then I'd have an error bar for each subject in the model error diff. figure

tid1 <- post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
  mutate(e2=abs(y-pred)) |> 
  summarise(y1=mean(y), pred1=mean(pred),mean_error=abs(y1-pred1)) |>
  group_by(id,condit,Model,Fit_Method) |> 
  summarise(mean_error=mean(mean_error)) |> 
  arrange(id,condit,Fit_Method) |>
  round_tibble(1)

best_id <- tid1 |> 
  group_by(id,condit,Fit_Method) |> mutate(best=ifelse(mean_error==min(mean_error),1,0))

lowest_error_model <- best_id %>%
  group_by(id, condit,Fit_Method) %>%
  summarise(Best_Model = Model[which.min(mean_error)],
            Lowest_error = min(mean_error),
            differential = min(mean_error) - max(mean_error)) %>%
  ungroup()

error_difference<- best_id %>%
  select(id, condit, Model,Fit_Method, mean_error) %>%
  pivot_wider(names_from = Model, values_from = c(mean_error)) %>%
  mutate(Error_difference = (ALM - EXAM))

full_comparison <- lowest_error_model |> left_join(error_difference, by=c("id","condit","Fit_Method"))  |> 
  group_by(condit,Fit_Method,Best_Model) |> mutate(nGrp=n(), model_rank = nGrp - rank(Error_difference) ) |> 
  arrange(Fit_Method,-Error_difference)

full_comparison |> filter(Fit_Method=="Test_Train") |> 
  ungroup() |>
  mutate(id = reorder(id, Error_difference)) %>%
  ggplot(aes(y=id,x=Error_difference,fill=Best_Model))+
  geom_col() +
  ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")+
  labs(fill="Best Model",x="Mean Model Error Difference (ALM - EXAM)",y="Participant")

# full_comparison |> filter(Fit_Method=="Test_Train") |> 
#   ungroup() |>
#   mutate(id = reorder(id, Error_difference)) |>
#   left_join(post_dat_avg |> filter(x==100) |> select(-x) |> ungroup(), by=c("id","condit")) |>
#   ggplot(aes(y=id,x=c,fill=Best_Model))+
#   stat_pointinterval(position=position_dodge(.1))

```

#### Subjects with biggest differential favoring ALM
```{r}
#| fig-height: 6
#| fig-width: 12
vAlm <- c(307,331,197); cAlm <- c(372,298,299)

post_dat_l |> filter(id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Subjects with biggest differential favoring ALM",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

```

#### Subjects with biggest differential favoring EXAM
```{r}
#| fig-height: 6
#| fig-width: 12
vAlm <- c(312,334,295); cAlm <- c(132,366,415)

post_dat_l |> filter(id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Subjects with biggest differential favoring EXAM",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

```

#### Subjects with no clear best model
```{r}
#| fig-height: 6
#| fig-width: 12
vAlm <- c(129,192,105); cAlm <- c(101, 109,134)

post_dat_l |> filter(id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Subjects with no clear best model",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

```

```{r}
#| fig-height: 8
#| fig-width: 12
#| 
vAlm <- c(29,70,68); cAlm <- c(73,49,128)

difAlm <-  post_dat_avg  |> filter(x==100,id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |>
    group_by(id, condit, Model, rank) %>%
      mutate(flab=paste0("Subject: ",id)) |>
    ggplot(aes(y=log(c), x = id,col=Model)) + 
  stat_pointinterval(position=position_dodge(.5)) +
   # ggh4x::facet_nested_wrap(condit~flab, axes = "all",scales="free_y",ncol=3) + 
      labs(title="c parameter - sbjs. with biggest diff. favoring ALM") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4)) +ylim(c(-11,-4))

vAlm <- c(23,155,184); cAlm <- c(119,85,175)

difExam <- post_dat_avg  |> filter(x==100,id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |>
    group_by(id, condit, Model, rank) %>%
      mutate(flab=paste0("Subject: ",id)) |>
    ggplot(aes(y=log(c), x = id,col=Model)) + 
  stat_pointinterval(position=position_dodge(.1)) +
   # ggh4x::facet_nested_wrap(condit~flab, axes = "all",scales="free_y",ncol=3) + 
      labs(title="c parameter - sbjs. with biggest diff. favoring EXAM") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4)) +ylim(c(-11,-4))

difAlm/difExam

vAlm <- c(29,70,68); cAlm <- c(73,49,128)

difAlm <-  post_dat_avg  |> filter(x==100,id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |>
    group_by(id, condit, Model, rank) %>%
      mutate(flab=paste0("Subject: ",id)) |>
    ggplot(aes(y=lr, x = id,col=Model)) + 
  stat_pointinterval(position=position_dodge(.5)) +
   # ggh4x::facet_nested_wrap(condit~flab, axes = "all",scales="free_y",ncol=3) + 
      labs(title="lr parameter - sbjs. with biggest diff. favoring ALM") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4))+ylim(c(0,12))

vAlm <- c(23,155,184); cAlm <- c(119,85,175)

difExam <- post_dat_avg  |> filter(x==100,id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |>
    group_by(id, condit, Model, rank) %>%
      mutate(flab=paste0("Subject: ",id)) |>
    ggplot(aes(y=lr, x = id,col=Model)) + 
  stat_pointinterval(position=position_dodge(.1)) +
   # ggh4x::facet_nested_wrap(condit~flab, axes = "all",scales="free_y",ncol=3) + 
      labs(title="lr parameter - sbjs. with biggest diff. favoring EXAM") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4)) +ylim(c(0,12))

difAlm/difExam
```

```{r fig.width=11, fig.height=13}

# 
# 
# full_comparison |> filter(Fit_Method=="Test_Train") |>
#   ungroup() |>
#   mutate(id = reorder(id, Error_difference)) %>%
#   ggplot(aes(y=id,x=Error_difference,fill=Best_Model))+
#   geom_col()+
#   ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")
#

# d <- testAvg |> left_join(full_comparison, by=c("id","condit")) |> filter(Fit_Method=="Test_Train")
# 
# 
# 
# d |> ggplot(aes(x=vb,y=vx,fill=condit)) + stat_bar + facet_wrap(Fit_Method~Best_Model2,ncol=2)
# 
# d |> 
#   group_by(condit,Fit_Method,Best_Model) |> 
#   mutate(nGrp2=n()) |>
#   filter(abs(Error_difference)>15) |> 
#   ggplot(aes(x=vb,y=dist,fill=condit)) + 
#   stat_bar + facet_wrap(Fit_Method~Best_Model,ncol=2)
# 
# d |> group_by(condit,Fit_Method,Best_Model) %>% tally() |> mutate(n=n/6)
# 
# d |> group_by(condit,Fit_Method,Best_Model) |> filter(abs(Error_difference)>15) |> tally() |> mutate(n=n/6)
# 
# d |> group_by(condit,Fit_Method) |> mutate(m=mean(Error_difference), 
#                                               sd=sd(Error_difference), 
#                                               n=n()/6,se=sd/sqrt(n)) |>
#    group_by(condit,Fit_Method,Best_Model) |> 
#  # filter(abs(Error_difference)>(2.5*se)) |> 
#   ggplot(aes(x=vb,y=dist,fill=condit)) + 
#   stat_bar + facet_wrap(Fit_Method~Best_Model,ncol=2)
#

```

```{r fig.width=11, fig.height=13}
#| eval: false
#| include: false

posterior <- post_dat |> group_by(id, condit, Model, Fit_Method,rank) |> 
    slice_head(n=1) |> select(id,condit,Model,Fit_Method,rank,c,lr, mean_error)
head(posterior)

# Combine the MAP density values with your original posterior data frame
posterior_with_density <- left_join(posterior, posterior_densities, by = c("id", "condit", "Model", "Fit_Method"))

head(posterior_with_density)

posterior %>%
  group_by(id, condit, Model, Fit_Method) %>%
  do({
    data = .
    c_density = density(data$c)
    lr_density = density(data$lr)

    # Find the index of the MAP estimate in your density objects
    map_c_index = which.min(abs(c_density$x - data$c[data$rank == 1]))
    map_lr_index = which.min(abs(lr_density$x - data$lr[data$rank == 1]))

    data.frame(
      c_density_at_MAP = c_density$y[map_c_index],
      lr_density_at_MAP = lr_density$y[map_lr_index]
    )
  }) -> densities_at_map

# Combine the MAP density values with your original posterior data frame
posterior_with_density <- left_join(posterior, posterior_densities, by = c("id", "condit", "Model", "Fit_Method"))

head(posterior_with_density)

map_id <- post_dat |> group_by(id, condit, Model, Fit_Method) |> 
  filter(rank==1) |> slice_head(n=1) |> select(id,condit,Fit_Method,Model,c,lr,mean_error,rank) |>
  left_join(id_n,by=(c("id","condit")))

map_id <- map_id |> 
  mutate(n = fifelse(Fit_Method == "Test", nTest, fifelse(Fit_Method == "Train", nTrain, nTotal)),
         BIC = -2 * log(1/mean_error) + log(n)*2) |>
  group_by(id,Fit_Method) |>
  mutate(Best_Model=Model[which.min(BIC)])

bic_id <- map_id %>%
  group_by(id, condit, Fit_Method) %>%
  arrange(Model) %>%
  summarize(BIC_diff = diff(BIC),
            .groups = 'drop')

 bic_id |> filter(Fit_Method=="Test_Train") |>
     ungroup() |>
   mutate(id = reorder(id, BIC_diff),decreasing=FALSE) %>%
   ggplot(aes(y=id,x=BIC_diff))+
  geom_col()+
  ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")

bic_id |> group_by(condit,Fit_Method) |> summarise(mean(BIC_diff))

```

### To add to appendix

```{r}

post_tabs$agg_x_full |> flextable::tabulator(rows=c("Fit_Method","x"), columns=c("condit","Model"),
                       `X` = as_paragraph(mean_error)) |> as_flextable()

# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred)) |> 
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method,x) |> 
#   summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |> 
#   round_tibble(1) |> pander::pandoc.table()

# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |>
#   mutate(e2=abs(y-pred)) |>
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method) |>
#   summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |>
#   round_tibble(1) |> pander::pandoc.table()
# 
# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred),Fit_Method=rename_fm(Fit_Method)) |> 
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method) |> 
#   summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |> 
#   select(-y,-pred) |>
#   arrange(condit,Fit_Method) |>
#   round_tibble(1) |> pander::pandoc.table()
# 
# 
# 
# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred),Fit_Method=rename_fm(Fit_Method)) |> 
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method,x) |> 
#   summarise(y2=mean(y1), pred2=mean(pred1)) |> 
#   group_by(condit,Model,Fit_Method) |> 
#   summarise(y=mean(y2), pred=mean(pred2),mean_error=abs(y-pred)) |> 
#   select(-y,-pred) |>
#   arrange(condit,Fit_Method) |>
#   round_tibble(1) |> pander::pandoc.table()
# 
# 
# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred),Fit_Method=rename_fm(Fit_Method)) |> 
#   summarise(y1=mean(y), pred1=mean(pred),mean_error=abs(y1-pred1)) |>
#   group_by(id,condit,Model,Fit_Method) |> 
#   summarise(mean_error=mean(mean_error)) |> 
#   group_by(condit,Model,Fit_Method) |>
#   summarise(mean_error=mean(mean_error)) |> 
#   arrange(condit,Fit_Method) |>
#   round_tibble(1) |> pander::pandoc.table()

```

{{< include ../Discussion.qmd >}}

```{r fig.width=11, fig.height=10}
#| label: fig-htw-best-model-old
#| eval: false
#| include: false
#| fig-cap: Difference in model errors for each participant, with models fit to both train and test data. Positive values favor EXAM, while negative values favor ALM. 
#| fig-height: 10
#| fig-width: 11
best_id_x <- post_dat_avg |> group_by(id,condit,Model,Fit_Method,x) |>
  mutate(e2=(y-pred)) |>
  summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2),abs_me=abs(mean_error)) |>
  group_by(id,condit,Fit_Method,x) |> mutate(best=ifelse(mean_error==min(mean_error),1,0)) |>
  group_by(id,condit,Fit_Method,Model) |> mutate(n_best=sum(best))

best_id <- best_id_x |> group_by(id,condit,Fit_Method,Model) |>
  summarise(mean_error=mean(mean_error), n_best=dplyr::first(n_best),abs_me=mean(abs_me))

lowest_error_model <- best_id %>%
  group_by(id, condit,Fit_Method) %>%
  summarise(Best_Model = Model[which.min(mean_error)],
            Best_Model2 = Model[which.min(abs_me)],
            n_best = n_best[which.min(mean_error)],
            Lowest_error = min(mean_error),
            differential = min(mean_error) - max(mean_error)) %>%
  ungroup()

error_difference <- best_id %>%
  select(id, condit, Model,Fit_Method, mean_error,abs_me) %>%
  pivot_wider(names_from = Model, values_from = c(abs_me,mean_error)) %>%
  mutate(Error_difference = (mean_error_ALM - mean_error_EXAM), abs_error_dif = (abs_me_ALM - abs_me_EXAM))

full_comparison <- lowest_error_model |> left_join(error_difference, by=c("id","condit","Fit_Method"))  |> 
  group_by(condit,Fit_Method,Best_Model) |> mutate(nGrp=n(), model_rank = nGrp - rank(Error_difference) ) |> 
  arrange(Fit_Method,-Error_difference)

full_comparison |> filter(Fit_Method=="Test_Train") |> 
  ungroup() |>
  mutate(id = reorder(id, abs_error_dif)) %>%
  ggplot(aes(y=id,x=abs_error_dif,fill=Best_Model2))+
  geom_col()+
  ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")+
  labs(fill="Best Model",x="Mean Model Error Difference (ALM - EXAM)",y="Participant")

```