### 2024-02-24 Sections/Discussion.qmd
reorganize sections

# General Discussion

## Comparison to Project 1

### Differences between the tasks

There are a number of differences between Project 1's Hit The Target (HTT), and Project 2's Hit The Wall (HTW) tasks.

- Task Space Complexity: In HTW, the task space is also almost perfectly smooth, at least for the continuous feedback subjects, if they throw 100 units too hard, they'll be told that they were 100 units too hard. Whereas in HTT,  it was possible to produce xy velocity combinations that were technically closer to the empirical solution space than other throws, but which resulted in worse feedback due to striking the barrier.

- Perceptual Distinctiveness: HTT offers perceptually distinct varied conditions that directly relate to the task's demands, which may increase the sallience between training positions encounted by the varied group. In contrast, HTW's varied conditions differ only in the numerical values displayed, lacking the same level of perceptual differentiation. Conversely in HTW, the only difference between conditions for the varied group are the numbers displayed at the top of the screen which indicate the current target band(e.g. 800-1000, or 1000-1200)

- In HTW, our primary testing stage of interest has no feedback, whereas in HTT testing always included feedback (the intermittent testing in HTT expt 1 being the only exception). Of course, we do collect testing with feedback data at the end of HTW, but we haven't focused on that data at all in our modelling work thus far. It's also interesting to recall that the gap between varied and constant in HTW does seem to close substantially in the testing-with-feedback stage. The difference between no-feedback and feedback testing might be relevant if the benefits of variation have anything to do with improving subsequent learning (as opposed to subsequent immediate performance), OR if the benefits of constant training rely on having the most useful anchor, having the most useful anchor might be a lot less helpful if you're getting feedback from novel positions and can thus immediately begin to form position-specific anchors for the novelties, rather than relying on a training anchor.

- HTW and HTT both have a similar amount of training trials (~200), and thus the constant groups acquire a similar amount of experience with their single position/velocity in both experiments. However, the varied conditions in both HTT experiments train on 2 positions, whereas the varied group in HTW trains on 3 velocity bands. This means that in HTT the varied group gets half as much experience on any one position as the constant group, and in HTW they only get 1/3 as much experience in any one position. There are likely myriad ways in which this might impact the success of the varied group regardless of how you think the benefits of variation might be occurring, e.g. maybe they also need to develop a coherent anchor, maybe they need more experience in order to extract a function, or more experience in order to properly learn to tune their c parameter.