### 2024-02-24 Sections/Results.qmd
remove code from main Sections

---
title: "Results"
author: "Thomas Gorman"
toc: true
toc-depth: 5
toc-expand: true
number-sections: true
number-depth: 5
strip-comments: true
echo: false
#toc: true
# toc-title: Contents
# toc-depth: 4
# toc-expand: 1
# execute:
#   echo: false
---

## Experiment 1

{{< include ../Analysis/E1_test.qmd >}}

## Experiment 2

{{< include ../Analysis/E2_test.qmd >}}

## Experiment 3

{{< include ../Analysis/E3_test.qmd >}}