### 2024-03-07 Analysis/combo_test.qmd
combined results. Model deviation predictions. More hybrid fits.

---
title: "HTW All Exp Testing"
categories: [Analyses, R, Bayesian]
page-layout: full
toc: false
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,knitr,ggpattern,ggh4x)
conflict_prefer_all("dplyr", quiet = TRUE)
options(scipen = 999)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model"), ~ source(here::here(paste0("Functions/", .x, ".R"))))

e1 <- readRDS(here("data/e1_08-21-23.rds")) 
e2 <- readRDS(here("data/e2_08-04-23.rds")) 
e3 <- readRDS(here("data/e3_08-04-23.rds"))

# combine all 3 experiments
d <- rbind(e1,e2,e3)

```

## Testing Deviation
```{r}
#| fig-height: 12
#| fig-width: 11

d |> filter(expMode2=="Test") |> ggplot(aes(x = vb, y = dist,fill=condit)) +
  stat_bar

d |> filter(expMode2=="Test") |> ggplot(aes(x = condit, y = dist,fill=vb)) +
  stat_bar + facet_wrap(~Exp)

d |> filter(expMode2=="Test") |> ggplot(aes(x = condit, y = dist,fill=vb)) +
  stat_bar + facet_wrap(fb~bandOrder)

k1 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_dist = sd(dist) / sqrt(n()),dist = mean(dist))

ep1 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_dist = sd(dist) / sqrt(n()),dist = mean(dist)) |>
  ggplot(aes(x = condit, y = dist, fill = vb, pattern=bandType)) +
   geom_bar_pattern(stat = "identity", 
                    position = position_dodge()) + 
  geom_errorbar(aes(ymin = dist - se_dist, ymax = dist + se_dist),position = position_dodge()) +
  facet_wrap2(~fb*bandOrder,axes="all") +
  labs(title="Testing Deviation",pattern = "Band Type") +
   scale_pattern_manual(values = c("stripe", "circle"))

ep2 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_dist = sd(dist) / sqrt(n()),dist = mean(dist)) |>
  ggplot(aes(x = vb, y = dist, fill = condit, pattern=bandType)) +
   geom_bar_pattern(stat = "identity", 
                    position = position_dodge()) + 
  geom_errorbar(aes(ymin = dist - se_dist, ymax = dist + se_dist),position = position_dodge()) +
  facet_wrap2(~fb*bandOrder,axes="all") +
  labs(title="Testing Deviation",pattern = "Band Type") +
   scale_pattern_manual(values = c("stripe", "circle"))

ep1/ep2

```

## Testing Vx
```{r}
#| fig-height: 12
#| fig-width: 11

ep1 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_vx = sd(vx) / sqrt(n()),vx = mean(vx)) |>
  ggplot(aes(x = condit, y = vx, fill = vb, pattern=bandType)) +
   geom_bar_pattern(stat = "identity", 
                    position = position_dodge()) + 
  geom_errorbar(aes(ymin = vx - se_vx, ymax = vx + se_vx),position = position_dodge()) +
  facet_wrap2(~fb*bandOrder,axes="all") +
  labs(title="Testing Vx",pattern = "Band Type") +
   scale_pattern_manual(values = c("stripe", "circle"))

ep2 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_vx = sd(vx) / sqrt(n()),vx = mean(vx)) |>
  ggplot(aes(x = vb, y = vx, fill = condit, pattern=bandType)) +
   geom_bar_pattern(stat = "identity", 
                    position = position_dodge()) + 
  geom_errorbar(aes(ymin = vx - se_vx, ymax = vx + se_vx),position = position_dodge()) +
  facet_wrap2(~fb*bandOrder,axes="all") +
  labs(title="Testing Vx",pattern = "Band Type") +
   scale_pattern_manual(values = c("stripe", "circle"))

ep1/ep2

```