### 2024-03-11 Analysis/combo_train.qmd

```{r}
pacman::p_load(dplyr,purrr,tidyr,here,conflicted,ggpmisc)
walk(c("dplyr"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))

e1 <- readRDS(here("data/e1_08-21-23.rds")) 
e2 <- readRDS(here("data/e2_08-04-23.rds")) 
e3 <- readRDS(here("data/e3_08-04-23.rds")) 
d <- rbind(e1,e2,e3)

```

```{r}
#| fig-height: 13
#| fig-width: 12

learn_curve_plot <- function(df, x_var, y_var, color_var, facet_var = NULL, groupVec, nbins, labels = FALSE, y_label=NULL) {

  if (is.null(y_label)) {
    y_label <- deparse(substitute(y_var))
  }
  df |> 
    group_by(!!!syms(groupVec)) |> 
    mutate(Trial_Bin = cut( {{x_var}}, breaks = seq(1, max({{x_var}}),length.out=nbins+1),include.lowest = TRUE, labels = labels)) |> 
    ggplot(aes(x = Trial_Bin, y = {{ y_var }}, color = {{ color_var }})) +
    stat_summary(aes(color = {{ color_var }}), geom = "line", fun = mean) +
    stat_summary(geom = "errorbar", fun.data = mean_se, width = .4, alpha = .7) +
    #facet_wrap(vars({{facet_var}}), scales = 'free_y') + 
    labs(y = y_label) + # Set the y axis label dynamically
    scale_x_continuous(breaks = seq(1, nbins + 1)) 
}

nb=6
e1 |> filter(expMode=="train") %>% 
    learn_curve_plot(gt.train,dist,condit,facet_var=vb,groupVec = c("id", "condit","vb"),nbins=nb) + facet_wrap(~vb)

d |> filter(expMode=="train") %>% 
    learn_curve_plot(gt.train,dist,condit,facet_var=c("Exp","vb"),groupVec = c("id", "condit","vb"),nbins=nb) + facet_wrap(~bandOrder*fb*vb,ncol=3)

# d |> group_by(id,condit,vb) |> ggplot(aes(x=gt.train,y=dist,col=condit)) + geom_smooth() + 
#     facet_wrap(~bandOrder*fb*vb,ncol=3)
# 
# d |> group_by(gt.train,condit) |> ggplot(aes(x=gt.train,y=dist,col=condit)) + geom_smooth(method="loess") + 
#     facet_wrap(~bandOrder*fb*vb,ncol=3)

```

```{r}
#| fig-height: 13
#| fig-width: 12
nb=3
formula <- y ~ poly(x,3,raw=TRUE)
formula <- y ~ poly(x,2,raw=TRUE)

d  |> filter(expMode=="train") %>% group_by(id,gt.train,condit,vb) |>
    select(id,condit,bandOrder,fb,gt.train,vb,expMode,dist) |>
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, nb + 1), include.lowest = TRUE, labels = FALSE)) |> 
    ggplot(aes(x=Trial_Bin,y=dist,col=condit)) + 
    geom_smooth(method="lm",formula=formula) + 
    facet_wrap(~bandOrder*fb*vb,ncol=3)

d  |> filter(expMode=="train") %>% group_by(id,gt.train,condit,vb) |>
    select(id,condit,bandOrder,fb,gt.train,vb,expMode,dist) |>
    ggplot(aes(x=gt.train,y=dist,col=condit)) + 
    geom_smooth(method="lm",formula=formula) + 
    facet_wrap(~bandOrder*fb*vb,ncol=3)

d  |> filter(expMode=="train",gt.train<=60) %>% group_by(id,gt.train,condit,vb) |>
    ggplot(aes(x=gt.train,y=dist,col=condit)) + 
    geom_smooth(method="lm",formula=formula) + 
    facet_wrap(~bandOrder*fb*vb,ncol=3) + 
  stat_poly_eq(geom = "text", aes(label = after_stat(eq.label)),
               label.x = c(60, 0), label.y = c(-0.1, 400.1), hjust = "inward",
               formula = formula)

micmen.formula <- y ~ SSmicmen(x, Vm, K)
d  |> filter(expMode=="train",gt.train<=60) %>% group_by(id,gt.train,condit,vb) |>
    ggplot(aes(x=gt.train,y=dist,col=condit)) + 
    geom_smooth(method="nls",formula = micmen.formula,se=FALSE) + 
    facet_wrap(~bandOrder*fb*vb,ncol=3)

d |>
  filter(expMode == "train", gt.train <= 60) |>
  group_by(id, gt.train, condit, vb) |>
  ggplot(aes(x = gt.train, y = dist, col = condit)) +
  geom_smooth(method = "nls", formula = micmen.formula, se = FALSE, method.args = list(start = list(Vm = 1, K = 1))) +
  facet_wrap(~bandOrder * fb * vb, ncol = 3)

#bf(dist ~ betaMu + (alphaMu - betaMu) * exp(-exp(gammaMu) * gt.train)

exp_model_formula <- y ~ a * exp(b * x)
d |> filter(expMode == "train", gt.train <= 60) %>%
  group_by(id, gt.train, condit, vb) %>%
  ggplot(aes(x = gt.train, y = dist, col = condit)) +
  geom_smooth(method = "nls", formula = exp_model_formula, se = FALSE, method.args = list(start = list(a = 1, b = 0.1))) +
  facet_wrap(~bandOrder * fb * vb, ncol = 3)

nb=3
e1 |> group_by(id) |> 
    select(id,condit,bandOrder,gt.train,vb,expMode2,dist) |>
    mutate(Block=case_when(expMode2=="Train" ~ cut(gt.train,breaks=seq(1,max(gt.train), length.out=nb+1),include.lowest=TRUE,labels=FALSE),expMode2=="Test" ~ nb+1)) %>%
     ggplot(aes(x=Block,y=dist,col=condit)) + 
    geom_smooth(method="lm", formula= formula) + 
    facet_wrap(~vb)

```

```{r}
#| fig-height: 13
#| fig-width: 12
three_param_exp_formula <- y ~ a + (b - a) * exp(-c * x)
d |> filter(expMode == "train", gt.train <= 60) %>%
  group_by(id, gt.train, condit, vb) %>%
  ggplot(aes(x = gt.train, y = dist, col = condit)) +
  geom_smooth(method = "nls", 
              formula = three_param_exp_formula, 
              method.args = list(start = list(a = 100, b = 500, c = 0.01)), # Starting values for a, b, c
              se = FALSE) +
  facet_wrap(~bandOrder * fb * vb, ncol = 3)

two_param_exp_formula <- y ~ a  * exp(-c * x)
d |> filter(expMode == "train") %>%
  group_by(id, gt.train, condit, vb) %>%
  ggplot(aes(x = gt.train, y = dist, col = condit)) +
  geom_smooth(method = "nls", 
              formula = two_param_exp_formula, 
              method.args = list(start = list(a = 500, c = 0.09)), # Starting values for a, b, c
              se = FALSE) +
  facet_wrap(~bandOrder * fb * vb, ncol = 3) 
```

```{r}
#| fig-height: 13
#| fig-width: 12
#| 
filtered_data <- d |> 
  filter(expMode == "train", gt.train <= 60)

# Fit the model
nls_fit <- nls(dist ~ a + (b - a) * exp(-c * gt.train), 
               data = filtered_data, 
               start = list(a = 100, b = 500, c = 0.01))

# Extract coefficients
coeffs <- coef(nls_fit)

# Create label (customize as needed)
label_text <- sprintf("y = %.2f + (%.2f - %.2f) * exp(-%.2f * x)", coeffs["a"], coeffs["b"], coeffs["a"], coeffs["c"])

# Plot with custom label
filtered_data %>%
  ggplot(aes(x = gt.train, y = dist, col = condit)) +
  geom_smooth(method = "nls", 
              formula = y ~ a + (b - a) * exp(-c * x), 
              method.args = list(start = list(a = 100, b = 500, c = 0.01)),
              se = FALSE) +
  ggpp::annotate("text", x = Inf, y = Inf, label = label_text, hjust = 1, vjust = 1, size = 3.5) +
  facet_wrap(~bandOrder * fb * vb, ncol = 3)

library(ggpmisc)

d |> filter(expMode == "train", gt.train <= 60) %>%
  group_by(id, gt.train, condit, vb) %>%
  ggplot(aes(x = gt.train, y = dist, col = condit)) +
  geom_smooth(method = "nlsLM",
              formula = y ~ three_param_exp_formula(x, a, b, c),
              method.args = list(start = list(a = 100, b = 500, c = 0.01)),
              se = FALSE) +
  stat_equation(aes(label = after_stat(eq.label)),
                formula = y ~ three_param_exp_formula(x, a, b, c),
                parse = TRUE) +
  facet_wrap(~bandOrder * fb * vb, ncol = 3)

```

# Join the labels back to the original data
filtered_data <- left_join(filtered_data, labels_df, by = c("condit", "bandOrder", "fb"))

# Plot with annotations
ggplot(filtered_data, aes(x = gt.train, y = dist, col = condit)) +
  geom_smooth(method = "nls", formula = y ~ a + (b - a) * exp(-c * x), 
              method.args = list(start = list(a = 100, b = 500, c = 0.01)), se = FALSE) +
  facet_wrap(~bandOrder * fb * vb, ncol = 3) +
  geom_text(data = labels_df, aes(label = label, x = 50, y = Inf), hjust = 0.5, vjust = 1, size = 3, inherit.aes = FALSE)

```{r}
#| fig-height: 16
#| fig-width: 12

fit_model_and_return_label <- function(data, condit, bandOrder, fb) {
  # Default starting parameters
  start_params <- list(a = 100, b = 500, c = 0.01)

  # Adjust starting parameters for the problematic condition
  if (condit == "Constant" && fb == "Ordinal") {
    start_params <- list(a = 100, b = 250, c = 0.01)
  }
  tryCatch({
    model <- nls(dist ~ a + (b - a) * exp(-c * gt.train), data = data, start = start_params)
    coeffs <- coef(model)
    label=sprintf("y = %.2f + (%.2f - %.2f) * exp(-%.2f * x)", coeffs["a"], coeffs["b"], coeffs["a"], coeffs["c"])
    data.frame(label = label, a = coeffs["a"], b = coeffs["b"], c = coeffs["c"])
  }, error = function(e) {
    # Return NA or some indication of failure
    return(data.frame(label = NA, a = NA, b = NA, c = NA))
  })
}

labels_df <- filtered_data %>%
  group_by(condit, bandOrder, fb,vb) %>%
  nest() %>%
  mutate(result = map(data, ~fit_model_and_return_label(.x, condit, bandOrder, fb)),
         data = NULL) %>%
  unnest(result) %>%
  ungroup()

head(labels_df)

generate_predictions <- function(x, a, b, c) {
  return(a + (b - a) * exp(-c * x))
}

plot_data <- filtered_data %>%
  left_join(labels_df, by = c("condit", "bandOrder", "fb"))

# Define a sequence of x values for generating model predictions (adjust as needed)
x_vals <- seq(min(filtered_data$gt.train), max(filtered_data$gt.train), length.out = 100)

# Create a new data frame for model predictions
predictions_data <- labels_df %>%
  mutate(x = list(x_vals)) %>%
  unnest(x) %>%
  mutate(y = pmap_dbl(list(a, b, c, x), ~ generate_predictions(..4, ..1, ..2, ..3)))

# Plot
ggplot() +
  #geom_point(data = plot_data, aes(x = gt.train, y = dist, color = condit), alpha = 0.6) +
  geom_line(data = predictions_data, aes(x = x, y = y, color = condit), size = 1) +
  facet_wrap(~ bandOrder * fb , ncol = 2, scales = "free_y") +
  labs(x = "GT Train", y = "Dist") +
  theme_minimal()

ggplot(filtered_data, aes(x = gt.train, y = dist, color = condit)) +
  geom_point() +
  stat_function(fun = function(x) with(labels_df[1, ], a + (b - a) * exp(-c * x)),
                linetype = "dashed", color = "black", size = 1) +
  facet_grid(fb ~ bandOrder, labeller = label_both) +
  labs(title = "Model Fits of Learning Curves",
       x = "Training Generation",
       y = "Distance",
       color = "Condition") +
  theme_bw()
```

### create learning models for condit and varied groups.

We can model the relation between performance and the number of practice
trials as a power law function, or exponential function. Aggregatign
over ids in dst. The models predict dist as an exponential decay
function of trial number. Band is an additional predictor.

$$
f_p(t) = \alpha + \beta t^{r} \enspace 
$$

$$
f_e(t) = \alpha + \beta e^{rt} \enspace 
$$

```{r}

# fit exponential decay model as a function of trial number

fit_exp <- function(trial,dist,input){
    # fit exponential decay model as a function of trial number, band is an additional predictor
    fit <- nls(dist ~ yf + (y0-yf) * exp(-r*trial) + beta2*input, start = list(yf = 300, y0 = 364, beta2=0, r = .1), data = data.frame(trial=trial,dist=dist,input=input))

    # extract parameters
    alpha <- coef(fit)[1]
    beta <- coef(fit)[2]
    beta2 <- coef(fit)[3]
    r <- coef(fit)[4]
    sigma_e <- summary(fit)$sigma

    # compute negative log likelihood
    nllh <- negative_llh_exp(dist, trial, alpha, beta, r, sigma_e)

    # return parameters and negative log likelihood
    return(list(alpha=alpha,beta=beta,beta2=beta2,r=r,sigma_e=sigma_e,nllh=nllh))
}

# Compute group averages for dist over trial and band. dst

avgTrain <- dst %>% group_by(id,condit,trial,band,input) %>% summarise(dist=mean(dist)) %>% ungroup() %>% group_by(condit,trial,band) %>% summarise(dist=mean(dist)) %>% ungroup()

ggplot(avgTrain,aes(x=trial,y=dist)) + geom_line(aes(group=band,color=band)) +facet_grid(~condit)

avgTrain %>% filter(condit=="Constant") %>% nls(dist ~ yf + (y0-yf) * exp(-r*trial), start = list(yf = 120, y0 = 364, r = .1), data = .) %>% summary()

avgTrain %>% filter(condit=="Constant") %>% nls(dist ~ SSasymp(trial, yf, y0, log_alpha),data=.)

fit_condit <- avgTrain %>% group_by(condit) %>% do(fit_exp(trial=.$trial,dist=.$dist,input=.$input))

```

Interpretation of improvement_model:

The intercept represents the performance when all factors are at their reference levels (Constant condition, original category order, and continuous feedback type).
Subjects in the Varied condition improved at a slower rate than those in the Constant condition, as the coefficient for the interaction term conditVaried:trial_band is -2.37284, with a t-value of -6.940.
Subjects in the Varied condition with reversed category order showed a greater decrease in performance, as the coefficient for the interaction term conditVaried:bandOrderrev is -43.67731, with a t-value of -2.323.
Other significant factors and interactions include trial_band, bandOrderrev, trial_band:bandOrderrev, and trial_band:fbordinal.
Interpretation of final_performance_model:

The intercept represents the final performance when all factors are at their reference levels (Constant condition, original category order, and continuous feedback type).
Subjects in the Varied condition had a better final performance than those in the Constant condition, with a coefficient of 109.73 and a t-value of 4.362.
The interaction between the Varied condition and reversed category order (conditVaried:bandOrderrev) had a negative impact on the final performance, with a coefficient of -92.75 and a t-value of -2.342.
The interaction between the Varied condition and ordinal feedback type (conditVaried:fbordinal) also had a negative impact on the final performance, with a coefficient of -85.44 and a t-value of -2.079.
In summary, subjects in the Varied condition improved at a slower rate during training but achieved a better final performance level compared to those in the Constant condition. The reversed category order and ordinal feedback type in the Varied condition showed negative impacts on both improvement rate and final performance.

### Exponential learning model

```{r}
#| eval: false
library(dplyr)
library(tidyr)
library(nls.multstart)
exp_fun <- function(a, b, c, x) {
  a * (1 - exp(-b * x)) + c
}
exp_models <- dst %>%
  nest(-id) %>%
  mutate(model = map(data, ~ nls_multstart(dist ~ exp_fun(a, b, c, trial_band),
                                           data = .x,
                                           iter = 500,
                                           start_lower = c(a = 0, b = 0, c = 0),
                                           start_upper = c(a = 5000, b = 1, c = 5000)))) %>%
  unnest(c(a = map_dbl(model, ~ coef(.x)['a']),
           b = map_dbl(model, ~ coef(.x)['b']),
           c = map_dbl(model, ~ coef(.x)['c'])))
group_averages <- exp_models %>%
  group_by(condit, bandOrder, fb) %>%
  summarise(a_avg = mean(a), b_avg = mean(b), c_avg = mean(c))
aic_improvement <- AIC(improvement_model)
aic_final_performance <- AIC(final_performance_model)
exp_models <- exp_models %>%
  mutate(aic = map_dbl(model, AIC))

aic_exp_avg <- exp_models %>%
  summarise(aic_avg = mean(aic))
```