### 2024-03-13 Analysis/e3.qmd

---
title: "HTW E3"
categories: [Analyses, R, Bayesian]
code-fold: true
cold-tools: true
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  stringr, here,conflicted, patchwork, knitr,kableExtra)

walk(c("brms","dplyr","bayestestR"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","org_functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))

e3 <- readRDS(here("data/e3_08-04-23.rds")) |> 
    mutate(trainCon=case_when(
    bandOrder=="Original" ~ "800",
    bandOrder=="Reverse" ~ "600",
    TRUE ~ NA_character_
    ), trainCon=as.numeric(trainCon))

e3Sbjs <- e3 |> group_by(id,condit,bandOrder) |> summarise(n=n())
testE3 <- e3 |> filter(expMode2 == "Test")

```

```{r}

nbins=6
trainE3 <-  e3 |> filter(expMode2=="Train") |> group_by(id,condit,bandOrder, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
# bayesian comparison of condits in training
trainE3_max <- trainE3 |> filter(Trial_Bin == nbins, bandInt==trainCon)

bmm_e3_train <- trainE3_max %>% 
  brm(dist ~ condit*bandOrder, 
      file=here("data/model_cache/e3_train_deviation"),
      data = .,
      iter = 2000,
      chains = 4,
      control = list(adapt_delta = .94, max_treedepth = 13))

#bayestestR::describe_posterior(bayes_comp)

mtr3 <- as.data.frame(describe_posterior(bmm_e3_train, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mtr3) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

mtr3 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  tibble::remove_rownames() |> 
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(escape=F,booktabs=T) %>% kable_styling(font_size=8)

```

```{r}
#| label: fig-e3-train-dev
#| fig-cap: E3. Deviations from target band during testing without feedback stage.

trainE3 |> ggplot(aes(x = Trial_Bin, y = dist, color = condit)) +
    stat_summary(geom = "line", fun = mean) +
    stat_summary(geom = "errorbar", fun.data = mean_se, width = .4, alpha = .7) +
    ggh4x::facet_nested_wrap(~bandOrder*vb,ncol=3)+
    scale_x_continuous(breaks = seq(1, nbins + 1)) +
    theme(legend.title=element_blank()) + 
    labs(y = "Deviation", x="Training Block")

```

```{r}
#options(brms.backend="cmdstanr",mc.cores=4)
modelFile <- paste0(here::here("data/model_cache/"), "e3_dist_Cond_Type_RFint")
bmtd3 <- brm(dist ~ condit * bandType*bandOrder + (1|id*bandInt), 
    data=testE3, file=modelFile,
    iter=5000,chains=4, control = list(adapt_delta = .94, max_treedepth = 13))

#bayestestR::describe_posterior(bmtd)

mted3 <- as.data.frame(describe_posterior(bmtd3, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mted3) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

mted3 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  tibble::remove_rownames() |> 
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(escape=F,booktabs=T) %>% kable_styling(font_size=8)

```

```{r}
#| label: fig-e3-test-dev
#| fig-cap: E3. Deviations from target band during testing without feedback stage. 
pe3td <- testE3 |>  ggplot(aes(x = vb, y = dist,fill=condit)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
    facet_wrap(~bandOrder,ncol=1) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation From Target")

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

}

pe3ce <- bmtd3 |> emmeans( ~condit *bandOrder*bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Absolute Deviation From Band", x="Band Type") + 
 facet_wrap(~bandOrder,ncol=1)

pe3td + pe3ce

```

### Discimination

```{r}
#| label: tbl-e3-bmm-vx
#| tbl-cap: "Experiment 3. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"

e3_vxBMM <- brm(vx ~ condit * bandOrder * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e3_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
GetModelStats(e3_vxBMM ) |> kable(escape=F,booktabs=T, caption="Fit to all 6 bands") %>% kable_styling(font_size=8)

cd3 <- get_coef_details(e3_vxBMM, "conditVaried")
sc3 <- get_coef_details(e3_vxBMM, "bandInt")
intCoef3 <- get_coef_details(e3_vxBMM, "conditVaried:bandInt")

```

```{r}
#| label: fig-e3-bmm-vx
#| fig-cap: Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands. 
#| fig-height: 19
#| fig-width: 12

pe3tv <- testE3 %>% group_by(id,vb,condit,bandOrder) |> plot_distByCondit() + ggh4x::facet_nested_wrap(bandOrder~condit,scale="free_x")

pe3vce <- e3_vxBMM |> emmeans( ~condit* bandOrder* bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects(bandInt) +
  facet_wrap(~bandOrder,ncol=1) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE3$vb), 
                     limits = c(0, 1400)) + 
  theme(legend.title=element_blank()) + 
  labs(y="Velcoity", x="Band")

fe <- fixef(e3_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e3_vxBMM)[,1]["bandInt"]
fixed_effect_interaction <- fixef(e3_vxBMM)[,1]["conditVaried:bandInt"]

re <- data.frame(ranef(e3_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e3Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction*(condit=="Varied"),slope = Estimate + adjust )

pid_den3 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density") +
  facet_wrap(~bandOrder,ncol=1)

pid_slopes3 <- re |>  mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
     theme(legend.title=element_blank()) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_wrap2(bandOrder~condit,axes="all",scales="free_y")

pe3tv / (pe3vce + pid_den3 + pid_slopes3)

```

```{r}

```