### 2024-03-16 Analysis/combo_test.qmd
control for train models and figures

---
title: "HTW All Exp Testing"
categories: [Analyses, R, Bayesian]
page-layout: full
toc: false
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,knitr,ggpattern,ggh4x)
conflict_prefer_all("dplyr", quiet = TRUE)
options(scipen = 999)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model"), ~ source(here::here(paste0("Functions/", .x, ".R"))))

e1 <- readRDS(here("data/e1_08-21-23.rds")) 
e2 <- readRDS(here("data/e2_08-04-23.rds")) 
e3 <- readRDS(here("data/e3_08-04-23.rds"))

# combine all 3 experiments
d <- rbind(e1,e2,e3)

d <- d |> 
    mutate(trainCon=case_when(
    bandOrder=="Original" ~ "800",
    bandOrder=="Reverse" ~ "600",
    TRUE ~ NA_character_
    ), trainCon=as.numeric(trainCon))

nbins=5
train <-  d |> filter(expMode2=="Train") |> group_by(id,condit,fb,bandOrder, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
train_max <- train |> filter(Trial_Bin == nbins, bandInt==trainCon)
train_avg <- train_max |> group_by(id,condit,fb,bandOrder) |> summarise(train_end = mean(dist))
train_avg2 <- train_max |> select(id,condit,fb,bandOrder,expMode2,vb,bandInt,dist,vx)

test2 <- d |> filter(expMode2=="Test") |> 
  select(id,condit,fb,bandOrder,expMode2,vb,bandInt,dist,vx) |> 
  rbind(train_avg2) |>
  left_join(train_avg, by=c("id","condit", "fb", "bandOrder"))

test <- d |> filter(expMode2=="Test") |> left_join(train_avg, by=c("id","condit", "fb", "bandOrder")) |>
  select(id,condit,bandType,bandInt,vb,vx,dist,train_end,fb,bandOrder)

```

```{r}
ltest|> group_by(id,condit) |> pivot_longer(c("dist","train_end"),names_to="var",values_to="value") |> 
  ggplot(aes(x=var,y=value, fill=condit)) + stat_bar + facet_wrap(~var)

test2 |> ggplot(aes(x=bandInt,y=dist,fill=expMode2)) + stat_bar + facet_wrap(~condit+fb+bandOrder)

test2 |> mutate(train_end_q = ntile(train_end,4)) |>  
  ggplot(aes(x=bandInt,y=dist,fill=expMode2)) + stat_bar + facet_wrap(~condit+train_end_q+bandOrder)

test2 |> mutate(train_end_q = ntile(train_end,4)) |>  
  ggplot(aes(x=condit,y=dist,fill=vb)) + stat_bar + 
  facet_wrap(~expMode2+train_end_q+bandOrder,scales="free")

test2 |> mutate(train_end_q = ntile(train_end,4)) |>  
  ggplot(aes(x=vb,y=dist,fill=condit)) + stat_bar + 
  facet_wrap(~expMode2+train_end_q+bandOrder,scales="free")

test2 |> mutate(train_end_q = ntile(train_end,4)) |>  
  ggplot(aes(x=train_end_q,y=dist,fill=condit)) + stat_bar + 
  facet_wrap(~expMode2+vb+bandOrder,scales="free")

test |> ggplot(aes(x=train_end,y=dist,fill=condit)) + 
  stat_summary(geom = "line", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
    facet_wrap(~vb) +
  labs(x="Band", y="Deviation From Target")

test |> filter(bandType=="Extrapolation") |>
  ggplot(aes(x=train_end,y=dist,fill=condit,col=condit)) + 
  #geom_point() +
  geom_smooth(method="loess") +
    facet_nested_wrap(~bandOrder+vb+fb) +
  labs(x="Band", y="Deviation From Target")

test |> filter(bandType=="Extrapolation") |>
  ggplot(aes(x=train_end,y=dist,fill=condit,col=condit)) + 
  #geom_point() +
  geom_smooth() +
    facet_nested_wrap(~bandOrder+vb+fb) +
  labs(x="Band", y="Deviation From Target")

test |> filter(bandType=="Extrapolation") |>
  ggplot(aes(x=train_end,y=dist,fill=condit,col=condit)) + 
  #geom_point() +
  geom_smooth(method="lm") +
    facet_nested_wrap(~bandOrder+vb) +
  labs(x="Band", y="Deviation From Target")

# create quartiles for train_end
test |> group_by(condit,vb,fb,bandOrder) |>  filter(bandType=="Extrapolation") |>
  mutate(train_end_q = ntile(train_end,4)) |> 
  ggplot(aes(x=train_end_q,y=dist,fill=condit)) + 
 stat_bar + 
    ggh4x::facet_wrap2(bandOrder~vb~fb) +
  labs(x="Band", y="Deviation From Target")

test |> group_by(condit,vb,fb,bandOrder) |>  #filter(bandType=="Extrapolation") |>
  mutate(train_end_q = ntile(train_end,4)) |> 
  ggplot(aes(x=condit,y=vx,fill=vb)) +
  stat_bar + 
    ggh4x::facet_wrap2(~train_end_q+bandOrder+fb)

```

## Testing Deviation
```{r}
#| fig-height: 12
#| fig-width: 11

d |> filter(expMode2=="Test") |> ggplot(aes(x = vb, y = dist,fill=condit)) +
  stat_bar

d |> filter(expMode2=="Test") |> ggplot(aes(x = condit, y = dist,fill=vb)) +
  stat_bar + facet_wrap(~Exp)

d |> filter(expMode2=="Test") |> ggplot(aes(x = condit, y = dist,fill=vb)) +
  stat_bar + facet_wrap(fb~bandOrder)

k1 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_dist = sd(dist) / sqrt(n()),dist = mean(dist))

ep1 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_dist = sd(dist) / sqrt(n()),dist = mean(dist)) |>
  ggplot(aes(x = condit, y = dist, fill = vb, pattern=bandType)) +
   geom_bar_pattern(stat = "identity", 
                    position = position_dodge()) + 
  geom_errorbar(aes(ymin = dist - se_dist, ymax = dist + se_dist),position = position_dodge()) +
  facet_wrap2(~fb*bandOrder,axes="all") +
  labs(title="Testing Deviation",pattern = "Band Type") +
   scale_pattern_manual(values = c("stripe", "circle"))

ep2 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_dist = sd(dist) / sqrt(n()),dist = mean(dist)) |>
  ggplot(aes(x = vb, y = dist, fill = condit, pattern=bandType)) +
   geom_bar_pattern(stat = "identity", 
                    position = position_dodge()) + 
  geom_errorbar(aes(ymin = dist - se_dist, ymax = dist + se_dist),position = position_dodge()) +
  facet_wrap2(~fb*bandOrder,axes="all") +
  labs(title="Testing Deviation",pattern = "Band Type") +
   scale_pattern_manual(values = c("stripe", "circle"))

ep1/ep2

```

## Testing Vx
```{r}
#| fig-height: 12
#| fig-width: 11

ep1 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_vx = sd(vx) / sqrt(n()),vx = mean(vx)) |>
  ggplot(aes(x = condit, y = vx, fill = vb, pattern=bandType)) +
   geom_bar_pattern(stat = "identity", 
                    position = position_dodge()) + 
  geom_errorbar(aes(ymin = vx - se_vx, ymax = vx + se_vx),position = position_dodge()) +
  facet_wrap2(~fb*bandOrder,axes="all") +
  labs(title="Testing Vx",pattern = "Band Type") +
   scale_pattern_manual(values = c("stripe", "circle"))

ep2 <- d |>
  filter(expMode2 == "Test") |>
  group_by(condit, bandType, fb, bandOrder, vb) |>  # No need for 'id' if not used later
  summarise(se_vx = sd(vx) / sqrt(n()),vx = mean(vx)) |>
  ggplot(aes(x = vb, y = vx, fill = condit, pattern=bandType)) +
   geom_bar_pattern(stat = "identity", 
                    position = position_dodge()) + 
  geom_errorbar(aes(ymin = vx - se_vx, ymax = vx + se_vx),position = position_dodge()) +
  facet_wrap2(~fb*bandOrder,axes="all") +
  labs(title="Testing Vx",pattern = "Band Type") +
   scale_pattern_manual(values = c("stripe", "circle"))

ep1/ep2

```