### 2024-03-17 Analysis/e1.qmd
e2 and e3 stats. re-render

---
title: "HTW E1"
author:
  - name: Thomas Gorman
    corresponding: true
categories: [Analyses, R, Bayesian]
toc: false
format-links: false
page-layout: full
cache: true
code-fold: true
cold-tools: true
lightbox: true
# execute: 
#   cache: true
keep-md: true
format:
  html: default
  # hikmah-manuscript-pdf:
  #   echo: false
  #   output-file: "hikmah-e1.pdf"
  #   mainfont: "Linux Libertine O"
  #   mainfontoptions:
  #     - "Numbers=Proportional"
  #     - "Numbers=OldStyle"
  #   mathfont: "Libertinus Math"
  # hikmah-pdf:
  #   echo: false
  #   cite-method: citeproc
  #   output-file: "hikmah.pdf"
  #   mainfont: "Linux Libertine O"
  #   mainfontoptions:
  #     - "Numbers=Proportional"
  #     - "Numbers=OldStyle"
  #   sansfont: "IBM Plex Sans"
  #   mathfont: "Libertinus Math"
  #   geometry:
  #     - top=1in     
  #     - bottom=1in  
  #     - left=1in    
  #     - right=1in  
  #     - heightrounded
  #   # linestretch: 1.05
  #   include-in-header:
  #     # The \everydisplay thing makes all math blocks use \small font
  #     text: |
  #       \usepackage{mathtools}
  #       \everydisplay\expandafter{\the\everydisplay\small}
  #       \usepackage{float}
  #       \floatplacement{figure}{H}
  #       \floatplacement{table}{H}
  # pdf: 
  #   echo: false
  #   documentclass: article
  #   papersize: a4
  #   geometry:
  #     - left=2cm
  #     - right=2cm
  #     - top=2cm
  #     - bottom=2cm
  # ipynb: default
  hugo-md:
    echo: false
    html-math-method: mathjax
    output-dir: "../_working"
    output-file: e1-hugo.md
  gfm:
    echo: true
    output-file: e1-gfm.md
  # docusaurus-md:
  #   echo: true
  #   output-file: e1-docusaurus.md
#prefer-html: false
---

```{r}

pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  stringr, here,conflicted, patchwork, knitr,kableExtra)
#options(brms.backend="cmdstanr",mc.cores=4)
walk(c("brms","dplyr","bayestestR"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","org_functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
e1 <- readRDS(here("data/e1_08-21-23.rds")) 
e1Sbjs <- e1 |> group_by(id,condit) |> summarise(n=n())
testE1 <- e1 |> filter(expMode2 == "Test")
nbins=5
trainE1 <-  e1 |> filter(expMode2=="Train") |> group_by(id,condit, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
trainE1_max <- trainE1 |> filter(Trial_Bin == nbins, bandInt==800)
trainE1_avg <- trainE1_max |> group_by(id,condit) |> summarise(avg = mean(dist))
```

### Analyses Strategy

All data processing and statistical analyses were performed in R version 4.32 @rcoreteamLanguageEnvironmentStatistical2020. To assess differences between groups, we used Bayesian Mixed Effects Regression. Model fitting was performed with the brms package in R @burknerBrmsPackageBayesian2017, and descriptive stats and tables were extracted with the BayestestR package @makowskiBayestestRDescribingEffects2019. Mixed effects regression enables us to take advantage of partial pooling, simultaneously estimating parameters at the individual and group level. Our use of Bayesian, rather than frequentist methods allows us to directly quantify the uncertainty in our parameter estimates, as well as avoiding convergence issues common to the frequentist analogues of our mixed models.

Each model was set to run with 4 chains, 5000 iterations per chain, with the first 2500 discarded as warmup chains. Rhat values were within an acceptable range, with values \<=1.02 (see appendix for diagnostic plots). We used uninformative priors for the fixed effects of the model (condition and velocity band), and weakly informative Student T distributions for for the random effects. For each model, we report 1)  the mean values of the posterior distribution for the parameters of interest, 2) the lower and upper credible intervals (CrI), and the probability of direction value (pd).

| Group Comparison         | Code                                                        | Data                 |
| ------------------------ | ----------------------------------------------------------- | -------------------- |
| End of Training Accuracy | `` brm(dist ~ condit) ``                                    | Final Training Block |
| Test Accuracy            | `` brm(dist ~ condit * bandType + (1|id) + (1|bandInt) `` | All Testing trials   |
| Band Discrimination      | `` brm(vx ~ condit * band +(1 + bandInt|id) ``             | All Testing Trials   |
\

In each experiment we compare varied and constant conditions in terms of 1) accuracy in the final training block; 2) testing accuracy as a function of  band type (trained vs. extrapolation bands); 3) extent of discrimination between all six testing bands. We quantified accuracy as the absolute deviation between the response velocity and the nearest boundary of the target  band. Thus, when the target band was velocity 600-800, throws of 400, 650, and 900 would result in deviation values of 200, 0, and 100, respectively. The degree of discrimination between bands was index by fitting a linear model predicting the response velocity as a function of the target velocity. Participants who reliably discriminated between velocity bands tended to haves slope values \~1, while participants who made throws irrespective of the current target band would have slopes \~0.

```{=tex}
\begin{equation}
dist_{ij} = \beta_0 + \beta_1 \cdot condit_{ij} + \beta_2 \cdot band_{ij} + \beta_3 \cdot condit_{ij} \cdot band_{ij} + b_{0i} + b_{1i} \cdot band_{ij} + \epsilon_{ij}
\end{equation}
```

### Results

```{r}
#| label: fig-e1-train-dev
#| fig-cap: E1. Deviations from target band across training blocks. 
#| eval: FALSE
p1 <- trainE1 |> ggplot(aes(x = Trial_Bin, y = dist, color = condit)) +
    stat_summary(geom = "line", fun = mean) +
    stat_summary(geom = "errorbar", fun.data = mean_se, width = .4, alpha = .7) +
    facet_wrap(~vb)+
    scale_x_continuous(breaks = seq(1, nbins + 1)) +
    theme(legend.title=element_blank()) + 
    labs(y = "Deviation", x="Training Block") 
#ggsave(here("Assets/figs/e1_train_deviation.png"), p1, width = 8, height = 4,bg="white")

```

::: {#fig-e1-train-dev}

![](../Assets/figs/e1_train_deviation.png)

E1. Deviations from target band during training
:::

```{r}
##| label: tbl-e1-train-dist
##| tbl-cap: "Experiment 1 - Learning curves. "
##| output: asis

bmm_e1_train<- trainE1_max %>% 
  brm(dist ~ condit, 
      file=here("data/model_cache/e1_train_deviation"),
      data = .,
      iter = 2000,
      chains = 4,
      control = list(adapt_delta = .94, max_treedepth = 13))
mtr1 <- as.data.frame(describe_posterior(bmm_e1_train, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mtr1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

# mtr1 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
#   tibble::remove_rownames() |> 
#   mutate(Term = stringr::str_remove(Term, "b_")) |>
#    kable(booktabs = TRUE)

cdtr1 <- get_coef_details(bmm_e1_train, "conditVaried")
```

::: {#tbl-e1-train-dist}

| Term         | Estimate | 95% CrI Lower | 95% CrI Upper |  pd |
|:-------------|---------:|--------------:|--------------:|----:|
| Intercept    |   106.34 |         95.46 |        117.25 |   1 |
| conditVaried |    79.64 |         57.92 |        101.63 |   1 |

: **Experiment 1 - End of training performance**. The Intercept represents the average of the baseline (constant condition), and the conditVaried coefficient reflects the difference between the constant and varied groups. A larger positive estimates indicates a greater deviation (lower accuracy) for the varied group.  {.sm}

:::

\

*Training*. @fig-e1-train-dev displays the average deviations across training blocks for the varied group, which trained on three velocity bands, and the constant group, which trained on one velocity band. To compare the training conditions at the end of training, we analyzed performance on the 800-1000 velocity band, which both groups trained on. The full model results are shown in Table 1. The varied group had a significantly greater deviation than the constant group in the final training block, ($B$ = `r cdtr1$estimate`, 95% CrI \[`r cdtr1$conf.low`, `r cdtr1$conf.high`\]; pd = `r cdtr1$pd`).

```{r}
##| label: tbl-e1-bmm-dist
##| tbl-cap: "E1. Training vs. Extrapolation"
#| 
modelFile <- paste0(here::here("data/model_cache/"), "e1_dist_Cond_Type_RF_2")
bmtd <- brm(dist ~ condit * bandType + (1|bandInt) + (1|id), 
    data=testE1, file=modelFile,
    iter=5000,chains=4, control = list(adapt_delta = .94, max_treedepth = 13))

# mted1 <- as.data.frame(describe_posterior(bmtd, centrality = "Mean"))[, c(1,2,4,5,6)]
# colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

# r_bandInt_params <- get_variables(bmtd)[grepl("r_bandInt", get_variables(bmtd))]
# posterior_summary(bmtd,variable=r_bandInt_params)
# 
# r_bandInt_params <- get_variables(bmtd)[grepl("r_id:bandInt", get_variables(bmtd))]
# posterior_summary(bmtd,variable=r_bandInt_params)

# mted1 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
#   tibble::remove_rownames() |> 
#   mutate(Term = stringr::str_remove(Term, "b_")) |> kable(booktabs = TRUE)
cdted1 <- get_coef_details(bmtd, "conditVaried")
cdted2 <-get_coef_details(bmtd, "bandTypeExtrapolation")
cdted3 <-get_coef_details(bmtd, "conditVaried:bandTypeExtrapolation")
```

::: {#tbl-e1-bmm-dist}

| Term                               | Estimate | 95% CrI Lower | 95% CrI Upper |  pd |
|:------------------------|-----------:|-----------:|-----------:|-----------:|
| Intercept                          |   152.55 |         70.63 |        229.85 | 1.0 |
| conditVaried                       |    39.00 |        -21.10 |        100.81 | 0.9 |
| bandTypeExtrapolation              |    71.51 |         33.24 |        109.60 | 1.0 |
| conditVaried:bandTypeExtrapolation |    66.46 |         32.76 |         99.36 | 1.0 |

: **Experiment 1 testing accuracy**. Main effects of condition and band type (training vs. extrapolation), and the interaction between the two factors. Larger coefficient estimates indicate larger deviations from the baselines (constant & trained bands) - and a positive interaction coefficient indicates disproporionate deviation for the varied condition on the extrapolation bands {.striped .hover .sm}
:::

*Testing.* To compare accuracy between groups in the testing stage, we fit a Bayesian mixed effects model predicting deviation from the target band as a function of training condition (varied vs. constant) and band type (trained vs. extrapolation), with random intercepts for participants and bands. The model results are shown in @tbl-e1-bmm-dist. The main effect of training condition was not significant ($B$ = `r cdted1$estimate`, 95% CrI \[`r cdted1$conf.low`, `r cdted1$conf.high`\]; pd = `r cdted1$pd`). The extrapolation testing items had a significantly greater deviation than the training bands (β = `r cdted2$estimate`, 95% CrI \[`r cdted2$conf.low`, `r cdted2$conf.high`\]; pd = `r cdted2$pd`). Most importantly, the interaction between training condition and band type was significant ($B$ = `r cdted3$estimate`, 95% CrI \[`r cdted3$conf.low`, `r cdted3$conf.high`\]; pd = `r cdted3$pd`), As shown in @fig-e1-test-dev,  the varied group had disproportionately larger deviations compared to the constant group in the extrapolation bands.

```{r}
#| label: fig-e1-test-dev
#| fig-cap: E1. Deviations from target band during testing without feedback stage. 
#| eval: FALSE
pe1td <- testE1 |>  ggplot(aes(x = vb, y = dist,fill=condit)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation From Target")

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

}

pe1ce <- bmtd |> emmeans( ~condit + bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Absolute Deviation From Band", x="Band Type")

p2 <- (pe1td + pe1ce) + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e1_test-dev.png"), p2, width=8, height=4, bg="white")

```

::: {#fig-e1-test-dev}

![](../Assets/figs/e1_test-dev.png)

E1. A) Deviations from target band during testing without feedback stage. B) Estimated marginal means for the interaction between training condition and band type. Error bars represent 95% confidence intervals.

:::

```{r}
##| label: tbl-e1-bmm-vx
##| tbl-cap: "Experiment 1. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))

#GetModelStats(e1_vxBMM) |> kable(booktabs = TRUE)

cd1 <- get_coef_details(e1_vxBMM, "conditVaried")
sc1 <- get_coef_details(e1_vxBMM, "bandInt")
intCoef1 <- get_coef_details(e1_vxBMM, "conditVaried:bandInt")
```

\

::: {#tbl-e1-bmm-vx}

| Term         | Estimate | 95% CrI Lower | 95% CrI Upper |   pd |
|:-------------|---------:|--------------:|--------------:|-----:|
| Intercept    |   408.55 |        327.00 |        490.61 | 1.00 |
| conditVaried |   164.05 |         45.50 |        278.85 | 1.00 |
| Band         |     0.71 |          0.62 |          0.80 | 1.00 |
| condit\*Band |    -0.14 |         -0.26 |         -0.01 | 0.98 |

Experiment 1. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band {.striped .hover .sm}
:::

Finally, to assess the ability of both conditions to discriminate between velocity bands, we fit a model predicting velocity as a function of training condition and velocity band, with random intercepts and random slopes for each participant. See @tbl-e1-bmm-vx for the full model results. The estimated coefficient for training condition (β = `r cd1$estimate`, 95% CrI \[`r cd1$conf.low`, `r cd1$conf.high`\]) suggests that the varied group tends to produce harder throws than the constant group, but is not in and of itself useful for assessing discrimination. Most relevant to the issue of discrimination is the slope on Velocity Band (β = `r sc1$estimate`, 95% CrI \[`r sc1$conf.low`, `r sc1$conf.high`\]). Although the median slope does fall underneath the ideal of value of 1, the fact that the 95% credible interval does not contain 0 provides strong evidence that participants exhibited some discrimination between bands. The estimate for the interaction between slope and condition (β = `r intCoef1$estimate`, 95% CrI \[`r intCoef1$conf.low`, `r intCoef1$conf.high`\]), suggests that the discrimination was somewhat modulated by training condition, with the varied participants showing less sensitivity between bands than the constant condition. This difference is depicted visually in @fig-e1-test-vx.

```{r}
#| label: fig-e1-test-vx
#| fig-cap: Experiment 1. Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands. 
#| fig-height: 16
#| fig-width: 12
#| eval: FALSE

pe1tv <- testE1 %>% group_by(id,vb,condit) |> plot_distByCondit()

pe1vce <- e1_vxBMM |> emmeans( ~condit + bandInt,re_formula=NA, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects(bandInt) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE1$vb), 
                     limits = c(0, 1400)) + 
  scale_y_continuous(expand=expansion(add=100),breaks=round(seq(0,2000,by=200),2)) +
  theme(legend.title=element_blank()) + 
  labs(y="Velcoity", x="Band")

fe <- fixef(e1_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e1_vxBMM)[,1]["bandInt"]
fixed_effect_interaction <- fixef(e1_vxBMM)[,1]["conditVaried:bandInt"]

re <- data.frame(ranef(e1_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e1Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction*(condit=="Varied"),slope = Estimate + adjust )

pid_den1 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density")

pid_slopes1 <- re |>  mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
     theme(legend.title=element_blank(), 
           axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_wrap2(~condit,axes="all",scales="free_y")

p3 <- pe1tv / (pe1vce + pid_den1 + pid_slopes1) + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e1_test-vx.png"), p3,width=9,height=11, bg="white",dpi=600)
p3
```

{{< pagebreak >}}

::: {#fig-e1-test-vx}

![](../Assets/figs/e1_test-vx.png)

Experiment 1. Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands.

:::

{{< pagebreak >}}

## E1 Discussion

In Experiment 1, we investigated how variability in training influenced participants' ability learn and extrapolate in a visuomotor task. Our findings that training with variable conditions  rresulted in lower final training performance is consistent with much of the prior researchon the influence of training variability [@ravivHowVariabilityShapes2022; @soderstromLearningPerformanceIntegrative2015], and is particularly unsurprising in the present work, given that the constant group received three times the amount of training on the velocity band common to the two conditions.

More importantly, the varied training group exhibited significantly larger deviations from the target velocity bands during the testing phase, particularly for the extrapolation bands that were not encountered by either condition during training.

## References

::: {#refs}
:::