### 2024-03-17 Analysis/e2.qmd

---
title: "Experiment 2"
categories: [Analyses, R, Bayesian]
toc: false
format-links: false
page-layout: full
cache: true
code-fold: true
cold-tools: true
lightbox: true
# execute: 
#   cache: true
keep-md: true
format:
  html: default
  hugo-md:
    echo: false
    html-math-method: mathjax
    output-file: e2-hugo.md
  gfm:
    echo: true
    output-file: e2-gfm.md
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  stringr, here,conflicted, patchwork, knitr,kableExtra)
#options(brms.backend="cmdstanr",mc.cores=4)
walk(c("brms","dplyr","bayestestR"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","org_functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
e2 <- readRDS(here("data/e2_08-04-23.rds")) 
e2Sbjs <- e2 |> group_by(id,condit) |> summarise(n=n())
testE2 <- e2 |> filter(expMode2 == "Test")
nbins=5
trainE2 <-  e2 |> filter(expMode2=="Train") |> group_by(id,condit, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
trainE2_max <- trainE2 |> filter(Trial_Bin == nbins, bandInt==600)
```

@fig-design-e2 illustrates the design of Experiment 2. The stages of the experiment (i.e. training, testing no-feedback, test with feedback), are identical to that of Experiment 1. The only change is that Experiment 2 participants train, and then test, on bands in the reverse order of Experiment 1 (i.e. training on the softer bands; and testing on the harder bands). 
```{dot}
//| label: fig-design-e2
//| fig-cap: "Experiment 2 Design. Constant and Varied participants complete different training conditions. The training and testing bands are the reverse of Experiment 1. "
//| fig-width: 8.0
//| fig-height: 2.5
//| fig-responsive: false
//| column: screen-inset-right
//| eval: FALSE 
digraph {
  graph [layout = dot, rankdir = LR]
  node [shape = rectangle, style = filled]
  data1 [label = " Varied Training \n100-300\n350-550\n600-800", fillcolor = "#FF0000"]
  data2 [label = " Constant Training \n600-800", fillcolor = "#00A08A"]
  Test3 [label = "    Final Test \n  Novel With Feedback  \n800-1000\n1000-1200\n1200-1400", fillcolor = "#ECCBAE"]
  data1 -> Test1
  data2 -> Test1
  subgraph cluster {
    label = "Test Phase \n(Counterbalanced Order)"
    Test1 [label = "Test  \nNovel Bands  \n800-1000\n1000-1200\n1200-1400", fillcolor = "#ECCBAE"]
    Test2 [label = "  Test \n  Varied Training Bands  \n100-300\n350-550\n600-800", fillcolor = "#ECCBAE"]
    Test1 -> Test2
  }
  Test2 -> Test3
}
```

::: {#fig-design-e2}

![](../Assets/figs/e2_design.png)

Experiment 2 Design. Constant and Varied participants complete different training conditions. The training and testing bands are the reverse of Experiment 1.
:::

```{r}
#| label: fig-e2-train-dev
#| fig-cap: E2. Deviations from target band across training blocks. 
#| eval: FALSE
p1 <- trainE2 |> ggplot(aes(x = Trial_Bin, y = dist, color = condit)) +
    stat_summary(geom = "line", fun = mean) +
    stat_summary(geom = "errorbar", fun.data = mean_se, width = .4, alpha = .7) +
    facet_wrap(~vb)+
    scale_x_continuous(breaks = seq(1, nbins + 1)) +
    theme(legend.title=element_blank()) + 
    labs(y = "Deviation", x="Training Block") 
#ggsave(here("Assets/figs/e2_train_deviation.png"), p1, width = 8, height = 4,bg="white")

```

::: {#fig-e2-train-dev}

![](../Assets/figs/e2_train_deviation.png)

E2. Deviations from target band across training blocks. 
:::

```{r}
bmm_e2_train <- trainE2_max %>% 
  brm(dist ~ condit, 
      file=here("data/model_cache/e2_train_deviation"),
      data = .,
      iter = 2000,
      chains = 4,
      control = list(adapt_delta = .94, max_treedepth = 13))

mtr2 <- as.data.frame(describe_posterior(bmm_e2_train, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mtr2) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

cdtr2 <- get_coef_details(bmm_e2_train, "conditVaried")
# mtr2 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
#   tibble::remove_rownames() |> 
#   mutate(Term = stringr::str_remove(Term, "b_")) |>
#   kable(escape=F,booktabs=T) 
```

::: {#tbl-e2-train-dist}

| Term         | Estimate | 95% CrI Lower | 95% CrI Upper |  pd |
|:-------------|---------:|--------------:|--------------:|----:|
| Intercept    |    91.01 |         80.67 |        101.26 |   1 |
| conditVaried |    36.15 |         16.35 |         55.67 |   1 |

: **Experiment 2 - End of training performance**. The Intercept represents the average of the baseline (constant condition), and the conditVaried coefficient reflects the difference between the constant and varied groups. A larger positive coefficient indicates a greater deviation (lower accuracy) for the varied group.  {.sm}

:::

\

*Training*. @fig-e2-train-dev displays the average deviations across training blocks for both training groups. To compare the training conditions at the end of training, we analyzed performance on the 600-800 velocity band, which both groups trained on. The full model results are shown in Table 1. The varied group had a significantly greater deviation than the constant group in the final training block, ($B$ = `r cdtr2$estimate`, 95% CrI \[`r cdtr2$conf.low`, `r cdtr2$conf.high`\]; pd = `r cdtr2$pd`).

```{r}
modelFile <- paste0(here::here("data/model_cache/"), "e2_dist_Cond_Type_RF_2")
bmtd2 <- brm(dist ~ condit * bandType + (1|bandInt) + (1|id), 
    data=testE2, file=modelFile,
    iter=5000,chains=4, control = list(adapt_delta = .94, max_treedepth = 13))

# mted2 <- as.data.frame(describe_posterior(bmtd2, centrality = "Mean"))[, c(1,2,4,5,6)]
# colnames(mted2) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")
# mted2 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
#   tibble::remove_rownames() |> 
#   mutate(Term = stringr::str_remove(Term, "b_")) |>
#   kable(booktabs=TRUE)

cd2ted1 <- get_coef_details(bmtd2, "conditVaried")
cd2ted2 <-get_coef_details(bmtd2, "bandTypeExtrapolation")
cd2ted3 <-get_coef_details(bmtd2, "conditVaried:bandTypeExtrapolation")
```

::: {#tbl-e2-bmm-dist}

| Term                               | Estimate | 95% CrI Lower | 95% CrI Upper |   pd |
|:-----------------------------------|---------:|--------------:|--------------:|-----:|
| Intercept                          |   190.91 |        125.03 |        259.31 | 1.00 |
| conditVaried                       |   -20.58 |        -72.94 |         33.08 | 0.78 |
| bandTypeExtrapolation              |    38.09 |         -6.94 |         83.63 | 0.95 |
| conditVaried:bandTypeExtrapolation |    82.00 |         41.89 |        121.31 | 1.00 |

: **Experiment 2 testing accuracy**. Main effects of condition and band type (training vs. extrapolation), and the interaction between the two factors. Larger coefficient estimates indicate larger deviations from the baselines (constant & trained bands) - and a positive interaction coefficient indicates disproporionate deviation for the varied condition on the extrapolation bands {.striped .hover .sm}
:::

*Testing.* To compare accuracy between groups in the testing stage, we fit a Bayesian mixed effects model predicting deviation from the target band as a function of training condition (varied vs. constant) and band type (trained vs. extrapolation), with random intercepts for participants and bands. The model results are shown in @tbl-e2-bmm-dist. The main effect of training condition was not significant ($B$ = `r cd2ted1$estimate`, 95% CrI \[`r cd2ted1$conf.low`, `r cd2ted1$conf.high`\]; pd = `r cd2ted1$pd`). The extrapolation testing items had a significantly greater deviation than the training bands (β = `r cd2ted2$estimate`, 95% CrI \[`r cd2ted2$conf.low`, `r cd2ted2$conf.high`\]; pd = `r cd2ted2$pd`). Most importantly, the interaction between training condition and band type was significant ($B$ = `r cd2ted3$estimate`, 95% CrI \[`r cd2ted3$conf.low`, `r cd2ted3$conf.high`\]; pd = `r cd2ted3$pd`), As shown in @fig-e2-test-dev,  the varied group had disproportionately larger deviations compared to the constant group in the extrapolation bands.

```{r}
#| label: fig-e2-test-dev
#| fig-cap: E2. Deviations from target band during testing without feedback stage. 
#| eval: FALSE
pe2td <- testE2 |>  ggplot(aes(x = vb, y = dist,fill=condit)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation From Target")

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) 
}

pe2ce <- bmtd2 |> emmeans( ~condit + bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Absolute Deviation From Band", x="Band Type")

p2 <- (pe2td + pe2ce) + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e2_test-dev.png"), p2, width=8, height=4, bg="white")
```

::: {#fig-e2-test-dev}

![](../Assets/figs/e2_test-dev.png)

E2. A) Deviations from target band during testing without feedback stage. B) Estimated marginal means for the interaction between training condition and band type. Error bars represent 95% confidence intervals.
:::

### Discimination

```{r}
##| label: tbl-e2-bmm-vx
##| tbl-cap: "Experiment 2. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"

e2_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e2_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))

#GetModelStats(e2_vxBMM ) |> kable(escape=F,booktabs=T, caption="Fit to all 6 bands")

cd2 <- get_coef_details(e2_vxBMM, "conditVaried")
sc2 <- get_coef_details(e2_vxBMM, "bandInt")
intCoef2 <- get_coef_details(e2_vxBMM, "conditVaried:bandInt")
```

::: {#tbl-e2-bmm-vx}

| Term         | Estimate | 95% CrI Lower | 95% CrI Upper |   pd |
|:-------------|---------:|--------------:|--------------:|-----:|
| Intercept    |   362.64 |        274.85 |        450.02 | 1.00 |
| conditVaried |    -8.56 |       -133.97 |        113.98 | 0.55 |
| Band         |     0.71 |          0.58 |          0.84 | 1.00 |
| condit\*Band |    -0.06 |         -0.24 |          0.13 | 0.73 |

Experiment 2. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band {.striped .hover .sm}
:::

See @tbl-e2-bmm-vx for the full model results. The estimated coefficient for training condition (β = `r cd2$estimate`, 95% CrI \[`r cd2$conf.low`, `r cd2$conf.high`\]) suggests that the varied group tends to produce harder throws than the constant group, but is not in and of itself useful for assessing discrimination. Most relevant to the issue of discrimination is the slope on Velocity Band (β = `r sc2$estimate`, 95% CrI \[`r sc2$conf.low`, `r sc2$conf.high`\]). Although the median slope does fall underneath the ideal of value of 1, the fact that the 95% credible interval does not contain 0 provides strong evidence that participants exhibited some discrimination between bands. The estimate for the interaction between slope and condition (β = `r intCoef2$estimate`, 95% CrI \[`r intCoef2$conf.low`, `r intCoef2$conf.high`\]), suggests that the discrimination was somewhat modulated by training condition, with the varied participants showing less sensitivity between bands than the constant condition. This difference is depicted visually in @fig-e2-test-vx.

```{r}
#| label: e2_test-vx
#| fig-cap: Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands. 
#| fig-height: 16
#| fig-width: 12
#| eval: FALSE

pe2tv <- testE2 %>% group_by(id,vb,condit) |> plot_distByCondit()

pe2vce <- e2_vxBMM |> emmeans( ~condit + bandInt,re_formula=NA, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects(bandInt) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE2$vb), 
                     limits = c(0, 1400)) + 
scale_y_continuous(expand=expansion(add=100),breaks=round(seq(0,2000,by=200),2)) +
  theme(legend.title=element_blank()) + 
  labs(y="Velcoity", x="Band")

fe <- fixef(e2_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e2_vxBMM)[,1]["bandInt"]
fixed_effect_interaction <- fixef(e2_vxBMM)[,1]["conditVaried:bandInt"]

re <- data.frame(ranef(e2_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e2Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction*(condit=="Varied"),slope = Estimate + adjust )

pid_den2 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density")

pid_slopes2 <- re |>  mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
      theme(legend.title=element_blank(), 
        axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_wrap2(~condit,axes="all",scales="free_y")

p3 <- pe2tv / (pe2vce + pid_den2 + pid_slopes2) + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e2_test-vx.png"), p3,width=9,height=11, bg="white",dpi=600)
#p3
```

{{< pagebreak >}}

::: {#fig-e2-test-vx}

![](../Assets/figs/e2_test-vx.png)

Experiment 2. Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands.

:::

## E2 Discussion