### 2024-03-18 Analysis/e2.qmd
e2/e3 updates

---
title: "Experiment 2"
categories: [Analyses, R, Bayesian]
toc: false
format-links: false
page-layout: full
cache: true
code-fold: true
cold-tools: true
lightbox: true
# execute: 
#   cache: true
keep-md: true
format:
  html: default
  hugo-md:
    echo: false
    html-math-method: mathjax
    output-file: e2-hugo.md
  gfm:
    echo: true
    output-file: e2-gfm.md
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  stringr, here,conflicted, patchwork, knitr,kableExtra)
#options(brms.backend="cmdstanr",mc.cores=4)
walk(c("brms","dplyr","bayestestR"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","org_functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
e2 <- readRDS(here("data/e2_08-04-23.rds")) 
e2Sbjs <- e2 |> group_by(id,condit) |> summarise(n=n())
testE2 <- e2 |> filter(expMode2 == "Test")
nbins=5
trainE2 <-  e2 |> filter(expMode2=="Train") |> group_by(id,condit, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
trainE2_max <- trainE2 |> filter(Trial_Bin == nbins, bandInt==600)
```

@fig-design-e2 illustrates the design of Experiment 2. The stages of the experiment (i.e. training, testing no-feedback, test with feedback), are identical to that of Experiment 1. The only change is that Experiment 2 participants train, and then test, on bands in the reverse order of Experiment 1 (i.e. training on the softer bands; and testing on the harder bands). 
```{dot}
//| label: fig-design-e2
//| fig-cap: "Experiment 2 Design. Constant and Varied participants complete different training conditions. The training and testing bands are the reverse of Experiment 1. "
//| fig-width: 8.0
//| fig-height: 2.5
//| fig-responsive: false
//| column: screen-inset-right
//| eval: FALSE 
digraph {
  graph [layout = dot, rankdir = LR]
  node [shape = rectangle, style = filled]
  data1 [label = " Varied Training \n100-300\n350-550\n600-800", fillcolor = "#FF0000"]
  data2 [label = " Constant Training \n600-800", fillcolor = "#00A08A"]
  Test3 [label = "    Final Test \n  Novel With Feedback  \n800-1000\n1000-1200\n1200-1400", fillcolor = "#ECCBAE"]
  data1 -> Test1
  data2 -> Test1
  subgraph cluster {
    label = "Test Phase \n(Counterbalanced Order)"
    Test1 [label = "Test  \nNovel Bands  \n800-1000\n1000-1200\n1200-1400", fillcolor = "#ECCBAE"]
    Test2 [label = "  Test \n  Varied Training Bands  \n100-300\n350-550\n600-800", fillcolor = "#ECCBAE"]
    Test1 -> Test2
  }
  Test2 -> Test3
}
```

::: {#fig-design-e2}

![](../Assets/figs/e2_design.png)

Experiment 2 Design. Constant and Varied participants complete different training conditions. The training and testing bands are the reverse of Experiment 1.
:::

```{r}
#| label: fig-e2-train-dev
#| fig-cap: E2. Deviations from target band across training blocks. 
#| eval: FALSE
p1 <- trainE2 |> ggplot(aes(x = Trial_Bin, y = dist, color = condit)) +
    stat_summary(geom = "line", fun = mean) +
    stat_summary(geom = "errorbar", fun.data = mean_se, width = .4, alpha = .7) +
    facet_wrap(~vb)+
    scale_x_continuous(breaks = seq(1, nbins + 1)) +
    theme(legend.title=element_blank()) + 
    labs(y = "Deviation", x="Training Block") 
#ggsave(here("Assets/figs/e2_train_deviation.png"), p1, width = 8, height = 4,bg="white")

```

::: {#fig-e2-train-dev}

![](../Assets/figs/e2_train_deviation.png)

E2. Deviations from target band across training blocks. 
:::

```{r}
bmm_e2_train <- trainE2_max %>% 
  brm(dist ~ condit, 
      file=here("data/model_cache/e2_train_deviation"),
      data = .,
      iter = 2000,
      chains = 4,
      control = list(adapt_delta = .94, max_treedepth = 13))

mtr2 <- as.data.frame(describe_posterior(bmm_e2_train, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mtr2) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

cdtr2 <- get_coef_details(bmm_e2_train, "conditVaried")
# mtr2 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
#   tibble::remove_rownames() |> 
#   mutate(Term = stringr::str_remove(Term, "b_")) |>
#   kable(escape=F,booktabs=T) 
```

::: {#tbl-e2-train-dist}

| Term         | Estimate | 95% CrI Lower | 95% CrI Upper |  pd |
|:-------------|---------:|--------------:|--------------:|----:|
| Intercept    |    91.01 |         80.67 |        101.26 |   1 |
| conditVaried |    36.15 |         16.35 |         55.67 |   1 |

: **Experiment 2 - End of training performance**. The Intercept represents the average of the baseline (constant condition), and the conditVaried coefficient reflects the difference between the constant and varied groups. A larger positive coefficient indicates a greater deviation (lower accuracy) for the varied group.  {.sm}

:::

\

*Training*. @fig-e2-train-dev presents the deviations across training blocks for both constant and varied training groups. We again compared training performance on the band common to both groups (600-800). The full model results are shown in Table 1. The varied group had a significantly greater deviation than the constant group in the final training block, ( $\beta$ = `r cdtr2$estimate`, 95% CrI \[`r cdtr2$conf.low`, `r cdtr2$conf.high`\]; pd = `r cdtr2$pd`).

```{r}
modelFile <- paste0(here::here("data/model_cache/"), "e2_dist_Cond_Type_RF_2")
bmtd2 <- brm(dist ~ condit * bandType + (1|bandInt) + (1|id), 
    data=testE2, file=modelFile,
    iter=5000,chains=4, control = list(adapt_delta = .94, max_treedepth = 13))

# mted2 <- as.data.frame(describe_posterior(bmtd2, centrality = "Mean"))[, c(1,2,4,5,6)]
# colnames(mted2) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")
# mted2 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
#   tibble::remove_rownames() |> 
#   mutate(Term = stringr::str_remove(Term, "b_")) |>
#   kable(booktabs=TRUE)

cd2ted1 <- get_coef_details(bmtd2, "conditVaried")
cd2ted2 <-get_coef_details(bmtd2, "bandTypeExtrapolation")
cd2ted3 <-get_coef_details(bmtd2, "conditVaried:bandTypeExtrapolation")
```

::: {#tbl-e2-bmm-dist}

| Term                               | Estimate | 95% CrI Lower | 95% CrI Upper |   pd |
|:-----------------------------------|---------:|--------------:|--------------:|-----:|
| Intercept                          |   190.91 |        125.03 |        259.31 | 1.00 |
| conditVaried                       |   -20.58 |        -72.94 |         33.08 | 0.78 |
| bandTypeExtrapolation              |    38.09 |         -6.94 |         83.63 | 0.95 |
| conditVaried:bandTypeExtrapolation |    82.00 |         41.89 |        121.31 | 1.00 |

: **Experiment 2 testing accuracy**. Main effects of condition and band type (training vs. extrapolation), and the interaction between the two factors. Larger coefficient estimates indicate larger deviations from the baselines (constant & trained bands) - and a positive interaction coefficient indicates disproporionate deviation for the varied condition on the extrapolation bands {.striped .hover .sm}
:::

*Testing.* The analysis of testing accuracy examined deviations from the target band as influenced by training condition (Varied vs. Constant) and band type (training vs. extrapolation bands). The results, summarized in @tbl-e2-bmm-dist, reveal no significant main effect of training condition ($\beta$ = `r cd2ted1$estimate`, 95% CrI \[`r cd2ted1$conf.low`, `r cd2ted1$conf.high`\]; pd = `r cd2ted1$pd`). However, the interaction between training condition and band type was significant ($\beta$ = `r cd2ted3$estimate`, 95% CrI \[`r cd2ted3$conf.low`, `r cd2ted3$conf.high`\]; pd = `r cd2ted3$pd`), with the varied group showing disproportionately larger deviations compared to the constant group on the extrapolation bands (see @fig-e2-test-dev).

```{r}
#| label: fig-e2-test-dev
#| fig-cap: E2. Deviations from target band during testing without feedback stage. 
#| eval: FALSE
pe2td <- testE2 |>  ggplot(aes(x = vb, y = dist,fill=condit)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation From Target")

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) 
}

pe2ce <- bmtd2 |> emmeans( ~condit + bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Absolute Deviation From Band", x="Band Type")

p2 <- (pe2td + pe2ce) + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e2_test-dev.png"), p2, width=8, height=4, bg="white")
```

::: {#fig-e2-test-dev}

![](../Assets/figs/e2_test-dev.png)

E2. A) Deviations from target band during testing without feedback stage. B) Estimated marginal means for the interaction between training condition and band type. Error bars represent 95% confidence intervals.
:::

### Discimination

```{r}
##| label: tbl-e2-bmm-vx
##| tbl-cap: "Experiment 2. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"

e2_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e2_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))

#GetModelStats(e2_vxBMM ) |> kable(escape=F,booktabs=T, caption="Fit to all 6 bands")

cd2 <- get_coef_details(e2_vxBMM, "conditVaried")
sc2 <- get_coef_details(e2_vxBMM, "bandInt")
intCoef2 <- get_coef_details(e2_vxBMM, "conditVaried:bandInt")
```

::: {#tbl-e2-bmm-vx}

| Term         | Estimate | 95% CrI Lower | 95% CrI Upper |   pd |
|:-------------|---------:|--------------:|--------------:|-----:|
| Intercept    |   362.64 |        274.85 |        450.02 | 1.00 |
| conditVaried |    -8.56 |       -133.97 |        113.98 | 0.55 |
| Band         |     0.71 |          0.58 |          0.84 | 1.00 |
| condit\*Band |    -0.06 |         -0.24 |          0.13 | 0.73 |

Experiment 2. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band {.striped .hover .sm}
:::

Finally, to assess the ability of both conditions to discriminate between velocity bands, we fit a model predicting velocity as a function of training condition and velocity band, with random intercepts and random slopes for each participant. The full model results are shown in @tbl-e2-bmm-vx. The overall slope on target velocity band predictor was significantly positive, ($\beta$ = `r sc2$estimate`, 95% CrI \[`r sc2$conf.low`, `r sc2$conf.high`\]; pd= `r sc2$pd`), indicating that participants exhibited discrimination between bands. The interaction between slope and condition was not significant, ($\beta$ = `r intCoef2$estimate`, 95% CrI \[`r intCoef2$conf.low`, `r intCoef2$conf.high`\]; pd= `r intCoef2$pd`), suggesting that the two conditions did not differ in their ability to discriminate between bands (see @fig-e2-test-vx).

```{r}
#| label: e2_test-vx
#| fig-cap: Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands. 
#| fig-height: 16
#| fig-width: 12
#| eval: FALSE

pe2tv <- testE2 %>% group_by(id,vb,condit) |> plot_distByCondit()

pe2vce <- e2_vxBMM |> emmeans( ~condit + bandInt,re_formula=NA, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects(bandInt) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE2$vb), 
                     limits = c(0, 1400)) + 
scale_y_continuous(expand=expansion(add=100),breaks=round(seq(0,2000,by=200),2)) +
  theme(legend.title=element_blank()) + 
  labs(y="Velcoity", x="Band")

fe <- fixef(e2_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e2_vxBMM)[,1]["bandInt"]
fixed_effect_interaction <- fixef(e2_vxBMM)[,1]["conditVaried:bandInt"]

re <- data.frame(ranef(e2_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e2Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction*(condit=="Varied"),slope = Estimate + adjust )

pid_den2 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density")

pid_slopes2 <- re |>  mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
      theme(legend.title=element_blank(), 
        axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_wrap2(~condit,axes="all",scales="free_y")

p3 <- pe2tv / (pe2vce + pid_den2 + pid_slopes2) + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e2_test-vx.png"), p3,width=9,height=11, bg="white",dpi=600)
#p3
```

{{< pagebreak >}}

::: {#fig-e2-test-vx}

![](../Assets/figs/e2_test-vx.png)

Experiment 2. Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands.

:::

## E2 Discussion

Experiment 2 extended the findings of Experiment 1 by examining the effects of training variability on extrapolation performance in a visuomotor function learning task, but with reversed training and testing bands. Similar to Experiment 1, the Varied group exhibited poorer performance during training and testing. However unlike experiment 1, the Varied group did not show a significant difference in discrimination between bands.