### 2024-03-20 Sections/combo1.qmd
create combo file and re-render

---
title: combo1
author: "Thomas Gorman"
toc: true
toc-depth: 5
toc-expand: true
number-sections: true
number-depth: 5
strip-comments: true
cold-tools: true
echo: false
format: 
  html:
    column: 
      - screen-inset-right
  hugo-md:
    echo: false
    html-math-method: mathjax
    output-file: htw-combo1-hugo.md
  # hikmah-manuscript-pdf:
  #   echo: false
  #   output-file: "hikmah-m_combo.pdf"
  #   mainfont: "Linux Libertine O"
  #   mainfontoptions:
  #     - "Numbers=Proportional"
  #     - "Numbers=OldStyle"
  #   mathfont: "Libertinus Math"
  # hikmah-pdf:
  #   echo: false
  #   cite-method: citeproc
  #   output-file: "hikmah_combo.pdf"
  #   mainfont: "Linux Libertine O"
  #   mainfontoptions:
  #     - "Numbers=Proportional"
  #     - "Numbers=OldStyle"
  #   sansfont: "IBM Plex Sans"
  #   mathfont: "Libertinus Math"
  #   geometry:
  #     - top=1in     
  #     - bottom=1in  
  #     - left=1in    
  #     - right=1in  
  #     - heightrounded
  #   # linestretch: 1.05
  #   include-in-header:
  #     # The \everydisplay thing makes all math blocks use \small font
  #     text: |
  #       \usepackage{mathtools}
  #       \everydisplay\expandafter{\the\everydisplay\small}
  #       \usepackage{float}
  #       \floatplacement{figure}{H}
  #       \floatplacement{table}{H}
  gfm:
    echo: true
    output-file: e1-gfm.md
#toc: true
# toc-title: Contents
# toc-depth: 4
# toc-expand: 1
# execute:
#   echo: false
---

## Introduction

{{< include ../Sections/Intro.qmd >}}

## Methods

{{< include ../Sections/Methods.qmd >}}

## Experiment 1

{{< include ../Analysis/e1.qmd >}}

## Experiment 2

{{< include ../Analysis/e2.qmd >}}

## Experiment 3

{{< include ../Analysis/e3.qmd >}}

## Results Summary

## Overall Summary of Experiments 1-3

Across three experiments, we investigated the impact of training variability on learning, transfer, and extrapolation in a visuomotor function learning task. In Experiment 1, participants in the varied training condition, who experienced a wider range of velocity bands during training, showed lower accuracy at the end of training compared to those in the constant training condition. Crucially, during the testing phase, the varied group exhibited significantly larger deviations from the target velocity bands, particularly for the extrapolation bands that were not encountered during training. The varied group also showed less discrimination between velocity bands, as evidenced by shallower slopes when predicting response velocity from target velocity band.

Experiment 2 extended these findings by reversing the order of the training and testing bands. Similar to Experiment 1, the varied group demonstrated poorer performance during both training and testing phases. However, unlike Experiment 1, the varied group did not show a significant difference in discrimination between bands compared to the constant group.

In Experiment 3, we introduced an ordinal feedback system during training, in contrast to the continuous feedback provided in the previous experiments. Participants were assigned to both an order condition (original or reverse) and a training condition (constant or varied). The varied condition showed larger deviations at the end of training, consistent with the previous experiments. Interestingly, there was a significant interaction between training condition and band order, with the varied condition showing greater accuracy in the reverse order condition. During testing, the varied group once again exhibited larger deviations, particularly for the extrapolation bands. The reverse order conditions showed smaller deviations compared to the original order conditions. Discrimination between velocity bands was poorer for the varied group in the original order condition, but not in the reverse order condition.

In summary, these experiments consistently demonstrated that training with greater variability led to lower accuracy during both training and testing phases, with the detrimental effect being more pronounced for extrapolation items. The varied training condition also tended to show less discrimination between velocity bands, although this effect was modulated by the order of the training and testing bands. These findings suggest that while variable training can hinder initial learning, it may have differential effects on transfer and extrapolation depending on the specific task structure and the range of stimuli encountered during training and testing.

## Computational Model

{{< include ../Model/htw_model.qmd >}}

## References

::: {#refs}
:::