### 2024-03-20 Analysis/test_pdf.qmd
e1 cognitive modelling - exploring model comparison via brms models predicting residuals (resid ~ Model * id); re-render

---
title: HTW Model
author: Thomas Gorman
date:  last-modified #"`r Sys.Date()`"
page-layout: full
categories: [Modeling, ALM, EXAM, R]
lightbox: true
toc: false
# code-fold: true
# code-tools: true
execute: 
  warning: false
  eval: true
  echo: false
format:
  html: default
  # Manuscript-y PDF output
  pdf:
    output-file: "manuscript-ms.pdf"
    # Custom fonts
    mainfont: "Linux Libertine O"
    mainfontoptions:
      - "Numbers=Proportional"
      - "Numbers=OldStyle"
    mathfont: "Libertinus Math"
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools,ggdist, kableExtra)
#conflict_prefer_all("dplyr", quiet = TRUE)
walk(c("flextable","dplyr"), conflict_prefer_all, quiet = TRUE)

options(scipen = 999)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
```

### Modelling Results

```{r}
#| cache: false

ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
nbins <- 3

fd <- readRDS(here("data/e1_08-21-23.rds"))
test <- fd |> filter(expMode2 == "Test") 
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

trainAvg <- fd |> filter(expMode2 == "Train") |> group_by(id) |> 
  mutate(tr=trial,x=vb,Block=case_when(expMode2=="Train" ~ cut(tr,breaks=seq(1,max(tr), length.out=nbins+1),include.lowest=TRUE,labels=FALSE),
                                         expMode2=="Test" ~ 4)) |> 
  group_by(id,condit,vb,x,Block) |> 
  summarise(dist=mean(dist),y=mean(vx))

input_layer <<- output_layer <<-  c(100,350,600,800,1000,1200)
ids2 <- c(1,66,36)
file_name <- "n_iter_200_ntry_300_5354"
#file_name <- "n_iter_400_ntry_100_2944"

ind_fits <- map(list.files(here(paste0('data/abc_reject/'),file_name),full.names=TRUE), readRDS)
ind_fits_df <- ind_fits |> map(~list(dat=.x[[1]], Model = .x[["Model"]], Fit_Method=.x[["Fit_Method"]]))
ind_fits_df <- ind_fits_df |> map(~rbindlist(.x$dat) |> mutate(Model = .x$Model, Fit_Method = .x$Fit_Method)) |> rbindlist() 
```

```{r}
#| cache: false

generate_data <- function(Model, post_samples, data, num_samples = 1, return_dat = "train_data, test_data") {
  # Filter data for the specific id without invalidating selfref
  sbj_data <- copy(data[id == post_samples$id[1]])
  simulation_function <- ifelse(Model == "EXAM", full_sim_exam, full_sim_alm)

  target_data <- switch(return_dat,
                        "test_data" = copy(sbj_data[expMode2 == "Test"]),
                        "train_data" = copy(sbj_data[expMode2 == "Train"]),
                        "train_data, test_data" = copy(sbj_data[expMode2 %in% c("Test", "Train")]))

  post_samples <- post_samples[order(mean_error)][1:num_samples, .(c, lr, mean_error, rank = .I)]

  simulated_data_list <- lapply(1:nrow(post_samples), function(i) {
    params <- post_samples[i]
    sim_data <- simulation_function(sbj_data, params$c, params$lr, input_layer = input_layer, 
                                    output_layer = output_layer, return_dat = return_dat)
    sim_data_dt <- data.table(id = sbj_data$id[1], condit = sbj_data$condit[1], 
                              expMode2 = target_data$expMode2, Model = Model,tr=target_data$tr,
                              y = target_data$y, x = target_data$x, c = params$c, 
                              lr = params$lr, mean_error = params$mean_error, rank = i,
                              pred = sim_data)
    return(sim_data_dt)
  })

  result_dt <- rbindlist(simulated_data_list)
  setcolorder(result_dt, c("id", "condit", "expMode2","tr", "c", "lr", "x", "y", "pred"))
  return(result_dt)
}

#future::plan(multisession)

nestSbjModelFit <- ind_fits_df %>% nest(.by=c(id,Model,Fit_Method))

# organize test data predictions
# post_dat <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
#    generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 50, return_dat="test_data")
#    })) |> 
#   select(Fit_Method,pp,-data) |>  
#   unnest(pp) |>  filter(expMode2=="Test") |> as.data.table()
# 
# saveRDS(post_dat, here("data/model_cache/post_dat.rds"))

post_dat <- readRDS(here("data/model_cache/post_dat.rds"))

post_dat_avg <- post_dat |> group_by(id, condit, Model, Fit_Method, x, c, lr, rank) |> 
  mutate(error2 = y - pred) |>
  summarise(y = mean(y), pred = mean(pred), error = y - pred, error2=mean(error2)) |> as.data.table()

setorder(post_dat_avg, id, x, rank)
post_dat_l <- melt(post_dat_avg, id.vars = c("id", "condit", "Model", "Fit_Method", "x", "c", "lr", "rank","error"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val")
post_dat_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))]
setorder(post_dat_l, id, Resp)
#rm(post_dat_avg)

post_dat_l <- post_dat_l |> mutate(dist = case_when(
    val >= x & val <= x + 200 ~ 0,                 
    val < x ~ abs(x - val),                       
    val > x + 200 ~ abs(val - (x + 200)),           
    TRUE ~ NA_real_                                 
  ))

# organize training data predictions
# pd_train <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
#    generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 20, return_dat="train_data")
#    })) |>
#   select(Fit_Method,pp,-data) |>
#   unnest(pp) |> as.data.table() |> filter(expMode2=="Train")

#saveRDS(pd_train, here("data/model_cache/pd_train.rds"))

pd_train <- readRDS(here("data/model_cache/pd_train.rds"))

nbins <- 3
pd_train <- pd_train |> group_by(id,condit,Model,Fit_Method) |>
  mutate(Block=cut(tr,breaks=seq(1,max(tr), length.out=nbins+1),include.lowest=TRUE,labels=FALSE))
setorder(pd_train, id, x,Block, rank)

pd_train_l <- reshape2::melt(pd_train, id.vars = c("id", "condit", "Model","Block", "Fit_Method", "x", "c", "lr", "rank"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val") |> as.data.table()
pd_train_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))] 
setorder(pd_train_l, id,Block, Resp)

pd_train_l <- pd_train_l  |>
  mutate(dist = case_when(
    val >= x & val <= x + 200 ~ 0,                 
    val < x ~ abs(x - val),                       
    val > x + 200 ~ abs(val - (x + 200)),           
    TRUE ~ NA_real_                                 
  ))

#plan(sequential)

```

### Group level aggregations

```{r}
#| label: tbl-htw-modelError
#| tbl-cap: "Mean model errors predicting testing data, aggregated over all participants and velocity bands. Note that Fit Method refers to how model parameters were optimized, while error values reflect mean absolute error for the 6 testing bands"
#| tbl-subcap: 
#|   - "Cars"
#|   - "Pressure"
#| layout-ncol: 2

post_tabs <- abc_tables(post_dat,post_dat_l)
train_tab <- abc_train_tables(pd_train,pd_train_l)

t1 <- post_tabs$agg_pred_full |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) |>
  flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"),
                       `ME` = as_paragraph(mean_error)) |> as_flextable()

#post_tabs$agg_pred_full |> pander::pandoc.table()

t2 <- train_tab$agg_pred_full |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) |>
  flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"),
                       `ME` = as_paragraph(mean_error)) |> as_flextable() |> set_caption("f1")

t1
t2

```

## Markdown subtable

::: {#tbl-panel3334 layout-ncol=2}

`r t1`
: First Table {#tbl-first}

`r t2`
: Second Table {#tbl-second}

Main Caption
:::

```{r}
#| message: false
#| label: tbl-counts
#| tbl-pos: h
#| tbl-cap: "Mean model errors "
#| tbl-subcap:
#|   - "Full datasets"
#|   - "Intersection of samples with all labels available"
#| layout-ncol: 2
format_table <- function(tab) {
  tab %>%
    kable(format.args = list(big.mark = ',')) %>%
    kable_styling(font_size = 10)
}

post_tabs <- abc_tables(post_dat,post_dat_l)

# post_tabs$agg_pred_full |> 
#   mutate(Fit_Method=rename_fm(Fit_Method)) |>
#   flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"),
#                        `ME` = as_paragraph(mean_error)) |> as_flextable()

#post_tabs$agg_pred_full |> pander::pandoc.table()

post_tabs$agg_pred_full |> format_table()

# train_tab$agg_pred_full |> 
#   mutate(Fit_Method=rename_fm(Fit_Method)) |>
#   flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"),
#                        `ME` = as_paragraph(mean_error)) |> as_flextable()

train_tab$agg_pred_full |> format_table()

```

## pander

```{r}
#| label: tbl-example
#| tbl-cap: "Example"
#| tbl-subcap: 
#|   - "Cars"
#|   - "Pressure"
#| layout-ncol: 1

pander::pander(train_tab$agg_pred_full, caption="Train")
pander::pander(post_tabs$agg_pred_full, caption="Test")

```

## kbl

```{r}
#| label: tbl-dist-emmeans2b
#| tbl-cap: Conditional Means and Contrasts
#| tbl-subcap: ["Conditional Means", "Contrasts"]
#| layout-ncol: 2
post_tabs$agg_pred_full |> 
  kbl()

train_tab$agg_pred_full |> kbl()
```

## gt

```{r}
#| label: tbl-dist-emmeans344
#| tbl-cap: Conditional Means and Contrasts
#| tbl-subcap: ["Conditional Means", "Contrasts"]
#| layout-ncol: 2

library(scales)
library(glue)
library(gt)
library(gtExtras)

gt_linebreak <- function(x) {
  if (fmt_out == "latex") {
    x %>% 
      str_replace_all("\\[", "{[}") %>% 
      str_replace_all("\\]", "{]}") %>% 
      kableExtra::linebreak(align = "c")
  } else {
    str_replace(x, "\\n", "<br>")
  }
}

fmt_markdown_latex <- function(x, columns) {
  if (fmt_out == "latex") {
    fmt_passthrough(x, columns = columns, escape = FALSE)
  } else {
    fmt_markdown(x, columns = columns)
  }
}

gt_pct <- function(x, page_width = 6) {
  if (knitr::pandoc_to("latex")) {
    paste0(page_width * (x / 100), "in")
  } else {
    gt::pct(x)
  }
}

post_tabs$agg_pred_full |> gt() |> tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>%
  tab_options(
    quarto.disable_processing = FALSE
  )

train_tab$agg_pred_full |> gt() |> tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>%
  tab_options(
    quarto.disable_processing = FALSE
  )

```

## Markdown subtable

::: {#tbl-panel layout-ncol=2}
| Col1 | Col2 | Col3 |
|------|------|------|
| A    | B    | C    |
| E    | F    | G    |
| A    | G    | G    |

: First Table {#tbl-first}

| Col1 | Col2 | Col3 |
|------|------|------|
| A    | B    | C    |
| E    | F    | G    |
| A    | G    | G    |

: Second Table {#tbl-second}

Main Caption
:::

```{r}
#| label: tbl-dist-emmea7
#| tbl-cap: Conditional Means and Contrasts
#| tbl-subcap: ["Conditional Means", "Contrasts"]
#| layout-ncol: 2

train_tab$agg_pred_full |> pander::pandoc.table()
post_tabs$agg_pred_full |> pander::pandoc.table()
```

```{r}
#| label: tbl-dist-emmeans5
#| tbl-cap: Conditional Means and Contrasts
#| tbl-subcap: ["Conditional Means", "Contrasts"]
#| layout-ncol: 2

train_tab$agg_pred_full |> pander::pandoc.table(style="grid", caption="test cap1")
post_tabs$agg_pred_full |> pander::pandoc.table(style="grid")
```

```{r}
#| label: tbl-dist-emmeans4
#| tbl-cap: Conditional Means and Contrasts
#| tbl-subcap: ["Conditional Means", "Contrasts"]
#| layout-ncol: 2

train_tab$agg_pred_full |> pander::pandoc.table(style="rmarkdown", caption="test cap1")
post_tabs$agg_pred_full |> pander::pandoc.table(style="rmarkdown", caption="test cap2")

```

```{r}

# format_dataframe <- function(df) {
#   df_split <- split(df, df$Group)
#   pander_return <- pander::pandoc.table.return(TRUE)
#   tables <- lapply(df_split, function(x) {
#    # x$Group <- NULL  # Remove the Group column since it's already used for splitting
#     pander(x, style = 'rmarkdown', split.table = Inf)
#   })
#   pander::pandoc.table.return(pander_return)
#   return(tables)
# }
# 
# formatted_tables <- format_dataframe(train_tab$agg_pred_full )
# 
# for (table in formatted_tables) {
#   cat(table, sep = '\n\n')
# }

```