### 2024-03-26 Model/htw_model.qmd
group exam vs alm figure

---
title: HTW Modeling 
author: Thomas Gorman
#date:  #"`r Sys.Date()`"
page-layout: full
categories: [Modeling, ALM, EXAM, R]
lightbox: true
toc: false
format-links: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
# format:
#   html: default
#   hugo-md:
#     echo: false
#     html-math-method: mathjax
#     output-file: e1-hugo.md
  # gfm:
  #   echo: true
  #   output-file: e1-gfm.md
prefer-html: true
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
               stringr,future,furrr, knitr, reactable, flextable,ggstance, htmltools,ggdist,ggh4x,brms,tidybayes,bayestestR, gt)
#conflict_prefer_all("dplyr", quiet = TRUE)
walk(c("flextable","dplyr"), conflict_prefer_all, quiet = TRUE)
#options(brms.backend="cmdstanr",mc.cores=4)
options(digits=3, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model", "prep_model_data","org_functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))

```

```{r}
#| cache: false

invisible(list2env(load_sbj_data(), envir = .GlobalEnv))
invisible(list2env(load_e1(), envir = .GlobalEnv))

e1Sbjs <- e1 |> group_by(id,condit) |> summarise(n=n())
```

# Modeling Approach

In project 1, I applied model-based techniques to quantify and control for the similarity between training and testing experience, which in turn enabled us to account for the difference between varied and constant training via an extended version of a similarity based generalization model. In project 2, I will go a step further, implementing a full process model capable of both 1) producing novel responses and 2) modeling behavior in both the learning and testing stages of the experiment. For this purpose, we will apply the associative learning model (ALM) and the EXAM model of function learning [@deloshExtrapolationSineQua1997]. ALM is a simple connectionist learning model which closely resembles Kruschke's ALCOVE model [@kruschkeALCOVEExemplarbasedConnectionist1992], with modifications to allow for the generation of continuous responses.

## ALM & Exam Description

ALM is a localist neural network model [@pageConnectionistModellingPsychology2000a], with each input node corresponding to a particular stimulus, and each output node corresponding to a particular response value. The units in the input layer activate as a function of their Gaussian similarity to the input stimulus. So, for example, an input stimulus of value 55 would induce maximal activation of the input unit tuned to 55. Depending on the value of the generalization parameter, the nearby units (e.g. 54 and 56; 53 and 57) may also activate to some degree. ALM is structured with input and output nodes that correspond to regions of the stimulus space, and response space, respectively. The units in the input layer activate as a function of their similarity to a presented stimulus. As was the case with the exemplar-based models, similarity in ALM is exponentially decaying function of distance. The input layer is fully connected to the output layer, and the activation for any particular output node is simply the weighted sum of the connection weights between that node and the input activations. The network then produces a response by taking the weighted average of the output units (recall that each output unit has a value corresponding to a particular response). During training, the network receives feedback which activates each output unit as a function of its distance from the ideal level of activation necessary to produce the correct response. The connection weights between input and output units are then updated via the standard delta learning rule, where the magnitude of weight changes are controlled by a learning rate parameter. The EXAM model is an extension of ALM, with the same learning rule and representational scheme for input and output units. The primary difference is that EXAM includes a linear extrapolation mechanism for generating novel responses during testing, a modification necessary to account for human extrapolation patterns in past research [@deloshExtrapolationSineQua1997 @mcdanielPredictingTransferPerformance2009 @brownUnderestimationLinearFunction2017]. Although this extrapolation rule departs from a strictly similarity-based generalization mechanism, EXAM is distinct from pure rule-based models in that it remains constrained by the weights learned during training.

See @tbl-alm-exam for a full specification of the equations that define ALM and EXAM.

::: column-page-inset-right
|                    | **ALM Response Generation**                                        |                                                                                               |
|------------------|-----------------------------|-------------------------|
| Input Activation   | $a_i(X) = \frac{e^{-c(X-X_i)^2}}{\sum_{k=1}^M e^{-c(X-X_k)^2}}$    | Input nodes activate as a function of Gaussian similarity to stimulus                         |
| Output Activation  | $O_j(X) = \sum_{k=1}^M w_{ji} \cdot a_i(X)$                        | Output unit $O_j$ activation is the weighted sum of input activations and association weights |
| Output Probability | $P[Y_j|X] = \frac{O_j(X)}{\sum_{k=1}^M O_k(X)}$                    | The response, $Y_j$ probabilites computed via Luce's choice rule                              |
| Mean Output        | $m(X) = \sum_{j=1}^L Y_j \cdot \frac{O_j(x)}{\sum_{k=1}^M O_k(X)}$ | Weighted average of probabilities determines response to X                                    |
|                    | **ALM Learning**                                                   |                                                                                               |
| Feedback           | $f_j(Z) = e^{-c(Z-Y_j)^2}$                                         | feedback signal Z computed as similarity between ideal response and observed response         |
| magnitude of error | $\Delta_{ji}=(f_{j}(Z)-o_{j}(X))a_{i}(X)$                          | Delta rule to update weights.                                                                 |
| Update Weights     | $w_{ji}^{new}=w_{ji}+\eta\Delta_{ji}$                              | Updates scaled by learning rate parameter $\eta$.                                             |
|                    | **EXAM Extrapolation**                                             |                                                                                               |
| Instance Retrieval | $P[X_i|X] = \frac{a_i(X)}{\sum_{k=1}^M a_k(X)}$                    | Novel test stimulus $X$ activates input nodes $X_i$                                           |
| Slope Computation  | $S =$ $\frac{m(X_{1})-m(X_{2})}{X_{1}-X_{2}}$                      | Slope value, $S$ computed from nearest training instances                                     |
| Response           | $E[Y|X_i] = m(X_i) + S \cdot [X - X_i]$                            | ALM response $m(X_i)$ adjusted by slope.                                                      |

: ALM & EXAM Equations {#tbl-alm-exam}
:::

## Model Fitting

To fit ALM and EXAM to our participant data, we employ a similar method to @mcdanielPredictingTransferPerformance2009, wherein we examine the performance of each model after being fit to various subsets of the data. Each model was fit to the data in with separate procedures: 1) fit to maximize predictions of the testing data, 2) fit to maximize predictions of both the training and testing data, 3) fit to maximize predictions of the just the training data. We refer to this fitting manipulations as "Fit Method" in the tables and figures below. It should be emphasized that for all three fit methods, the ALM and EXAM models behave identically - with weights updating only during the training phase.Models to were fit separately to the data of each individual participant. The free parameters for both models are the generalization ($c$) and learning rate ($lr$) parameters. Parameter estimation was performed using approximate bayesian computation (ABC), which we describe in detail below.

::: {.callout}
**{{< fa regular lightbulb >}} Approximate Bayesian Computation**

To estimate parameters, we used approximate bayesian computation (ABC), enabling us to obtain an estimate of the posterior distribution of the generalization and learning rate parameters for each individual. ABC belongs to the class of simulation-based inference methods [@cranmerFrontierSimulationbasedInference2020], which have begun being used for parameter estimation in cognitive modeling relatively recently [@turnerTutorialApproximateBayesian2012; @turnerBayesianAnalysisSimulationbased2016; @kangasraasioParameterInferenceComputational2019]. Although they can be applied to any model from which data can be simulated, ABC methods are most useful for complex models that lack an explicit likelihood function (e.g. many neural network and evidence accumulation models).

The general ABC procedure is to 1) define a prior distribution over model parameters. 2) sample candidate parameter values, $\theta^*$, from the prior. 3) Use $\theta^*$ to generate a simulated dataset, $Data_{sim}$. 4) Compute a measure of discrepancy between the simulated and observed datasets, $discrep$($Data_{sim}$, $Data_{obs}$). 5) Accept $\theta^*$ if the discrepancy is less than the tolerance threshold, $\epsilon$, otherwise reject $\theta^*$. 6) Repeat until desired number of posterior samples are obtained.

Although simple in the abstract, implementations of ABC require researchers to make a number of non-trivial decisions as to i) the discrepancy function between observed and simulated data, ii) whether to compute the discrepancy between trial level data, or a summary statistic of the datasets, iii) the value of the minimum tolerance $\epsilon$ between simulated and observed data. For the present work, we follow the guidelines from previously published ABC tutorials [@turnerTutorialApproximateBayesian2012; @farrellComputationalModelingCognition2018]. For the test stage, we summarized datasets with mean velocity of each band in the observed dataset as $V_{obs}^{(k)}$ and in the simulated dataset as $V_{sim}^{(k)}$, where $k$ represents each of the six velocity bands. For computing the discrepancy between datasets in the training stage, we aggregated training trials into three equally sized blocks (separately for each velocity band in the case of the varied group). After obtaining the summary statistics of the simulated and observed datasets, the discrepancy was computed as the mean of the absolute difference between simulated and observed datasets (@eq-discrep-test and @eq-discrep-train). For the models fit to both training and testing data, discrepancies were computed for both stages, and then averaged together.

::: column-page-inset-left
$$
discrep_{Test}(Data_{sim}, Data_{obs}) = \frac{1}{6} \sum_{k=1}^{6} |V_{obs}^{(k)} - V_{sim}^{(k)}|
$$ {#eq-discrep-test}

$$
\begin{aligned} \\
discrep_{Train,constant}(Data_{sim}, Data_{obs}) = \frac{1}{N_{blocks}} \sum_{j=1}^{N_{blocks}} |V_{obs,constant}^{(j)} - V_{sim,constant}^{(j)}| \\ \\
discrep_{Train,varied}(Data_{sim}, Data_{obs}) = \frac{1}{N_{blocks} \times 3} \sum_{j=1}^{N_{blocks}} \sum_{k=1}^{3} |V_{obs,varied}^{(j,k)} - V_{sim,varied}^{(j,k)}|
\end{aligned}
$$ {#eq-discrep-train}
:::

The final component of our ABC implementation is the determination of an appropriate value of $\epsilon$. The setting of $\epsilon$ exerts strong influence on the approximated posterior distribution. Smaller values of $\epsilon$ increase the rejection rate, and improve the fidelity of the approximated posterior, while larger values result in an ABC sampler that simply reproduces the prior distribution. Because the individual participants in our dataset differed substantially in terms of the noisiness of their data, we employed an adaptive tolerance setting strategy to tailor $\epsilon$ to each individual. The initial value of $\epsilon$ was set to the overall standard deviation of each individuals velocity values. Thus, sampled parameter values that generated simulated data within a standard deviation of the observed data were accepted, while worse performing parameters were rejected. After every 300 samples the tolerance was allowed to increase only if the current acceptance rate of the algorithm was less than 1%. In such cases, the tolerance was shifted towards the average discrepancy of the 5 best samples obtained thus far. To ensure the acceptance rate did not become overly permissive, $\epsilon$ was also allowed to decrease every time a sample was accepted into the posterior.
:::

For each of the 156 participants from Experiment 1, the ABC algorithm was run until 200 samples of parameters were accepted into the posterior distribution. Obtaining this number of posterior samples required an average of 205,000 simulation runs per participant. Fitting each combination of participant, Model (EXAM & ALM), and fitting method (Test only, Train only, Test & Train) required a total of 192 million simulation runs. To facilitate these intensive computational demands, we used the Future Package in R [@bengtssonUnifyingFrameworkParallel2021], allowing us to parallelize computations across a cluster of ten M1 iMacs, each with 8 cores.

### Modelling Results

#### Group level Patterns

```{r}
#| label: tbl-htw-modelError-e1
#| tbl-cap: "Models errors predicting empirical data -  aggregated over all participants, posterior parameter values,  and velocity bands. Note that Fit Method refers to the subset of the data that the model was trained on, while Task Stage refers to the subset of the data that the model was evaluated on."
post_tabs <- abc_tables(post_dat,post_dat_l)
train_tab <- abc_train_tables(pd_train,pd_train_l)

e1_tab <- rbind(post_tabs$agg_pred_full |> mutate("Task Stage"="Test"), train_tab$agg_pred_full |> mutate("Task Stage"="Train")) |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) 
  # |>
  # flextable::tabulator(rows=c("stage","Fit_Method","Model"), columns=c("condit"),
  #                      `ME` = as_paragraph(mean_error)) |> as_flextable()

e1_tab %>%
  group_by(`Task Stage`, Fit_Method, Model, condit) %>%
  summarize(ME = mean(mean_error), .groups = "drop") %>%
  pivot_wider(
    names_from = c(Model, condit),
    values_from = ME,
    names_sep = "_"  # Add this line to specify the separator for column names
  ) %>%
  rename("Fit Method" = Fit_Method) %>%
  gt() %>%
  # tab_header(
  #   title = "Mean Error by Fit Method, Task Stage, and Condition",
  #   subtitle = "Comparison across Models"
  # ) %>%
  cols_move_to_start(columns = c(`Task Stage`)) %>%
  cols_label(
    `Task Stage` = "Task Stage"
  ) %>%
  fmt_number(
    columns = starts_with("ALM") | starts_with("EXAM"),
    decimals = 2
  ) %>%
  tab_spanner_delim(delim = "_") %>%
  tab_style(
    style = cell_fill(color = "white"),
     locations = cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_style(
    style = cell_borders(sides = "top", color = "black", weight = px(1)),
    locations = cells_column_labels()
  ) %>%
  tab_options(
    column_labels.font.size = 10,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.font.size = 10, 
    quarto.disable_processing = TRUE
  ) #%>%
  # tab_footnote(
  #   footnote = "Note: Mean errors are calculated as the absolute difference between the predicted and actual values.",
  #   locations = cells_title(groups = "subtitle")
  # ) %>%
  # sub_missing(missing_text = "")

#  post_dat  |> group_by(condit,Model,Fit_Method,x) |> 
#     mutate(e2=abs(dist-pred_dist)) |> 
#     summarise(dist=mean(dist), pred=mean(pred_dist), mean_error=mean(e2)) |>
#     group_by(condit,Model,Fit_Method) |> 
#     summarise(mean_error=mean(mean_error)) |> 
#     round_tibble(1) |> 
#   mutate(Fit_Method=rename_fm(Fit_Method)) |>
#   flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit"),
#                        `ME` = as_paragraph(mean_error)) |> as_flextable()

# post_tabs$agg_pred_full_bt |> 
#    pander::pandoc.table(style="rmarkdown", split.table=Inf)
#   
# post_tabs$agg_pred_full |> 
#    pander::pandoc.table(style="rmarkdown", split.table=Inf)
# 
# post_tabs$agg_x_full_bt |> filter((Fit_Method=="Test_Train")) |>
#   mutate(Fit_Method=(Fit_Method)) |>
#   rename("Band"=x, "Condition"=condit, "Model Error"=mean_error) |>
#   pander::pandoc.table(style="rmarkdown", split.table=Inf)

#  post_dat |> group_by(id,condit,Model,Fit_Method) |> summarise(y=mean(y), pred=mean(pred), ae=abs(y-pred)) |>
#     group_by(id,condit,Model,Fit_Method) |> 
#     summarise(mean_error=mean(ae)) |> 
# ggplot(aes(x=mean_error,fill=condit,y=interaction(Model,Fit_Method))) + stat_halfeye()
# 
#  post_dat |> group_by(id,condit,Model,Fit_Method,x) |> summarise(y=mean(y), pred=mean(pred), ae=abs(y-pred)) |>
#     group_by(id,condit,Model,Fit_Method,x) |> 
#     summarise(mean_error=mean(ae)) |> 
#   ggplot(aes(x=mean_error,fill=condit,y=interaction(Model,Fit_Method))) + 
#   stat_halfeye() +
#   facet_wrap(~x)

```

```{r}
#| label: fig-htw-post-dist
#| fig-cap: "Posterior Distributions of $c$ and $lr$ parameters. Points represent median values, thicker intervals represent 66% credible intervals and thin intervals represent 95% credible intervals around the median. Note that the y axes of the plots for the c parameter are scaled logarithmically."
#| fig-height: 6
#| fig-width: 10

c_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=log(c), x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.2)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="c parameter") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4))

lr_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=lr, x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.4)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="learning rate parameter") +
  theme(legend.title = element_blank(), legend.position = "none",plot.title=element_text(hjust=.5))
c_post + lr_post
```

```{r}
#| label: fig-htw-resid-pred
#| fig-cap: "Model residuals for each combination of training condition, fit method, and model. Residuals reflect the difference between observed and predicted values. Lower values indicate better model fit. Note that y axes are scaled differently between facets. A) Residuals predicting each block of the training data. B) Residuals predicting each band during the testing stage. Bolded bars indicate bands that were trained, non-bold bars indicate extrapolation bands."
#| fig-height: 14
#| fig-width: 11
##| layout: [[45,-5, 45], [100]]
##| fig-subcap: ["Model Residuals - training data", "Model Residuals - testing data","Full posterior predictive distributions vs. observed data from participants."]
train_resid <- pd_train |> group_by(id,condit,Model,Fit_Method, Block,x) |> 
  summarise(y=mean(y), pred=mean(pred), mean_error=abs(y-pred)) |>
  group_by(id,condit,Model,Fit_Method,Block) |>
  summarise(mean_error=mean(mean_error)) |>
  ggplot(aes(x=interaction(Block,Model), y = mean_error, fill=factor(Block))) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, scales="free",ncol=2) +
   scale_x_discrete(guide = "axis_nested") +
  scale_fill_manual(values=c("gray10","gray50","gray92"))+
  labs(title="Model Residual Errors - Training Stage", y="RMSE", x= "Model",fill="Training Block") +
  theme(legend.position="top")

test_resid <-  post_dat |> 
   group_by(id,condit,x,Model,Fit_Method,bandType) |>
    summarise(y=mean(y), pred=mean(pred), error=abs(y-pred)) |> 
  mutate(vbLab = factor(paste0(x,"-",x+200))) |>
  ggplot(aes(x = Model, y = abs(error), fill=vbLab,col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
  stat_bar + 
  #scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  scale_color_manual(values = c("black" = "black"), guide = "none") +
  scale_size_manual(values = c("black" = .5), guide = "none") +
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, axes = "all",ncol=2,scale="free") +
  labs(title="Model Residual Errors - Testing Stage",y="RMSE", x="Velocity Band")

(train_resid / test_resid) +
  #plot_layout(heights=c(1,1.5)) & 
  plot_annotation(tag_levels = list(c('A','B')),tag_suffix = ') ') 
```

The posterior distributions of the $c$ and $lr$ parameters are shown @fig-htw-post-dist, and model predictions are shown alongside the empirical data in @fig-cm-vx-pat. There were substantial individual differences in the posteriors of both parameters, with the within-group individual differences generally swamped any between-group or between-model differences. The magnitude of these individual differences remains even if we consider only the single best parameter set for each subject.

We used the posterior distribution of $c$ and $lr$ parameters to generate a posterior predictive distribution of the observed data for each participant, which then allows us to compare the empirical data to the full range of predictions from each model. Aggregated residuals are displayed in @fig-htw-resid-pred. The pattern of training stage residual errors are unsurprising across the combinations of models and fitting method . Differences in training performance between ALM and EXAM  are generally minor (the two models have identical learning mechanisms). The differences in the magnitude of residuals across the three fitting methods are also straightforward, with massive errors for the 'fit to Test Only' model, and the smallest errors for the 'fit to train only' models. It is also noteworthy that the residual errors are generally larger for the first block of training, which is likely due to the initial values of the ALM weights being unconstrained by whatever initial biases participants tend to bring to the task. Future work may explore the ability of the models to capture more fine grained aspects of the learning trajectories. However for the present purposes, our primary interest is in the ability of ALM and EXAM to account for the testing patterns while being constrained, or not constrained, by the training data. All subsequent analyses and discussion will thus focus on the testing stage.

The residuals of the model predictions for the testing stage (@fig-htw-resid-pred) also show an unsurprising pattern across fitting methods - with models fit only to the test data showing the best performance, followed by models fit to both training and test data, and with models fit only to the training data showing the worst performance (note that y axes are scaled different between plots). Although EXAM tends to perform better for both Constant and Varied participants (see also @fig-ee-e1), the relative advantage of EXAM is generally larger for the Constant group - a pattern consistent across all three fitting methods. The primary predictive difference between ALM and EXAM is made clear in @fig-cm-vx-pat, which directly compares the observed data against the posterior predictive distributions for both models. Regardless of how the models are fit, only EXAM can capture the pattern where participants are able to discriminate all 6 target bands.

-   \*\* does EXAM do better for the Constant group because the constant group performs better? Or does training with a single example encourage an exam sort of strategy?

```{r}
#| label: fig-cm-vx-pat
#| fig-cap: "Empirical data and Model predictions for mean velocity across target bands. Fitting methods (Test Only, Test & Train, Train Only) - are separated across rows, and Training Condition (Constant vs. Varied) are separated by columns. Each facet contains the predictions of ALM and EXAM, alongside the observed data."
#| fig-height: 10
#| fig-width: 10

post_dat_l |> 
  group_by(id,condit, Fit_Method,Resp,bandType,x,vb) |> 
 summarize(vx=median(val)) |> 
 #left_join(testAvgE1, by=join_by(id,condit,x==bandInt)) |>
 ggplot( aes(x=Resp,y=vx, fill=vb,col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
  stat_bar + 
    facet_wrap(~rename_fm(Fit_Method)+condit, ncol=2,strip.position = "top", scales = "free_x") +
        scale_color_manual(values = c("black" = "black"), guide = "none") +
  scale_size_manual(values = c("black" = .5), guide = "none") +
    theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside",
         legend.position = "none",plot.title = element_text(hjust=.50),
         axis.title.x = element_blank(),
         plot.margin = unit(c(10,0,0,0), "pt")) + 
         labs(title="Model Predictions - Experiment 1 Data", y="Vx")
```

```{r}
#| label: fig-ee-e1
#| fig-caption: conditional and marginal effects
#| fig-height: 9
#| fig-width: 10

# library(brms)
# options(brms.backend="cmdstanr",mc.cores=4)

pdl <- post_dat_l |> rename("bandInt"=x) |> left_join(testAvgE1,by=c("id","condit","bandInt")) |> 
  filter(rank<=1,Fit_Method=="Test_Train", !(Resp=="Observed")) |> mutate(aerror = abs(error))

# aerror is model error, which is predicted by Model(ALM vs. EXAM) & condit (Constant vs. Varied)
e1_ee_brm_ae <- brm(data=pdl,
  aerror ~  Model * condit + (1+bandInt|id), 
  file = paste0(here("data/model_cache/e1_ae_modelCond_RFint.rds")),
  chains=4,silent=1, iter=2000, control=list(adapt_delta=0.92, max_treedepth=11))

bct_e1 <- as.data.frame(bayestestR::describe_posterior(e1_ee_brm_ae, centrality = "Mean")) %>%
  select(1,2,4,5,6) %>%
  setNames(c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")) %>%
  mutate(across(where(is.numeric), \(x) round(x, 2))) %>%
  tibble::remove_rownames() %>%
  mutate(Term = stringr::str_remove(Term, "b_")) %>%
  kable(booktabs = TRUE)

#wrap_plots(plot(conditional_effects(e1_ee_brm_ae),points=FALSE,plot=FALSE))

p1 <- plot(conditional_effects(e1_ee_brm_ae, effects="condit"),points=FALSE, plot=FALSE)$condit + 
  ggplot2::xlab("Condition") +ylab("Model Error")
p2 <- plot(conditional_effects(e1_ee_brm_ae, effects="Model"),points=FALSE, plot=FALSE)$Model + 
  labs(x="Model",y=NULL)
p3 <- plot(conditional_effects(e1_ee_brm_ae, effects="Model:condit"),points=FALSE, plot=FALSE)$`Model:condit` + 
  scale_color_manual(values=wes_palette("Darjeeling1")) +
  labs(x="Model",y=NULL,fill=NULL,col=NULL) + theme(legend.position="right")

p_ce_1 <- (p1 + p2+ p3) + plot_annotation(tag_levels = c('A'), tag_suffix=".")

bm1 <- get_coef_details(e1_ee_brm_ae, "conditVaried")
bm2 <- get_coef_details(e1_ee_brm_ae, "ModelEXAM")
bm3 <- get_coef_details(e1_ee_brm_ae, "ModelEXAM:conditVaried")

posterior_estimates <- as.data.frame(e1_ee_brm_ae) %>%
  select(starts_with("b_")) %>%
  setNames(c("Intercept", "ModelEXAM", "conditVaried", "ModelEXAM_conditVaried"))

library(tidybayes)
library(emmeans)

constant_EXAM <- posterior_estimates$Intercept + posterior_estimates$ModelEXAM
varied_EXAM <- posterior_estimates$Intercept + posterior_estimates$ModelEXAM + posterior_estimates$conditVaried + posterior_estimates$ModelEXAM_conditVaried
comparison_EXAM <- constant_EXAM - varied_EXAM
summary_EXAM <- bayestestR::describe_posterior(comparison_EXAM, centrality = "Mean")

# e1_ee_brm_ae |> emmeans(pairwise ~ Model * condit, re_formula=NULL)
# e1_ee_brm_ae |> emmeans(pairwise ~ Model * condit, re_formula=NA)

  # full set of Model x condit contrasts

# ALM - EXAM
btw_model <- e1_ee_brm_ae |> emmeans(pairwise~ Model | condit, re_formula=NULL)  |> 
  pluck("contrasts") |> 
  gather_emmeans_draws() |> 
  group_by(contrast,.draw,condit) |> summarise(value=mean(.value), n=n())

# btw_model |> ggplot(aes(x=value,y=contrast,fill=condit)) +stat_halfeye()

# Constant - Varied
emm_condit <- e1_ee_brm_ae |> emmeans(~ condit | Model, re_formula = NULL)
btw_con <- emm_condit |>  pairs() |> gather_emmeans_draws() |> 
  group_by(contrast,.draw, Model) |> summarise(value=mean(.value), n=n()) 
# btw_con |> ggplot(aes(x=value,y=Model,fill=Model)) +stat_halfeye()

p_em_1<- e1_ee_brm_ae |> emmeans(pairwise~ Model*condit, re_formula=NA)  |> 
  pluck("contrasts") |>
  gather_emmeans_draws() |> 
  group_by(contrast,.draw) |> summarise(value=mean(.value), n=n()) |> 
  filter(!(contrast %in% c("ALM Constant - EXAM Constant","ALM Constant - EXAM Varied","ALM Varied - EXAM Varied ", "EXAM Constant - ALM Varied" ))) |> 
  ggplot(aes(x=value,y=contrast,fill=contrast)) +stat_halfeye() + labs(x="Model Error Difference",y="Contrast") + theme(legend.position="none")

p_ce_1 / p_em_1

```

To quantitatively assess whether the differences in performance between models,  we fit a bayesian regressions predicting the errors of the posterior predictions of each models as a function of the Model (ALM vs. EXAM) and training condition (Constant vs. Varied).

Model errors were significantly lower for EXAM  ($\beta$ = `r bm1$estimate`, 95% CrI \[`r bm1$conf.low`, `r bm1$conf.high`\], pd = `r bm1$pd`)  than ALM.  There was also a significant interaction between Model and Condition ($\beta$ = `r bm3$estimate`, 95% CrI \[`r bm3$conf.low`, `r bm3$conf.high`\], pd = `r bm3$pd`), indicating that the advantage of EXAM over ALM was significantly greater for the constant group. To assess whether EXAM predicts constant performance significantly better for Constant than for Varied subjects, we calculated the difference in model error between the Constant and Varied conditions specifically for EXAM. The results indicated that the model error for EXAM was significantly lower in the Constant condition compared to the Varied condition, with a mean difference of `r summary_EXAM$Mean` (95% CrI \[`r summary_EXAM$CI_low`, `r summary_EXAM$CI_high`\], pd = `r summary_EXAM$pd`).

```{r}
#| label: tbl-htw-modelError-e23
#| tbl-cap: "Models errors predicting empirical data -  aggregated over all participants, posterior parameter values,  and velocity bands. Note that Fit Method refers to the subset of the data that the model was trained on, while Task Stage refers to the subset of the data that the model was evaluated on."

e2_model <- load_e2()
post_tabs2 <- abc_tables(e2_model$post_dat,e2_model$post_dat_l)
train_tab2 <- abc_train_tables(e2_model$pd_train,e2_model$pd_train_l)

pdl2 <- e2_model$post_dat_l |> rename("bandInt"=x) |> filter(rank<=1,Fit_Method=="Test_Train", !(Resp=="Observed")) |> mutate(aerror = abs(error))

e2_tab <- rbind(post_tabs2$agg_pred_full |>
 mutate("Task Stage"="Test"), train_tab2$agg_pred_full |> 
 mutate("Task Stage"="Train")) |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) 
  # |>
  # flextable::tabulator(rows=c("Task Stage","Fit_Method","Model"), columns=c("condit"),
  #                      `ME` = as_paragraph(mean_error)) |> as_flextable()
e3_model <- load_e3()
post_tabs3 <- abc_tables(e3_model$post_dat,e3_model$post_dat_l)
train_tab3 <- abc_train_tables(e3_model$pd_train,e3_model$pd_train_l)

pdl3 <- e3_model$post_dat_l |> rename("bandInt"=x) |> filter(rank<=1,Fit_Method=="Test_Train", !(Resp=="Observed")) |> mutate(aerror = abs(error))

e3_tab <- rbind(post_tabs3$agg_pred_full |> 
  mutate("Task Stage"="Test"), train_tab3$agg_pred_full |> mutate("Task Stage"="Train")) |> 
  mutate(Fit_Method=rename_fm(Fit_Method)) 
  # |>
  # flextable::tabulator(rows=c("Task Stage","Fit_Method","Model"), columns=c("condit"),
  #                      `ME` = as_paragraph(mean_error)) |> as_flextable()

e23_tab <- rbind(e2_tab |> mutate(Exp="E2"), e3_tab |> mutate(Exp="E3")) 
gt_table <- e23_tab %>%
  pivot_wider(
    names_from = c(Exp, Model, condit),
    values_from = mean_error,
    names_glue = "{Exp}_{Model}_{condit}"
  ) %>%
  arrange(Fit_Method, `Task Stage`) %>%
  gt() %>%
  # tab_header(
  #   title = "Mean Error by Fit Method, Task Stage, and Experiment",
  #   subtitle = "Comparison across Conditions and Models"
  # ) %>%
  cols_move_to_start(columns = `Task Stage`) %>%
  cols_label(`Task Stage` = "Task Stage") %>%
  fmt_number(columns = matches("E2|E3"), decimals = 1) %>%
  tab_spanner_delim(delim = "_") %>%
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_borders(sides = "top", color = "black", weight = px(1))
    ),
    locations = cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_options(
    column_labels.font.size = 10,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    table.font.size = 10,
    quarto.disable_processing = TRUE
  ) #%>%
  # tab_footnote(
  #   footnote = "Note: Mean errors are calculated as the absolute difference between the predicted and actual values.",
  #   locations = cells_title(groups = "subtitle")
  # )
gt_table
```

```{r}
#| label: tbl-htw-ee-e23
#| tbl-cap: "Results of Bayesian Regression models predicting model error as a function of Model (ALM vs. EXAM), Condition (Constant vs. Varied), and the interaction between Model and Condition. The values represent the estimate coefficient for each term, with 95% credible intervals in brackets.  The intercept reflects the baseline of ALM and Constant. The other estimates indicate deviations from the baseline for the EXAM mode and varied condition. Lower values indicate better model fit."

e2_ee_brm_ae <- brm(data=pdl2,
  aerror ~  Model * condit + (1+bandInt|id), 
  file = paste0(here("data/model_cache/e2_ae_modelCond_RFint.rds")),
  chains=4,silent=1, iter=2000, control=list(adapt_delta=0.92, max_treedepth=11))

bm1_e2 <- get_coef_details(e2_ee_brm_ae, "conditVaried")
bm2_e2 <- get_coef_details(e2_ee_brm_ae, "ModelEXAM")
bm3_e2 <- get_coef_details(e2_ee_brm_ae, "ModelEXAM:conditVaried")
# bct_e2 <- as.data.frame(bayestestR::describe_posterior(e2_ee_brm_ae, centrality = "Mean")) %>%
#   select(1,2,4,5,6) %>%
#   setNames(c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")) %>%
#   mutate(across(where(is.numeric), \(x) round(x, 2))) %>%
#   tibble::remove_rownames() %>%
#   mutate(Term = stringr::str_remove(Term, "b_")) %>%
#   kable(booktabs = TRUE)

e3_ee_brm_ae <- brm(data=pdl3,
  aerror ~  Model * condit*bandOrder + (1+bandInt|id), 
  file = paste0(here("data/model_cache/e3_ae_modelCondBo_RFint2.rds")),
  chains=4,silent=1, iter=2000, control=list(adapt_delta=0.92, max_treedepth=11))

bm1_e3 <- get_coef_details(e3_ee_brm_ae, "conditVaried")
bm2_e3  <- get_coef_details(e3_ee_brm_ae, "ModelEXAM")
bm3_e3  <- get_coef_details(e3_ee_brm_ae, "ModelEXAM:conditVaried")

# bct_e3  <- as.data.frame(bayestestR::describe_posterior(e3_ee_brm_ae, centrality = "Mean")) %>%
#   select(1,2,4,5,6) %>%
#   setNames(c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")) %>%
#   mutate(across(where(is.numeric), \(x) round(x, 2))) %>%
#   tibble::remove_rownames() %>%
#   mutate(Term = stringr::str_remove(Term, "b_")) %>%
#   kable(booktabs = TRUE)

msa <- modelsummary::modelsummary(list("Exp 1."=e1_ee_brm_ae,"Exp 2."=e2_ee_brm_ae,"Exp 3. "=e3_ee_brm_ae), 
    gof_map = NA,
    fmt=1,
    estimate = "{estimate}",
    statistic="[{conf.low}, {conf.high}]", 
    coef_omit = "^sigma|^sd|^cor",
    coef_rename=function(x) stringr::str_remove(x, "b_"), output="gt")

msa

#library(modelsummary)
# modelsummary::modelsummary(get_estimates(e1_ee_brm_ae, test = c("pd")),gof_map = NA)
#modelsummary::get_estimates(e1_ee_brm_ae,test="pd")
```

```{r}
#| label: fig-e2_e3_ae
#| fig-cap: "Conditional effects of Model and Condition on Model Error for Experiment 2 and 3 data."
#| 
#| fig-height: 5
#| fig-width: 9

#wrap_plots(plot(conditional_effects(e1_ee_brm_ae),points=FALSE,plot=FALSE))
p1 <- plot(conditional_effects(e2_ee_brm_ae, effects="condit"),points=FALSE, plot=FALSE)$condit + 
  ggplot2::xlab("Condition") +ylab("Model Error") + labs(title="E2. Model Error")
p2 <- plot(conditional_effects(e2_ee_brm_ae, effects="Model"),points=FALSE, plot=FALSE)$Model + 
  labs(x="Model",y=NULL)
p3 <- plot(conditional_effects(e2_ee_brm_ae, effects="Model:condit"),points=FALSE, plot=FALSE)$`Model:condit` + 
  scale_color_manual(values=wes_palette("Darjeeling1")) +
  labs(x="Model",y=NULL,fill=NULL,col=NULL) + theme(legend.position="right")

 p_e2 <- (p1 + p2+ p3)

#wrap_plots(plot(conditional_effects(e3_ee_brm_ae),points=FALSE,plot=FALSE))

p_e3 <- plot(conditional_effects(e3_ee_brm_ae, 
                         effects = "Model:condit", 
                         conditions=make_conditions(e3_ee_brm_ae,vars=c("bandOrder"))),
     points=FALSE,plot=FALSE)$`Model:condit` + 
     labs(x="Model",y="Model Error", title="E3. Model Error") + 
     theme(legend.position="right") + 
     scale_color_manual(values=wes_palette("Darjeeling1"))

p1 <- plot(conditional_effects(e3_ee_brm_ae, effects="condit"),points=FALSE, plot=FALSE)$condit + 
  ggplot2::xlab("Condition") +ylab("Model Error")
p2 <- plot(conditional_effects(e3_ee_brm_ae, effects="Model"),points=FALSE, plot=FALSE)$Model + 
  labs(x="Model",y=NULL)
p3 <- plot(conditional_effects(e3_ee_brm_ae, effects="Model:condit"),points=FALSE, plot=FALSE)$`Model:condit` + 
  scale_color_manual(values=wes_palette("Darjeeling1")) +
  labs(x="Model",y=NULL,fill=NULL,col=NULL) + theme(legend.position="right")

 #p2 <- (p1 + p2+ p3)
 (p_e2 / p_e3) + plot_annotation(tag_levels = c('A'), tag_suffix=".")

```

```{r}
#| fig-height: 8
#| fig-width: 10
#| fig-cap: "Empirical data and Model predictions from Experiment 2 and 3 for the testing stage. Observed data is shown on the right. Bolded bars indicate bands that were trained, non-bold bars indicate extrapolation bands.  "

rbind(e2_model$post_dat_l |> filter( Fit_Method=="Test_Train") |> 
  group_by(id,condit, Fit_Method,Resp,bandType,x,vb) |> 
 summarize(vx=median(val)) |> mutate(Exp="E2",bandOrder="Reverse"), 
 e3_model$post_dat_l |> filter( Fit_Method=="Test_Train") |> 
  group_by(id,condit, Fit_Method,Resp,bandType,x,vb,bandOrder) |>
  summarize(vx=median(val)) |> mutate(Exp="E3")) |>
  ggplot( aes(x=condit,y=vx, fill=vb,col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) +
  stat_bar + 
    facet_nested_wrap(~Exp+bandOrder+Resp, strip.position = "top", scales = "free_x") +
    scale_color_manual(values = c("black" = "black"), guide = "none") +
  scale_size_manual(values = c("black" = .7), guide = "none") +
    theme(panel.spacing = unit(0, "lines"), 
        #  strip.background = element_blank(),
        #  strip.placement = "outside",
         legend.position = "none",plot.title = element_text(hjust=.50),
         axis.title.x = element_blank(),
         plot.margin = unit(c(20,0,0,0), "pt")) + 
         labs(title="Model Predictions Experiment 2 & 3", y="vx")

```

```{r}
#| label: fig-htw-best-model
#| fig-cap: Difference in model errors for each participant, with models fit to both train and test data. Positive values favor EXAM, while negative values favor ALM. 
#| fig-height: 12
#| fig-width: 11

tid1 <- post_dat  |> mutate(Exp="E1",bandOrder="Original") |> select(-pred_dist, -dist) |>
  rbind(e2_model$post_dat |> mutate(Exp="E2",bandOrder="Reverse")) |>
  rbind(e3_model$post_dat |> mutate(Exp="E3")) |>
  filter(Fit_Method=="Test_Train") |>
  group_by(id,condit,Model,Fit_Method,x, Exp) |> 
    mutate(e2=abs(y-pred)) |> 
    summarise(y1=median(y), pred1=median(pred),mean_error=abs(y1-pred1)) |>
    group_by(id,condit,Model,Fit_Method,Exp) |> 
    summarise(mean_error=mean(mean_error)) |> 
    arrange(id,condit,Fit_Method) |>
    round_tibble(1)

best_id <- tid1 |> 
  group_by(id,condit,Fit_Method) |> 
  mutate(best=ifelse(mean_error==min(mean_error),1,0))

lowest_error_model <- best_id %>%
  group_by(id, condit,Fit_Method, Exp) %>%
  summarise(Best_Model = Model[which.min(mean_error)],
            Lowest_error = min(mean_error),
            differential = min(mean_error) - max(mean_error)) %>%
  ungroup()

error_difference<- best_id %>%
  select(id, condit, Model,Fit_Method, mean_error) %>%
  pivot_wider(names_from = Model, values_from = c(mean_error)) %>%
  mutate(Error_difference = (ALM - EXAM))

full_comparison <- lowest_error_model |> 
  left_join(error_difference, by=c("id","condit","Fit_Method"))  |> 
  group_by(condit,Fit_Method,Best_Model) |> 
  mutate(nGrp=n(), model_rank = nGrp - rank(Error_difference) ) |> 
  arrange(Fit_Method,-Error_difference)

full_comparison |> 
  filter(Fit_Method=="Test_Train") |> 
  ungroup() |>
  mutate(id = reorder(id, Error_difference)) %>%
  ggplot(aes(y=id,x=Error_difference,fill=Best_Model))+
  geom_col() +
  #ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")+
  ggh4x::facet_nested_wrap(~condit+Exp,scales="free") + 
  labs(fill="Best Model",
  x="Mean Model Error Difference (ALM - EXAM)",
  y="Participant")

```

### Accounting for individual patterns

To more accurately assess the relative abilities of ALM and EXAM to capture important empirical patterns - we will now examine the predictions of both models for the subset of individual participants shown in @fig-htw-indv-pred. Panel A presents three varied and constant participants who demonstrated a reasonable degree of discrimination between the 6 velocity bands during testing.

-   \*\* comment on the different ways ALM can completely fail to mimic discrimination patterns (sbj. 35; sbj. 137),and on how it can sometimes partially succeed (sbj. 11; 14,74)

-   \*\* comment on how EXAM can somtimes mimic non-monotonic spacing between bands due to associative stregth from training (i.e. subject 47)

-   \*\* compare c values to slope parameters from the statistical models earlier in paper

```{r  fig.width=11, fig.height=10}
#| label: fig-htw-indv-pred
#| fig-cap: Model predictions alongside observed data for a subset of individual participants. A) 3 constant and 3 varied participants fit to both the test and training data. B) 3 constant and 3 varied subjects fit to only the trainign data. Bolded bars indicate bands that were trained, non-bold bars indicate extrapolation bands. 
#| fig-height: 11
#| fig-width: 12

cId_tr <- c(137, 181, 11)
vId_tr <- c(14, 193, 47)
cId_tt <- c(11, 93, 35)
vId_tt <- c(1,14,74)
cId_new <- c(175, 185, 134, 155, 68, 93, 74, 134)
# filter(id %in% (filter(bestTestEXAM,group_rank<=9, Fit_Method=="Test")

# e3 sbjs - 245, 219, 264, 280, 249, 274

testIndv <- post_dat_l |> filter(id %in% c(cId_tt,vId_tt,cId_new), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=vb, col=ifelse(bandType=="Trained","black",NA),size=ifelse(bandType=="Trained","black",NA))) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  scale_color_manual(values = c("black" = "black"), guide = FALSE) +
  scale_size_manual(values = c("black" = .5), guide = FALSE) + 
  labs(title="Individual Participant fits from Test & Train Fitting Method",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

testIndv

```