### 2024-03-28 Sections/Discussion.qmd
add description of 0 point

# General Discussion

Here is a proposed outline for the Discussion section of the manuscript, based on the information provided:

Discussion
I. Overview of the study's aims and main findings
   A. Investigating the influence of training variability on learning, generalization, and extrapolation in a visuomotor function learning task
   B. Summary of the main results from Experiments 1 and 2

II. Interpretation of the results
    A. Lower training performance in the varied conditions
       1. Consistency with prior research on the influence of training variability
       2. Explanation of the effect in the context of the current study design
    B. Impaired extrapolation accuracy in the varied conditions
       1. Significance of the interaction between training condition and band type
       2. Implications for the role of variability in promoting generalization and extrapolation
    C. Reduced discrimination between velocity bands in the varied conditions
       1. Interpretation of the slopes in the velocity band analysis
       2. Potential mechanisms underlying the reduced discrimination in varied training

III. Integration with existing literature
     A. Comparison of findings with previous research on variability in function learning tasks
        1. DeLosh et al. (1997) study on the influence of variability (density) on function learning
        2. van Dam & Ernst (2015) study on variability in a visuomotor learning task
     B. Implications for theories and models of function learning and extrapolation
        1. Rule-based models (e.g., polynomial hypothesis testing, log-polynomial adaptive regression)
        2. Similarity-based models (e.g., Associative Learning Model, Extrapolation-Association Model)

IV. Limitations and future directions
    A. Potential limitations of the study design and methodology
    B. Suggestions for future research to address these limitations and extend the findings
       1. Investigating the influence of variability across different function types and task domains
       2. Examining the role of feedback and training duration on the effects of variability

V. Conclusion
   A. Recap of the main findings and their implications for understanding the role of variability in function learning and extrapolation
   B. Emphasis on the need for further research to develop a comprehensive framework for optimizing training variability in educational and instructional settings

## Overall Summary of Experiments 1-3

Across three experiments, we investigated the impact of training variability on learning, transfer, and extrapolation in a visuomotor function learning task. In Experiment 1, participants in the varied training condition, who experienced a wider range of velocity bands during training, showed lower accuracy at the end of training compared to those in the constant training condition. Crucially, during the testing phase, the varied group exhibited significantly larger deviations from the target velocity bands, particularly for the extrapolation bands that were not encountered during training. The varied group also showed less discrimination between velocity bands, as evidenced by shallower slopes when predicting response velocity from target velocity band.

Experiment 2 extended these findings by reversing the order of the training and testing bands. Similar to Experiment 1, the varied group demonstrated poorer performance during both training and testing phases. However, unlike Experiment 1, the varied group did not show a significant difference in discrimination between bands compared to the constant group.

In Experiment 3, we introduced an ordinal feedback system during training, in contrast to the continuous feedback provided in the previous experiments. Participants were assigned to both an order condition (original or reverse) and a training condition (constant or varied). The varied condition showed larger deviations at the end of training, consistent with the previous experiments. Interestingly, there was a significant interaction between training condition and band order, with the varied condition showing greater accuracy in the reverse order condition. During testing, the varied group once again exhibited larger deviations, particularly for the extrapolation bands. The reverse order conditions showed smaller deviations compared to the original order conditions. Discrimination between velocity bands was poorer for the varied group in the original order condition, but not in the reverse order condition.

In summary, these experiments consistently demonstrated that training with greater variability led to lower accuracy during both training and testing phases, with the detrimental effect being more pronounced for extrapolation items. The varied training condition also tended to show less discrimination between velocity bands, although this effect was modulated by the order of the training and testing bands. These findings suggest that while variable training can hinder initial learning, it may have differential effects on transfer and extrapolation depending on the specific task structure and the range of stimuli encountered during training and testing.

## Modelling Interpretation
- why does Constant do better 
    - kind of task that permits for prior knowledge about 0
    - learning - end of training
- what does it suggest that the constant group was disproportionately well explained by the EXAM model?
-

### Comparison to DeLosh 1997 & van Dam & Ernst 2015
- DeLosh used way larger numbers of unique stimuli. 
- van Dam had more complex input dimension (spikiness). Which lacks any sense of 0 point.

### Comparison to General Benefits of Variability Literature

### Limitations
- amount of training
- no mechanism to account for sequence effects in models
- more rigorous model comparison to account for exams greater complexity
- knowledge of function

## Comparison to Project 1

### Differences between the tasks

There are a number of differences between Project 1's Hit The Target (HTT), and Project 2's Hit The Wall (HTW) tasks.

- Task Space Complexity: In HTW, the task space is also almost perfectly smooth, at least for the continuous feedback subjects, if they throw 100 units too hard, they'll be told that they were 100 units too hard. Whereas in HTT,  it was possible to produce xy velocity combinations that were technically closer to the empirical solution space than other throws, but which resulted in worse feedback due to striking the barrier.

- Perceptual Distinctiveness: HTT offers perceptually distinct varied conditions that directly relate to the task's demands, which may increase the sallience between training positions encounted by the varied group. In contrast, HTW's varied conditions differ only in the numerical values displayed, lacking the same level of perceptual differentiation. Conversely in HTW, the only difference between conditions for the varied group are the numbers displayed at the top of the screen which indicate the current target band(e.g. 800-1000, or 1000-1200)

- In HTW, our primary testing stage of interest has no feedback, whereas in HTT testing always included feedback (the intermittent testing in HTT expt 1 being the only exception). Of course, we do collect testing with feedback data at the end of HTW, but we haven't focused on that data at all in our modelling work thus far. It's also interesting to recall that the gap between varied and constant in HTW does seem to close substantially in the testing-with-feedback stage. The difference between no-feedback and feedback testing might be relevant if the benefits of variation have anything to do with improving subsequent learning (as opposed to subsequent immediate performance), OR if the benefits of constant training rely on having the most useful anchor, having the most useful anchor might be a lot less helpful if you're getting feedback from novel positions and can thus immediately begin to form position-specific anchors for the novelties, rather than relying on a training anchor.

- HTW and HTT both have a similar amount of training trials (~200), and thus the constant groups acquire a similar amount of experience with their single position/velocity in both experiments. However, the varied conditions in both HTT experiments train on 2 positions, whereas the varied group in HTW trains on 3 velocity bands. This means that in HTT the varied group gets half as much experience on any one position as the constant group, and in HTW they only get 1/3 as much experience in any one position. There are likely myriad ways in which this might impact the success of the varied group regardless of how you think the benefits of variation might be occurring, e.g. maybe they also need to develop a coherent anchor, maybe they need more experience in order to extract a function, or more experience in order to properly learn to tune their c parameter.