### 2024-04-04 Sections/combo1.qmd

---
title: "combo1"
author: "Thomas Gorman"
toc: true
toc-depth: 5
toc-expand: true
number-sections: true
number-depth: 5
strip-comments: true
cold-tools: true
echo: false
format: 
  html:
    column: 
      - screen-inset-right
  hugo-md:
    echo: false
    html-math-method: mathjax
    output-file: htw-combo1-hugo.md
  # hikmah-manuscript-pdf:
  #   echo: false
  #   output-file: "hikmah-m_combo.pdf"
  #   mainfont: "Linux Libertine O"
  #   mainfontoptions:
  #     - "Numbers=Proportional"
  #     - "Numbers=OldStyle"
  #   mathfont: "Libertinus Math"
  # hikmah-pdf:
  #   echo: false
  #   cite-method: citeproc
  #   output-file: "hikmah_combo.pdf"
  #   mainfont: "Linux Libertine O"
  #   mainfontoptions:
  #     - "Numbers=Proportional"
  #     - "Numbers=OldStyle"
  #   sansfont: "IBM Plex Sans"
  #   mathfont: "Libertinus Math"
  #   geometry:
  #     - top=1in     
  #     - bottom=1in  
  #     - left=1in    
  #     - right=1in  
  #     - heightrounded
  #   # linestretch: 1.05
  #   include-in-header:
  #     # The \everydisplay thing makes all math blocks use \small font
  #     text: |
  #       \usepackage{mathtools}
  #       \everydisplay\expandafter{\the\everydisplay\small}
  #       \usepackage{float}
  #       \floatplacement{figure}{H}
  #       \floatplacement{table}{H}
  gfm:
    echo: true
    output-file: e1-gfm.md
#toc: true
# toc-title: Contents
# toc-depth: 4
# toc-expand: 1
# execute:
#   echo: false
---

## Introduction

{{< include ../Sections/Intro.qmd >}}

## Methods

{{< include ../Sections/Methods.qmd >}}

## Experiment 1

{{< include ../Analysis/e1.qmd >}}

## Experiment 2

{{< include ../Analysis/e2.qmd >}}

## Experiment 3

{{< include ../Analysis/e3.qmd >}}

## Results Summary

## Computational Model

{{< include ../Model/htw_model.qmd >}}

## General Discussion

{{< include ../Sections/Discussion.qmd >}}

## References

::: {#refs}
:::