### 2024-04-09 Sections/Results.qmd
fix crossref - render combo pages

---
title: Results
author: "Thomas Gorman"
toc: true
toc-location: right
# number-sections: true
# number-depth: 5
# strip-comments: true
# cold-tools: true
echo: false
#format: 
  # html:
  #   column: 
  #     - screen-inset-right
  # hugo-md:
  #   echo: false
  #   html-math-method: mathjax
  #   output-file: full-results-hugo.md
  # gfm:
  #   echo: true
  #   output-file: e1-gfm.md
#toc: true
# toc-title: Contents
# toc-depth: 4
# toc-expand: 1
# execute:
#   echo: false
---

# Full

## Experiment 1

{{< include ../Analysis/e1.qmd >}}

## Experiment 2

{{< include ../Analysis/e2.qmd >}}

## Experiment 3

{{< include ../Analysis/e3.qmd >}}

## Results Summary

## Overall Summary of Experiments 1-3

Across three experiments, we investigated the impact of training variability on learning and extrapolation in a visuomotor function learning task.

In Experiment 1, participants in the varied training condition, who experienced a wider range of velocity bands during training, showed lower accuracy at the end of training compared to those in the constant training condition. Crucially, during the testing phase, the varied group exhibited significantly larger deviations from the target velocity bands, particularly for the extrapolation bands that were not encountered during training. The varied group also showed less discrimination between velocity bands, as evidenced by shallower slopes when predicting response velocity from target velocity band.

Experiment 2 extended these findings by reversing the order of the training and testing bands. Similar to Experiment 1, the varied group demonstrated poorer performance during both training and testing phases. However, unlike Experiment 1, the varied group did not show a significant difference in discrimination between bands compared to the constant group.

In Experiment 3, we provided only ordinal feedback  during training, in contrast to the continuous feedback provided in the previous experiments. Participants were assigned to both an order condition (original or reverse) and a training condition (constant or varied). The varied condition showed larger deviations at the end of training, consistent with the previous experiments. Interestingly, there was a significant interaction between training condition and band order, with the varied condition showing greater accuracy in the reverse order condition. During testing, the varied group once again exhibited larger deviations, particularly for the extrapolation bands. The reverse order conditions showed smaller deviations compared to the original order conditions. Discrimination between velocity bands was poorer for the varied group in the original order condition, but not in the reverse order condition.

## References

::: {#refs}
:::